### 2022-08-18 paper/paper.tex
Initial commit of paper

\documentclass[sigplan,10pt,anonymous,review]{acmart}
\settopmatter{printfolios=true,printccs=false,printacmref=false}

\usepackage[british]{babel}

\usepackage{mathpartir}

%\usepackage{unicode-math}
%\setmathfont{latinmodern-math.otf}
%\setmathfont[version=bold, FakeBold=2]{latinmodern-math.otf}

\usepackage{xcolor}

\usepackage{academicons}

\usepackage{tikz}
\usepackage{pgfplots}
\pgfplotsset{compat=1.16}

\usepackage{enumitem}

\setlength{\abovecaptionskip}{0pt}
\setlength{\textfloatsep}{18pt plus 2pt minus 4pt}

\usepackage{listings}
\definecolor{isarblue}{HTML}{006699}
\definecolor{isargreen}{HTML}{009966}
\lstdefinelanguage{isabelle}{%
    keywords=[1]{type_synonym,datatype,fun,abbreviation,definition,proof,lemma,theorem,corollary},
    keywordstyle=[1]\bfseries\color{isarblue},
    keywords=[2]{where,assumes,shows,and},
    keywordstyle=[2]\bfseries\color{isargreen},
    keywords=[3]{if,then,else,case,of,SOME,let,in,O},
    keywordstyle=[3]\color{isarblue},
}
\lstset{%
  language=isabelle,
  escapeinside={&}{&},
  columns=fixed,
  extendedchars,
  basewidth={0.5em,0.45em},
  basicstyle=\ttfamily,
  mathescape,
}
\makeatletter
\lst@AddToHook{OnEmptyLine}{\vspace{-0.4\baselineskip}}
\makeatother

\usepackage{calc}

\newcommand{\textover}[3][l]{%
 % #1 is the alignment, default l
 % #2 is the text to be printed
 % #3 is the text for setting the width
 \makebox[\widthof{#3}][#1]{#2}%
}

\hyphenation{Isa-belle}

\pagestyle{plain} % turn on page numbers in llncs

% % LuaLatex on Arxiv: https://tex.stackexchange.com/questions/372154/lualatex-how-to-produce-pdf-acceptable-by-arxiv 
% \hypersetup{%
%   pdfcreator = {},
%   pdfproducer = {}
% }
% \pdfvariable suppressoptionalinfo \numexpr 1+2+4+8+16+32+64+128+256+512 \relax

\begin{document}

\setcounter{tocdepth}{1}

\title{Verifying a Tableau Prover for Multi-Level-Syllogistics}
\author{Lukas Stevens}
\orcid{0000-0003-0222-6858}
\affiliation{%
  \institution{Technical University of Munich}
  \department{Department of Informatics}
  \streetaddress{Boltzmannstr. 3}
  \city{Garching}
  \postcode{85748}
  \country{Germany}
}
\email{lukas.stevens@in.tum.de}

\begin{abstract}
  TODO
\end{abstract}

\maketitle

\section{Introduction}
TODO

\subsection{Related work}
TODO

\subsection{Contributions}
TODO

\subsection{Notation}
Isabelle/HOL~\cite{isabelle} conforms to everyday mathematical notation for the most part.
For the benefit of the reader that is unfamiliar with Isabelle/HOL, we establish notation and in particular some essential datatypes together with their primitive operations that are specific to Isabelle/HOL.
We write \lstinline!t :: 'a! to specify that the term \lstinline!t! has the type \lstinline!'a! and \lstinline!'a $\Rightarrow$ 'b! for the type of a total function from \lstinline!'a! to \lstinline!'b!.
The types for booleans, natural numbers, and in integers are \lstinline!bool!, \lstinline!nat!, and \lstinline!int!, respectively.
Sets with elements of type \lstinline!'a! have the type \lstinline!'a set!.
Analogously, we use \lstinline!'a list! to describe lists, which are constructed as the empty list \lstinline![]! or with the infix constructor \lstinline!#!, and are appended with the infix operator \lstinline!@!.
The function \lstinline!set! converts a list into a set.
For optional values, Isabelle/HOL offers the type \lstinline!option! where a term \lstinline!opt :: 'a option! is either \lstinline!None! or \lstinline!Some a! with \lstinline!a :: 'a!.
Finally, we remark that \textbf{iff} is equivalent to \lstinline!$=$! on type \lstinline!bool! and \lstinline!$\equiv$! is definitional equality of the meta-logic of Isabelle/HOL, which is called Isabelle/Pure.

\section{A Semantics for Sets\label{sec:semantics}}
TODO

\section{Conclusion}
TODO

\begin{acks}
  TODO
\end{acks}

\bibliographystyle{ACM-Reference-Format}
\bibliography{sources}

\end{document}

### 2022-08-22 paper/paper.tex
Outline and some listings

@@
 \lstdefinelanguage{isabelle}{%
-    keywords=[1]{type_synonym,datatype,fun,abbreviation,definition,proof,lemma,theorem,corollary},
+    keywords=[1]{type_synonym,datatype,fun,function,abbreviation,definition,proof,lemma,theorem,corollary},
     keywordstyle=[1]\bfseries\color{isarblue},
@@
 
-\pagestyle{plain} % turn on page numbers in llncs
-
 % % LuaLatex on Arxiv: https://tex.stackexchange.com/questions/372154/lualatex-how-to-produce-pdf-acceptable-by-arxiv 
@@
 
-\title{Verifying a Tableau Prover for Multi-Level-Syllogistics}
+\title{Towards a Verified Tableau Prover for Multi-Level-Syllogistics}
 \author{Lukas Stevens}
@@
 \subsection{Related work}
-TODO
+AFP entry~\cite{hybrid_logic_afp} and workshop paper in Isabelle 2022 workshop.
 
@@
 is definitional equality of the meta-logic of Isabelle/HOL, which is called Isabelle/Pure.
+TODO: Add arrows and bracket notation for Pure formulae
+\section{A Semantics for Sets\label{sec:semantics}}
+\begin{lstlisting}
+datatype 'a fm =
+  Atom 'a | Neg ('a fm)
+| And ('a fm) ('a fm) | Or ('a fm) ('a fm)
+\end{lstlisting}
 
-\section{A Semantics for Sets\label{sec:semantics}}
+% TODO: Think about introducing notation for single
+\begin{lstlisting}
+datatype 'a pset_term =
+  $\emptyset$ | Var 'a | Single ('a pset_term)
+| 'a pset_term $\sqcup_\text{s}$ 'a pset_term
+| 'a pset_term $\sqcap_\text{s}$ 'a pset_term
+| 'a pset_term $-_\text{s}$ 'a pset_term
+\end{lstlisting}
+
+\begin{lstlisting}
+datatype 'a pset_atom =
+  'a pset_term $\in_\text{s}$ 'a pset_term
+| 'a pset_term $\approx_\text{s}$ 'a pset_term
+\end{lstlisting}
+
+\begin{lstlisting}
+type_synonym 'a pset_fm = 'a pset_atom fm
+type_synonym 'a branch = 'a pset_fm list
+\end{lstlisting}
+
+\begin{lstlisting}
+definition vars_fm :: 'a pset_fm $\Rightarrow$ 'a set where
+  vars_fm $\phi$ $\equiv$ $\bigcup$(set_pset_atom ` atoms $\phi$)
+\end{lstlisting}
+
+\begin{lstlisting}
+definition vars_branch :: 'a branch $\Rightarrow$ 'a set 
+  vars_branch b $\equiv$ $\bigcup$(vars_fm ` set b)
+\end{lstlisting}
+
+\begin{lstlisting}
+fun I$_\text{st}$ :: ('a $\Rightarrow$ V) $\Rightarrow$ 'a pset_term $\Rightarrow$ V where
+  I$_\text{st}$ v $\emptyset$
+| I$_\text{st}$ v (Var x) = v x
+| I$_\text{st}$ v (Single s) = vset {I$_\text{st}$ v s}
+| I$_\text{st}$ v (s $\sqcup_\text{s}$ t) = I$_\text{st}$ v s $\sqcup$ I$_\text{st}$ v t 
+| I$_\text{st}$ v (s $\sqcap_\text{s}$ t) = I$_\text{st}$ v s $\sqcap$ I$_\text{st}$ v t 
+| I$_\text{st}$ v (s $-_\text{s}$ t) = I$_\text{st}$ v s $-$ I$_\text{st}$ v t 
+\end{lstlisting}
+
+\begin{lstlisting}
+fun I$_\text{sa}$ :: ('a $\Rightarrow$ V) $\Rightarrow$ 'a pset_atom $\Rightarrow$ V where
+  I$_\text{sa}$ v (s $\in_\text{s}$ t) $\longleftrightarrow$ I$_\text{st}$ v s $\in$ elts (I$_\text{st}$ v t)
+| I$_\text{sa}$ v (s $\approx_\text{s}$ t) $\longleftrightarrow$ I$_\text{st}$ v s = I$_\text{st}$ v t
+\end{lstlisting}
+
+\begin{lstlisting}
+fun subterms :: 'a pset_term $\Rightarrow$ 'a pset_term set
+where
+  subterms $\emptyset$ = {$\emptyset$}
+| subterms (Var x) = {Var x}
+| subterms (s $\sqcup_\text{s}$ t) = {s $\sqcup_\text{s}$ t}
+    $\cup$ subterms s $\cup$ subterms t
+| subterms (s $\sqcap_\text{s}$ t) = {s $\sqcap_\text{s}$ t}
+    $\cup$ subterms s $\cup$ subterms t
+| subterms (s $-_\text{s}$ t) = {s $-\text{s}$ t}
+    $\cup$ subterms s $\cup$ subterms t
+| subterms (Single t) = {Single t} $\cup$ subterms t
+\end{lstlisting}
+
+\begin{lstlisting}
+fun subterms_atom ::
+      'a pset_atom $\Rightarrow$ 'a pset_term set where
+  subterms_atom (s $\in_\text{s}$ t) = subterms s $\cup$ subterms t
+| subterms_atom (s $\approx_\text{s}$ t) = subterms s $\cup$ subterms t
+\end{lstlisting}
+
+\begin{lstlisting}
+definition subterms_fm ::
+    'a pset_fm $\Rightarrow$ 'a pset_term set where
+  subterms_fm $\phi$ $\equiv$ $\bigcup$(subterms_atom ` atoms $\phi$)
+\end{lstlisting}
+
+\begin{lstlisting}
+definition subterms_branch ::
+    'a branch $\Rightarrow$ 'a pset_term set where
+  subterms_branch b $\equiv$ $\bigcup$(subterms_fm ` set b)
+\end{lstlisting}
+
+\section{A Tableau Calculus for MLSS}
+Interesting: they forgot one rule.
+
+\section{Completeness of the Calculus}
+\begin{lstlisting}
+definition lin_sat b $\equiv$
+  $\forall$b'. lextends b' b $\rightarrow$ set b' $\subseteq$ set b
+\end{lstlisting}
+
+\begin{lstlisting}
+definition sat b $\equiv$
+  lin_sat b $\land$ ($\nexists$bs'. fextends bs' b)
+\end{lstlisting}
+
+\subsection{Building a Model: Realisation of a Branch}
+Idea: keep this short. The formalisation matches the pen-and-paper proof quite closely.
+Show the results in Realisation of an Open Branch.
+
+\subsection{Formalising the Key Technical Lemma}
+Only one quite trivial invariant needed
+
+\section{An Abstract Specification of the Decision Procedure}
+Think about using the notion of coherent (or is this wfbranch?)
+\begin{figure*}
+  \centering
+\begin{lstlisting}
+function mlss_proc_branch :: 'a branch $\Rightarrow$ bool where
+  $\neg$ lin_sat b $\Longrightarrow$
+  mlss_proc_branch = mlss_proc_branch ((SOME b'. lextends b' b $\land$ set b $\subset$ set (b' @ b)) @ b)
+|  $\llbracket$ $\neg$ sat b; bopen b; lin_sat b $\rrbracket$ $\Longrightarrow$
+   mlss_proc_branch b = ($\forall$b' $\in$ (SOME bs. fextends bs b). mlss_proc_branch (b' @ b))
+| $\llbracket$ lin_sat b; sat b $\rrbracket$ $\Longrightarrow$ mlss_proc_branch b = bclosed b
+| $\llbracket$ lin_sat b; bclosed b $\rrbracket$ $\Longrightarrow$ mlss_proc_branch b = True
+
+definition mlss_proc :: 'a pset_fm $\Rightarrow$ bool where
+  mlss_proc $\phi$ $\equiv$ mlss_proc_branch [$\phi$]
+\end{lstlisting}
+\caption{Definition of the function \lstinline|mlss_proc_branch| and \lstinline|mlss_proc|.}
+\end{figure*}
+
+\subsection{Termination}
+Maybe interesting: technical lemma that allows to obtain the point at which a parameter was introduced.
+\subsection{Soundness}
+Interesting: The model changes because auf new parameters.
+Maybe move this to the chapter on the calculus?
+\subsection{Lifting the Completeness of the Calculus}
 TODO
+
+\section{Future Work}
+Challenges in obtaining an executable specification: membership cycle test unnecessary, fast branch membership test, proof reconstruction.
+Additional constructs such as finiteness and pairs.
+Optimisations for the algorithm.
 

### 2022-08-23 paper/paper.tex
Continue

@@
 
+\usepackage{todonotes}
+
 \begin{document}
@@
 \begin{abstract}
-  TODO
+  \todo[inline]{Write the abstract}
 \end{abstract}
@@
 
+\listoftodos{}
+
 \section{Introduction}
-TODO
+\todo[inline]{Write the introduction}
 
@@
 \subsection{Contributions}
-TODO
+\todo[inline]{Write the contributions}
 
@@
 is definitional equality of the meta-logic of Isabelle/HOL, which is called Isabelle/Pure.
-TODO: Add arrows and bracket notation for Pure formulae
-\section{A Semantics for Sets\label{sec:semantics}}
-\begin{lstlisting}
-datatype 'a fm =
-  Atom 'a | Neg ('a fm)
-| And ('a fm) ('a fm) | Or ('a fm) ('a fm)
-\end{lstlisting}
-
-% TODO: Think about introducing notation for single
+\todo[inline]{Add arrows and bracket notation for Pure formulae}
+\section{Syntax and Semantics of MLSS\label{sec:semantics}}
+\subsection{Syntax}
+At the heart of \textbf{MLSS}, we have the type of set terms which includes the empty set and variables as well as the operations union, intersection, set difference, and \lstinline!Single!, which represents a singleton set containing the argument.
+\todo[inline]{Syntax for \lstinline!Single!?}
 \begin{lstlisting}
@@
 \end{lstlisting}
-
+We can combine two set terms to form a set atom by using the membership or the equality operator.
 \begin{lstlisting}
@@
 \end{lstlisting}
+With the above operators we can also represent the subset operator \lstinline!$\sqsubseteq_\text{s}$!
+and enumerate finite sets as \lstinline!$s \subseteq_\text{s} t$!
+is equivalent to \lstinline!s $\sqcup_\text{s}$ t $\approx_\text{s}$ t!
+and a finite set of elements \lstinline!{t$_1$,t$_2$,$\ldots$,t$_k$}!
+can be expressed by \lstinline!Single t$_1$ $\sqcup_\text{s}$ $\ldots$ $\sqcup_\text{s}$ Single t$_k$!.
 
+The unquantified fragment \textbf{MLSS} of set theory is then formed by combining set atoms with propositional logical connectives.
 \begin{lstlisting}
+datatype (atoms: 'a) fm =
+  Atom 'a | Neg ('a fm)
+| And ('a fm) ('a fm) | Or ('a fm) ('a fm)
+
 type_synonym 'a pset_fm = 'a pset_atom fm
-type_synonym 'a branch = 'a pset_fm list
 \end{lstlisting}
-
+Note that Isabelle's datatype package automatically defines a function \lstinline!atoms :: 'a fm $\Rightarrow$ 'a set! that allows us to obtain all set atoms in the formula.
+Similarly, it defines a function \lstinline!set_pset_atom :: 'a pset_atom $\Rightarrow$ 'a set! with which we can extract all the variable names in a set atom.
+We combine these functions in order to define \lstinline!vars_fm! that extracts all the variable names occuring in a set formula.
 \begin{lstlisting}
@@
 \end{lstlisting}
+In addition to the variable names that appear in a formula, we will later also need to obtain those set terms that occur as subterms in a formula. 
+We first define a function that collects the subterms of a set term.
+\begin{lstlisting}
+fun subterms ::
+    'a pset_term $\Rightarrow$ 'a pset_term set where
+  subterms $\emptyset$ = {$\emptyset$}
+| subterms (Var x) = {Var x}
+| subterms (s $\sqcup_\text{s}$ t) =
+    {s $\sqcup_\text{s}$ t} $\cup$ subterms s $\cup$ subterms t
+| subterms (s $\sqcap_\text{s}$ t) =
+    {s $\sqcap_\text{s}$ t} $\cup$ subterms s $\cup$ subterms t
+| subterms (s $-_\text{s}$ t) =
+    {s $-_\text{s}$ t} $\cup$ subterms s $\cup$ subterms t
+| subterms (Single t) =
+    {Single t} $\cup$ subterms t
+\end{lstlisting}
+The subterms of an atom are then the subterms of those terms that make up the atom.
+\begin{lstlisting}
+fun subterms_atom ::
+      'a pset_atom $\Rightarrow$ 'a pset_term set where
+  subterms_atom (s $\in_\text{s}$ t) =
+    subterms s $\cup$ subterms t
+| subterms_atom (s $\approx_\text{s}$ t) =
+    subterms s $\cup$ subterms t
+\end{lstlisting}
+Ultimately, we lift this to formulae by using \lstinline!atoms! again.
+\begin{lstlisting}
+definition subterms_fm ::
+    'a pset_fm $\Rightarrow$ 'a pset_term set where
+  subterms_fm $\phi$ $\equiv$ $\bigcup$(subterms_atom ` atoms $\phi$)
+\end{lstlisting}
+The last function that we define computes the subformulae of a formula.
+\begin{lstlisting}
+fun subfms :: 'a fm $\Rightarrow$ 'a fm set where
+  subfms (Atom a) = {Atom a}
+| subfms (And p q) =
+    {And p q} $\cup$ subfms p $\cup$ subfms q
+| subfms (Or p q) =
+    {Or p q} $\cup$ subfms p $\cup$ subfms q
+| subfms (Neg q) = {Neg q} $\cup$ subfms q
+\end{lstlisting}
 
-\begin{lstlisting}
-definition vars_branch :: 'a branch $\Rightarrow$ 'a set 
-  vars_branch b $\equiv$ $\bigcup$(vars_fm ` set b)
-\end{lstlisting}
+\subsection{Semantics}
+Formulae in \textbf{MLSS} contain arbitrarily nested set terms.
+We account for this arbitrary nesting by basing the semantics on the von Neumann hierarchy $\mathcal{V}$ of sets that are defined inductively as
+  \[
+    \begin{array}{rclr}
+      \mathcal{V}_0 & = & \emptyset, \\
+      \mathcal{V}_{\alpha + 1} & = & \mathcal{P}(\mathcal{V}_\alpha) & \text{for each ordinal $\alpha$}, \\
+      \mathcal{V}_{\lambda} & = & \bigcup_{\mu < \lambda} \mathcal{V}_\mu & \text{for each limit ordinal $\lambda$}, \\
+      \mathcal{V} & = & \bigcup_\alpha \mathcal{V}_\alpha, \\
+    \end{array}
+  \]
+where $\mathcal{P}(S)$ is the powerset of $S$.
+The sets in $\mathcal{V}$ are well-founded meaning that there can be no membership cycle in $\mathcal{V}$.
+
+In Isabelle/HOL, which is simply typed, this definition is not accepted, though.
+A way to work around this limitiation is provided by Lawrence in an AFP entry~\cite{zfc_in_hol_afp}, which adds $\mathcal{V}$ to Isabelle/HOL by way of axiomatisation.
+An article by Lawrence~\cite{zfc_in_hol_afp} provides some further context to this formalisation.
 
@@
 fun I$_\text{st}$ :: ('a $\Rightarrow$ V) $\Rightarrow$ 'a pset_term $\Rightarrow$ V where
-  I$_\text{st}$ v $\emptyset$
-| I$_\text{st}$ v (Var x) = v x
-| I$_\text{st}$ v (Single s) = vset {I$_\text{st}$ v s}
-| I$_\text{st}$ v (s $\sqcup_\text{s}$ t) = I$_\text{st}$ v s $\sqcup$ I$_\text{st}$ v t 
-| I$_\text{st}$ v (s $\sqcap_\text{s}$ t) = I$_\text{st}$ v s $\sqcap$ I$_\text{st}$ v t 
-| I$_\text{st}$ v (s $-_\text{s}$ t) = I$_\text{st}$ v s $-$ I$_\text{st}$ v t 
+  I$_\text{st}$ M $\emptyset$
+| I$_\text{st}$ M (Var x) = M x
+| I$_\text{st}$ M (Single s) = vset {I$_\text{st}$ M s}
+| I$_\text{st}$ M (s $\sqcup_\text{s}$ t) = I$_\text{st}$ M s $\sqcup$ I$_\text{st}$ M t 
+| I$_\text{st}$ M (s $\sqcap_\text{s}$ t) = I$_\text{st}$ M s $\sqcap$ I$_\text{st}$ M t 
+| I$_\text{st}$ M (s $-_\text{s}$ t) = I$_\text{st}$ M s $-$ I$_\text{st}$ M t 
 \end{lstlisting}
@@
 fun I$_\text{sa}$ :: ('a $\Rightarrow$ V) $\Rightarrow$ 'a pset_atom $\Rightarrow$ V where
-  I$_\text{sa}$ v (s $\in_\text{s}$ t) $\longleftrightarrow$ I$_\text{st}$ v s $\in$ elts (I$_\text{st}$ v t)
-| I$_\text{sa}$ v (s $\approx_\text{s}$ t) $\longleftrightarrow$ I$_\text{st}$ v s = I$_\text{st}$ v t
+  I$_\text{sa}$ M (s $\in_\text{s}$ t) $\longleftrightarrow$ I$_\text{st}$ M s $\in$ elts (I$_\text{st}$ M t)
+| I$_\text{sa}$ M (s $\approx_\text{s}$ t) $\longleftrightarrow$ I$_\text{st}$ M s = I$_\text{st}$ M t
 \end{lstlisting}
 
+\section{A Tableau Calculus for MLSS}
 \begin{lstlisting}
-fun subterms :: 'a pset_term $\Rightarrow$ 'a pset_term set
-where
-  subterms $\emptyset$ = {$\emptyset$}
-| subterms (Var x) = {Var x}
-| subterms (s $\sqcup_\text{s}$ t) = {s $\sqcup_\text{s}$ t}
-    $\cup$ subterms s $\cup$ subterms t
-| subterms (s $\sqcap_\text{s}$ t) = {s $\sqcap_\text{s}$ t}
-    $\cup$ subterms s $\cup$ subterms t
-| subterms (s $-_\text{s}$ t) = {s $-\text{s}$ t}
-    $\cup$ subterms s $\cup$ subterms t
-| subterms (Single t) = {Single t} $\cup$ subterms t
-\end{lstlisting}
-
-\begin{lstlisting}
-fun subterms_atom ::
-      'a pset_atom $\Rightarrow$ 'a pset_term set where
-  subterms_atom (s $\in_\text{s}$ t) = subterms s $\cup$ subterms t
-| subterms_atom (s $\approx_\text{s}$ t) = subterms s $\cup$ subterms t
-\end{lstlisting}
-
-\begin{lstlisting}
-definition subterms_fm ::
-    'a pset_fm $\Rightarrow$ 'a pset_term set where
-  subterms_fm $\phi$ $\equiv$ $\bigcup$(subterms_atom ` atoms $\phi$)
+type_synonym 'a branch = 'a pset_fm list
 \end{lstlisting}
@@
 
-\section{A Tableau Calculus for MLSS}
+\begin{lstlisting}
+definition vars_branch :: 'a branch $\Rightarrow$ 'a set 
+  vars_branch b $\equiv$ $\bigcup$(vars_fm ` set b)
+\end{lstlisting}
+
 Interesting: they forgot one rule.
@@
   lin_sat b $\land$ ($\nexists$bs'. fextends bs' b)
+\end{lstlisting}
+
+\begin{lstlisting}
+definition wf_branch b $\equiv$ $\exists$$\phi$. extendss b [$\phi$]
+\end{lstlisting}
+
+\begin{lstlisting}
+definition params :: 'a branch $\Rightarrow$ 'a set where
+  params b $\equiv$ vars_branch b - vars_fm (last b)
+\end{lstlisting}
+
+\begin{lstlisting}
+definition params' :: 'a branch $\Rightarrow$ 'a set where
+  params' b $\equiv$
+    {c $\in$ params b. $\forall$t $\in$ subterms_fm (last b).
+      AT (Var c $\approx_\text{s}$ t) $\notin$ set b $\land$
+      AT (t $\approx_\text{s}$ Var c) $\notin$ set b} 
+\end{lstlisting}
+
+\begin{lstlisting}
+definition subterms_branch' ::
+    'a branch $\Rightarrow$ 'a pset_term set where
+  subterms_branch' b $\equiv$
+    subterms_fm (last b) $\cup$ 
+    Var ` (params b - params' b)
+\end{lstlisting}
+
+\begin{lstlisting}
+definition bgraph $\equiv$ $\ldots$ 
+\end{lstlisting}
+
+\begin{lstlisting}
+definition no_new_subterms b $\equiv$
+  $\forall$t $\in$ subterms_branch b. t $\notin$ Var ` params b
+    $\longrightarrow$ t $\in$ subterms_fm (last b))
 \end{lstlisting}
@@
 \subsection{Lifting the Completeness of the Calculus}
-TODO
+\todo[inline]{Do you even lift?}
 
@@
 \section{Conclusion}
-TODO
+\todo[inline]{Conclude}
 
 \begin{acks}
-  TODO
+  \todo[inline]{Acknowledge}
 \end{acks}

### 2022-08-24 paper/paper.tex
Finish semantics section

@@
 Formulae in \textbf{MLSS} contain arbitrarily nested set terms.
-We account for this arbitrary nesting by basing the semantics on the von Neumann hierarchy $\mathcal{V}$ of sets that are defined inductively as
+We account for this arbitrary nesting by basing the semantics on the von Neumann hierarchy $\mathcal{V}$ of sets that is defined inductively as
   \[
@@
 A way to work around this limitiation is provided by Lawrence in an AFP entry~\cite{zfc_in_hol_afp}, which adds $\mathcal{V}$ to Isabelle/HOL by way of axiomatisation.
-An article by Lawrence~\cite{zfc_in_hol_afp} provides some further context to this formalisation.
+An article by Lawrence~\cite{zfc_in_hol} provides some further context to this entry.
+The entry declares a type \lstinline!V! that is equipped with a function \lstinline!elts :: V $\Rightarrow$ V set! that maps a set to its elements.
+Even though we use the typed sets provided by Isabelle/HOL here, \lstinline!elts! actually returns a class.
+Only if the class is small enough, as indicated by the polymorphic predicate \lstinline!small :: 'a set $\Rightarrow$ bool!, it is indeed a set.
+Inversely, this predicate characterises those sets that can be turned into a set of type \lstinline!V! using the polymorphic function \lstinline!vset :: 'a set $\Rightarrow$ V!.
+The type \lstinline!V! comes with all the usual set operations such as equality ($=$), union ($\sqcup$), intersection ($\sqcap$), and difference ($-$).
+The empty set coincides with the ordinal $0$ so we denote it with \lstinline!0 :: V!.
 
+Equipped with the above, we define the interpretation function \lstinline!I$_\text{st}$!
+for set terms that interprets a set term with respect to a valuation function \lstinline!M :: 'a $\Rightarrow$ V!
+for variable.
 \begin{lstlisting}
 fun I$_\text{st}$ :: ('a $\Rightarrow$ V) $\Rightarrow$ 'a pset_term $\Rightarrow$ V where
-  I$_\text{st}$ M $\emptyset$
+  I$_\text{st}$ M $\emptyset$ = 0
 | I$_\text{st}$ M (Var x) = M x
@@
 \end{lstlisting}
-
+The interpretation function \lstinline!I$_\text{sa}$! for set atoms is straightforward as well.
 \begin{lstlisting}
@@
 \end{lstlisting}
+For formulae, we write \lstinline!M $\models$ $\phi$! for the judgement that $\phi$ holds under the valuation function \lstinline!M!.
+We omit the implementation of $\models$ for brevity.
 

### 2022-08-25 paper/paper.tex
Continue

@@
 \lstdefinelanguage{isabelle}{%
-    keywords=[1]{type_synonym,datatype,fun,function,abbreviation,definition,proof,lemma,theorem,corollary},
+    keywords=[1]{type_synonym,datatype,fun,function,abbreviation,definition,proof,lemma,theorem,corollary,inductive},
     keywordstyle=[1]\bfseries\color{isarblue},
@@
 \usepackage{todonotes}
+
+% Commands
+\newlength{\trianglewidth}
+\settowidth{\trianglewidth}{\(\triangleleft\)}
+
+\newcommand{\lefttrianglebar}{\mathrel{\tikz[baseline]{\draw (1ex, 0.75ex) -- (0, 1.25ex) -- (0, 0.25ex) -- cycle; \draw (0, 0.75ex) -- (1ex, 0.75ex);}}}
+\newcommand{\lefttriangle}{\mathrel{\tikz[baseline]{\draw (1ex, 0.75ex) -- (0, 1.25ex) -- (0, 0.25ex) -- cycle;}}}
+
+\newcommand{\sextends}{\lefttriangle}
+\newcommand{\fextends}{\lefttrianglebar}
+
+\newcommand{\unionS}{\sqcup_\text{s}}
+\newcommand{\interS}{\sqcap_\text{s}}
+\newcommand{\diffS}{-_\text{s}}
+\newcommand{\inS}{\in_\text{s}}
+\newcommand{\notinS}{\notin_\text{s}}
+\newcommand{\eqS}{\approx_\text{s}}
+\newcommand{\neqS}{{\not\approx}_\text{s}}
+\newcommand{\subseteqS}{\sqsubseteq_\text{s}}
+\newcommand{\Ist}{I$_\text{st}$}
+\newcommand{\Isa}{I$_\text{sa}$}
 
@@
   $\emptyset$ | Var 'a | Single ('a pset_term)
-| 'a pset_term $\sqcup_\text{s}$ 'a pset_term
-| 'a pset_term $\sqcap_\text{s}$ 'a pset_term
-| 'a pset_term $-_\text{s}$ 'a pset_term
+| 'a pset_term $\unionS$ 'a pset_term
+| 'a pset_term $\interS$ 'a pset_term
+| 'a pset_term $\diffS$ 'a pset_term
 \end{lstlisting}
@@
 datatype 'a pset_atom =
-  'a pset_term $\in_\text{s}$ 'a pset_term
-| 'a pset_term $\approx_\text{s}$ 'a pset_term
+  'a pset_term $\inS$ 'a pset_term
+| 'a pset_term $\eqS$ 'a pset_term
 \end{lstlisting}
-With the above operators we can also represent the subset operator \lstinline!$\sqsubseteq_\text{s}$!
-and enumerate finite sets as \lstinline!$s \subseteq_\text{s} t$!
-is equivalent to \lstinline!s $\sqcup_\text{s}$ t $\approx_\text{s}$ t!
+With the above operators we can also represent the subset operator \lstinline!$\subseteqS$!
+and enumerate finite sets as \lstinline!s $\subseteqS$ t!
+is equivalent to \lstinline!s $\unionS$ t $\eqS$ t!
 and a finite set of elements \lstinline!{t$_1$,t$_2$,$\ldots$,t$_k$}!
-can be expressed by \lstinline!Single t$_1$ $\sqcup_\text{s}$ $\ldots$ $\sqcup_\text{s}$ Single t$_k$!.
+can be expressed by \lstinline!Single t$_1$ $\unionS$ $\ldots$ $\unionS$ Single t$_k$!.
 
@@
 \end{lstlisting}
+We will sometimes drop the \lstinline!Atom! constructor to improve presentation.
+Additionally, we use the abbreviations \lstinline!s $\notinS$ t! and \lstinline!s $\neqS$ t! to denote \lstinline!Neg (Atom (s $\inS$ t))! and \lstinline!Neg (Atom (s $\eqS$ t))!, respectively.
 Note that Isabelle's datatype package automatically defines a function \lstinline!atoms :: 'a fm $\Rightarrow$ 'a set! that allows us to obtain all set atoms in the formula.
@@
 | subterms (Var x) = {Var x}
-| subterms (s $\sqcup_\text{s}$ t) =
-    {s $\sqcup_\text{s}$ t} $\cup$ subterms s $\cup$ subterms t
-| subterms (s $\sqcap_\text{s}$ t) =
-    {s $\sqcap_\text{s}$ t} $\cup$ subterms s $\cup$ subterms t
-| subterms (s $-_\text{s}$ t) =
-    {s $-_\text{s}$ t} $\cup$ subterms s $\cup$ subterms t
+| subterms (s $\unionS$ t) =
+    {s $\unionS$ t} $\cup$ subterms s $\cup$ subterms t
+| subterms (s $\interS$ t) =
+    {s $\interS$ t} $\cup$ subterms s $\cup$ subterms t
+| subterms (s $\diffS$ t) =
+    {s $\diffS$ t} $\cup$ subterms s $\cup$ subterms t
 | subterms (Single t) =
@@
       'a pset_atom $\Rightarrow$ 'a pset_term set where
-  subterms_atom (s $\in_\text{s}$ t) =
+  subterms_atom (s $\inS$ t) =
     subterms s $\cup$ subterms t
-| subterms_atom (s $\approx_\text{s}$ t) =
+| subterms_atom (s $\eqS$ t) =
     subterms s $\cup$ subterms t
@@
 
-Equipped with the above, we define the interpretation function \lstinline!I$_\text{st}$!
+Equipped with the above, we define the interpretation function \lstinline!I_$\text{st}$!
 for set terms that interprets a set term with respect to a valuation function \lstinline!M :: 'a $\Rightarrow$ V!
@@
 \begin{lstlisting}
-fun I$_\text{st}$ :: ('a $\Rightarrow$ V) $\Rightarrow$ 'a pset_term $\Rightarrow$ V where
-  I$_\text{st}$ M $\emptyset$ = 0
-| I$_\text{st}$ M (Var x) = M x
-| I$_\text{st}$ M (Single s) = vset {I$_\text{st}$ M s}
-| I$_\text{st}$ M (s $\sqcup_\text{s}$ t) = I$_\text{st}$ M s $\sqcup$ I$_\text{st}$ M t 
-| I$_\text{st}$ M (s $\sqcap_\text{s}$ t) = I$_\text{st}$ M s $\sqcap$ I$_\text{st}$ M t 
-| I$_\text{st}$ M (s $-_\text{s}$ t) = I$_\text{st}$ M s $-$ I$_\text{st}$ M t 
+fun &\Ist& :: ('a $\Rightarrow$ V) $\Rightarrow$ 'a pset_term $\Rightarrow$ V where
+  &\Ist& M $\emptyset$ = 0
+| &\Ist& M (Var x) = M x
+| &\Ist& M (Single s) = vset {&\Ist& M s}
+| &\Ist& M (s $\unionS$ t) = &\Ist& M s $\sqcup$ &\Ist& M t 
+| &\Ist& M (s $\interS$ t) = &\Ist& M s $\sqcap$ &\Ist& M t 
+| &\Ist& M (s $\diffS$ t) = &\Ist& M s $-$ &\Ist& M t 
 \end{lstlisting}
@@
 \begin{lstlisting}
-fun I$_\text{sa}$ :: ('a $\Rightarrow$ V) $\Rightarrow$ 'a pset_atom $\Rightarrow$ V where
-  I$_\text{sa}$ M (s $\in_\text{s}$ t) $\longleftrightarrow$ I$_\text{st}$ M s $\in$ elts (I$_\text{st}$ M t)
-| I$_\text{sa}$ M (s $\approx_\text{s}$ t) $\longleftrightarrow$ I$_\text{st}$ M s = I$_\text{st}$ M t
+fun &\Isa& :: ('a $\Rightarrow$ V) $\Rightarrow$ 'a pset_atom $\Rightarrow$ bool
+where
+  &\Isa& M (s $\inS$ t) $\longleftrightarrow$ &\Ist& M s $\in$ elts (&\Ist& M t)
+| &\Isa& M (s $\eqS$ t) $\longleftrightarrow$ &\Ist& M s = &\Ist& M t
 \end{lstlisting}
@@
 \section{A Tableau Calculus for MLSS}
+$A \fextends B \sextends C$
+
 \begin{lstlisting}
@@
   vars_branch b $\equiv$ $\bigcup$(vars_fm ` set b)
+\end{lstlisting}
+
+\begin{lstlisting}
+inductive bclosed :: 'a branch $\Rightarrow$ bool where
+  $\llbracket$ $\phi$ $\in$ set b; Neg $\phi$ $\in$ set b $\rrbracket$ $\Longrightarrow$ bclosed b
+| (t $\inS$ $\emptyset$) $\in$ set b $\Longrightarrow$ bclosed b
+| (t $\neqS$ t) $\in$ set b $\Longrightarrow$ bclosed b
+| $\llbracket$ member_cycle cs; set cs $\subseteq$ set b$\rrbracket$
+  $\Longrightarrow$ bclosed b
 \end{lstlisting}
@@
     {c $\in$ params b. $\forall$t $\in$ subterms_fm (last b).
-      AT (Var c $\approx_\text{s}$ t) $\notin$ set b $\land$
-      AT (t $\approx_\text{s}$ Var c) $\notin$ set b} 
+      AT (Var c $\eqS$ t) $\notin$ set b $\land$
+      AT (t $\eqS$ Var c) $\notin$ set b} 
 \end{lstlisting}

### 2022-08-26 paper/paper.tex
Continue with paper

@@
 \usepackage{xcolor}
-
-\usepackage{academicons}
 
@@
 \makeatother
+% This is a hack to allow escaping inside lstinline (https://tex.stackexchange.com/questions/43526/escaping-in-lstinline)
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd{\lsthk@TextStyle}{\let\lst@DefEsc\@empty}{}{}{\errmessage{failed to patch}}
+\makeatother
 
@@
 % Commands
+\newcommand{\MLSS}{\textbf{MLSS}}
 \newlength{\trianglewidth}
@@
 
-\newcommand{\sextends}{\lefttriangle}
-\newcommand{\fextends}{\lefttrianglebar}
+\newcommand{\sexpands}[2]{#1 $\lefttriangle$ #2}
+\newcommand{\fexpands}[2]{#1 $\lefttrianglebar$ #2}
+\newcommand{\expandss}[2]{#1 $\lefttriangle^*$ #2}
 
@@
 \newcommand{\Isa}{I$_\text{sa}$}
+
+\newcommand{\fmAnd}[2]{#1 $\boldsymbol{\land}$ #2}
+\newcommand{\fmOr}[2]{#1 $\boldsymbol{\lor}$ #2}
+\newcommand{\fmNeg}[1]{$\boldsymbol{\neg}$ #1}
+\newcommand{\fmAtom}{\textbf{A}}
 
@@
 \subsection{Syntax}
-At the heart of \textbf{MLSS}, we have the type of set terms which includes the empty set and variables as well as the operations union, intersection, set difference, and \lstinline!Single!, which represents a singleton set containing the argument.
+At the heart of \MLSS{}, we have the type of set terms which includes the empty set and variables as well as the operations union, intersection, set difference, and \lstinline!Single!, which represents a singleton set containing the argument.
 \todo[inline]{Syntax for \lstinline!Single!?}
@@
 
-The unquantified fragment \textbf{MLSS} of set theory is then formed by combining set atoms with propositional logical connectives.
+The unquantified fragment \MLSS{} of set theory is then formed by combining set atoms, which are injected into formulae with the constructor \lstinline!&\fmAtom&!, with propositional logical connectives.
+\todo[inline]{Cite Nipkow's Linear Quantifier Elimination?}
 \begin{lstlisting}
 datatype (atoms: 'a) fm =
-  Atom 'a | Neg ('a fm)
-| And ('a fm) ('a fm) | Or ('a fm) ('a fm)
+  &\fmAtom& 'a | &\fmNeg{('a fm)}&
+| &\fmAnd{'a fm}{'a fm}& | &\fmOr{'a fm}{'a fm}&
 
@@
 \end{lstlisting}
-We will sometimes drop the \lstinline!Atom! constructor to improve presentation.
-Additionally, we use the abbreviations \lstinline!s $\notinS$ t! and \lstinline!s $\neqS$ t! to denote \lstinline!Neg (Atom (s $\inS$ t))! and \lstinline!Neg (Atom (s $\eqS$ t))!, respectively.
-Note that Isabelle's datatype package automatically defines a function \lstinline!atoms :: 'a fm $\Rightarrow$ 'a set! that allows us to obtain all set atoms in the formula.
+We will sometimes drop the \lstinline!Atom! constructor to reduce clutter.
+Additionally, we use the abbreviations \lstinline!s $\notinS$ t!
+and \lstinline!s $\neqS$ t!
+to denote \lstinline!&\fmNeg{\fmAtom{} (s $\inS$ t)}&!
+and \lstinline!&\fmNeg{\fmAtom{} (s $\eqS$ t)}&!, respectively.
+Note that Isabelle's datatype package automatically defines a function \lstinline!atoms :: 'a fm $\Rightarrow$ 'a set! that allows us to obtain all set atoms in a formula.
 Similarly, it defines a function \lstinline!set_pset_atom :: 'a pset_atom $\Rightarrow$ 'a set! with which we can extract all the variable names in a set atom.
 We combine these functions in order to define \lstinline!vars_fm! that extracts all the variable names occuring in a set formula.
+\todo[inline]{Motivation: one rule introduces new variables}
 \begin{lstlisting}
@@
 We first define a function that collects the subterms of a set term.
+\todo[inline]{Motivation: saturating rules don't introduce subterms}
 \begin{lstlisting}
@@
 fun subfms :: 'a fm $\Rightarrow$ 'a fm set where
-  subfms (Atom a) = {Atom a}
-| subfms (And p q) =
-    {And p q} $\cup$ subfms p $\cup$ subfms q
-| subfms (Or p q) =
-    {Or p q} $\cup$ subfms p $\cup$ subfms q
-| subfms (Neg q) = {Neg q} $\cup$ subfms q
+  subfms (&\fmAtom& a) = {&\fmAtom& a}
+| subfms (&\fmAnd{p}{q}&) =
+    {&\fmAnd{p}{q}&} $\cup$ subfms p $\cup$ subfms q
+| subfms (&\fmOr{p}{q}&) =
+    {&\fmOr{p}{q}&} $\cup$ subfms p $\cup$ subfms q
+| subfms (&\fmNeg{q}&) = {&\fmNeg{q}&} $\cup$ subfms q
 \end{lstlisting}
@@
 \subsection{Semantics}
-Formulae in \textbf{MLSS} contain arbitrarily nested set terms.
+Formulae in \MLSS{} contain arbitrarily nested set terms.
 We account for this arbitrary nesting by basing the semantics on the von Neumann hierarchy $\mathcal{V}$ of sets that is defined inductively as
+\todo[inline]{Simplify definition by deleting cases?}
   \[
@@
 
-Equipped with the above, we define the interpretation function \lstinline!I_$\text{st}$!
+Equipped with the above, we define the interpretation function \lstinline!&\Ist&!
 for set terms that interprets a set term with respect to a valuation function \lstinline!M :: 'a $\Rightarrow$ V!
@@
 \section{A Tableau Calculus for MLSS}
-$A \fextends B \sextends C$
-
+We formalise the tableau calculus for \MLSS{} as described by \citeauthor{new_fast_tableau}~\cite{new_fast_tableau}.
+Inspired by the formalisation of a tableau calculus for Hybrid Logic by \citeauthor{hybrid_logic_afp}~\cite{hybrid_logic_afp}, we simply use a list to represent a branch of the tableau tree.
+Note that formulae are added to the front of the list during branch expansion so \lstinline!last b! for a branch \lstinline!b! is always the formula that we are trying to disprove with the tableau. 
 \begin{lstlisting}
@@
 \end{lstlisting}
-
+We now lift the functions that extract the variables respectively subterms of formulae to branches.
 \begin{lstlisting}
@@
   subterms_branch b $\equiv$ $\bigcup$(subterms_fm ` set b)
-\end{lstlisting}
 
-\begin{lstlisting}
 definition vars_branch :: 'a branch $\Rightarrow$ 'a set 
@@
 \end{lstlisting}
-
+In the standard tableau calculus for propositional logic, a branch is called \textit{closed} if it contains both the negation of a formula as well as the formula itself and it is called \textit{open} if it is not closed.
+We extend the notion of closedness with three additional rules; the first two are straightforward while the last one states that a branch is closed when the branch contains a membership cycle \lstinline!s $\inS$ t$_\text{1}$, t$_\text{1}$ $\inS$ t$_\text{2}$, $\ldots$, t$_\text{k}$ $\inS$ s!.
 \begin{lstlisting}
 inductive bclosed :: 'a branch $\Rightarrow$ bool where
-  $\llbracket$ $\phi$ $\in$ set b; Neg $\phi$ $\in$ set b $\rrbracket$ $\Longrightarrow$ bclosed b
+  $\llbracket$ $\phi$ $\in$ set b; &\fmNeg{$\phi$}& $\in$ set b $\rrbracket$ $\Longrightarrow$ bclosed b
 | (t $\inS$ $\emptyset$) $\in$ set b $\Longrightarrow$ bclosed b
@@
     $\Longrightarrow$ bclosed b
+
+abbreviation bopen b $\equiv$ $\neg$ bclosed b
+\end{lstlisting}
+A tableau is called \textit{closed} if all of its branches are closed.
+
+\subsection{Saturating Expansion Rules}
+The calculus considers two kinds of expansion rules: \textit{saturating} and \textit{fulfilling}.
+Whereas fulfilling expansion rules lead to the creation of new branches in the tableau, saturating rules only extend a branch \lstinline!b!
+with new formulae \lstinline[breaklines=true]!b' = [$\psi_1$,$\ldots$,$\psi_n$]!, which we denote by \lstinline!&\sexpands{b'}{b}&!.
+Furthermore, saturating rules are restricted such that no new subterm may be created by their application.
+For example, the rule
+\begin{center}
+\lstinline!(s $\inS$ t$_\text{1}$) $\in$ set b $\Longrightarrow$ &\sexpands{[s $\inS$ t$_\text{1}$ $\unionS$ t$_\text{2}$]}{b}&!
+\end{center}
+may only be used under the condition
+\begin{center}
+  \lstinline!t$_\text{1}$ $\unionS$ t$_\text{2}$ $\in$ subterms_fm (last b)!.
+\end{center}
+This ensures that repeated application of saturating rules eventually results in a \textit{linearly saturated} branch, i.e.\ a branch where no application of saturating rules would produce new formulae.
+In Isabelle, we express this as follows.
+\begin{lstlisting}
+definition lin_sat b $\equiv$
+  $\forall$b'. &\sexpands{b'}{b}& $\longrightarrow$ set b' $\subseteq$ set b
+\end{lstlisting}
+\todo[inline]{Reference figure and mention missing rule + example}
+
+\subsection{Fulfilling Expansion Rules\label{sec:fulfilling}}
+Since the number of branches doubles each time we apply a fulfilling rule, it is prudent to only apply them to open, linearly saturated branches. 
+To further prevent unnecessary branching, each rule has a \textit{subsumption} condition.
+
+\begin{lstlisting}
+definition sat b $\equiv$ lin_sat b $\land$ ($\nexists$bs'. &\fexpands{bs'}{b}&)
 \end{lstlisting}
 
-Interesting: they forgot one rule.
-
 \section{Completeness of the Calculus}
-\begin{lstlisting}
-definition lin_sat b $\equiv$
-  $\forall$b'. lextends b' b $\rightarrow$ set b' $\subseteq$ set b
-\end{lstlisting}
 
 \begin{lstlisting}
-definition sat b $\equiv$
-  lin_sat b $\land$ ($\nexists$bs'. fextends bs' b)
-\end{lstlisting}
-
-\begin{lstlisting}
-definition wf_branch b $\equiv$ $\exists$$\phi$. extendss b [$\phi$]
+definition wf_branch b $\equiv$ $\exists$$\phi$. &\expandss{b}{[$\phi$]}&
 \end{lstlisting}
@@
   $\neg$ lin_sat b $\Longrightarrow$
-  mlss_proc_branch = mlss_proc_branch ((SOME b'. lextends b' b $\land$ set b $\subset$ set (b' @ b)) @ b)
+  mlss_proc_branch = mlss_proc_branch ((SOME b'. &\sexpands{b'}{b}& $\land$ set b $\subset$ set (b' @ b)) @ b)
 |  $\llbracket$ $\neg$ sat b; bopen b; lin_sat b $\rrbracket$ $\Longrightarrow$
-   mlss_proc_branch b = ($\forall$b' $\in$ (SOME bs. fextends bs b). mlss_proc_branch (b' @ b))
+   mlss_proc_branch b = ($\forall$b' $\in$ (SOME bs. &\fexpands{bs}{b}&). mlss_proc_branch (b' @ b))
 | $\llbracket$ lin_sat b; sat b $\rrbracket$ $\Longrightarrow$ mlss_proc_branch b = bclosed b

### 2022-08-29 paper/paper.tex
Continue

@@
 
-\newcommand{\sexpands}[2]{#1 $\lefttriangle$ #2}
+\newcommand{\lexpands}[2]{#1 $\lefttriangle$ #2}
 \newcommand{\fexpands}[2]{#1 $\lefttrianglebar$ #2}
@@
 We first define a function that collects the subterms of a set term.
-\todo[inline]{Motivation: saturating rules don't introduce subterms}
+\todo[inline]{Motivation: linear rules don't introduce subterms}
 \begin{lstlisting}
@@
 The type \lstinline!V! comes with all the usual set operations such as equality ($=$), union ($\sqcup$), intersection ($\sqcap$), and difference ($-$).
-The empty set coincides with the ordinal $0$ so we denote it with \lstinline!0 :: V!.
+The empty set coincides with the ordinal $0$ so it is denoted by \lstinline!0 :: V!.
 
@@
 \end{lstlisting}
-In the standard tableau calculus for propositional logic, a branch is called \textit{closed} if it contains both the negation of a formula as well as the formula itself and it is called \textit{open} if it is not closed.
-We extend the notion of closedness with three additional rules; the first two are straightforward while the last one states that a branch is closed when the branch contains a membership cycle \lstinline!s $\inS$ t$_\text{1}$, t$_\text{1}$ $\inS$ t$_\text{2}$, $\ldots$, t$_\text{k}$ $\inS$ s!.
+In the standard tableau calculus for propositional logic, a branch is called \textit{closed} if it contains both the negation of a formula and the formula itself;
+conversely, it is called \textit{open} if it is not closed.
+For MLSS, we extend the notion of closedness with three additional rules; the first two are straightforward while the last one states that a branch is closed when the branch contains a membership cycle \lstinline!s $\inS$ t$_\text{1}$, t$_\text{1}$ $\inS$ t$_\text{2}$, $\ldots$, t$_\text{k}$ $\inS$ s!.
 \begin{lstlisting}
@@
 
-\subsection{Saturating Expansion Rules}
-The calculus considers two kinds of expansion rules: \textit{saturating} and \textit{fulfilling}.
-Whereas fulfilling expansion rules lead to the creation of new branches in the tableau, saturating rules only extend a branch \lstinline!b!
-with new formulae \lstinline[breaklines=true]!b' = [$\psi_1$,$\ldots$,$\psi_n$]!, which we denote by \lstinline!&\sexpands{b'}{b}&!.
-Furthermore, saturating rules are restricted such that no new subterm may be created by their application.
-For example, the rule
+\subsection{Linear Expansion Rules}
+The calculus considers two kinds of branch expansion rules: \textit{linear} and \textit{branching} rules.
+As the name suggests, branching rules lead to the creation of new branches in the tableau while linear rules only extend a branch \lstinline!b!
+with new formulae \lstinline[breaklines=true]!b' = [$\psi_1$,$\ldots$,$\psi_n$]!, which we denote by \lstinline!&\lexpands{b'}{b}&!.
+Furthermore, linear rules are restricted such that no new subterm may be created by their application:
+for example, the rule
 \begin{center}
-\lstinline!(s $\inS$ t$_\text{1}$) $\in$ set b $\Longrightarrow$ &\sexpands{[s $\inS$ t$_\text{1}$ $\unionS$ t$_\text{2}$]}{b}&!
+\lstinline!(s $\inS$ t$_\text{1}$) $\in$ set b $\Longrightarrow$ &\lexpands{[s $\inS$ t$_\text{1}$ $\unionS$ t$_\text{2}$]}{b}&!
 \end{center}
@@
 \end{center}
-This ensures that repeated application of saturating rules eventually results in a \textit{linearly saturated} branch, i.e.\ a branch where no application of saturating rules would produce new formulae.
+\todo[inline]{Explain this restriction in more detail.}
+This ensures that repeated application of linear rules eventually results in a \textit{linearly saturated} branch, i.e.\ a branch where no application of linear rules would produce new formulae.
 In Isabelle, we express this as follows.
@@
 definition lin_sat b $\equiv$
-  $\forall$b'. &\sexpands{b'}{b}& $\longrightarrow$ set b' $\subseteq$ set b
+  $\forall$b'. &\lexpands{b'}{b}& $\longrightarrow$ set b' $\subseteq$ set b
 \end{lstlisting}
@@
 
-\subsection{Fulfilling Expansion Rules\label{sec:fulfilling}}
-Since the number of branches doubles each time we apply a fulfilling rule, it is prudent to only apply them to open, linearly saturated branches. 
+\subsection{Branching Rules\label{sec:branching}}
+Since the number of branches increases when applying branching rules, it is prudent to only apply them to open, linearly saturated branches.
 To further prevent unnecessary branching, each rule has a \textit{subsumption} condition. 
+Note that even branching rules are restricted such that they never create new subterms, with exception of the last rule which adds a fresh variable.
+\todo[inline]{Reference figure}
 
@@
 \section{Completeness of the Calculus}
-
+\begin{itemize}
+  \item Basic definitions and graph
+  \item Graph is basically a syntactic model.
+\end{itemize}
 \begin{lstlisting}
@@
 \section{An Abstract Specification of the Decision Procedure}
-Think about using the notion of coherent (or is this wfbranch?)
+\todo[inline]{Mention coherence lemma?}
 \begin{figure*}
@@
   $\neg$ lin_sat b $\Longrightarrow$
-  mlss_proc_branch = mlss_proc_branch ((SOME b'. &\sexpands{b'}{b}& $\land$ set b $\subset$ set (b' @ b)) @ b)
+  mlss_proc_branch = mlss_proc_branch ((SOME b'. &\lexpands{b'}{b}& $\land$ set b $\subset$ set (b' @ b)) @ b)
 |  $\llbracket$ $\neg$ sat b; bopen b; lin_sat b $\rrbracket$ $\Longrightarrow$
@@
 Maybe move this to the chapter on the calculus?
+\todo[inline]{Show \lstinline!memberCycle! case? This requires \lstinline!member_seq!.}
 \subsection{Lifting the Completeness of the Calculus}

### 2022-08-30 paper/paper.tex
Continue

@@
 %\setmathfont[version=bold, FakeBold=2]{latinmodern-math.otf}
+
+% Record brackets
+\usepackage{stmaryrd}
 
@@
 conversely, it is called \textit{open} if it is not closed.
-For MLSS, we extend the notion of closedness with three additional rules; the first two are straightforward while the last one states that a branch is closed when the branch contains a membership cycle \lstinline!s $\inS$ t$_\text{1}$, t$_\text{1}$ $\inS$ t$_\text{2}$, $\ldots$, t$_\text{k}$ $\inS$ s!.
+For MLSS, we extend the notion of closedness with three additional rules; the first two are straightforward while the last one states that a branch is closed when the branch contains a membership cycle
+\begin{center}
+\lstinline!s $\inS$ t$_\text{1}$, t$_\text{1}$ $\inS$ t$_\text{2}$, $\ldots$, t$_\text{k}$ $\inS$ s!.
+\end{center}
 \begin{lstlisting}
@@
 \section{Completeness of the Calculus}
-\begin{itemize}
-  \item Basic definitions and graph
-  \item Graph is basically a syntactic model.
-\end{itemize}
-\begin{lstlisting}
-definition wf_branch b $\equiv$ $\exists$$\phi$. &\expandss{b}{[$\phi$]}&
-\end{lstlisting}
-
+Once a branch \lstinline!b! is saturated and open at the same time, we are unable to expand the branch further in our calculus.
+Thus, to establish completeness of the calculus, we must construct a model \lstinline!M! such that \lstinline!M $\models$ $\phi$! for all formulae \lstinline!$\phi$ $\in$ set b!.
+Our model is defined in terms of the following three collections.
 \begin{lstlisting}
@@
   params b $\equiv$ vars_branch b - vars_fm (last b)
-\end{lstlisting}
 
-\begin{lstlisting}
 definition params' :: 'a branch $\Rightarrow$ 'a set where
@@
       AT (t $\eqS$ Var c) $\notin$ set b}
-\end{lstlisting}
 
-\begin{lstlisting}
 definition subterms_branch' ::
@@
 \end{lstlisting}
+The goal is to construct a syntactic model that is derived from the membership literals \lstinline!(s $\inS$ t) $\in$ set b!.
+To this end, we construct a graph whose vertices are set terms and there is an edge from \lstinline!s! to \lstinline!t! in the graph if, and only if, \lstinline!(s $\inS$ t) $\in$ set b!.
+Note that we use \citeauthor{graph_theory_afp}'s graph library~\cite{graph_theory_afp} where a graph is represented as a record of vertices, arcs (directed edges), and two functions \lstinline!tail!
+and \lstinline!head!
+that map an arc to its source respectively target vertex.
+\begin{lstlisting}
+definition bgraph b $\equiv$
+  let vs = Var ` params' b $\cup$ subterms_branch' b
+  in $\llparenthesis$ verts = vs,
+     &$\phantom{\llparenthesis}$& arcs = {(s, t). (s $\inS$ t) $\in$ set b},
+     &$\phantom{\llparenthesis}$& tail = fst, head = snd $\rrparenthesis$
+\end{lstlisting}
+The model is defined in terms of a \textit{realization} function relative to this graph.
+Observe that the vertices are partitioned into two sets, a collection of parameters \lstinline!Var ` params' b!
+and the rest of the set terms \lstinline!subterms_branch' b!, which are treated differently by the realization.
+Terms in the latter set are evaluated in accordance to the structure of the graph, i.e.\ the realization of a vertex is defined as the union of the realizations of the parent vertices. 
+For the former set, we fix a function \lstinline!I!
+that assigns the elements \lstinline!x $\in$ Var ` params' b!
+pairwise distinct sets with cardinality greater than that of the vertices and then return the singleton set \lstinline!vset {I x}!.
+The cardinality constraint ensures that realizations are distinct between those two sets.  
+\begin{lstlisting}
+abbreviation parents G s $\equiv$ {u. u $\rightarrow_\text{G}$ s}
 
+function realize :: 'a pset_term $\Rightarrow$ V where
+  x $\in$ Var ` params' b
+  $\Longrightarrow$ realize x = vset {I x}
+| x $\in$ subterms_branch' b
+  $\Longrightarrow$ realize t =
+  &$\phantom{\Longrightarrow}$&   vset {realize ` parents (bgraph b) s}
+| x $\notin$ verts G $\Longrightarrow$ realize x = 0
+\end{lstlisting}
+Since the function is recursive, we need to ensure that it terminates for it to be well-defined.
+The assumption that \lstinline!b! is open implies that there are no membership cycles and thus \lstinline!bgraph b! is acyclic.
+Together with the fact that \lstinline!bgraph b! is finite, we conclude that we can only recurse a finite number of times until we encounter a vertex that has no parents.
+
+\subsection{Formalising the Key Technical Lemma}
 \begin{lstlisting}
-definition bgraph $\equiv$ $\ldots$ 
+fun is_literal :: 'a fm $\Rightarrow$ bool where
+  is_literal (&\fmAtom& _) = True
+| is_literal (&\fmNeg{(\fmAtom{} \_)}&) = True
+| is_literal _ = False
 \end{lstlisting}
+\begin{lstlisting}
+lemma
+  assumes wf_branch b
+  assumes c $\in$ params' b
+      and t $\in$ params_subterms b
+  shows (Var c $\eqS$ t) $\notin$ set b
+    and (t $\eqS$ Var c) $\notin$ set b
+    and (t $\inS$ Var c) $\notin$ set b
+\end{lstlisting}
+Only one quite trivial invariant needed
 
@@
 
+\begin{lstlisting}
+definition wf_branch b $\equiv$ $\exists$$\phi$. &\expandss{b}{[$\phi$]}&
+\end{lstlisting}
+
 \subsection{Building a Model: Realisation of a Branch}
@@
 Show the results in Realisation of an Open Branch.
-
-\subsection{Formalising the Key Technical Lemma}
-Only one quite trivial invariant needed
 

### 2022-08-31 paper/paper.tex
Continue with paper

@@
   escapeinside={&}{&},
-  columns=fixed,
+  columns=\lst@ifdisplaystyle{fullflexible}\else{fixed}\fi,,
   extendedchars,
@@
 \newcommand{\notinS}{\notin_\text{s}}
-\newcommand{\eqS}{\approx_\text{s}}
-\newcommand{\neqS}{{\not\approx}_\text{s}}
+\newcommand{\eqS}{=_\text{s}}
+\newcommand{\neqS}{\neq_\text{s}}
 \newcommand{\subseteqS}{\sqsubseteq_\text{s}}
@@
 At the heart of \MLSS{}, we have the type of set terms which includes the empty set and variables as well as the operations union, intersection, set difference, and \lstinline!Single!, which represents a singleton set containing the argument.
-\todo[inline]{Syntax for \lstinline!Single!?}
+\todo[inline]{Syntax for \lstinline!Single!?, overloading of \lstinline!vars_term!, \lstinline!vars_atom!}
 \begin{lstlisting}
-datatype 'a pset_term =
+datatype (vars: 'a) pset_term =
   $\emptyset$ | Var 'a | Single ('a pset_term)
@@
 \begin{lstlisting}
-datatype 'a pset_atom =
+datatype (vars: 'a) pset_atom =
   'a pset_term $\inS$ 'a pset_term
@@
 Note that Isabelle's datatype package automatically defines a function \lstinline!atoms :: 'a fm $\Rightarrow$ 'a set! that allows us to obtain all set atoms in a formula.
-Similarly, it defines a function \lstinline!set_pset_atom :: 'a pset_atom $\Rightarrow$ 'a set! with which we can extract all the variable names in a set atom.
-We combine these functions in order to define \lstinline!vars_fm! that extracts all the variable names occuring in a set formula.
-\todo[inline]{Motivation: one rule introduces new variables}
+We combine these functions in order to define \lstinline!vars! that extracts all the variable names occuring in a set formula.
+\todo[inline]{Motivation: one rule introduces new variables, Mention vars polymorphic}
 \begin{lstlisting}
-definition vars_fm :: 'a pset_fm $\Rightarrow$ 'a set where
-  vars_fm $\phi$ $\equiv$ $\bigcup$(set_pset_atom ` atoms $\phi$)
+definition vars :: 'a pset_fm $\Rightarrow$ 'a set where
+  vars $\phi$ $\equiv$ $\bigcup$(vars ` atoms $\phi$)
 \end{lstlisting}
@@
 \begin{lstlisting}
-fun subterms ::
-    'a pset_term $\Rightarrow$ 'a pset_term set where
+fun subterms :: 'a pset_term $\Rightarrow$ 'a pset_term set
+where
   subterms $\emptyset$ = {$\emptyset$}
@@
 \begin{lstlisting}
-fun subterms_atom ::
-      'a pset_atom $\Rightarrow$ 'a pset_term set where
-  subterms_atom (s $\inS$ t) =
-    subterms s $\cup$ subterms t
-| subterms_atom (s $\eqS$ t) =
-    subterms s $\cup$ subterms t
+fun subterms :: 'a pset_atom $\Rightarrow$ 'a pset_term set
+where
+  subterms (s $\inS$ t) = subterms s $\cup$ subterms t
+| subterms (s $\eqS$ t) = subterms s $\cup$ subterms t
 \end{lstlisting}
@@
 \begin{lstlisting}
-definition subterms_fm ::
+definition subterms ::
     'a pset_fm $\Rightarrow$ 'a pset_term set where
-  subterms_fm $\phi$ $\equiv$ $\bigcup$(subterms_atom ` atoms $\phi$)
+  subterms $\phi$ $\equiv$ $\bigcup$(subterms ` atoms $\phi$)
 \end{lstlisting}
@@
 \begin{lstlisting}
-definition subterms_branch ::
+definition subterms ::
     'a branch $\Rightarrow$ 'a pset_term set where
-  subterms_branch b $\equiv$ $\bigcup$(subterms_fm ` set b)
+  subterms b $\equiv$ $\bigcup$(subterms ` set b)
 
-definition vars_branch :: 'a branch $\Rightarrow$ 'a set 
-  vars_branch b $\equiv$ $\bigcup$(vars_fm ` set b)
+definition vars :: 'a branch $\Rightarrow$ 'a set 
+  vars b $\equiv$ $\bigcup$(vars ` set b)
 \end{lstlisting}
@@
 \begin{center}
-  \lstinline!t$_\text{1}$ $\unionS$ t$_\text{2}$ $\in$ subterms_fm (last b)!.
+  \lstinline!t$_\text{1}$ $\unionS$ t$_\text{2}$ $\in$ subterms (last b)!.
 \end{center}
@@
 Thus, to establish completeness of the calculus, we must construct a model \lstinline!M! such that \lstinline!M $\models$ $\phi$! for all formulae \lstinline!$\phi$ $\in$ set b!.
-Our model is defined in terms of the following three collections.
+The model is defined in terms of the following three collections.
+\todo[inline]{explain wits, explain that wits' are those whose value is not fixed by literals in b}
 \begin{lstlisting}
-definition params :: 'a branch $\Rightarrow$ 'a set where
-  params b $\equiv$ vars_branch b - vars_fm (last b)
+definition wits :: 'a branch $\Rightarrow$ 'a set where
+  wits b $\equiv$ vars b - vars (last b)
 
-definition params' :: 'a branch $\Rightarrow$ 'a set where
-  params' b $\equiv$
-    {c $\in$ params b. $\forall$t $\in$ subterms_fm (last b).
+definition wits' :: 'a branch $\Rightarrow$ 'a set where
+  wits' b $\equiv$
+    {c $\in$ wits b. $\forall$t $\in$ subterms (last b).
       AT (Var c $\eqS$ t) $\notin$ set b $\land$
@@
   subterms_branch' b $\equiv$
-    subterms_fm (last b) $\cup$ 
-    Var ` (params b - params' b)
+    subterms (last b) $\cup$ 
+    Var ` (wits b - wits' b)
 \end{lstlisting}
-The goal is to construct a syntactic model that is derived from the membership literals \lstinline!(s $\inS$ t) $\in$ set b!.
+We aim to construct a syntactic model that is derived from the membership literals \lstinline!(s $\inS$ t) $\in$ set b!.
 To this end, we construct a graph whose vertices are set terms and there is an edge from \lstinline!s! to \lstinline!t! in the graph if, and only if, \lstinline!(s $\inS$ t) $\in$ set b!.
@@
 definition bgraph b $\equiv$
-  let vs = Var ` params' b $\cup$ subterms_branch' b
+  let vs = Var ` wits' b $\cup$ subterms_branch' b
   in $\llparenthesis$ verts = vs,
@@
 The model is defined in terms of a \textit{realization} function relative to this graph.
-Observe that the vertices are partitioned into two sets, a collection of parameters \lstinline!Var ` params' b!
+Observe that the vertices are partitioned into two sets, a collection of witnesses \lstinline!Var ` wits' b!
 and the rest of the set terms \lstinline!subterms_branch' b!, which are treated differently by the realization.
 Terms in the latter set are evaluated in accordance to the structure of the graph, i.e.\ the realization of a vertex is defined as the union of the realizations of the parent vertices. 
-For the former set, we fix a function \lstinline!I!
-that assigns the elements \lstinline!x $\in$ Var ` params' b!
-pairwise distinct sets with cardinality greater than that of the vertices and then return the singleton set \lstinline!vset {I x}!.
-The cardinality constraint ensures that realizations are distinct between those two sets.  
+For the former set, we choose a function \lstinline!I! that assigns the elements \lstinline!x $\in$ Var ` wits' b! pairwise distinct sets with cardinality greater than that of the vertices.
+We can always choose such a function since we assume an infinite universe of variables.
+Then, we return the singleton set \lstinline!vset {I x}!, which, together with the cardinality constraint, ensures that realizations are distinct between those witnesses and the set terms.  
 \begin{lstlisting}
@@
 function realize :: 'a pset_term $\Rightarrow$ V where
-  x $\in$ Var ` params' b
-  $\Longrightarrow$ realize x = vset {I x}
+  x $\in$ Var ` wits' b $\Longrightarrow$ realize x = vset {I x}
 | x $\in$ subterms_branch' b
@@
 The assumption that \lstinline!b! is open implies that there are no membership cycles and thus \lstinline!bgraph b! is acyclic.
-Together with the fact that \lstinline!bgraph b! is finite, we conclude that we can only recurse a finite number of times until we encounter a vertex that has no parents.
+Now, consider the ancestors \lstinline!av! of a vertex \lstinline!v!, which are those vertices from which we can reach \lstinline!v!.
+Due to acyclicity, it holds that the ancestors \lstinline!ap! of a parent \lstinline!p! of \lstinline!v! are a subset of \lstinline!av!.
+They are also a proper subset of \lstinline!av! since \lstinline!p! is an ancestor of \lstinline!v! but not of \lstinline!p! itself.
+Together with the fact that \lstinline!bgraph b! is finite, it follows that the cardinality of the ancestors decreases in each recursive call thus proving the termination of \lstinline!realize!.
 
@@
 \begin{lstlisting}
-lemma
+lemma lemma_2:
   assumes wf_branch b
-  assumes c $\in$ params' b
-      and t $\in$ params_subterms b
+  assumes c $\in$ wits' b and t $\in$ subterms b
   shows (Var c $\eqS$ t) $\notin$ set b
@@
 definition no_new_subterms b $\equiv$
-  $\forall$t $\in$ subterms_branch b. t $\notin$ Var ` params b
-    $\longrightarrow$ t $\in$ subterms_fm (last b))
+  $\forall$t $\in$ subterms b. t $\notin$ Var ` wits b
+    $\longrightarrow$ t $\in$ subterms (last b))
 \end{lstlisting}
@@
 \subsection{Building a Model: Realisation of a Branch}
-Idea: keep this short. The formalisation matches the pen-and-paper proof quite closely.
-Show the results in Realisation of an Open Branch.
+Let \lstinline!b! be an open and saturated branch, i.e.\ it holds that \lstinline!sat b $\land$ bopen b!;
+then, the following propositions hold:
+\begin{itemize}
+  \item We have \lstinline!realize s $=$ realize t! if \lstinline!s $\eqS$ t! is in \lstinline!b!.
+  \item We have \lstinline!realize s $\neq$ realize t! if \lstinline!s $\neqS$ t! is in \lstinline!b!.
+  \item We have \lstinline!realize s $\in$ elts (realize t)! if it holds that \lstinline!(s $\inS$ t)! is in \lstinline!b!. 
+  \item We have \lstinline!realize s $\notin$ elts (realize t)! if it holds that \lstinline!(s $\notinS$ t)! is in \lstinline!b!. 
+\end{itemize}
+
+We extend this to the set term operators.
+Let \lstinline!b! be an open and saturated branch.
+\begin{itemize}
+  \item If \lstinline!$\emptyset$ $\in$ subterms b!, then \lstinline!realize $\emptyset$ = 0!.
+  \item Let \lstinline!$\star_\text{s}$ $\in$ {$\unionS$, $\diffS$, $\interS$}!. If \lstinline!s $\star_\text{s}$ t $\in$ subterms b!, then
+    \begin{center}
+    \lstinline!realize (s $\star_\text{s}$ t) = realize s $\star$ realize t!.
+    \end{center}
+  \item If \lstinline!Single t $\in$ subterms b!, then
+    \begin{center}
+    \lstinline!realize (Single t) = vset {realize t}!.
+    \end{center}
+\end{itemize}
 
@@
 \subsection{Soundness}
-Interesting: The model changes because auf new parameters.
+Interesting: The model changes because auf new witnesses.
 Maybe move this to the chapter on the calculus?

### 2022-09-01 paper/paper.tex
Continue

@@
 We first define a function that collects the subterms of a set term.
-\todo[inline]{Motivation: linear rules don't introduce subterms}
+\todo[inline]{Motivation: linear rules don't introduce subterms, \lstinline!subterms b $\approx$ subterms (last b)!}
 \begin{lstlisting}
@@
 conversely, it is called \textit{open} if it is not closed.
-For MLSS, we extend the notion of closedness with three additional rules; the first two are straightforward while the last one states that a branch is closed when the branch contains a membership cycle
+For \MLSS{}, we extend the notion of closedness with three additional rules; the first two are straightforward while the last one states that a branch is closed when the branch contains a membership cycle
 \begin{center}
@@
 
-definition subterms_branch' ::
+definition subterms' ::
     'a branch $\Rightarrow$ 'a pset_term set where
-  subterms_branch' b $\equiv$
+  subterms' b $\equiv$
     subterms (last b) $\cup$ 
@@
 We aim to construct a syntactic model that is derived from the membership literals \lstinline!(s $\inS$ t) $\in$ set b!.
-To this end, we construct a graph whose vertices are set terms and there is an edge from \lstinline!s! to \lstinline!t! in the graph if, and only if, \lstinline!(s $\inS$ t) $\in$ set b!.
+To this end, we construct a graph whose vertices are set terms and there is an edge from \lstinline!s! to \lstinline!t! in the graph if, and only if, \lstinline!s $\inS$ t! is in \lstinline!b!.
 Note that we use \citeauthor{graph_theory_afp}'s graph library~\cite{graph_theory_afp} where a graph is represented as a record of vertices, arcs (directed edges), and two functions \lstinline!tail!
@@
 definition bgraph b $\equiv$
-  let vs = Var ` wits' b $\cup$ subterms_branch' b
+  let vs = Var ` wits' b $\cup$ subterms' b
   in $\llparenthesis$ verts = vs,
@@
 Observe that the vertices are partitioned into two sets, a collection of witnesses \lstinline!Var ` wits' b!
-and the rest of the set terms \lstinline!subterms_branch' b!, which are treated differently by the realization.
+and the rest of the set terms \lstinline!subterms' b!, which are treated differently by the realization.
 Terms in the latter set are evaluated in accordance to the structure of the graph, i.e.\ the realization of a vertex is defined as the union of the realizations of the parent vertices. 
-For the former set, we choose a function \lstinline!I! that assigns the elements \lstinline!x $\in$ Var ` wits' b! pairwise distinct sets with cardinality greater than that of the vertices.
+For the former set, we choose a function \lstinline!I! that assigns the witnesses in \lstinline!Var ` wits' b! pairwise distinct sets with cardinality greater than that of the vertices.
 We can always choose such a function since we assume an infinite universe of variables.
@@
   x $\in$ Var ` wits' b $\Longrightarrow$ realize x = vset {I x}
-| x $\in$ subterms_branch' b
+| x $\in$ subterms' b
   $\Longrightarrow$ realize t =
@@
 \end{lstlisting}
+
+\todo[inline]{Rename \lstinline!lemma_2!}
 \begin{lstlisting}
@@
 
-\subsection{Building a Model: Realisation of a Branch}
-Let \lstinline!b! be an open and saturated branch, i.e.\ it holds that \lstinline!sat b $\land$ bopen b!;
-then, the following propositions hold:
-\begin{itemize}
-  \item We have \lstinline!realize s $=$ realize t! if \lstinline!s $\eqS$ t! is in \lstinline!b!.
+\subsection{Realisation of an Open Branch}
+Remember that for completeness, we need to show that the realization function for an open and saturated branch \lstinline!b! actually constitutes a model for all formualae in the branch.
+We start by verifying that the realization function models all literals in the branch; more formally, the following propositions hold: 
+\begin{enumerate}[label=(\roman*)]
+  \item\label{it:at_mem} We have \lstinline!realize s $\in$ elts (realize t)! if it holds that \lstinline!s $\inS$ t! is in \lstinline!b!. 
+  \item\label{it:at_eq} We have \lstinline!realize s $=$ realize t! if \lstinline!s $\eqS$ t! is in \lstinline!b!.
   \item We have \lstinline!realize s $\neq$ realize t! if \lstinline!s $\neqS$ t! is in \lstinline!b!.
-  \item We have \lstinline!realize s $\in$ elts (realize t)! if it holds that \lstinline!(s $\inS$ t)! is in \lstinline!b!. 
-  \item We have \lstinline!realize s $\notin$ elts (realize t)! if it holds that \lstinline!(s $\notinS$ t)! is in \lstinline!b!. 
-\end{itemize}
+  \item We have \lstinline!realize s $\notin$ elts (realize t)! if it holds that \lstinline!s $\notinS$ t! is in \lstinline!b!. 
+\end{enumerate}
+To illustrate the usefulness of \lstinline!lemma_2!, we prove Proposition~\ref{it:at_eq}.
+The proofs of all propositions translate well into Isabelle so we refer to the original paper~\cite{new_fast_tableau} for the remaining proofs. 
+\begin{proof}[Proof of Proposition~\ref{it:at_eq}]
+  Assume that \lstinline!s $\eqS$ t! is in \lstinline!b!.
+  If there exists a \lstinline!c $\in$ wits' b! where \lstinline!s = Var c! or \lstinline!t = Var c!, we arrive at a contradiction due to \lstinline!lemma_2!.
+  Therefore, both \lstinline!s $\in$ subterms' b! and \lstinline!t $\inS$ subterms' b! must hold.
+  Now, assume for contradiction that \lstinline!realize s $\neq$ realize t! which implies that \lstinline!elts (realize s) $\neq$ elts (realize t)!.
+  Without loss of generality --- the other case is symmetric ---, we obtain an \lstinline!a! such that \lstinline!a $\in$ elts (realize s)! and \lstinline!a $\notin$ elts (realize t)!. 
+  Furthermore, we obtain \lstinline!u! such that \lstinline!a! is the realization of \lstinline!u!;
+  that is, we have \lstinline!a = realize u!.
+  Since \lstinline!s $\in$ subterms' b!, the definition of \lstinline!realize!
+gives us that \lstinline!u $\rightarrow_\text{bgraph b}$ s!
+holds, which, in turn, yields that \lstinline!u $\inS$ s!
+must be in \lstinline!b!.
+  Together with the assumption \lstinline!(s $\eqS$ t) $\in$ b! and the openness of \lstinline!b!, it follows that \lstinline!u $\inS$ t! must also be in \lstinline!b!.
+  But then we have
+\begin{lstlisting}[columns=fixed]
+realize u $\in$ elts (realize t)
+      = a $\in$ elts (realize t)
+\end{lstlisting}
+  using Proposition~\ref{it:at_mem}, which stands in contradiction to the assumption \lstinline!a $\notin$ elts (realize t)!.
+\end{proof}
 
-We extend this to the set term operators.
-Let \lstinline!b! be an open and saturated branch.
-\begin{itemize}
-  \item If \lstinline!$\emptyset$ $\in$ subterms b!, then \lstinline!realize $\emptyset$ = 0!.
-  \item Let \lstinline!$\star_\text{s}$ $\in$ {$\unionS$, $\diffS$, $\interS$}!. If \lstinline!s $\star_\text{s}$ t $\in$ subterms b!, then
+The results on literals can now be lowered to set terms.
+\begin{enumerate}[label=(\alph*)]
+  \item\label{it:empty} If \lstinline!$\emptyset$ $\in$ subterms b!, then \lstinline!realize $\emptyset$ = 0!.
+  \item\label{it:op} Let \lstinline!$\star_\text{s}$ $\in$ {$\unionS$, $\diffS$, $\interS$}!. If the term \lstinline!s $\star_\text{s}$ t! occurs in \lstinline!subterms b!, then
     \begin{center}
@@
     \end{center}
-\end{itemize}
+\end{enumerate}
+\begin{proof}[Proof of Proposition~\ref{it:empty}]
+  Since \lstinline!b!
+is open, we know \lstinline!s $\inS$ $\emptyset$!
+is not in \lstinline!b!
+for any \lstinline!s!, implying that the term \lstinline!$\emptyset$!
+has no parents in \lstinline!bgraph b!.
+  Moreover, the assumption \lstinline!$\emptyset$ $\in$ subterms b! can be strengthened to \lstinline!$\emptyset$ $\in$ subterms' b! as the term \lstinline!$\emptyset$! is not a witness. 
+  Ultimately, we obtain \lstinline!realize $\emptyset$ = 0! by definition of \lstinline!realize!.
+\end{proof}
+\begin{proof}[Proof of Proposition~\ref{it:op} for \lstinline!$\star_text{s}$ = $\diffS$!]
+  Test
+\end{proof}
 
@@
 function mlss_proc_branch :: 'a branch $\Rightarrow$ bool where
-  $\neg$ lin_sat b $\Longrightarrow$
-  mlss_proc_branch = mlss_proc_branch ((SOME b'. &\lexpands{b'}{b}& $\land$ set b $\subset$ set (b' @ b)) @ b)
-|  $\llbracket$ $\neg$ sat b; bopen b; lin_sat b $\rrbracket$ $\Longrightarrow$
-   mlss_proc_branch b = ($\forall$b' $\in$ (SOME bs. &\fexpands{bs}{b}&). mlss_proc_branch (b' @ b))
+  $\neg$ lin_sat b
+  $\Longrightarrow$ mlss_proc_branch = mlss_proc_branch ((SOME b'. &\lexpands{b'}{b}& $\land$ set b $\subset$ set (b' @ b)) @ b)
+| $\llbracket$ $\neg$ sat b; bopen b; lin_sat b $\rrbracket$
+  $\Longrightarrow$ mlss_proc_branch b = ($\forall$b' $\in$ (SOME bs. &\fexpands{bs}{b}&). mlss_proc_branch (b' @ b))
 | $\llbracket$ lin_sat b; sat b $\rrbracket$ $\Longrightarrow$ mlss_proc_branch b = bclosed b

### 2022-09-02 paper/paper.tex
Add todo
Continue

@@
 \section{Introduction}
-\todo[inline]{Write the introduction}
+\todo[inline]{Write the introduction, Remember that \MLSS{} is multi-level syllogistics with singleton}
 
@@
   \item\label{it:at_eq} We have \lstinline!realize s $=$ realize t! if \lstinline!s $\eqS$ t! is in \lstinline!b!.
-  \item We have \lstinline!realize s $\neq$ realize t! if \lstinline!s $\neqS$ t! is in \lstinline!b!.
-  \item We have \lstinline!realize s $\notin$ elts (realize t)! if it holds that \lstinline!s $\notinS$ t! is in \lstinline!b!. 
+  \item\label{it:af_eq} We have \lstinline!realize s $\neq$ realize t! if \lstinline!s $\neqS$ t! is in \lstinline!b!.
+  \item\label{it:af_mem} We have \lstinline!realize s $\notin$ elts (realize t)! if it holds that \lstinline!s $\notinS$ t! is in \lstinline!b!. 
 \end{enumerate}
@@
   Now, assume for contradiction that \lstinline!realize s $\neq$ realize t! which implies that \lstinline!elts (realize s) $\neq$ elts (realize t)!.
-  Without loss of generality --- the other case is symmetric ---, we obtain an \lstinline!a! such that \lstinline!a $\in$ elts (realize s)! and \lstinline!a $\notin$ elts (realize t)!. 
-  Furthermore, we obtain \lstinline!u! such that \lstinline!a! is the realization of \lstinline!u!;
-  that is, we have \lstinline!a = realize u!.
-  Since \lstinline!s $\in$ subterms' b!, the definition of \lstinline!realize!
-gives us that \lstinline!u $\rightarrow_\text{bgraph b}$ s!
-holds, which, in turn, yields that \lstinline!u $\inS$ s!
-must be in \lstinline!b!.
+  Without loss of generality --- the other case is symmetric ---, we obtain an \lstinline!a! such that
+\begin{lstlisting}
+  a $\in$ elts (realize s) $\land$ a $\notin$ elts (realize t). 
+\end{lstlisting}
+  Considering the fact that \lstinline!s $\in$ subterms' b!
+and the definition of \lstinline!realize!, we obtain an \lstinline!u!
+such that \lstinline!a = realize u!
+and \lstinline!u $\rightarrow_\text{bgraph b}$ s!.
+  This, in turn, yields that \lstinline!u $\inS$ s! must be in \lstinline!b!.
   Together with the assumption \lstinline!(s $\eqS$ t) $\in$ b! and the openness of \lstinline!b!, it follows that \lstinline!u $\inS$ t! must also be in \lstinline!b!.
@@
 \end{proof}
-\begin{proof}[Proof of Proposition~\ref{it:op} for \lstinline!$\star_text{s}$ = $\diffS$!]
-  Test
+\begin{proof}[Proof of Proposition~\ref{it:op} for $\star_\texttt{s} = -_\texttt{s}$]
+  For equality, it suffices to prove subset inclusion in both directions.
+
+  \textbf{Direction $\subseteq$. } Fix some \lstinline!e! in \lstinline!elts (realize (s $\diffS$ t))!.
+  Again, we obtain an \lstinline!u! where \lstinline!e = realize u! and \lstinline!u $\rightarrow_\text{bgraph b}$ (s $\diffS$ t)!.
+  We therefore have \lstinline!(u $\inS$ s $\diffS$ t) $\in$ set b!
+which, due to \lstinline!b!
+being saturated, implies that \lstinline!u $\inS$ s!
+is in \lstinline!b!
+while \lstinline!u $\inS$ t!
+is not.
+  By applying Proposition~\ref{it:at_mem} and~\ref{it:af_mem} we obtain
+\begin{lstlisting}
+  realize u $\in$ realize s $\land$ realize u $\notin$ realize t
+\end{lstlisting}
+  and together with \lstinline!e = realize u! we arrive at the goal 
+\begin{lstlisting}
+  e $\in$ elts (realize s - realize t).
+\end{lstlisting}
+
+\textbf{Direction $\supseteq$. } Let \lstinline[columns=fullflexible]!e $\in$ elts (realize s - realize t)! for some fixed \lstinline!e!.
+  As before, we obtain an \lstinline!u! such that \lstinline!e! is the realization of \lstinline!u!, i.e.\ \lstinline!e = realization u!, and
+\begin{lstlisting}
+  u $\rightarrow_\text{bgraph b}$ s $\land$ u $\nrightarrow_\text{bgraph b}$ t.
+\end{lstlisting}
+  An immediate consequence is that \lstinline!u $\inS$ s! is in \lstinline!b!.
+  Moreover, we claim that \lstinline!u $\notinS$ t! is also in \lstinline!b!.
+  To see that this is true, recall that \lstinline!b! is saturated and thus the subsumption condition of rule TODO needs to hold.
+  This implies that either \lstinline!u $\inS$ t!
+or \lstinline!u $\notinS$ t!
+is in \lstinline!b!
+but the latter being in \lstinline!b!
+would be a contradiction to \lstinline!u $\nrightarrow_\text{bgraph b}$ t!
+meaning that the former must hold.
+  Ultimately, we infer that \lstinline!u $\in$ s $\diffS$ t! is in \lstinline!b! by invoking saturation again.
+  The last step is to apply Proposition~\ref{it:at_mem} in order to obtain the goal
+  \begin{lstlisting}
+    realize u $\in$ elts (realize (s $\diffS$ t))
+  = e $\in$ elts (realize (s $\diffS$ t)).
+  \end{lstlisting}
 \end{proof}

### 2022-09-05 paper/paper.tex
Continue

@@
 \usepackage{enumitem}
-
-\setlength{\abovecaptionskip}{0pt}
-\setlength{\textfloatsep}{18pt plus 2pt minus 4pt}
 
@@
 \newcommand{\MLSS}{\textbf{MLSS}}
+
 \newlength{\trianglewidth}
 \settowidth{\trianglewidth}{\(\triangleleft\)}
-
 \newcommand{\lefttrianglebar}{\mathrel{\tikz[baseline]{\draw (1ex, 0.75ex) -- (0, 1.25ex) -- (0, 0.25ex) -- cycle; \draw (0, 0.75ex) -- (1ex, 0.75ex);}}}
@@
 \newcommand{\fexpands}[2]{#1 $\lefttrianglebar$ #2}
+\newcommand{\fexpandsW}[2]{#1 $\lefttrianglebar_\text{v}$ #2}
 \newcommand{\expandss}[2]{#1 $\lefttriangle^*$ #2}
@@
 The entry declares a type \lstinline!V! that is equipped with a function \lstinline!elts :: V $\Rightarrow$ V set! that maps a set to its elements.
+\todo[inline]{Use itemize for different functions}
 Even though we use the typed sets provided by Isabelle/HOL here, \lstinline!elts! actually returns a class.
@@
 Together with the fact that \lstinline!bgraph b! is finite, it follows that the cardinality of the ancestors decreases in each recursive call thus proving the termination of \lstinline!realize!.
+\todo[inline]{Überleitung}
 
@@
 
-The results on literals can now be lowered to set terms.
+\noindent The results on literals can now be lowered to set terms.
 \begin{enumerate}[label=(\alph*)]
@@
     \end{center}
-  \item If \lstinline!Single t $\in$ subterms b!, then
+  \item\label{it:single} If \lstinline!Single t $\in$ subterms b!, then
     \begin{center}
@@
   \textbf{Direction $\subseteq$. } Fix some \lstinline!e! in \lstinline!elts (realize (s $\diffS$ t))!.
-  Again, we obtain an \lstinline!u! where \lstinline!e = realize u! and \lstinline!u $\rightarrow_\text{bgraph b}$ (s $\diffS$ t)!.
+  Again, we obtain an \lstinline!u! with 
+\begin{lstlisting}
+  e = realize u $\land$ u $\rightarrow_\text{bgraph b}$ (s $\diffS$ t).
+\end{lstlisting}
   We therefore have \lstinline!(u $\inS$ s $\diffS$ t) $\in$ set b!
@@
     realize u $\in$ elts (realize (s $\diffS$ t))
-  = e $\in$ elts (realize (s $\diffS$ t)).
+    &\textover[r]{=}{realize}& e $\in$ elts (realize (s $\diffS$ t)).
   \end{lstlisting}
 \end{proof}
+
+\noindent The final step towards obtaining a proper model is to connect the realization function to the semantics as defined in Section~\ref{sec:semantics}.
+For set terms, we can use the Propositions~\ref{it:empty}--\ref{it:single} to prove the lemma below by induction.
+\begin{lstlisting}
+lemma
+  assumes t $\in$ subterms b
+  shows &\Ist& ($\lambda$x. realize (Var x)) t = realize t
+\end{lstlisting}
+Lifting the above result to formulae yields the coherence of \lstinline!b! as the original paper~\cite{new_fast_tableau} calls it.
+The proof is a tedious but straightforward induction on the the size of the formulae.
+\begin{lstlisting}
+lemma coherence:
+  assumes $\phi$ $\in$ set b
+  shows ($\lambda$x. realize (Var x)) $\models$ $\phi$
+\end{lstlisting}
+The coherence property finishes the proof of completeness of the calculus since it gives us a model for every formula in an open and saturated branch.
+
+\section{Soundness of the Calculus}
+Two aspects determine the soundness of the calculus: a closed branch must be unsatisfiable and all expansion rules must maintain satisfiability.
+We can express the former aspect in Isabelle as follows.
+\begin{lstlisting}[belowskip=0pt]
+lemma
+  assumes bclosed b shows $\exists$$\phi$ $\in$ set b. M $\not\models$ $\phi$
+\end{lstlisting}
+\begin{proof}
+  Assume for contradiction that \lstinline!b! is closed but we have a model, i.e.\ it holds that \lstinline!$\forall$$\phi$ $\in$ set b. M $\models$ $\phi$!.
+  We perform a case analysis on \lstinline!bclosed b!.
+  It is clear that, for any \lstinline!s!, neither does \lstinline!M! model \lstinline!s $\in$ $\emptyset$! nor \lstinline!s $\neqS$ s!.
+  Furthermore, no model can satisfy both \lstinline!$\phi$! and \lstinline!&\fmNeg{$\phi$}&! at the same time.
+  This leaves us with the case where \lstinline!b! contains a membership cycle
+\begin{lstlisting}
+  t$_\text{0}$ $\inS$ t$_\text{1}$, t$_\text{1}$ $\inS$ t$_\text{2}$, $\ldots$, t$_\text{k}$ $\inS$ t$_\text{0}$.
+\end{lstlisting}
+  By assumption, each atom of the membership cycle holds under the model \lstinline!M!; in other words, we have
+\begin{lstlisting}
+  &\Ist& M t$_\text{i}$ $\in$ elts (&\Ist& M t$_\text{i + 1 mod k + 1}$)
+\end{lstlisting}
+  for each \lstinline!i $\in$ {0,$\ldots$,k}!.
+  Therefore, the transitive closure of the membership relation \lstinline!{(x, y). x $\in$ elts y}! contains a cycle. 
+  This, however, is a contradiction since the membership relation of \lstinline!V! is well-founded (see Section~\ref{sec:semantics}) and well-foundedness is closed under transitive closure.
+\end{proof}
+
+\noindent We are left with showing that both linear and branching expansion rules preserve satisfiability.
+As for the linear rules, a straightforward proof by rule induction suffices to obtain the lemma below.
+\begin{lstlisting}
+lemma
+  assumes &\lexpands{b'}{b}& and $\phi$ $\in$ set b'
+  assumes $\bigwedge$$\psi$. $\psi$ $\in$ set b $\Longrightarrow$ M $\models$ $\psi$
+  shows M $\models$ $\phi$
+\end{lstlisting}
+The same would also hold for branching rules if it were not for the last rule that adds additional variables.
+Those variables need to be assigned specific values as we will see in the proof below.
+\begin{lstlisting}[belowskip=0pt]
+lemma
+  assumes &\fexpandsW{bs'}{b}&
+  assumes $\bigwedge$$\psi$. $\psi$ $\in$ set b $\Longrightarrow$ M $\models$ $\psi$
+  shows $\exists$M'. $\exists$b' $\in$ bs'.
+          $\forall$$\psi$ $\in$ set (b' @ b). M' $\models$ $\psi$
+\end{lstlisting}
+\begin{proof}
+  Since the last branching rule is applied here, it holds that
+\begin{lstlisting}
+  bs' = {[Var x $\inS$ s, Var x $\notinS$ t],
+         [Var x $\inS$ t, Var x $\notinS$ s]}
+\end{lstlisting}
+  for some fresh variable \lstinline!x!, i.e.\ \lstinline!x $\notin$ vars b! and in particular \lstinline!x $\notin$ vars s! as well as \lstinline!x $\notin$ vars t!.
+  Another consequence of applying the last rule is that \lstinline!s $\neqS$ t! is in \lstinline!b!, which, because \lstinline!M! is a model, implies that \lstinline!&\Ist& M s $\neq$ &\Ist& M t!.
+  This inequality manifests itself through some \lstinline!y! with
+\begin{lstlisting}
+  &\phantom{$\lor$}& y $\in$ elts (&\Ist& M s) $\land$ y $\notin$ elts (&\Ist& M t)
+  $\lor$ y $\in$ elts (&\Ist& M t) $\land$ y $\notin$ elts (&\Ist& M s).
+\end{lstlisting}
+  We now update the assignment \lstinline!M! such that it maps \lstinline!x! to \lstinline!y! to obtain the assignment \lstinline!M'!.
+  Note that values of both \lstinline!s! and \lstinline!t! are unchanged under this update as \lstinline!x $\notin$ vars s! and \lstinline!x $\notin$ vars t!.
+  Together with the our assumption about \lstinline!y!, we infer that 
+\end{proof}
+
+\section{Upper Bounding the Cardinality of a Branch}
 
@@
 \end{lstlisting}
-\caption{Definition of the function \lstinline|mlss_proc_branch| and \lstinline|mlss_proc|.}
+\caption{Definition of the functions \lstinline|mlss_proc_branch| and \lstinline|mlss_proc|.}
 \end{figure*}

### 2022-09-06 paper/paper.tex
Continue

@@
 We omit the implementation of $\models$ for brevity.
+\todo[inline]{Add implementation? Add satisfiability!}
 
@@
 A tableau is called \textit{closed} if all of its branches are closed.
+\todo[inline]{Move closed}
 
@@
 \end{lstlisting}
-The coherence property finishes the proof of completeness of the calculus since it gives us a model for every formula in an open and saturated branch.
+The coherence property finishes the proof of completeness of the calculus as it gives us a model for every formula in an open and saturated branch.
 
@@
 \begin{lstlisting}[belowskip=0pt]
-lemma
+lemma bclosed_sound:
   assumes bclosed b shows $\exists$$\phi$ $\in$ set b. M $\not\models$ $\phi$
@@
 \begin{lstlisting}
-lemma
+lemma lexpands_sound:
   assumes &\lexpands{b'}{b}& and $\phi$ $\in$ set b'
@@
 \end{lstlisting}
-The same would also hold for branching rules if it were not for the last rule that adds additional variables.
-Those variables need to be assigned specific values as we will see in the proof below.
+A similar argument would work for the branching rules if it were not for the last rule adding new variables.
+Those variables need to be assigned specific values, hence modifying the model, as we will see in the proof below.
 \begin{lstlisting}[belowskip=0pt]
-lemma
-  assumes &\fexpandsW{bs'}{b}&
+lemma fexpands_sound:
+  assumes &\fexpands{bs'}{b}&
   assumes $\bigwedge$$\psi$. $\psi$ $\in$ set b $\Longrightarrow$ M $\models$ $\psi$
@@
 \begin{proof}
-  Since the last branching rule is applied here, it holds that
+  We perform a case analysis on the derivation \lstinline!&\fexpands{bs'}{b}&!.
+  As all cases but that of the last rule are straightforward, we only consider this last case where we have 
 \begin{lstlisting}
@@
 \end{lstlisting}
-  for some fresh variable \lstinline!x!, i.e.\ \lstinline!x $\notin$ vars b! and in particular \lstinline!x $\notin$ vars s! as well as \lstinline!x $\notin$ vars t!.
+  for some fresh variable \lstinline!x!, i.e.\ \lstinline!x $\notin$ vars b! and in particular \lstinline!x $\notin$ vars s $\cup$ vars t!.
   Another consequence of applying the last rule is that \lstinline!s $\neqS$ t! is in \lstinline!b!, which, because \lstinline!M! is a model, implies that \lstinline!&\Ist& M s $\neq$ &\Ist& M t!.
@@
   Note that values of both \lstinline!s! and \lstinline!t! are unchanged under this update as \lstinline!x $\notin$ vars s! and \lstinline!x $\notin$ vars t!.
-  Together with the our assumption about \lstinline!y!, we infer that 
+  Considering our assumption about \lstinline!y! and the fact that \lstinline!&\Ist& M' (Var x) = y!, we infer that
+\begin{lstlisting}
+  $\exists$b' $\in$ bs'. $\forall$$\psi$ $\in$ set b'. M' $\models$ $\psi$.
+\end{lstlisting}
+  In other words, due to the disjunction, it holds that \lstinline!M'! is a model for one \lstinline!b' $\in$ bs!.
+  Thus, we are left with proving that \lstinline!M'! is a model for \lstinline!b!.
+  This is indeed the case since the assumption \lstinline!M $\models$ $\psi$!
+is equivalent to \lstinline!M' $\models$ $\psi$!
+for all \lstinline!$\psi$ $\in$ set b!
+given that \lstinline!x!
+is fresh with respect to b.
 \end{proof}
@@
 \section{Upper Bounding the Cardinality of a Branch}
+Soundness and completeness of the calculus are the building blocks with which we establish partial correctness of Algorithm~\ref{fig:alg}.
+Consequently, we still need to give a termination proof of Algorithm~\ref{fig:alg} to obtain total correctness of the algorithm.
+The core of our argument is that the number of distinct formulae in a branch that can be derived with the expansion rules is finite.
+In particular, we claim that the number of distinct formulae has the following upper bound.
+\begin{lstlisting}
+  &\phantom{$\leq$}& card (set b)
+  &\textover{$=$}{$\leq$}& card {$\psi$ $\in$ set b. $\neg$ is_literal $\psi$} +
+  &\phantom{$\leq$}& card {$\psi$ $\in$ set b. is_literal $\psi$}
+  $\leq$ 2 * card (subfms (last b)) +
+  &\phantom{$\leq$}& card {$\psi$ $\in$ set b. is_literal $\psi$}
+  $\leq$ 2 * card (subfms (last b)) +
+  &\phantom{$\leq$}& 16 * (card (subterms (last b)))$^4$
+\end{lstlisting}
+We justify both inequalities separately, starting with the first one whose justification hinges on an invariant:
+Its justification hinges on the invariant that the only new subformulae introduced by the expansion rules introduce
 
 \section{An Abstract Specification of the Decision Procedure}
-\todo[inline]{Mention coherence lemma?}
 \begin{figure*}
@@
 \end{lstlisting}
-\caption{Definition of the functions \lstinline|mlss_proc_branch| and \lstinline|mlss_proc|.}
+\caption{Definition of the functions \lstinline|mlss_proc_branch| and \lstinline|mlss_proc|.\label{fig:alg}}
 \end{figure*}
-
-\subsection{Termination}
-Maybe interesting: technical lemma that allows to obtain the point at which a parameter was introduced.
-\subsection{Soundness}
-Interesting: The model changes because auf new witnesses.
-Maybe move this to the chapter on the calculus?
-\todo[inline]{Show \lstinline!memberCycle! case? This requires \lstinline!member_seq!.}
-\subsection{Lifting the Completeness of the Calculus}
-\todo[inline]{Do you even lift?}
 

### 2022-09-07 paper/paper.tex
Continue

@@
     keywordstyle=[1]\bfseries\color{isarblue},
-    keywords=[2]{where,assumes,shows,and},
+    keywords=[2]{where,assumes,shows,and,fixes},
     keywordstyle=[2]\bfseries\color{isargreen},
@@
 
+% Autoref names
+\renewcommand{\sectionautorefname}{Section}
+\renewcommand{\subsectionautorefname}{Section}
+
 % Commands
@@
 \newcommand{\fmOr}[2]{#1 $\boldsymbol{\lor}$ #2}
-\newcommand{\fmNeg}[1]{$\boldsymbol{\neg}$ #1}
+\newcommand{\fmNegSymbol}{\boldsymbol{\neg}}
+\newcommand{\fmNeg}[1]{$\fmNegSymbol$ #1}
 \newcommand{\fmAtom}{\textbf{A}}
@@
 An article by Lawrence~\cite{zfc_in_hol} provides some further context to this entry.
-The entry declares a type \lstinline!V! that is equipped with a function \lstinline!elts :: V $\Rightarrow$ V set! that maps a set to its elements.
-\todo[inline]{Use itemize for different functions}
-Even though we use the typed sets provided by Isabelle/HOL here, \lstinline!elts! actually returns a class.
-Only if the class is small enough, as indicated by the polymorphic predicate \lstinline!small :: 'a set $\Rightarrow$ bool!, it is indeed a set.
+The entry declares a type \lstinline!V! that comes with the following functionality:
+\begin{itemize}
+  \item The function \lstinline!elts :: V $\Rightarrow$ V set! maps a set to its elements.
+      Note that even though we use the typed sets provided by Isabelle/HOL here, \lstinline!elts! actually returns a class.
+      This class only constitutes a set if the class is small enough.
+  \item The predicate \lstinline!small :: 'a set $\Rightarrow$ bool!, which is polymorphic in \lstinline!'a!, indicates whether a set is small.
 Inversely, this predicate characterises those sets that can be turned into a set of type \lstinline!V! using the polymorphic function \lstinline!vset :: 'a set $\Rightarrow$ V!.
-The type \lstinline!V! comes with all the usual set operations such as equality ($=$), union ($\sqcup$), intersection ($\sqcap$), and difference ($-$).
-The empty set coincides with the ordinal $0$ so it is denoted by \lstinline!0 :: V!.
+\item The usual set operations such as equality ($=$), union ($\sqcup$), intersection ($\sqcap$), and difference ($-$) are defined.
+\item Finally, the empty set coincides with the ordinal $0$ so it is denoted by \lstinline!0 :: V!.
+\end{itemize}
 
@@
 A tableau is called \textit{closed} if all of its branches are closed.
-\todo[inline]{Move closed}
 
@@
 that map an arc to its source respectively target vertex.
+\todo[inline]{Explain arrow notation}
 \begin{lstlisting}
@@
   Therefore, both \lstinline!s $\in$ subterms' b! and \lstinline!t $\inS$ subterms' b! must hold.
-  Now, assume for contradiction that \lstinline!realize s $\neq$ realize t! which implies that \lstinline!elts (realize s) $\neq$ elts (realize t)!.
+  Now, assume for contradiction that
+\begin{lstlisting}
+  &\phantom{$\Longrightarrow$}& realize s $\neq$ realize t
+  $\Longrightarrow$ elts (realize s) $\neq$ elts (realize t).
+\end{lstlisting}
   Without loss of generality --- the other case is symmetric ---, we obtain an \lstinline!a! such that
@@
 
-\noindent The final step towards obtaining a proper model is to connect the realization function to the semantics as defined in Section~\ref{sec:semantics}.
+\noindent The final step towards obtaining a proper model is to connect the realization function to the semantics as defined in \autoref{sec:semantics}.
 For set terms, we can use the Propositions~\ref{it:empty}--\ref{it:single} to prove the lemma below by induction.
@@
   Therefore, the transitive closure of the membership relation \lstinline!{(x, y). x $\in$ elts y}! contains a cycle. 
-  This, however, is a contradiction since the membership relation of \lstinline!V! is well-founded (see Section~\ref{sec:semantics}) and well-foundedness is closed under transitive closure.
+  This, however, is a contradiction since the membership relation of \lstinline!V! is well-founded (see \autoref{sec:semantics}) and well-foundedness is closed under transitive closure.
 \end{proof}
@@
 A similar argument would work for the branching rules if it were not for the last rule adding new variables.
-Those variables need to be assigned specific values, hence modifying the model, as we will see in the proof below.
+Those variables need to be assigned specific values; hence, we modify the model as we will see in the proof below.
 \begin{lstlisting}[belowskip=0pt]
@@
 \section{Upper Bounding the Cardinality of a Branch}
-Soundness and completeness of the calculus are the building blocks with which we establish partial correctness of Algorithm~\ref{fig:alg}.
-Consequently, we still need to give a termination proof of Algorithm~\ref{fig:alg} to obtain total correctness of the algorithm.
-The core of our argument is that the number of distinct formulae in a branch that can be derived with the expansion rules is finite.
-In particular, we claim that the number of distinct formulae has the following upper bound.
+Soundness and completeness of the calculus are the building blocks with which we establish partial correctness of \autoref{fig:alg}.
+Consequently, we still need to give a termination proof to ascertain total correctness.
+The core of our argument is that the number of distinct formulae that can be derived with the expansion rules is finite.
+In particular, we claim that the following upper bound holds.
 \begin{lstlisting}
@@
 \end{lstlisting}
-We justify both inequalities separately, starting with the first one whose justification hinges on an invariant:
-Its justification hinges on the invariant that the only new subformulae introduced by the expansion rules introduce
+We treat both inequalities separately, starting with the first one whose justification hinges on an invariant: 
+every new composite subformula that is introduced by an expansion rule is the negation of a subformula that already occurs in \lstinline!last b!.
+Written as a formula, we have
+\begin{lstlisting}
+  &\phantom{$\subseteq$}& {$\psi$ $\in$ set b. $\neg$ is_literal $\psi$}
+  $\subseteq$ subfms (last b) $\cup$ $\fmNegSymbol$ ` subfms (last b).
+\end{lstlisting}
+Confirming that this is an invariant is an easy exercise of going through each of the propositional rules in Table TODO and the first two branching rules in Table TODO.
+Thus, we have
+\begin{lstlisting}
+  &\phantom{$\leq$}& card {$\psi$ $\in$ set b. $\neg$ is_literal $\psi$}
+  $\leq$ card (subfms (last b)) +
+  &\phantom{$\leq$}& card ($\fmNegSymbol$ ` subfms (last b))
+  $\leq$ 2 * card (subfms (last b))
+\end{lstlisting}
+which is the first inequality we asserted.
+
+Concerning the second inequality, first observe that each literal contains an atom, implying that the set of literals is at most twice as large as the set of atoms in a branch. 
+In addition, an atom is formed by combining two subterms occuring in the branch with either \lstinline!$\inS$! or \lstinline!$\eqS$!, ultimately yielding
+\begin{lstlisting}
+  &\phantom{$\leq$}& card {$\psi$ $\in$ set b. is_literal $\psi$}
+  $\leq$ 2 * card ($\bigcup$(atoms ` set b))
+  $\leq$ 2 * 2 * (card (subterms b))$^\text{2}$.
+\end{lstlisting}
+Since no expansion rule introduces new subterms, we know the cardinality of the subterms in a branch \lstinline!b!
+is limited by the subterms of \lstinline!last b!
+plus the number of witnesses added to the branch.
+Regarding an upper bound for the witnesses, recognise that the last branching rule requires \lstinline!s $\neqS$ t! to be in \lstinline!b! for some \lstinline!s,t $\in$ subterms (last b)!.
+The subsumption condition then prohibits that the rule can be applied again for the combination of \lstinline!s!
+and \lstinline!t!, thus yielding an upper bound of \lstinline!(card (subterms (last b)))$^\text{2}$!.
+This allows us to finish the calculation that justifies the second inequality:
+\begin{lstlisting}
+  &\phantom{$\leq$}& 4 * (card (subterms b))$^\text{2}$
+  $\leq$ 4 * (card (subterms (last b))
+  &\phantom{$\leq$}&      + card (wits b))$^\text{2}$
+  $\leq$ 4 * (card (subterms (last b))
+  &\phantom{$\leq$}&      + (card (subterms (last b))$^\text{2}$)$^\text{2}$
+  $\leq$ 4 * 4 * (card (subterms (last b)))$^\text{4}$.
+\end{lstlisting}
 
@@
 
+\begin{lstlisting}
+lemma mlss_proc_branch_dom_if_wf_branch:
+  assumes wf_branch b
+  shows mlss_proc_branch_dom b
+\end{lstlisting}
+
+\begin{lstlisting}
+lemma mlss_proc_branch_complete:
+  fixes b :: "'a branch"
+  assumes wf_branch b
+  assumes $\neg$ mlss_proc_branch b
+  assumes infinite (UNIV :: 'a set)
+  shows $\exists$M. M $\models$ (last b)
+
+theorem mlss_proc_complete:
+  fixes $\phi$ :: 'a pset_fm
+  assumes $\neg$ mlss_proc $\phi$
+  assumes infinite (UNIV :: 'a set)
+  shows $\exists$M. M $\models$ $\phi$
+\end{lstlisting}
+
+\begin{lstlisting}
+lemma mlss_proc_branch_sound:
+  assumes wf_branch b
+  assumes $\forall$$\psi$ $\in$ set b. M $\models$ $\psi$
+  shows $\neg$ mlss_proc_branch b
+
+theorem mlss_proc_sound:
+  assumes M $\models$ $\phi$
+  shows $\neg$ mlss_proc $\phi$
+\end{lstlisting}
+
 \section{Future Work}

### 2022-09-08 paper/paper.tex
Continue

@@
 \subsection{Syntax}
-At the heart of \MLSS{}, we have the type of set terms which includes the empty set and variables as well as the operations union, intersection, set difference, and \lstinline!Single!, which represents a singleton set containing the argument.
+At the heart of \MLSS{}, we have the type of set terms which is the disjoint union of the empty set and variables as well as the operations union, intersection, set difference, and \lstinline!Single!, which represents a singleton set containing the argument.
 \todo[inline]{Syntax for \lstinline!Single!?, overloading of \lstinline!vars_term!, \lstinline!vars_atom!}
@@
 is equivalent to \lstinline!s $\unionS$ t $\eqS$ t!
-and a finite set of elements \lstinline!{t$_1$,t$_2$,$\ldots$,t$_k$}!
+and a finite set of elements \lstinline!{t$_\text{1}$,$\ldots$,t$_\text{k}$}!
 can be expressed by \lstinline!Single t$_1$ $\unionS$ $\ldots$ $\unionS$ Single t$_k$!.
@@
 \end{lstlisting}
-We will sometimes drop the \lstinline!Atom! constructor to reduce clutter.
+We will often drop the atom constructor \lstinline!&\fmAtom{}&! to reduce clutter.
 Additionally, we use the abbreviations \lstinline!s $\notinS$ t!
@@
 \begin{lstlisting}
-fun subterms :: 'a pset_term $\Rightarrow$ 'a pset_term set
-where
+fun subterms :: 'a pset_term
+             $\Rightarrow$ 'a pset_term set where
   subterms $\emptyset$ = {$\emptyset$}
@@
 \begin{lstlisting}
-fun subterms :: 'a pset_atom $\Rightarrow$ 'a pset_term set
-where
+fun subterms :: 'a pset_atom
+             $\Rightarrow$ 'a pset_term set where
   subterms (s $\inS$ t) = subterms s $\cup$ subterms t
@@
       This class only constitutes a set if the class is small enough.
-  \item The predicate \lstinline!small :: 'a set $\Rightarrow$ bool!, which is polymorphic in \lstinline!'a!, indicates whether a set is small.
-Inversely, this predicate characterises those sets that can be turned into a set of type \lstinline!V! using the polymorphic function \lstinline!vset :: 'a set $\Rightarrow$ V!.
+  \item The predicate \lstinline!small :: 'a set $\Rightarrow$ bool!, which is polymorphic in \lstinline!'a!, indicates whether a class is small.
+  \item The function \lstinline!vset :: 'a set $\Rightarrow$ V! turns a class into a set of type \lstinline!V! if it is small. 
 \item The usual set operations such as equality ($=$), union ($\sqcup$), intersection ($\sqcap$), and difference ($-$) are defined.
@@
 For \MLSS{}, we extend the notion of closedness with three additional rules; the first two are straightforward while the last one states that a branch is closed when the branch contains a membership cycle
-\begin{center}
-\lstinline!s $\inS$ t$_\text{1}$, t$_\text{1}$ $\inS$ t$_\text{2}$, $\ldots$, t$_\text{k}$ $\inS$ s!.
-\end{center}
+\begin{lstlisting}
+  t$_\text{0}$ $\inS$ t$_\text{1}$, t$_\text{1}$ $\inS$ t$_\text{2}$, $\ldots$, t$_\text{k}$ $\inS$ t$_\text{0}$.
+\end{lstlisting}
+
 \begin{lstlisting}
@@
 \subsection{Branching Rules\label{sec:branching}}
-Since the number of branches increases when applying branching rules, it is prudent to only apply them to open, linearly saturated branches.
+The number of branches increases when applying branching rules so it is prudent to only apply them to open, linearly saturated branches.
 To further prevent unnecessary branching, each rule has a \textit{subsumption} condition. 
@@
 
-\section{Completeness of the Calculus}
+\section{A Decision Procedure for MLSS}
+\begin{figure*}
+  \centering
+\begin{lstlisting}
+function (domintros) mlss_proc_branch :: 'a branch $\Rightarrow$ bool where
+  $\neg$ lin_sat b
+  $\Longrightarrow$ mlss_proc_branch = mlss_proc_branch ((SOME b'. &\lexpands{b'}{b}& $\land$ set b $\subset$ set (b' @ b)) @ b)
+| $\llbracket$ lin_sat b; bclosed b $\rrbracket$ $\Longrightarrow$ mlss_proc_branch b = True
+| $\llbracket$ $\neg$ sat b; bopen b; lin_sat b $\rrbracket$
+  $\Longrightarrow$ mlss_proc_branch b = ($\forall$b' $\in$ (SOME bs. &\fexpands{bs}{b}&). mlss_proc_branch (b' @ b))
+| $\llbracket$ lin_sat b; sat b $\rrbracket$ $\Longrightarrow$ mlss_proc_branch b = bclosed b
+
+definition mlss_proc :: 'a pset_fm $\Rightarrow$ bool where
+  mlss_proc $\phi$ $\equiv$ mlss_proc_branch [$\phi$]
+\end{lstlisting}
+\caption{Definition of the functions \lstinline|mlss_proc_branch| and \lstinline|mlss_proc|.\label{fig:alg}}
+\end{figure*}
+
+\noindent We abstractly specify the behaviour of the decision procedure, leaving an executable specification from which code can be generated to future work.
+The implementation as displayed in \autoref{fig:alg} uses a couple features of Isabelle's function package:
+instead of defining the function via pattern matching, we specify the equations of the function as conditional rewrite rules.
+This requires us to prove that the assumptions of the equations are non-overlapping, which is done by automation.
+The other concern is that a recursive function needs to terminate in order for it to be well-defined;
+nevertheless, the termination proof is separated from the definition of the function for modularity.
+The function package maintains the soundness of the definition by introducing a so-called domain predicate \lstinline!mlss_proc_branch_dom!
+and adding the assumption that the predicate holds for the argument to each equation of the function.
+To be able to reason about this predicate, we pass the \lstinline!domintros! flag which instructs the function package to derive introduction rules for it.
+
+\section{Completeness of the Calculus\label{sec:complete}}
 Once a branch \lstinline!b! is saturated and open at the same time, we are unable to expand the branch further in our calculus.
@@
 \end{lstlisting}
+\todo[inline]{Non-Overlapping!}
 Since the function is recursive, we need to ensure that it terminates for it to be well-defined.
@@
 \noindent The final step towards obtaining a proper model is to connect the realization function to the semantics as defined in \autoref{sec:semantics}.
-For set terms, we can use the Propositions~\ref{it:empty}--\ref{it:single} to prove the lemma below by induction.
+For set terms, we can use the Propositions~\ref{it:empty}--\ref{it:single} to prove the lemma below by induction on \lstinline!t!.
 \begin{lstlisting}
@@
 
-\section{Soundness of the Calculus}
+\section{Soundness of the Calculus\label{sec:sound}}
 Two aspects determine the soundness of the calculus: a closed branch must be unsatisfiable and all expansion rules must maintain satisfiability.
@@
 \noindent We are left with showing that both linear and branching expansion rules preserve satisfiability.
-As for the linear rules, a straightforward proof by rule induction suffices to obtain the lemma below.
+As for the linear rules, a straightforward proof by case analysis on \lstinline!&\lexpands{b'}{b}&! suffices to obtain the lemma below.
 \begin{lstlisting}
@@
   for some fresh variable \lstinline!x!, i.e.\ \lstinline!x $\notin$ vars b! and in particular \lstinline!x $\notin$ vars s $\cup$ vars t!.
-  Another consequence of applying the last rule is that \lstinline!s $\neqS$ t! is in \lstinline!b!, which, because \lstinline!M! is a model, implies that \lstinline!&\Ist& M s $\neq$ &\Ist& M t!.
+  Another consequence of the last rule being applied is that \lstinline!s $\neqS$ t!
+is in \lstinline!b!, which, because \lstinline!M!
+is a model, implies that \lstinline!&\Ist& M s $\neq$ &\Ist& M t!.
   This inequality manifests itself through some \lstinline!y! with
@@
 
-\section{Upper Bounding the Cardinality of a Branch}
-Soundness and completeness of the calculus are the building blocks with which we establish partial correctness of \autoref{fig:alg}.
+\section{Upper Bounding the Cardinality of a Branch\label{sec:upper}}
+Soundness and completeness of the calculus are the building blocks with which we establish partial correctness of the algorithm in \autoref{fig:alg}.
 Consequently, we still need to give a termination proof to ascertain total correctness.
@@
 
-\section{An Abstract Specification of the Decision Procedure}
-\begin{figure*}
-  \centering
-\begin{lstlisting}
-function mlss_proc_branch :: 'a branch $\Rightarrow$ bool where
-  $\neg$ lin_sat b
-  $\Longrightarrow$ mlss_proc_branch = mlss_proc_branch ((SOME b'. &\lexpands{b'}{b}& $\land$ set b $\subset$ set (b' @ b)) @ b)
-| $\llbracket$ $\neg$ sat b; bopen b; lin_sat b $\rrbracket$
-  $\Longrightarrow$ mlss_proc_branch b = ($\forall$b' $\in$ (SOME bs. &\fexpands{bs}{b}&). mlss_proc_branch (b' @ b))
-| $\llbracket$ lin_sat b; sat b $\rrbracket$ $\Longrightarrow$ mlss_proc_branch b = bclosed b
-| $\llbracket$ lin_sat b; bclosed b $\rrbracket$ $\Longrightarrow$ mlss_proc_branch b = True
-
-definition mlss_proc :: 'a pset_fm $\Rightarrow$ bool where
-  mlss_proc $\phi$ $\equiv$ mlss_proc_branch [$\phi$]
-\end{lstlisting}
-\caption{Definition of the functions \lstinline|mlss_proc_branch| and \lstinline|mlss_proc|.\label{fig:alg}}
-\end{figure*}
+\section{Total Correctness of the Decision Procedure\label{sec:correct}}
 

### 2022-09-09 paper/paper.tex
Continue

@@
 \hyphenation{Isa-belle}
+\hyphenation{tab-leau}
 
@@
 \newcommand{\fexpandsW}[2]{#1 $\lefttrianglebar_\text{v}$ #2}
-\newcommand{\expandss}[2]{#1 $\lefttriangle^*$ #2}
+\newcommand{\expandssSym}{\lefttriangle^*}
+\newcommand{\expandss}[2]{#1 $\expandssSym$ #2}
 
@@
 \begin{abstract}
-  \todo[inline]{Write the abstract}
 \end{abstract}
@@
 \section{Introduction}
-\todo[inline]{Write the introduction, Remember that \MLSS{} is multi-level syllogistics with singleton}
+\todo[inline]{\MLSS{} is multi-level syllogistics with singleton}
 
@@
 \subsection{Contributions}
-\todo[inline]{Write the contributions}
 
@@
 We first define a function that collects the subterms of a set term.
-\todo[inline]{Motivation: linear rules don't introduce subterms, \lstinline!subterms b $\approx$ subterms (last b)!}
+\todo[inline]{Motivation: linear rules don't introduce subterms, \\ \lstinline!subterms b $\approx$ subterms (last b)!}
 \begin{lstlisting}
@@
 
+\begin{lstlisting}
+definition wits :: 'a branch $\Rightarrow$ 'a set where
+  wits b $\equiv$ vars b - vars (last b)
+\end{lstlisting}
+
 \section{A Decision Procedure for MLSS}
@@
 \end{lstlisting}
-\caption{Definition of the functions \lstinline|mlss_proc_branch| and \lstinline|mlss_proc|.\label{fig:alg}}
+\caption{Definition of the decision procedure through the \lstinline|mlss_proc_branch| and \lstinline|mlss_proc|.\label{fig:alg}}
 \end{figure*}
@@
 This requires us to prove that the assumptions of the equations are non-overlapping, which is done by automation.
-The other concern is that a recursive function needs to terminate in order for it to be well-defined;
+The other concern is that Isabelle requires functions to be total so a recursive function needs to terminate in order for it to be well-defined;
 nevertheless, the termination proof is separated from the definition of the function for modularity.
-The function package maintains the soundness of the definition by introducing a so-called domain predicate \lstinline!mlss_proc_branch_dom!
-and adding the assumption that the predicate holds for the argument to each equation of the function.
-To be able to reason about this predicate, we pass the \lstinline!domintros! flag which instructs the function package to derive introduction rules for it.
+The function package maintains the soundness of the definition by introducing a so-called domain predicate \lstinline!mlss_proc_branch_dom! which characterises where the function terminates.
+Each equation of the function receives the assumption that the predicate holds for the argument.
+To be able to reason about this predicate, we pass the \lstinline!domintros! flag that instructs the function package to derive introduction rules for it.
+Before we go into more detail on how the termination is handled, we discuss the definition of \lstinline!mlss_proc_branch!.
+
+The purpose of the function is to take a branch and determine whether the branch can be expanded to a closed tableau.
+As stated before, we first use linear expansion rules in order to prevent premature branching;
+therefore, we recursively expand the branch with linear rules until the branch is linearly saturated.
+Note that we use Hilbert's $\varepsilon$-operator in the form of \lstinline!SOME! to choose some rule that actually adds new formulae to the branch. 
+As soon as the branch is linearly saturated, we terminate if the branch is closed as shown in the second equation.
+Otherwise, we choose an applicable branching rule and recursively check whether all newly created branches can be closed.
+The final equation applies once no further branch expansion is possible, in which case we just tests for closedness of the branch.
+
+Every unsatisfiable formula has a closed tableau so we can determine the satisfiability of a formula \lstinline!$\psi$!
+by passing the singleton branch \lstinline![$\psi$]!
+to \lstinline!mlss_proc_branch!.
+Because this is the only context that \lstinline!mlss_proc_branch! is used in, it only encounters branches that result from applying branch expansion rules to a singleton branch.
+We call this kind of branches \textit{well-formed} branches and define a predicate for it.
+In the definition, the expression \lstinline!&\expandss{b'}{b}&! denotes that \lstinline!b'! is one of the branches that results from applying (potentially zero) expansion rules to \lstinline!b!.
+\begin{lstlisting}
+definition wf_branch b $\equiv$ $\exists$$\phi$. &\expandss{b}{[$\phi$]}&
+\end{lstlisting}
+In particular, we will use this notion in \autoref{sec:bound} where derive an upper bound for the cardinality of a well-formed branch.
+This is then used in \autoref{sec:correct} to justify the termination of the decision procedure. 
+Before we come to that, though, we prove soundness and completeness in Section~\ref{sec:sound} and~\ref{sec:complete}, respectively.
+In \autoref{sec:correct}, we also show that both properties easily transfer to the decision procedure which, together with termination, give us total correctness.
 
@@
 \begin{lstlisting}
-definition wits :: 'a branch $\Rightarrow$ 'a set where
-  wits b $\equiv$ vars b - vars (last b)
-
 definition wits' :: 'a branch $\Rightarrow$ 'a set where
@@
 \todo[inline]{Non-Overlapping!}
-Since the function is recursive, we need to ensure that it terminates for it to be well-defined.
+Again, we need to ensure that the function terminates for it to be well-defined.
 The assumption that \lstinline!b! is open implies that there are no membership cycles and thus \lstinline!bgraph b! is acyclic.
@@
     $\longrightarrow$ t $\in$ subterms (last b))
-\end{lstlisting}
-
-\begin{lstlisting}
-definition wf_branch b $\equiv$ $\exists$$\phi$. &\expandss{b}{[$\phi$]}&
 \end{lstlisting}
@@
 
-\section{Upper Bounding the Cardinality of a Branch\label{sec:upper}}
+\section{Upper Bounding the Cardinality of a Branch\label{sec:bound}}
 Soundness and completeness of the calculus are the building blocks with which we establish partial correctness of the algorithm in \autoref{fig:alg}.
@@
 \section{Conclusion}
-\todo[inline]{Conclude}
 

### 2022-09-12 paper/paper.tex
Continue

@@
 Because this is the only context that \lstinline!mlss_proc_branch! is used in, it only encounters branches that result from applying branch expansion rules to a singleton branch.
-We call this kind of branches \textit{well-formed} branches and define a predicate for it.
+We call this kind of branches \textit{well-formed}.
 In the definition, the expression \lstinline!&\expandss{b'}{b}&! denotes that \lstinline!b'! is one of the branches that results from applying (potentially zero) expansion rules to \lstinline!b!.
@@
 Before we come to that, though, we prove soundness and completeness in Section~\ref{sec:sound} and~\ref{sec:complete}, respectively.
-In \autoref{sec:correct}, we also show that both properties easily transfer to the decision procedure which, together with termination, give us total correctness.
+In \autoref{sec:correct}, we also show that both properties easily transfer to the decision procedure which, together with termination, gives us total correctness.
 
@@
 
-theorem mlss_proc_complete:
-  fixes $\phi$ :: 'a pset_fm
-  assumes $\neg$ mlss_proc $\phi$
-  assumes infinite (UNIV :: 'a set)
-  shows $\exists$M. M $\models$ $\phi$
 \end{lstlisting}
@@
   shows $\neg$ mlss_proc_branch b
+\end{lstlisting}
+
+\begin{lstlisting}
+theorem mlss_proc_complete:
+  fixes $\phi$ :: 'a pset_fm
+  assumes $\neg$ mlss_proc $\phi$
+  assumes infinite (UNIV :: 'a set)
+  shows $\exists$M. M $\models$ $\phi$
 

### 2022-09-13 paper/paper.tex
Continue

@@
 Once a branch \lstinline!b! is saturated and open at the same time, we are unable to expand the branch further in our calculus.
-Thus, to establish completeness of the calculus, we must construct a model \lstinline!M! such that \lstinline!M $\models$ $\phi$! for all formulae \lstinline!$\phi$ $\in$ set b!.
-The model is defined in terms of the following three collections.
-\todo[inline]{explain wits, explain that wits' are those whose value is not fixed by literals in b}
+Thus, to establish completeness of the calculus, we must construct a model \lstinline!M! such that \lstinline!M $\models$ $\phi$! for all \lstinline!$\phi$ $\in$ set b!.
+
+At the core of the model, there is a \textit{realization} function that maps set terms to sets of type \lstinline!V!.
+A subset of the witnesses, which we call \textit{pure} witnesses, receives special treatment from the realization function for reasons that will become apparent in \autoref{sec:pwits}.
+The set terms of a branch can thus be partitioned into two collections as defined below.
 \begin{lstlisting}
@@
 \end{lstlisting}
-We aim to construct a syntactic model that is derived from the membership literals \lstinline!(s $\inS$ t) $\in$ set b!.
-To this end, we construct a graph whose vertices are set terms and there is an edge from \lstinline!s! to \lstinline!t! in the graph if, and only if, \lstinline!s $\inS$ t! is in \lstinline!b!.
+
+We aim to construct a syntactic model that is derived from the membership literals \lstinline!s $\inS$ t! in the branch.
+To this end, we construct a graph whose vertices are the disjoint union of the sets above and there is an edge from \lstinline!s!
+to \lstinline!t!
+in the graph if, and only if, \lstinline!s $\inS$ t!
+is in \lstinline!b!.
 Note that we use \citeauthor{graph_theory_afp}'s graph library~\cite{graph_theory_afp} where a graph is represented as a record of vertices, arcs (directed edges), and two functions \lstinline!tail!
@@
 that map an arc to its source respectively target vertex.
-\todo[inline]{Explain arrow notation}
 \begin{lstlisting}
@@
 \end{lstlisting}
-The model is defined in terms of a \textit{realization} function relative to this graph.
-Observe that the vertices are partitioned into two sets, a collection of witnesses \lstinline!Var ` wits' b!
-and the rest of the set terms \lstinline!subterms' b!, which are treated differently by the realization.
+
+The realization function is defined relative to this graph.
+As mentioned before, the pure witnesses \lstinline!Var ` wits' b! and the rest of the set terms \lstinline!subterms' b! are treated differently by the realization function.
 Terms in the latter set are evaluated in accordance to the structure of the graph, i.e.\ the realization of a vertex is defined as the union of the realizations of the parent vertices. 
@@
 We can always choose such a function since we assume an infinite universe of variables.
-Then, we return the singleton set \lstinline!vset {I x}!, which, together with the cardinality constraint, ensures that realizations are distinct between those witnesses and the set terms.  
+Then, we return the singleton set \lstinline!vset {I x}!, which, together with the cardinality constraint, ensures that realizations are distinct between pure witnesses themselves as well as between pure witnesses and set terms.
+The notation \lstinline!u $\rightarrow_\text{G}$ s! in the definition below indicates that there is an edge from \lstinline!u! to \lstinline!s! in the graph \lstinline!G!.
 \begin{lstlisting}
@@
   $\Longrightarrow$ realize t =
-  &$\phantom{\Longrightarrow}$&   vset {realize ` parents (bgraph b) s}
+  &\phantom{$\Longrightarrow$ }&  vset {realize ` parents (bgraph b) s}
 | x $\notin$ verts G $\Longrightarrow$ realize x = 0
 \end{lstlisting}
-\todo[inline]{Non-Overlapping!}
-Again, we need to ensure that the function terminates for it to be well-defined.
+
+Again, we need to ensure that the assumptions of the equations are non-overlapping and that function terminates.
+The former is taken care of by automation, leaving us to prove termination.
 The assumption that \lstinline!b! is open implies that there are no membership cycles and thus \lstinline!bgraph b! is acyclic.
@@
 Together with the fact that \lstinline!bgraph b! is finite, it follows that the cardinality of the ancestors decreases in each recursive call thus proving the termination of \lstinline!realize!.
-\todo[inline]{Überleitung}
 
-\subsection{Formalising the Key Technical Lemma}
+Before we prove that the realization function constitutes a model in \autoref{sec:real}, we will first explain the significance of the pure witnesses.
+
+\subsection{Characterisation of the Pure Witnesses\label{sec:pwits}}
+Recall that the pure witnesses for a branch \lstinline!b! are those witnesses that are not related to other subterms in \lstinline!last b! by equality.
+In the context of a well-formed branch, this characterisation can be strengthened to any set term and, in addition, we also get that there is no membership literal where a pure witness is on the right-hand side.
+Intuitively speaking, the realization of a pure witness does not depend on the realization of any other set term.
+\begin{lstlisting}
+lemma lemma_2:
+  assumes wf_branch b
+  assumes c $\in$ wits' b
+  shows (Var c $\eqS$ t) $\notin$ set b
+    and (t $\eqS$ Var c) $\notin$ set b
+    and (t $\inS$ Var c) $\notin$ set b
+\end{lstlisting}
+So why are pure witnesses treated differently?
+According to the definition of \lstinline!realize!, the pure witnesses would be evaluated to the empty set \lstinline!0 :: V!, were they not treated separately. 
+To see that this is a problem, consider the branch
+\begin{lstlisting}
+  b = [Var s $\neqS$ Var t, Var t $\neqS$ Var u]
+\end{lstlisting}
+which expands to several open and saturated branches, one of which is
+\begin{lstlisting}
+  [Var x $\inS$ Var s, Var x $\notinS$ Var t,
+   Var y $\inS$ Var t, Var y $\notinS$ Var u] @ b
+\end{lstlisting}
+for some fresh \lstinline!x! and \lstinline!y!.
+Assigning both \lstinline!Var x! and \lstinline!Var y! a value of \lstinline!0! would result in the realization of \lstinline!Var s! and \lstinline!Var t! becoming equal;
+therefore, the realization would not be a model for \lstinline!Var s $\neqS$ Var t!.
+To prevent this, we assign the pure witnesses pairwise different values.
+
 \begin{lstlisting}
@@
 
-\todo[inline]{Rename \lstinline!lemma_2!}
-\begin{lstlisting}
-lemma lemma_2:
-  assumes wf_branch b
-  assumes c $\in$ wits' b and t $\in$ subterms b
-  shows (Var c $\eqS$ t) $\notin$ set b
-    and (t $\eqS$ Var c) $\notin$ set b
-    and (t $\inS$ Var c) $\notin$ set b
-\end{lstlisting}
-Only one quite trivial invariant needed
-
 \begin{lstlisting}
@@
 
-\subsection{Realisation of an Open Branch}
-Remember that for completeness, we need to show that the realization function for an open and saturated branch \lstinline!b! actually constitutes a model for all formualae in the branch.
+\subsection{Realisation of an Open Branch\label{sec:real}}
+Remember that for completeness, we need to show that the realization function for an open and saturated branch \lstinline!b! actually constitutes a model for all formulae in the branch.
 We start by verifying that the realization function models all literals in the branch; more formally, the following propositions hold: 
@@
   assumes infinite (UNIV :: 'a set)
-  shows $\exists$M. M $\models$ (last b)
+  shows $\exists$M. M $\models$ last b
 

### 2022-09-14 paper/paper.tex
Continue

@@
 \begin{abstract}
+  Using Isabelle/HOL, we verify the state-of-the-art decision procedure for multi-level syllogistics with singleton (\MLSS{} for short), which is a ground fragment of set theory.
+  We formalise the syntax and semantics of \MLSS{} as well as a sound and complete tableau calculus for it.
+  We also provide an abstract specification of a decision procedure that applies the rules of the calculus exhaustively and prove its termination.
 \end{abstract}
@@
 The unquantified fragment \MLSS{} of set theory is then formed by combining set atoms, which are injected into formulae with the constructor \lstinline!&\fmAtom&!, with propositional logical connectives.
-\todo[inline]{Cite Nipkow's Linear Quantifier Elimination?}
+\todo[inline]{Cite Nipkow's Linear Quantifier Elimination?, literals}
 \begin{lstlisting}
@@
 type_synonym 'a pset_fm = 'a pset_atom fm
+\end{lstlisting}
+
+\begin{lstlisting}
+fun is_literal :: 'a fm $\Rightarrow$ bool where
+  is_literal (&\fmAtom& _) = True
+| is_literal (&\fmNeg{(\fmAtom{} \_)}&) = True
+| is_literal _ = False
 \end{lstlisting}
@@
 \end{lstlisting}
-\caption{Definition of the decision procedure through the \lstinline|mlss_proc_branch| and \lstinline|mlss_proc|.\label{fig:alg}}
+\caption{Definition of the decision procedure through \lstinline|mlss_proc_branch| and \lstinline|mlss_proc|.\label{fig:alg}}
 \end{figure*}
@@
 Intuitively speaking, the realization of a pure witness does not depend on the realization of any other set term.
-\begin{lstlisting}
+\begin{lstlisting}[label={lst:lemma_2}]
 lemma lemma_2:
@@
 \begin{lstlisting}
-  [Var x $\inS$ Var s, Var x $\notinS$ Var t,
+  [Var x $\neqS$ Var y,
+   Var x $\inS$ Var s, Var x $\notinS$ Var t,
    Var y $\inS$ Var t, Var y $\notinS$ Var u] @ b
@@
 for some fresh \lstinline!x! and \lstinline!y!.
-Assigning both \lstinline!Var x! and \lstinline!Var y! a value of \lstinline!0! would result in the realization of \lstinline!Var s! and \lstinline!Var t! becoming equal;
-therefore, the realization would not be a model for \lstinline!Var s $\neqS$ Var t!.
+Assigning both \lstinline!Var x! and \lstinline!Var y! a value of \lstinline!0! would contradict the literal \lstinline!Var x $\neqS$ Var y!.
 To prevent this, we assign the pure witnesses pairwise different values.
 
-\begin{lstlisting}
-fun is_literal :: 'a fm $\Rightarrow$ bool where
-  is_literal (&\fmAtom& _) = True
-| is_literal (&\fmNeg{(\fmAtom{} \_)}&) = True
-| is_literal _ = False
-\end{lstlisting}
-
-\begin{lstlisting}
-definition no_new_subterms b $\equiv$
-  $\forall$t $\in$ subterms b. t $\notin$ Var ` wits b
-    $\longrightarrow$ t $\in$ subterms (last b))
-\end{lstlisting}
+The proof of \hyperref[lst:lemma_2]{\lstinline!lemma_2!} is more technical than interesting so we refer the reader to the formalisation.
 
@@
 \end{enumerate}
-To illustrate the usefulness of \lstinline!lemma_2!, we prove Proposition~\ref{it:at_eq}.
+To illustrate the usefulness of \hyperref[lst:lemma_2]{\lstinline!lemma_2!}, we prove Proposition~\ref{it:at_eq}.
 The proofs of all propositions translate well into Isabelle so we refer to the original paper~\cite{new_fast_tableau} for the remaining proofs. 
@@
   Assume that \lstinline!s $\eqS$ t! is in \lstinline!b!.
-  If there exists a \lstinline!c $\in$ wits' b! where \lstinline!s = Var c! or \lstinline!t = Var c!, we arrive at a contradiction due to \lstinline!lemma_2!.
+  If there exists a \lstinline!c $\in$ wits' b! where \lstinline!s = Var c! or \lstinline!t = Var c!, we arrive at a contradiction due to \hyperref[lst:lemma_2]{\lstinline!lemma_2!}.
   Therefore, both \lstinline!s $\in$ subterms' b! and \lstinline!t $\inS$ subterms' b! must hold.
@@
 \begin{enumerate}[label=(\alph*)]
-  \item\label{it:empty} If \lstinline!$\emptyset$ $\in$ subterms b!, then \lstinline!realize $\emptyset$ = 0!.
+  \item\label{it:empty} It holds that \lstinline!realize $\emptyset$ = 0!.
   \item\label{it:op} Let \lstinline!$\star_\text{s}$ $\in$ {$\unionS$, $\diffS$, $\interS$}!. If the term \lstinline!s $\star_\text{s}$ t! occurs in \lstinline!subterms b!, then
@@
 \begin{proof}[Proof of Proposition~\ref{it:empty}]
-  Since \lstinline!b!
+  By definition, \lstinline!realize $\emptyset$! is equal to \lstinline!0! if \lstinline!$\emptyset$! is not a vertex of \lstinline!bgraph b!.
+  Otherwise, \lstinline!$\emptyset$! must be in \lstinline!subterms' b! as it is not a witness.
+  Moreover, since \lstinline!b!
 is open, we know \lstinline!s $\inS$ $\emptyset$!
@@
 has no parents in \lstinline!bgraph b!.
-  Moreover, the assumption \lstinline!$\emptyset$ $\in$ subterms b! can be strengthened to \lstinline!$\emptyset$ $\in$ subterms' b! as the term \lstinline!$\emptyset$! is not a witness. 
-  Ultimately, we obtain \lstinline!realize $\emptyset$ = 0! by definition of \lstinline!realize!.
+  Ultimately, we obtain \lstinline!realize $\emptyset$ = 0!.
 \end{proof}
@@
 \end{lstlisting}
-  We therefore have \lstinline!(u $\inS$ s $\diffS$ t) $\in$ set b!
-which, due to \lstinline!b!
-being saturated, implies that \lstinline!u $\inS$ s!
-is in \lstinline!b!
-while \lstinline!u $\inS$ t!
-is not.
+  We therefore have \lstinline!(u $\inS$ s $\diffS$ t) $\in$ set b! which, due to \lstinline!b! being saturated, implies that \lstinline!u $\inS$ s! and \lstinline!u $\notinS$ t! are in \lstinline!b!.
+  \todo[inline]{Explain latter of line above}
   By applying Proposition~\ref{it:at_mem} and~\ref{it:af_mem} we obtain
@@
 The proof is a tedious but straightforward induction on the the size of the formulae.
-\begin{lstlisting}
+\begin{lstlisting}[label={lst:coherence}]
 lemma coherence:
@@
 We can express the former aspect in Isabelle as follows.
-\begin{lstlisting}[belowskip=0pt]
+\begin{lstlisting}[belowskip=0pt,label={lst:bclosed_sound}]
 lemma bclosed_sound:
@@
 As for the linear rules, a straightforward proof by case analysis on \lstinline!&\lexpands{b'}{b}&! suffices to obtain the lemma below.
-\begin{lstlisting}
+\begin{lstlisting}[label={lst:lexpands_sound}]
 lemma lexpands_sound:
@@
 Those variables need to be assigned specific values; hence, we modify the model as we will see in the proof below.
-\begin{lstlisting}[belowskip=0pt]
-lemma fexpands_sound:
+\begin{lstlisting}[belowskip=0pt, label={lst:bexpands_sound}]
+lemma bexpands_sound:
   assumes &\fexpands{bs'}{b}&
@@
 \end{lstlisting}
+All in all, we get the following lemma.
+\begin{lstlisting}[label={lst:card_branch}]
+lemma card_wf_branch_ub:
+  assumes wf_branch b
+  shows card (set b)
+        $\leq$ 2 * card (subfms (last b)) +
+        &\phantom{$\leq$}& 16 * (card (subterms (last b)))$^\text{4}$
+\end{lstlisting}
 
 \section{Total Correctness of the Decision Procedure\label{sec:correct}}
-
-\begin{lstlisting}
+We first prove the termination of the procedure for well-formed branches, i.e.\ every well-formed branch is in the domain of \lstinline!mlss_proc_branch!.
+\begin{lstlisting}[belowskip=0pt]
 lemma mlss_proc_branch_dom_if_wf_branch:
@@
 \end{lstlisting}
+\begin{proof}
+  We proceed by strong induction on the difference between the upper bound due to \hyperref[lst:card_branch]{\lstinline!card_wf_branch_ub!} and \lstinline!card b!.
 
+  We only have one inductive case where the induction hypothesis holds for all well-formed branches \lstinline!b'!
+where the difference between the upper bound and the cardinality is smaller than that of \lstinline!b!.
+  We strengthen those assumptions to obtain
 \begin{lstlisting}
+  $\llbracket$ expandss b' b; set b $\subset$ set b' $\rrbracket$
+  $\Longrightarrow$ mlss_proc_branch_dom b'
+\end{lstlisting}
+for all \lstinline!b'!.
+The former strengthening is admissable since, by definition, every expansion of a well-formed branch is again well-formed while the latter ensures that the cardinality of \lstinline!b'!
+is larger than the cardinality of \lstinline!b!, ensuring that the difference to the upper bound decreases.
+
+Now, consider the case where \lstinline!b! is saturated.
+Then, the procedure terminates without recursive calls and we immediately get \lstinline!mlss_proc_branch_dom b! using the corresponding introduction rule for the domain predicate.
+
+If \lstinline!b! is not saturated, we can expand \lstinline!b! using a linear or a branching rule, which adds new formulae to the resulting branch(es).
+In the first case, we can thus, using \lstinline!SOME!, choose a branch \lstinline!b'! that linearly expands \lstinline!b!.
+Moreover, the domain predicate holds for \lstinline!b' @ b! since the assumptions of the modified induction hypothesis are fulfilled.
+Ultimately, we use the corresponding introduction rule for the domain predicate to obtain \lstinline!mlss_proc_branch b!.
+
+The case of a branching rule is analogous.
+\end{proof}
+
+\noindent The above lemma allows us to use the computation induction rule of \lstinline!mlss_proc_branch!
+on well-formed branches, which we do to prove soundness and completeness, starting with completeness.
+\begin{lstlisting}[belowskip=0pt]
 lemma mlss_proc_branch_complete:
-  fixes b :: "'a branch"
+  fixes b :: 'a branch
   assumes wf_branch b
@@
   shows $\exists$M. M $\models$ last b
+\end{lstlisting}
+\begin{proof}
+  We prove the completeness by computation induction on \lstinline!b!.
 
+  The case of the first and third equation, where the branch can be expanded, are immediately resolved with the induction hypothesis.
+
+  We evaluate \lstinline!mlss_proc_branch b! to \lstinline!True! in the case of the third equation, yielding a contradiction with the assumption \lstinline!$\neg$ mlss_proc_branch b!.
+
+  This leaves us with the final case where we assume that the branch is saturated and \lstinline!mlss_proc_branch b! evaluates to \lstinline!bclosed b!.
+  The branch, however, must be open due to the assumption \lstinline!$\neg$ mlss_proc_branch b!.
+  This allows us to use the Lemma~\hyperref[lst:coherence]{\lstinline!coherence!} to obtain the model
+\begin{lstlisting}
+  ($\lambda$x. realize (Var x)) $\models$ last b.
 \end{lstlisting}
+\end{proof}
 
-\begin{lstlisting}
+\noindent With completeness out of the way, only soundness remains to be proven. 
+\begin{lstlisting}[belowskip=0pt]
 lemma mlss_proc_branch_sound:
@@
 \end{lstlisting}
+\begin{proof}
+  We proceed by assuming \lstinline!mlss_proc_branch b! and deriving \lstinline!False!.
+  Using the assumptions, we can prove that there exists \lstinline!b'! such that
+\begin{lstlisting}
+  expandss b' b $\land$
+  ($\exists$M. $\forall$$\psi$ $\in$ set b'. M $\models$ $\psi$) $\land$ bclosed b'
+\end{lstlisting}
+  by computation induction on \lstinline!b!.
 
+  In the case where we choose a branching rule to expand \lstinline!b!
+with the branches in \lstinline!bs'!, we use Lemma~\hyperref[lst:lexpands_sound]{\lstinline!bexpands_sound!} to obtain \lstinline!b' $\in$ bs'!
+and \lstinline!M'!
+with
+\begin{lstlisting}
+  $\forall$$\psi$ $\in$ set (b' @ b). M' $\models$ $\psi$.
+\end{lstlisting}
+  Together with the induction hypothesis, we obtain a \lstinline!b''! where 
+\begin{lstlisting}
+  expandss b'' (b' @ b) $\land$
+  ($\exists$M. $\forall$$\psi$ $\in$ set b''. M $\models$ $\psi$) $\land$ bclosed b''.
+\end{lstlisting}
+  Due to the transitivity of \lstinline!expandss!, the above also holds if we replace \lstinline!b' @ b! with \lstinline!b!, which is our goal.
+
+  The case of a linear expansion is dealt with in an analogous manner using Lemma~\hyperref[lst:lexpands_sound]{\lstinline!lexpands_sound!}.
+
+  The two remaining cases are trivial as the branch is already closed, finishing the proof of the proposition.
+
+  This proposition, however, is a contradiction to the statement of Lemma~\hyperref[lst:bclosed_sound]{\lstinline!bclosed_sound!}: no closed branch can have a model. 
+\end{proof}
+
+\noindent To finish of the proof of total correctness, note that every singleton branch is trivially well-formed so termination, completeness, and soundless easily transfer to \lstinline!mlss_proc!.
 \begin{lstlisting}

### 2022-09-15 paper/paper.tex
Continue

@@
 \usepackage{enumitem}
+
+\usepackage{array}
+\usepackage{multirow}
 
@@
 \makeatother
+\newcommand{\lstinlinew}[2][]{\lstinline[#1]{#2}}
 % This is a hack to allow escaping inside lstinline (https://tex.stackexchange.com/questions/43526/escaping-in-lstinline)
@@
 \subsection{Contributions}
+Fill the proof of Lemma 2. More explicit proof of the cardinality of a branch.
 
@@
 \end{lstlisting}
+
+\begin{table*}
+  \lstset{
+    escapeinside={@}{@}
+  }
+  \newcommand{\ton}{\text{t}_\text{1}}
+  \newcommand{\ttw}{\text{t}_\text{2}}
+  \begin{tabular}{p{0.21\textwidth}cp{0.18\textwidth} p{0.01\textwidth} p{0.21\textwidth}cp{0.18\textwidth}}
+    \multicolumn{3}{c}{Propositional Rules} && \multicolumn{3}{c}{Rules for \lstinline!$\unionS$!} \\
+    \cmidrule{1-3}\cmidrule{5-7}
+    \lstinline!@\fmAnd{p}{q}@! & $\Longrightarrow$ & \lstinline!p, q! &&
+    \lstinline!s $\notinS$ $\ton$ $\unionS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$, s $\notinS$ $\ttw$! \\
+
+    \lstinline!@\fmNeg{(\fmOr{p}{q})}@! & $\Longrightarrow$ & \lstinline!@\fmNeg{p}@, @\fmNeg{q}@! &&
+    \lstinline!s $\inS$ $\ton$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$ $\unionS$ $\ttw$! \\
+
+    \lstinline!@\fmOr{p}{q}, \fmNeg{p}@! & $\Longrightarrow$ & \lstinline!q! &&
+    \lstinline!s $\in$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$ $\unionS$ $\ttw$! \\
+
+    \lstinline!@\fmOr{p}{q}, \fmNeg{q}@! & $\Longrightarrow$ & \lstinline!p! &&
+    \lstinline!s $\inS$ $\ton$ $\unionS$ $\ttw$, s $\notinS$ $\ton$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ttw$! \\
+
+    \lstinline!@\fmNeg{(\fmOr{p}{q})}@, p! & $\Longrightarrow$ & \lstinline!@\fmNeg{q}@! &&
+    \lstinline!s $\inS$ $\ton$ $\unionS$ $\ttw$, s $\notinS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$! \\
+
+    \lstinline!@\fmNeg{(\fmOr{p}{q})}@, q! & $\Longrightarrow$ & \lstinline!@\fmNeg{p}@! &&
+    \lstinline!s $\notinS$ $\ton$, s $\notinS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$ $\unionS$ $\ttw$! \\
+    \lstinline!@\fmNeg{(\fmNeg{p})}@! & $\Longrightarrow$ & \lstinline!p! &&&& \\[3ex]
+
+    \multicolumn{3}{c}{Rules for \lstinline!$\interS$!} && \multicolumn{3}{c}{Rules for \lstinline!$\diffS$!} \\
+    \cmidrule{1-3}\cmidrule{5-7}
+    \lstinline!s $\inS$ $\ton$ $\interS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$, s $\inS$ $\ttw$! &&
+    \lstinline!s $\inS$ $\ton$ $\diffS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$, s $\notinS$ $\ttw$! \\
+
+    \lstinline!s $\notinS$ $\ton$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$ $\interS$ $\ttw$! && 
+    \lstinline!s $\notinS$ $\ton$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$ $\diffS$ $\ttw$! \\
+
+    \lstinline!s $\notinS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$ $\interS$ $\ttw$! &&
+    \lstinline!s $\inS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$ $\diffS$ $\ttw$! \\
+
+    \lstinline!s $\notinS$ $\ton$ $\interS$ $\ttw$, s $\inS$ $\ton$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ttw$! && 
+    \lstinline!s $\notinS$ $\ton$ $\diffS$ $\ttw$, s $\inS$ $\ton$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ttw$! \\
+
+    \lstinline!s $\notinS$ $\ton$ $\interS$ $\ttw$, s $\inS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$! && 
+    \lstinline!s $\notinS$ $\ton$ $\diffS$ $\ttw$, s $\notinS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$! \\
+
+    \lstinline!s $\inS$ $\ton$, s $\inS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$ $\interS$ $\ttw$! &&
+    \lstinline!s $\inS$ $\ton$, s $\notinS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$ $\diffS$ $\ttw$! \\[3ex]
+
+    \multicolumn{3}{c}{Rules for \lstinline!Single!} && \multicolumn{3}{c}{Rules for \lstinline!$\eqS$!} \\
+    \cmidrule{1-3}\cmidrule{5-7}
+    & $\Longrightarrow$ & \lstinline!s $\inS$ Single s! &&
+    \lstinline!$\ton$ $\eqS$ $\ttw$, ($\fmNegSymbol$) a! & $\Longrightarrow$ & \lstinline!($\fmNegSymbol$) a[$\ttw$/$\ton$]! \\
+
+    \lstinline!s $\inS$ Single t! & $\Longrightarrow$ & \lstinline!s $\eqS$ t! &&
+    \lstinline!$\ton$ $\eqS$ $\ttw$, ($\fmNegSymbol$) a! & $\Longrightarrow$ & \lstinline!($\fmNegSymbol$) a[$\ton$/$\ttw$]! \\
+
+    \lstinline!s $\notinS$ Single t! & $\Longrightarrow$ & \lstinline!s $\neqS$ t! &&
+    \lstinline!s$_\text{1}$ $\inS$ t, s$_\text{2}$ $\inS$ t! & $\Longrightarrow$ & \lstinline!s$_\text{1}$ $\neqS$ s$_\text{2}$! \\
+  \end{tabular}
+  \bigskip
+  \caption{Linear expansion rules.\label{tab:lexpands}}
+\end{table*}
 \todo[inline]{Reference figure and mention missing rule + example}
@@
 \end{lstlisting}
+
+\begin{table*}
+  \lstset{
+    escapeinside={@}{@}
+  }
+  \newcommand{\ton}{\text{t}_\text{1}}
+  \newcommand{\ttw}{\text{t}_\text{2}}
+  \newcommand{\lpipe}{\rule[-0.4ex]{0.6pt}{2.3ex}}
+  \begin{tabular}{>{\centering\arraybackslash}p{0.28\textwidth} >{\centering\arraybackslash}p{0.3\textwidth} >{\centering\arraybackslash}p{0.3\textwidth}}
+    Rule & Precondition & Subsumption condition \\
+    \midrule
+    \begin{tabular}{@{}>{\raggedleft\arraybackslash}p{0.11\textwidth}c>{\raggedright\arraybackslash}p{0.11\textwidth}@{}}
+      \\
+      \midrule
+      \lstinline!p! & \lpipe & \lstinline!@\fmNeg{p}@! \\
+    \end{tabular}
+         & Test & \lstinline!p $\in$ set b $\lor$ @\fmNeg{p}@ $\in$ set b! \\
+
+    \begin{tabular}{@{}>{\raggedleft\arraybackslash}p{0.11\textwidth}c>{\raggedright\arraybackslash}p{0.11\textwidth}@{}}
+      \\
+      \midrule
+      \lstinline!@\fmNeg{p}@! & \lpipe & \lstinline!p! \\
+    \end{tabular}
+         & Test & \lstinline!@\fmNeg{p}@ $\in$ set b $\lor$ p $\in$ set b! \\
+
+    \begin{tabular}{@{}>{\raggedleft\arraybackslash}p{0.11\textwidth}c>{\raggedright\arraybackslash}p{0.11\textwidth}@{}}
+      \\
+      \midrule
+      \lstinline!s $\inS$ $\ton$! & \lpipe & \lstinline!s $\notinS$ $\ton$! \\
+    \end{tabular}
+         & Test &
+         \begin{tabular}{@{}rl@{}}
+                  & \lstinline!(s $\inS$ $\ton$) $\in$ set b! \\
+           $\lor$ & \lstinline!(s $\notinS$ $\ton$) $\in$ set b!
+         \end{tabular}
+         \\
+
+    \begin{tabular}{@{}>{\raggedleft\arraybackslash}p{0.11\textwidth}c>{\raggedright\arraybackslash}p{0.11\textwidth}@{}}
+      \\
+      \midrule
+      \lstinline!s $\inS$ $\ttw$! & \lpipe & \lstinline!s $\notinS$ $\ttw$! \\
+    \end{tabular}
+         & Test &
+         \begin{tabular}{@{}rl@{}}
+                  & \lstinline!(s $\inS$ $\ttw$) $\in$ set b! \\
+           $\lor$ & \lstinline!(s $\notinS$ $\ttw$) $\in$ set b! \\
+        \end{tabular}
+        \\
+
+    \begin{tabular}{@{}>{\raggedleft\arraybackslash}p{0.11\textwidth}c>{\raggedright\arraybackslash}p{0.11\textwidth}@{}}
+      \\
+      \midrule
+      \lstinline!s $\inS$ $\ttw$! & \lpipe & \lstinline!s $\notinS$ $\ttw$! \\
+    \end{tabular}
+         & Test &
+         \begin{tabular}{@{}rl@{}}
+                  & \lstinline!(s $\inS$ $\ttw$) $\in$ set b! \\
+           $\lor$ & \lstinline!(s $\notinS$ $\ttw$) $\in$ set b! \\
+         \end{tabular} \\
+    \begin{tabular}{@{}>{\raggedleft\arraybackslash}p{0.11\textwidth}c>{\raggedright\arraybackslash}p{0.11\textwidth}@{}}
+      \\
+      \midrule
+      \lstinline!Var x $\inS$ $\ton$! \lstinline!Var x $\notinS$ $\ttw$! & \multirow{2}{*}{\rule[-0.4ex]{0.6pt}{4.8ex}} & \lstinline!Var x $\notinS$ $\ton$! \lstinline!Var x $\inS$ $\ttw$! \\
+    \end{tabular}
+         & Test & Test \\
+  \end{tabular}
+  \bigskip
+  \caption{Branching expansion rules.\label{tab:bexpands}}
+\end{table*}
 
@@
 
-At the core of the model, there is a \textit{realization} function that maps set terms to sets of type \lstinline!V!.
-A subset of the witnesses, which we call \textit{pure} witnesses, receives special treatment from the realization function for reasons that will become apparent in \autoref{sec:pwits}.
+At the core of the model, there is a \textit{realisation} function that maps set terms to sets of type \lstinline!V!.
+A subset of the witnesses, which we call \textit{pure} witnesses, receives special treatment from the realisation function for reasons that will become apparent in \autoref{sec:pwits}.
 The set terms of a branch can thus be partitioned into two collections as defined below.
@@
 
-The realization function is defined relative to this graph.
-As mentioned before, the pure witnesses \lstinline!Var ` wits' b! and the rest of the set terms \lstinline!subterms' b! are treated differently by the realization function.
-Terms in the latter set are evaluated in accordance to the structure of the graph, i.e.\ the realization of a vertex is defined as the union of the realizations of the parent vertices. 
+The realisation function is defined relative to this graph.
+As mentioned before, the pure witnesses \lstinline!Var ` wits' b! and the rest of the set terms \lstinline!subterms' b! are treated differently by the realisation function.
+Terms in the latter set are evaluated in accordance to the structure of the graph, i.e.\ the realisation of a vertex is defined as the union of the realisations of the parent vertices. 
 For the former set, we choose a function \lstinline!I! that assigns the witnesses in \lstinline!Var ` wits' b! pairwise distinct sets with cardinality greater than that of the vertices.
 We can always choose such a function since we assume an infinite universe of variables.
-Then, we return the singleton set \lstinline!vset {I x}!, which, together with the cardinality constraint, ensures that realizations are distinct between pure witnesses themselves as well as between pure witnesses and set terms.
+Then, we return the singleton set \lstinline!vset {I x}!, which, together with the cardinality constraint, ensures that realisations are distinct between pure witnesses themselves as well as between pure witnesses and set terms.
 The notation \lstinline!u $\rightarrow_\text{G}$ s! in the definition below indicates that there is an edge from \lstinline!u! to \lstinline!s! in the graph \lstinline!G!.
@@
 
-function realize :: 'a pset_term $\Rightarrow$ V where
-  x $\in$ Var ` wits' b $\Longrightarrow$ realize x = vset {I x}
+function realise :: 'a pset_term $\Rightarrow$ V where
+  x $\in$ Var ` wits' b $\Longrightarrow$ realise x = vset {I x}
 | x $\in$ subterms' b
-  $\Longrightarrow$ realize t =
-  &\phantom{$\Longrightarrow$ }&  vset {realize ` parents (bgraph b) s}
-| x $\notin$ verts G $\Longrightarrow$ realize x = 0
+  $\Longrightarrow$ realise t =
+  &\phantom{$\Longrightarrow$ }&  vset {realise ` parents (bgraph b) s}
+| x $\notin$ verts G $\Longrightarrow$ realise x = 0
 \end{lstlisting}
@@
 They are also a proper subset of \lstinline!av! since \lstinline!p! is an ancestor of \lstinline!v! but not of \lstinline!p! itself.
-Together with the fact that \lstinline!bgraph b! is finite, it follows that the cardinality of the ancestors decreases in each recursive call thus proving the termination of \lstinline!realize!.
+Together with the fact that \lstinline!bgraph b! is finite, it follows that the cardinality of the ancestors decreases in each recursive call thus proving the termination of \lstinline!realise!.
 
-Before we prove that the realization function constitutes a model in \autoref{sec:real}, we will first explain the significance of the pure witnesses.
+Before we prove that the realisation function constitutes a model in \autoref{sec:real}, we will first explain the significance of the pure witnesses.
 
@@
 In the context of a well-formed branch, this characterisation can be strengthened to any set term and, in addition, we also get that there is no membership literal where a pure witness is on the right-hand side.
-Intuitively speaking, the realization of a pure witness does not depend on the realization of any other set term.
+Intuitively speaking, the realisation of a pure witness does not depend on the realisation of any other set term.
 \begin{lstlisting}[label={lst:lemma_2}]
@@
 So why are pure witnesses treated differently?
-According to the definition of \lstinline!realize!, the pure witnesses would be evaluated to the empty set \lstinline!0 :: V!, were they not treated separately. 
+According to the definition of \lstinline!realise!, the pure witnesses would be evaluated to the empty set \lstinline!0 :: V!, were they not treated separately. 
 To see that this is a problem, consider the branch
@@
 \subsection{Realisation of an Open Branch\label{sec:real}}
-Remember that for completeness, we need to show that the realization function for an open and saturated branch \lstinline!b! actually constitutes a model for all formulae in the branch.
-We start by verifying that the realization function models all literals in the branch; more formally, the following propositions hold: 
+Remember that for completeness, we need to show that the realisation function for an open and saturated branch \lstinline!b! actually constitutes a model for all formulae in the branch.
+We start by verifying that the realisation function models all literals in the branch; more formally, the following propositions hold: 
 \begin{enumerate}[label=(\roman*)]
-  \item\label{it:at_mem} We have \lstinline!realize s $\in$ elts (realize t)! if it holds that \lstinline!s $\inS$ t! is in \lstinline!b!. 
-  \item\label{it:at_eq} We have \lstinline!realize s $=$ realize t! if \lstinline!s $\eqS$ t! is in \lstinline!b!.
-  \item\label{it:af_eq} We have \lstinline!realize s $\neq$ realize t! if \lstinline!s $\neqS$ t! is in \lstinline!b!.
-  \item\label{it:af_mem} We have \lstinline!realize s $\notin$ elts (realize t)! if it holds that \lstinline!s $\notinS$ t! is in \lstinline!b!. 
+  \item\label{it:at_mem} We have \lstinline!realise s $\in$ elts (realise t)! if it holds that \lstinline!s $\inS$ t! is in \lstinline!b!. 
+  \item\label{it:at_eq} We have \lstinline!realise s $=$ realise t! if \lstinline!s $\eqS$ t! is in \lstinline!b!.
+  \item\label{it:af_eq} We have \lstinline!realise s $\neq$ realise t! if \lstinline!s $\neqS$ t! is in \lstinline!b!.
+  \item\label{it:af_mem} We have \lstinline!realise s $\notin$ elts (realise t)! if it holds that \lstinline!s $\notinS$ t! is in \lstinline!b!. 
 \end{enumerate}
@@
 \begin{lstlisting}
-  &\phantom{$\Longrightarrow$}& realize s $\neq$ realize t
-  $\Longrightarrow$ elts (realize s) $\neq$ elts (realize t).
+  &\phantom{$\Longrightarrow$}& realise s $\neq$ realise t
+  $\Longrightarrow$ elts (realise s) $\neq$ elts (realise t).
 \end{lstlisting}
@@
 \begin{lstlisting}
-  a $\in$ elts (realize s) $\land$ a $\notin$ elts (realize t). 
+  a $\in$ elts (realise s) $\land$ a $\notin$ elts (realise t). 
 \end{lstlisting}
   Considering the fact that \lstinline!s $\in$ subterms' b!
-and the definition of \lstinline!realize!, we obtain an \lstinline!u!
-such that \lstinline!a = realize u!
+and the definition of \lstinline!realise!, we obtain an \lstinline!u!
+such that \lstinline!a = realise u!
 and \lstinline!u $\rightarrow_\text{bgraph b}$ s!.
@@
 \begin{lstlisting}[columns=fixed]
-  realize u $\in$ elts (realize t)
-        = a $\in$ elts (realize t)
+  realise u $\in$ elts (realise t)
+        = a $\in$ elts (realise t)
 \end{lstlisting}
-  using Proposition~\ref{it:at_mem}, which stands in contradiction to the assumption \lstinline!a $\notin$ elts (realize t)!.
+  using Proposition~\ref{it:at_mem}, which stands in contradiction to the assumption \lstinline!a $\notin$ elts (realise t)!.
 \end{proof}
@@
 \begin{enumerate}[label=(\alph*)]
-  \item\label{it:empty} It holds that \lstinline!realize $\emptyset$ = 0!.
+  \item\label{it:empty} It holds that \lstinline!realise $\emptyset$ = 0!.
   \item\label{it:op} Let \lstinline!$\star_\text{s}$ $\in$ {$\unionS$, $\diffS$, $\interS$}!. If the term \lstinline!s $\star_\text{s}$ t! occurs in \lstinline!subterms b!, then
     \begin{center}
-    \lstinline!realize (s $\star_\text{s}$ t) = realize s $\star$ realize t!.
+    \lstinline!realise (s $\star_\text{s}$ t) = realise s $\star$ realise t!.
     \end{center}
@@
     \begin{center}
-    \lstinline!realize (Single t) = vset {realize t}!.
+    \lstinline!realise (Single t) = vset {realise t}!.
     \end{center}
@@
 \begin{proof}[Proof of Proposition~\ref{it:empty}]
-  By definition, \lstinline!realize $\emptyset$! is equal to \lstinline!0! if \lstinline!$\emptyset$! is not a vertex of \lstinline!bgraph b!.
+  By definition, \lstinline!realise $\emptyset$! is equal to \lstinline!0! if \lstinline!$\emptyset$! is not a vertex of \lstinline!bgraph b!.
   Otherwise, \lstinline!$\emptyset$! must be in \lstinline!subterms' b! as it is not a witness.
@@
 has no parents in \lstinline!bgraph b!.
-  Ultimately, we obtain \lstinline!realize $\emptyset$ = 0!.
+  Ultimately, we obtain \lstinline!realise $\emptyset$ = 0!.
 \end{proof}
@@
 
-  \textbf{Direction $\subseteq$. } Fix some \lstinline!e! in \lstinline!elts (realize (s $\diffS$ t))!.
+  \textbf{Direction $\subseteq$. } Fix some \lstinline!e! in \lstinline!elts (realise (s $\diffS$ t))!.
   Again, we obtain an \lstinline!u! with 
 \begin{lstlisting}
-  e = realize u $\land$ u $\rightarrow_\text{bgraph b}$ (s $\diffS$ t).
+  e = realise u $\land$ u $\rightarrow_\text{bgraph b}$ (s $\diffS$ t).
 \end{lstlisting}
@@
 \begin{lstlisting}
-  realize u $\in$ realize s $\land$ realize u $\notin$ realize t
+  realise u $\in$ realise s $\land$ realise u $\notin$ realise t
 \end{lstlisting}
-  and together with \lstinline!e = realize u! we arrive at the goal 
+  and together with \lstinline!e = realise u! we arrive at the goal 
 \begin{lstlisting}
-  e $\in$ elts (realize s - realize t).
+  e $\in$ elts (realise s - realise t).
 \end{lstlisting}
 
-\textbf{Direction $\supseteq$. } Let \lstinline[columns=fullflexible]!e $\in$ elts (realize s - realize t)! for some fixed \lstinline!e!.
-  As before, we obtain an \lstinline!u! such that \lstinline!e! is the realization of \lstinline!u!, i.e.\ \lstinline!e = realization u!, and
+\textbf{Direction $\supseteq$. } Let \lstinline[columns=fullflexible]!e $\in$ elts (realise s - realise t)! for some fixed \lstinline!e!.
+  As before, we obtain an \lstinline!u! such that \lstinline!e! is the realisation of \lstinline!u!, i.e.\ \lstinline!e = realisation u!, and
 \begin{lstlisting}
@@
   \begin{lstlisting}
-    realize u $\in$ elts (realize (s $\diffS$ t))
-    &\textover[r]{=}{realize}& e $\in$ elts (realize (s $\diffS$ t)).
+    realise u $\in$ elts (realise (s $\diffS$ t))
+    &\textover[r]{=}{realise}& e $\in$ elts (realise (s $\diffS$ t)).
   \end{lstlisting}
@@
 
-\noindent The final step towards obtaining a proper model is to connect the realization function to the semantics as defined in \autoref{sec:semantics}.
+\noindent The final step towards obtaining a proper model is to connect the realisation function to the semantics as defined in \autoref{sec:semantics}.
 For set terms, we can use the Propositions~\ref{it:empty}--\ref{it:single} to prove the lemma below by induction on \lstinline!t!.
@@
   assumes t $\in$ subterms b
-  shows &\Ist& ($\lambda$x. realize (Var x)) t = realize t
+  shows &\Ist& ($\lambda$x. realise (Var x)) t = realise t
 \end{lstlisting}
@@
   assumes $\phi$ $\in$ set b
-  shows ($\lambda$x. realize (Var x)) $\models$ $\phi$
+  shows ($\lambda$x. realise (Var x)) $\models$ $\phi$
 \end{lstlisting}
@@
 \begin{lstlisting}
-  ($\lambda$x. realize (Var x)) $\models$ last b.
+  ($\lambda$x. realise (Var x)) $\models$ last b.
 \end{lstlisting}
@@
 
-\noindent To finish of the proof of total correctness, note that every singleton branch is trivially well-formed so termination, completeness, and soundless easily transfer to \lstinline!mlss_proc!.
+\noindent To finish of the proof of total correctness, note that every singleton branch is trivially well-formed so termination, completeness, and soundness easily transfer to \lstinline!mlss_proc!.
 \begin{lstlisting}

### 2022-09-16 paper/paper.tex
Continue

@@
 \end{lstlisting}
+
 In addition to the variable names that appear in a formula, we will later also need to obtain those set terms that occur as subterms in a formula. 
@@
 \subsection{Linear Expansion Rules}
-The calculus considers two kinds of branch expansion rules: \textit{linear} and \textit{branching} rules.
-As the name suggests, branching rules lead to the creation of new branches in the tableau while linear rules only extend a branch \lstinline!b!
-with new formulae \lstinline[breaklines=true]!b' = [$\psi_1$,$\ldots$,$\psi_n$]!, which we denote by \lstinline!&\lexpands{b'}{b}&!.
-Furthermore, linear rules are restricted such that no new subterm may be created by their application:
-for example, the rule
-\begin{center}
-\lstinline!(s $\inS$ t$_\text{1}$) $\in$ set b $\Longrightarrow$ &\lexpands{[s $\inS$ t$_\text{1}$ $\unionS$ t$_\text{2}$]}{b}&!
-\end{center}
-may only be used under the condition
-\begin{center}
-  \lstinline!t$_\text{1}$ $\unionS$ t$_\text{2}$ $\in$ subterms (last b)!.
-\end{center}
-\todo[inline]{Explain this restriction in more detail.}
-This ensures that repeated application of linear rules eventually results in a \textit{linearly saturated} branch, i.e.\ a branch where no application of linear rules would produce new formulae.
-In Isabelle, we express this as follows.
-\begin{lstlisting}
-definition lin_sat b $\equiv$
-  $\forall$b'. &\lexpands{b'}{b}& $\longrightarrow$ set b' $\subseteq$ set b
-\end{lstlisting}
-
 \begin{table*}
@@
   \begin{tabular}{p{0.21\textwidth}cp{0.18\textwidth} p{0.01\textwidth} p{0.21\textwidth}cp{0.18\textwidth}}
+    \toprule
     \multicolumn{3}{c}{Propositional Rules} && \multicolumn{3}{c}{Rules for \lstinline!$\unionS$!} \\
@@
     & $\Longrightarrow$ & \lstinline!s $\inS$ Single s! &&
-    \lstinline!$\ton$ $\eqS$ $\ttw$, ($\fmNegSymbol$) a! & $\Longrightarrow$ & \lstinline!($\fmNegSymbol$) a[$\ttw$/$\ton$]! \\
+    \lstinline!$\ton$ $\eqS$ $\ttw$, l! & $\Longrightarrow$ & \lstinline!l{$\ttw$/$\ton$}! \\
 
     \lstinline!s $\inS$ Single t! & $\Longrightarrow$ & \lstinline!s $\eqS$ t! &&
-    \lstinline!$\ton$ $\eqS$ $\ttw$, ($\fmNegSymbol$) a! & $\Longrightarrow$ & \lstinline!($\fmNegSymbol$) a[$\ton$/$\ttw$]! \\
+    \lstinline!$\ton$ $\eqS$ $\ttw$, l! & $\Longrightarrow$ & \lstinline!l{$\ton$/$\ttw$}! \\
 
@@
     \lstinline!s$_\text{1}$ $\inS$ t, s$_\text{2}$ $\inS$ t! & $\Longrightarrow$ & \lstinline!s$_\text{1}$ $\neqS$ s$_\text{2}$! \\
+    \bottomrule
   \end{tabular}
@@
 \end{table*}
-\todo[inline]{Reference figure and mention missing rule + example}
+The calculus considers two kinds of branch expansion rules: \textit{linear} and \textit{branching} rules.
+As the name suggests, branching rules lead to the creation of new branches in the tableau while linear rules only extend a branch \lstinline!b!
+with new formulae \lstinline[breaklines=true]!b' = [$\psi_1$,$\ldots$,$\psi_n$]!, which we denote by \lstinline!&\lexpands{b'}{b}&!.
+\autoref{tab:lexpands} shows all linear expansion rules.
+Note that, in the first two rules for \lstinline!$\eqS$!, \lstinline!l! is a literal occuring in the branch.
+Furthermore, the term-for-term substitution \lstinline!l{s/t}! is restricted to the top-level set terms of the literal;
+for example, given the literal
+\begin{lstlisting}
+  l = &\fmNeg{((s $\unionS$ u) $\diffS$ s $\eqS$ s $\unionS$ u)}&
+\end{lstlisting}
+we have
+\begin{lstlisting}
+  &\phantom{=}& (&\fmNeg{((s $\unionS$ u) $\diffS$ s $\eqS$ s $\unionS$ u)}&){t/s $\unionS$ u}
+  = &\fmNeg{((s $\unionS$ u) $\diffS$ s $\eqS$ t)}&.
+\end{lstlisting}
+The perhaps more crucial restriction of the linear rules is that that no new subterm may be created by their application;
+for instance, the second rule for \lstinline!$\unionS$!
+\begin{lstlisting}
+  s $\inS$ t$_\text{1}$ $\Longrightarrow$ s $\inS$ t$_\text{1}$ $\unionS$ t$_\text{2}$,
+\end{lstlisting}
+which formally represents 
+\begin{lstlisting}
+  (s $\inS$ t$_\text{1}$) $\in$ set b $\Longrightarrow$ &\lexpands{[s $\inS$ t$_\text{1}$ $\unionS$ t$_\text{2}$]}{b}&,
+\end{lstlisting}
+may only be used under the condition
+\begin{lstlisting}
+  t$_\text{1}$ $\unionS$ t$_\text{2}$ $\in$ subterms (last b).
+\end{lstlisting}
+The purpose of this restriction is to prevent unbounded expansion of the branch.
+In fact, we give an explicit upper bound for the number of formulae in a branch in \autoref{sec:bound}.
+
+Due to boundedness, repeated application of linear rules eventually results in a \textit{linearly saturated} branch, i.e.\ a branch where no application of linear rules would produce new formulae.
+\begin{lstlisting}
+definition lin_sat b $\equiv$
+  $\forall$b'. &\lexpands{b'}{b}& $\longrightarrow$ set b' $\subseteq$ set b
+\end{lstlisting}
+
+Finally, we remark that the original paper~\cite{new_fast_tableau} is missing the last propositional rules dealing with double negation.
+This rule is required for completeness, though, considering that the branch
+\begin{lstlisting}
+  [&\fmNeg{\fmNeg{\fmNeg{p}}}&, p, &\fmAnd{\fmNeg{\fmNeg{\fmNeg{p}}}}{p}&]
+\end{lstlisting}
+is saturated --- neither linear nor branching rules apply --- and open but there is obviously no model for the initial formula \lstinline!&\fmAnd{\fmNeg{\fmNeg{\fmNeg{p}}}}{p}&!.
 
 \subsection{Branching Rules\label{sec:branching}}
-The number of branches increases when applying branching rules so it is prudent to only apply them to open, linearly saturated branches.
-To further prevent unnecessary branching, each rule has a \textit{subsumption} condition. 
-Note that even branching rules are restricted such that they never create new subterms, with exception of the last rule which adds a fresh variable.
-\todo[inline]{Reference figure}
-
-\begin{lstlisting}
-definition sat b $\equiv$ lin_sat b $\land$ ($\nexists$bs'. &\fexpands{bs'}{b}&)
-\end{lstlisting}
-
-\begin{lstlisting}
-definition wits :: 'a branch $\Rightarrow$ 'a set where
-  wits b $\equiv$ vars b - vars (last b)
-\end{lstlisting}
-
 \begin{table*}
@@
   \newcommand{\lpipe}{\rule[-0.4ex]{0.6pt}{2.3ex}}
-  \begin{tabular}{>{\centering\arraybackslash}p{0.28\textwidth} >{\centering\arraybackslash}p{0.3\textwidth} >{\centering\arraybackslash}p{0.3\textwidth}}
+  \begin{tabular}{>{\centering\arraybackslash}p{0.28\textwidth} >{\centering\arraybackslash}p{0.34\textwidth} >{\centering\arraybackslash}p{0.3\textwidth}}
+    \toprule
     Rule & Precondition & Subsumption condition \\
@@
     \end{tabular}
-         & Test & \lstinline!p $\in$ set b $\lor$ @\fmNeg{p}@ $\in$ set b! \\
+         & \lstinline!@\fmOr{p}{q}@ $\in$ set b! & \lstinline!p $\in$ set b $\lor$ @\fmNeg{p}@ $\in$ set b! \\
 
@@
     \end{tabular}
-         & Test & \lstinline!@\fmNeg{p}@ $\in$ set b $\lor$ p $\in$ set b! \\
+         & \lstinline!@\fmNeg{(\fmAnd{p}{q})}@ $\in$ set b!
+         & \lstinline!@\fmNeg{p}@ $\in$ set b $\lor$ p $\in$ set b! \\
 
@@
     \end{tabular}
-         & Test &
+         &
+         \begin{tabular}{c}
+           \lstinline!(s $\inS$ $\ton$ $\unionS$ $\ttw$) $\in$ set  b! \\
+           \lstinline!$\ton$ $\unionS$ $\ttw$ $\in$ subterms (last b)! \\
+         \end{tabular}
+         &
          \begin{tabular}{@{}rl@{}}
@@
     \end{tabular}
-         & Test &
+         &
+         \begin{tabular}{c}
+           \lstinline!(s $\inS$ $\ton$ $\interS$ $\ttw$) $\in$ set  b! \\
+           \lstinline!($\ton$ $\interS$ $\ttw$) $\in$ subterms (last b)! \\
+         \end{tabular}
+         &
          \begin{tabular}{@{}rl@{}}
@@
     \end{tabular}
-         & Test &
+         &
+         \begin{tabular}{c}
+           \lstinline!(s $\inS$ $\ton$ $\diffS$ $\ttw$) $\in$ set  b! \\
+           \lstinline!($\ton$ $\diffS$ $\ttw$) $\in$ subterms (last b)! \\
+         \end{tabular}
+         &
          \begin{tabular}{@{}rl@{}}
@@
          \end{tabular} \\
+         \addlinespace[0.6em]
+
     \begin{tabular}{@{}>{\raggedleft\arraybackslash}p{0.11\textwidth}c>{\raggedright\arraybackslash}p{0.11\textwidth}@{}}
@@
     \end{tabular}
-         & Test & Test \\
+         & 
+         \begin{tabular}{c}
+           \lstinline!($\ton$ $\neqS$ $\ttw$) $\in$ set b! \\
+           \lstinline!$\ton$ $\in$ subterms (last b)! \\
+           \lstinline!$\ttw$ $\in$ subterms (last b)! \\
+           \lstinline!x $\notin$ vars b! \\
+         \end{tabular}
+         &
+         \begin{tabular}{r@{\hspace{0.5em}}l}
+
+           \lstinline!$\exists$s.! & \lstinline!(s $\inS$ $\ton$) $\in$ set b $\land$! \\
+                                   & \lstinline!(s $\notinS$ $\ttw$) $\in$ set b! \\
+           \multicolumn{2}{c}{$\lor$} \\
+           \lstinline!$\exists$s.! & \lstinline!(s $\notinS$ $\ton$) $\in$ set b $\land$! \\
+                                   & \lstinline!(s $\inS$ $\ttw$) $\in$ set b! \\
+         \end{tabular}
+         \\
+         \bottomrule
   \end{tabular}
@@
 \end{table*}
+After running out of linear rules apply, only the branching rules shown in \autoref{tab:bexpands} remain.
+A rule is applicable if its \textit{precondition} is met and, to prevent unnecessary branching, it is not subsumed as indicated by the \textit{subsumption condition}.
+These rules create multiple branches in the tableau so we represent the different possibilities \lstinline!bs'! to expand a branch \lstinline!b! as a set and write \lstinline!&\fexpands{bs'}{b}&!. 
+Accordingly, we get a new branch \lstinline!b' @ b! in the tableau for each \lstinline!b' $\in$ set bs'!.
+
+A linearly saturated branch where no further branching is possible is called a \textit{saturated} branch. 
+\begin{lstlisting}
+definition sat b $\equiv$ lin_sat b $\land$ ($\nexists$bs'. &\fexpands{bs'}{b}&)
+\end{lstlisting}
+
+Note that even branching rules are defined such that they never create new subterms with exception of the last rule, which adds a new variable to the branch.
+These variables serve to manifest an inequality so we call them \textit{witnesses}.
+\begin{lstlisting}
+definition wits :: 'a branch $\Rightarrow$ 'a set where
+  wits b $\equiv$ vars b - vars (last b)
+\end{lstlisting}
 
@@
 \end{lstlisting}
-Confirming that this is an invariant is an easy exercise of going through each of the propositional rules in Table TODO and the first two branching rules in Table TODO.
+Confirming that this is an invariant is an easy exercise of going through each of the propositional rules in \autoref{tab:lexpands} and the first two branching rules in \autoref{tab:bexpands}.
 Thus, we have

### 2022-09-19 paper/paper.tex
Continue

@@
 \subsection{Related work}
-AFP entry~\cite{hybrid_logic_afp} and workshop paper in Isabelle 2022 workshop.
+AFP entry~\cite{hybrid_logic_afp} and article~\cite{hybrid_logic} and workshop paper in Isabelle 2022 workshop.
+A Verified Proof of Craig Interpolation for Basic Modal Logic via Tableaux in Lean.
+A Tableau for Ecumenical Propositional Logic.
+Automatization of the semantic tableau method.
+Formalisation de la logique temporelle dans le coq.
+\todo[inline]{Look at all articles of From}
 
@@
 For the benefit of the reader that is unfamiliar with Isabelle/HOL, we establish notation and in particular some essential datatypes together with their primitive operations that are specific to Isabelle/HOL.
+
 We write \lstinline!t :: 'a!
@@
 and \lstinline!'a $\Rightarrow$ 'b!
-for the type of a total function from \lstinline!'a!
+for the space of total functions from \lstinline!'a!
 to \lstinline!'b!.
-The types for booleans, natural numbers, and in integers are \lstinline!bool!, \lstinline!nat!, and \lstinline!int!, respectively.
+
 Sets with elements of type \lstinline!'a! have the type \lstinline!'a set!.
-Analogously, we use \lstinline!'a list!
-to describe lists, which are constructed as the empty list \lstinline![]!
+The cardinality of a set is determined with \lstinline!card! and \lstinline!f ` A! denotes the image of \lstinline!A! under \lstinline!f!.
+
+Analogously to sets, we use \lstinline!'a list!
+to describe the type of lists, which are constructed as the empty list \lstinline![]!
 or with the infix constructor \lstinline!#!, and are appended with the infix operator \lstinline!@!.
 The function \lstinline!set! converts a list into a set.
-For optional values, Isabelle/HOL offers the type \lstinline!option! where a term \lstinline!opt :: 'a option! is either \lstinline!None! or \lstinline!Some a! with \lstinline!a :: 'a!.
-Finally, we remark that \textbf{iff} is equivalent to \lstinline!$=$!
-on type \lstinline!bool!
+
+We remark that \textbf{iff} is equivalent to \lstinline!$=$!
+on type of booleans \lstinline!bool!
 and \lstinline!$\equiv$!
 is definitional equality of the meta-logic of Isabelle/HOL, which is called Isabelle/Pure.
-\todo[inline]{Add arrows and bracket notation for Pure formulae}
+The other important constant of the meta-logic, apart from \lstinline!$\equiv$!, is the meta-implication \lstinline!$\Longrightarrow$!.
+A chain of implications
+\begin{lstlisting}
+  A$_\text{1}$ $\Longrightarrow$ $\cdots$ $\Longrightarrow$ A$_\text{k}$ $\Longrightarrow$ C
+\end{lstlisting}
+can be denoted more concisely by
+\begin{lstlisting}
+  $\llbracket$ A$_\text{1}$;$\,\ldots\,$;A$_\text{k}$ $\rrbracket$ $\Longrightarrow$ C.
+\end{lstlisting}
+
 \section{Syntax and Semantics of MLSS\label{sec:semantics}}
 \subsection{Syntax}
-At the heart of \MLSS{}, we have the type of set terms which is the disjoint union of the empty set and variables as well as the operations union, intersection, set difference, and \lstinline!Single!, which represents a singleton set containing the argument.
-\todo[inline]{Syntax for \lstinline!Single!?, overloading of \lstinline!vars_term!, \lstinline!vars_atom!}
+At the heart of \MLSS{}, we have the type of set terms which is the disjoint union of the empty set and variables as well as the operations union, intersection, set difference, and the singleton set represented by the constructor \lstinline!Single!.
+We keep the type of variables abstract by making it a parameter of the set term data type.
+The only restriction on the type of variables is that it needs to be infinite.
+Isabelle's data type package automatically defines a function that gives us the set of variables in a set term, which we name \lstinline!vars!.
+In what follows, we will overload the function \lstinline!vars! to also work on set atoms, formulae, and branches.
 \begin{lstlisting}
@@
 With the above operators we can also represent the subset operator \lstinline!$\subseteqS$!
-and enumerate finite sets as \lstinline!s $\subseteqS$ t!
+and enumerate finite sets: \lstinline!s $\subseteqS$ t!
 is equivalent to \lstinline!s $\unionS$ t $\eqS$ t!
@@
 
-The unquantified fragment \MLSS{} of set theory is then formed by combining set atoms, which are injected into formulae with the constructor \lstinline!&\fmAtom&!, with propositional logical connectives.
-\todo[inline]{Cite Nipkow's Linear Quantifier Elimination?, literals}
+We use the propositional fragment of formulae due to \citeauthor{lqe}~\cite{lqe} with set atoms as propositional atoms to form the unquantified fragment \MLSS{} of set theory.
 \begin{lstlisting}
@@
 type_synonym 'a pset_fm = 'a pset_atom fm
-\end{lstlisting}
-
-\begin{lstlisting}
-fun is_literal :: 'a fm $\Rightarrow$ bool where
-  is_literal (&\fmAtom& _) = True
-| is_literal (&\fmNeg{(\fmAtom{} \_)}&) = True
-| is_literal _ = False
 \end{lstlisting}
@@
 and \lstinline!&\fmNeg{\fmAtom{} (s $\eqS$ t)}&!, respectively.
-Note that Isabelle's datatype package automatically defines a function \lstinline!atoms :: 'a fm $\Rightarrow$ 'a set! that allows us to obtain all set atoms in a formula.
-We combine these functions in order to define \lstinline!vars! that extracts all the variable names occuring in a set formula.
-\todo[inline]{Motivation: one rule introduces new variables, Mention vars polymorphic}
+
+Similarly to the function \lstinline!vars!, the automatically defined function \lstinline!atoms :: 'a fm $\Rightarrow$ 'a set! retrieves all set atoms in a formula.
+We combine these functions to extract all the variables occuring in a set formula.
 \begin{lstlisting}
@@
 
-In addition to the variable names that appear in a formula, we will later also need to obtain those set terms that occur as subterms in a formula. 
-We first define a function that collects the subterms of a set term.
-\todo[inline]{Motivation: linear rules don't introduce subterms, \\ \lstinline!subterms b $\approx$ subterms (last b)!}
+In addition to the variable that appear in a formula, we will later also need to obtain those set terms that occur as subterms in a formula. 
+Like \lstinline!vars!, we overload the constant \lstinline!subterms!.
 \begin{lstlisting}
@@
 \end{lstlisting}
-The last function that we define computes the subformulae of a formula.
+
+Finally, we consider the subformulae of a formula and define a function that computes them.
 \begin{lstlisting}
@@
 \end{lstlisting}
-
+The \textit{literals} are a type of subformulae of particular importance so we define a predicate for them.
+\begin{lstlisting}
+fun is_literal :: 'a fm $\Rightarrow$ bool where
+  is_literal (&\fmAtom& _) = True
+| is_literal (&\fmNeg{(\fmAtom{} \_)}&) = True
+| is_literal _ = False
+\end{lstlisting}
 \subsection{Semantics}
 Formulae in \MLSS{} contain arbitrarily nested set terms.
-We account for this arbitrary nesting by basing the semantics on the von Neumann hierarchy $\mathcal{V}$ of sets that is defined inductively as
-\todo[inline]{Simplify definition by deleting cases?}
+We account for this nesting by basing the semantics on the von Neumann hierarchy $\mathcal{V}$ of sets that is defined inductively as
   \[
@@
 A way to work around this limitiation is provided by Lawrence in an AFP entry~\cite{zfc_in_hol_afp}, which adds $\mathcal{V}$ to Isabelle/HOL by way of axiomatisation.
-An article by Lawrence~\cite{zfc_in_hol} provides some further context to this entry.
+Two articles~\cite{zfc_in_hol1,zfc_in_hol2} have been written that provide further context to this entry.
 The entry declares a type \lstinline!V! that comes with the following functionality:
@@
 for set terms that interprets a set term with respect to a valuation function \lstinline!M :: 'a $\Rightarrow$ V!
-for variable.
+for variables.
 \begin{lstlisting}
@@
 \end{lstlisting}
+
 For formulae, we write \lstinline!M $\models$ $\phi$! for the judgement that $\phi$ holds under the valuation function \lstinline!M!.
-We omit the implementation of $\models$ for brevity.
-\todo[inline]{Add implementation? Add satisfiability!}
+The implementation of $\models$ coincides with the interpretation function of \citeauthor{lqe}~\cite{lqe}, so omit it for brevity.
+As usual, a formula \lstinline!$\phi$! is called \textit{satisfiable} if there exists a model \lstinline!$M$! with \lstinline!$M$ $\models$ $\phi$!.
+Otherwise, \lstinline!$\phi$! is called \textit{unsatisfiable}.
 
@@
 We formalise the tableau calculus for \MLSS{} as described by \citeauthor{new_fast_tableau}~\cite{new_fast_tableau}.
-Inspired by the formalisation of a tableau calculus for Hybrid Logic by \citeauthor{hybrid_logic_afp}~\cite{hybrid_logic_afp}, we simply use a list to represent a branch of the tableau tree.
+Inspired by the formalisation of a tableau calculus for hybrid logic by \citeauthor{hybrid_logic_afp}~\cite{hybrid_logic_afp}, we simply use a list to represent a branch of the tableau tree.
 Note that formulae are added to the front of the list during branch expansion so \lstinline!last b! for a branch \lstinline!b! is always the formula that we are trying to disprove with the tableau. 
@@
 \end{lstlisting}
-In the standard tableau calculus for propositional logic, a branch is called \textit{closed} if it contains both the negation of a formula and the formula itself;
+In the standard tableau calculus for propositional logic as \citeauthor{tableau}~\cite{tableau} describes it, a branch is called \textit{closed} if it contains both the negation of a formula and the formula itself;
 conversely, it is called \textit{open} if it is not closed.
@@
 Note that, in the first two rules for \lstinline!$\eqS$!, \lstinline!l! is a literal occuring in the branch.
-Furthermore, the term-for-term substitution \lstinline!l{s/t}! is restricted to the top-level set terms of the literal;
+Furthermore, the term-for-term substitution \lstinline!l{s/t}! is restricted to the top-level set terms of \lstinline!l!;
 for example, given the literal
@@
 
-Due to boundedness, repeated application of linear rules eventually results in a \textit{linearly saturated} branch, i.e.\ a branch where no application of linear rules would produce new formulae.
+Due to boundedness, repeated expansion with linear rules eventually results in a \textit{linearly saturated} branch, i.e.\ a branch where no application of linear rules would produce new formulae.
 \begin{lstlisting}
@@
 
-Finally, we remark that the original paper~\cite{new_fast_tableau} is missing the last propositional rules dealing with double negation.
+Finally, we remark that the original paper~\cite{new_fast_tableau} is missing the last propositional rule dealing with double negation.
 This rule is required for completeness, though, considering that the branch
@@
 \end{lstlisting}
-is saturated --- neither linear nor branching rules apply --- and open but there is obviously no model for the initial formula \lstinline!&\fmAnd{\fmNeg{\fmNeg{\fmNeg{p}}}}{p}&!.
+is saturated --- neither linear nor branching rules apply --- and open but there clearly is no model for the initial formula \lstinline!&\fmAnd{\fmNeg{\fmNeg{\fmNeg{p}}}}{p}&!.
 
@@
          \begin{tabular}{c}
-           \lstinline!(s $\inS$ $\ton$ $\interS$ $\ttw$) $\in$ set  b! \\
+           \lstinline!(s $\inS$ $\ton$) $\in$ set  b! \\
            \lstinline!($\ton$ $\interS$ $\ttw$) $\in$ subterms (last b)! \\
@@
          \begin{tabular}{c}
-           \lstinline!(s $\inS$ $\ton$ $\diffS$ $\ttw$) $\in$ set  b! \\
+           \lstinline!(s $\inS$ $\ton$) $\in$ set  b! \\
            \lstinline!($\ton$ $\diffS$ $\ttw$) $\in$ subterms (last b)! \\
@@
 
+\todo[inline]{First give intuition then technical details}
 \noindent We abstractly specify the behaviour of the decision procedure, leaving an executable specification from which code can be generated to future work.
@@
 \subsection{Realisation of an Open Branch\label{sec:real}}
+\todo[inline]{Reformulate with initial formula}
 Remember that for completeness, we need to show that the realisation function for an open and saturated branch \lstinline!b! actually constitutes a model for all formulae in the branch.
@@
 \end{lstlisting}
-  Without loss of generality --- the other case is symmetric ---, we obtain an \lstinline!a! such that
-\begin{lstlisting}
-  a $\in$ elts (realise s) $\land$ a $\notin$ elts (realise t). 
+  Without loss of generality --- the other case is symmetric ---, we obtain an \lstinline!e! such that
+\begin{lstlisting}
+  e $\in$ elts (realise s) $\land$ e $\notin$ elts (realise t). 
 \end{lstlisting}
@@
 and the definition of \lstinline!realise!, we obtain an \lstinline!u!
-such that \lstinline!a = realise u!
+such that \lstinline!e = realise u!
 and \lstinline!u $\rightarrow_\text{bgraph b}$ s!.
   This, in turn, yields that \lstinline!u $\inS$ s! must be in \lstinline!b!.
-  Together with the assumption \lstinline!(s $\eqS$ t) $\in$ b! and the openness of \lstinline!b!, it follows that \lstinline!u $\inS$ t! must also be in \lstinline!b!.
+  Together with the assumption \lstinline!(s $\eqS$ t) $\in$ set b! and the saturation of \lstinline!b!, it follows that \lstinline!u $\inS$ t! must also be in \lstinline!b!.
   But then we have
@@
   realise u $\in$ elts (realise t)
-        = a $\in$ elts (realise t)
-\end{lstlisting}
-  using Proposition~\ref{it:at_mem}, which stands in contradiction to the assumption \lstinline!a $\notin$ elts (realise t)!.
+        = e $\in$ elts (realise t)
+\end{lstlisting}
+  using Proposition~\ref{it:at_mem}, which stands in contradiction to the assumption \lstinline!e $\notin$ elts (realise t)!.
 \end{proof}
@@
 \end{lstlisting}
-  We therefore have \lstinline!(u $\inS$ s $\diffS$ t) $\in$ set b! which, due to \lstinline!b! being saturated, implies that \lstinline!u $\inS$ s! and \lstinline!u $\notinS$ t! are in \lstinline!b!.
-  \todo[inline]{Explain latter of line above}
+  We therefore have \lstinline!(u $\inS$ s $\diffS$ t) $\in$ set b! which, due to \lstinline!b! being saturated, implies that \lstinline!u $\inS$ s! is in \lstinline!b!.
+  Furthermore, the fifth branching rule (see \autoref{tab:bexpands}) must have its subsumption condition met due to saturation of \lstinline!b!, meaning that either one of \lstinline!u $\inS$ t!
+or \lstinline!u $\notinS$ t!
+must be in the branch.
+  Only the latter, however, can be the case since the branch is open.
   By applying Proposition~\ref{it:at_mem} and~\ref{it:af_mem} we obtain
@@
   Moreover, we claim that \lstinline!u $\notinS$ t! is also in \lstinline!b!.
-  To see that this is true, recall that \lstinline!b! is saturated and thus the subsumption condition of rule TODO needs to hold.
+  To see that this is true, recall that \lstinline!b! is saturated and thus the subsumption condition of the fifth branching rule must hold.
   This implies that either \lstinline!u $\inS$ t!
@@
 The coherence property finishes the proof of completeness of the calculus as it gives us a model for every formula in an open and saturated branch.
+\todo[inline]{In particular initial formula is satisfied}
 
@@
 \end{lstlisting}
-  for each \lstinline!i $\in$ {0,$\ldots$,k}!.
-  Therefore, the transitive closure of the membership relation \lstinline!{(x, y). x $\in$ elts y}! contains a cycle. 
-  This, however, is a contradiction since the membership relation of \lstinline!V! is well-founded (see \autoref{sec:semantics}) and well-foundedness is closed under transitive closure.
+  for each \lstinline!i $\in$ {0,$\,\ldots\,$,k}!.
+  Therefore, the transitive closure of the membership relation \lstinline!{(x, y). x $\in$ elts y}$^\text{+}$! contains a cycle. 
+  This, however, is a contradiction because the membership relation of \lstinline!V! is well-founded (see \autoref{sec:semantics}) and well-foundedness is closed under transitive closure.
 \end{proof}
@@
 A similar argument would work for the branching rules if it were not for the last rule adding new variables.
-Those variables need to be assigned specific values; hence, we modify the model as we will see in the proof below.
+Those variables need to be assigned specific values; hence, we modify the model as shown in the proof below.
 \begin{lstlisting}[belowskip=0pt, label={lst:bexpands_sound}]
@@
 plus the number of witnesses added to the branch.
-Regarding an upper bound for the witnesses, recognise that the last branching rule requires \lstinline!s $\neqS$ t! to be in \lstinline!b! for some \lstinline!s,t $\in$ subterms (last b)!.
+Note that the last branching rule requires \lstinline!s $\neqS$ t! to be in \lstinline!b! for some \lstinline!s,t $\in$ subterms (last b)!.
 The subsumption condition then prohibits that the rule can be applied again for the combination of \lstinline!s!
-and \lstinline!t!, thus yielding an upper bound of \lstinline!(card (subterms (last b)))$^\text{2}$!.
+and \lstinline!t!, thus yielding an upper bound of \lstinline!(card (subterms (last b)))$^\text{2}$!
+for the witnesses.
 This allows us to finish the calculation that justifies the second inequality:
@@
 where the difference between the upper bound and the cardinality is smaller than that of \lstinline!b!.
-  We strengthen those assumptions to obtain
+  We strengthen the assumptions of the induction hypothesis to obtain
 \begin{lstlisting}
@@
 The former strengthening is admissable since, by definition, every expansion of a well-formed branch is again well-formed while the latter ensures that the cardinality of \lstinline!b'!
-is larger than the cardinality of \lstinline!b!, ensuring that the difference to the upper bound decreases.
+is larger than the cardinality of \lstinline!b!, implying that the difference to the upper bound decreases.
 

### 2022-09-21 paper/paper.tex
Continue

@@
 
-\title{Towards a Verified Tableau Prover for Multi-Level-Syllogistics}
+\title{Towards a Verified Tableau Prover for a Ground Fragment of Set Theory}
 \author{Lukas Stevens}
@@
 \section{Introduction}
-\todo[inline]{\MLSS{} is multi-level syllogistics with singleton}
-
-\subsection{Related work}
+In contrast to pen-and-paper proofs where we expect the reader to fill in trivial details, interactive theorem provers (ITPs) require us to justify each proof step rigorously.
+Manually reducing each proof step down to the axioms would be impractical, though;
+hence, ITPs usually offer a wide array of automated methods to deal with the proof goals that arise during interactive use.
+
+In Isabelle/HOL, there are specialised procedures for dealing with e.g.\ natural numbers, linear arithmetic, and metric spaces.
+Some of these procedures have been verified in Isabelle/HOL such as a procedure for presburger arithmetic~\cite{presburger} that was later extended to mixed real-integer arithmetic~\cite{arithmetic}.
+This procedure, though, uses reflection to work on goals in Isabelle/HOL, which, during execution, either sacrifices speed by going through the simplifier or soundness by trusting generated code.
+More recently, \citeauthor{orders}~\cite{orders} presented a verified decision procedure for orders that produces certificates. 
+This approach offers efficient execution by using generated code as well as soundness because the certificates are replayed through Isabelle's inference kernel.
+
+The focus of this paper is another ubiquitous structure in mathematics, namely sets.
+In particular, we consider a ground fragment of set theory which \citeauthor{new_fast_tableau}~\cite{new_fast_tableau} call multi-level syllogistics with singleton (\MLSS{}).
+The fragment includes the usual set operations of union, intersection, difference, membership, equality and, in addition, it allows the construction of singleton sets.
+Due to the omnipresence of sets, we think that automation for this fragment will benefit the users of Isabelle, especially when integrated into the simplifier like the aforementioned order solver.
+We also expect that the generation of certificates will be straightforward since \MLSS{} admits a tableau calculus.
+
+\subsection{Contributions}
+In this paper, we present a formalisation in Isabelle/HOL of a tableau calculus for \MLSS{} due to \citeauthor{new_fast_tableau}~\cite{new_fast_tableau}\cite[Chapter 14]{set_theory}.
+We prove soundness and completeness of the calculus and give an abstract specification of a decision procedure that applies the rules of the calculus exhaustively.
+To obtain total correctness of the procedure, we prove its termination.
+The formalisation follows the paper closely but gives a more thorough account of some important details:
+\begin{itemize}
+  \item We deliver the omitted proof of Lemma~2 in the paper~\cite{new_fast_tableau}, which is key for the completeness proof of the calculus.
+  \item The formal proof of completeness lead to the discovery that the calculus was missing a rule for eliminating double negation. 
+  \item We derive an explicit upper bound for the number of formulae in a branch of the tableau.
+\end{itemize}
+
+All in all, the abstract specification of the decision procedure provides a stepping stone towards a refinement to an executable specification from which code can be generated.
+
+\subsection{Related Work}
 AFP entry~\cite{hybrid_logic_afp} and article~\cite{hybrid_logic} and workshop paper in Isabelle 2022 workshop.
@@
 Formalisation de la logique temporelle dans le coq.
+Procedure was later generalised to cartesian product etc.
+Procedure was also generate to add monotone functions.
 \todo[inline]{Look at all articles of From}
 
-\subsection{Contributions}
-Fill the proof of Lemma 2. More explicit proof of the cardinality of a branch.
+First time MLSS~\cite{mlss_first}.
+MLSS is in NP~\cite{mlss_np}.
+First MLSS Tableau~\cite{mlss_first_tableau}.
+Beckert~\cite{tableau_quantifier_free}, Diploma~\cite{tableau_quantifier_free_diploma}.
+Tableau-based MLSS with functions and quantification~\cite{mlss_functions_quantification}.
+MLSS with monotone functions~\cite{mlss_monotone_functions}, reduction to MLSS.
+Another generalisation with cartesion product and some map constructs~\cite{mlss_cartesian_map}.
 
@@
     \lstinline!s $\notinS$ Single t! & $\Longrightarrow$ & \lstinline!s $\neqS$ t! &&
-    \lstinline!s$_\text{1}$ $\inS$ t, s$_\text{2}$ $\inS$ t! & $\Longrightarrow$ & \lstinline!s$_\text{1}$ $\neqS$ s$_\text{2}$! \\
+    \lstinline!s$_\text{1}$ $\inS$ t, s$_\text{2}$ $\notinS$ t! & $\Longrightarrow$ & \lstinline!s$_\text{1}$ $\neqS$ s$_\text{2}$! \\
     \bottomrule
@@
 
-\todo[inline]{First give intuition then technical details}
-\noindent We abstractly specify the behaviour of the decision procedure, leaving an executable specification from which code can be generated to future work.
+\noindent The mechanics of the decision procedure are characteristical for a procedure based on a tableau calculus:
+it determines the satisfiability of a given formula \lstinline!$\phi$! by determining whether the formula has a closed tableau.
+More specifically, it initialises the tableau with the singleton branch \lstinline![$\phi$]! and checks whether this branch can be expanded to a closed tableau.
+
+We abstractly specify the behaviour of the decision procedure, leaving an executable specification from which code can be generated to future work.
 The implementation as displayed in \autoref{fig:alg} uses a couple features of Isabelle's function package:
@@
 nevertheless, the termination proof is separated from the definition of the function for modularity.
-The function package maintains the soundness of the definition by introducing a so-called domain predicate \lstinline!mlss_proc_branch_dom! which characterises where the function terminates.
+The function package maintains the soundness of the definition by introducing a so-called domain predicate \lstinline!mlss_proc_branch_dom!
+which characterises for which arguments the function terminates.
 Each equation of the function receives the assumption that the predicate holds for the argument.
@@
 
-The purpose of the function is to take a branch and determine whether the branch can be expanded to a closed tableau.
+The purpose of the function is to determine whether a given branch can be expanded to a closed tableau.
 As stated before, we first use linear expansion rules in order to prevent premature branching;
-therefore, we recursively expand the branch with linear rules until the branch is linearly saturated.
+to this end, we recursively expand the branch with linear rules until the branch is linearly saturated.
 Note that we use Hilbert's $\varepsilon$-operator in the form of \lstinline!SOME! to choose some rule that actually adds new formulae to the branch. 
@@
 
-Every unsatisfiable formula has a closed tableau so we can determine the satisfiability of a formula \lstinline!$\psi$!
-by passing the singleton branch \lstinline![$\psi$]!
-to \lstinline!mlss_proc_branch!.
-Because this is the only context that \lstinline!mlss_proc_branch! is used in, it only encounters branches that result from applying branch expansion rules to a singleton branch.
+The procedure \lstinline!mlss_proc! then calls \lstinline!mlss_proc_branch! with a singleton branch \lstinline![$\phi$]! to determine the satisfiability of a given formula \lstinline!$\phi$!.
+
+This means that \lstinline!mlss_proc_branch! is only applied to branches that result from applying the expansion rules.
 We call this kind of branches \textit{well-formed}.
@@
 \end{lstlisting}
-In particular, we will use this notion in \autoref{sec:bound} where derive an upper bound for the cardinality of a well-formed branch.
+
+In particular, we will use this notion in \autoref{sec:bound} where we derive an upper bound for the cardinality of a well-formed branch.
 This is then used in \autoref{sec:correct} to justify the termination of the decision procedure. 
@@
 \section{Completeness of the Calculus\label{sec:complete}}
-Once a branch \lstinline!b! is saturated and open at the same time, we are unable to expand the branch further in our calculus.
-Thus, to establish completeness of the calculus, we must construct a model \lstinline!M! such that \lstinline!M $\models$ $\phi$! for all \lstinline!$\phi$ $\in$ set b!.
-
+For completeness of the calculus, we need to show that every unsatisfiable formula has a closed tableau or, conversely, that the formula is satisfiable if there is a saturated and open branch in the tableau.
+To facilitate inductive reasoning, we show a stronger statement by constructing a model \lstinline!M! such that \lstinline!M $\models$ $\phi$! for all \lstinline!$\phi$ $\in$ set b!.
 At the core of the model, there is a \textit{realisation} function that maps set terms to sets of type \lstinline!V!.
@@
 \subsection{Realisation of an Open Branch\label{sec:real}}
-\todo[inline]{Reformulate with initial formula}
 Remember that for completeness, we need to show that the realisation function for an open and saturated branch \lstinline!b! actually constitutes a model for all formulae in the branch.
 We start by verifying that the realisation function models all literals in the branch; more formally, the following propositions hold:
-\begin{enumerate}[label=(\roman*)]
+\begin{enumerate}[label={(\roman*)}]
   \item\label{it:at_mem} We have \lstinline!realise s $\in$ elts (realise t)! if it holds that \lstinline!s $\inS$ t! is in \lstinline!b!. 
@@
 The coherence property finishes the proof of completeness of the calculus as it gives us a model for every formula in an open and saturated branch.
-\todo[inline]{In particular initial formula is satisfied}
 
 \section{Soundness of the Calculus\label{sec:sound}}
-Two aspects determine the soundness of the calculus: a closed branch must be unsatisfiable and all expansion rules must maintain satisfiability.
-We can express the former aspect in Isabelle as follows.
+A tableau calculus is sound if, for any closed tableau, the corresponding formula is unsatisfiable. 
+We prove the following two properties to establish soundness.
+\begin{enumerate}[label={(\roman*)}]
+  \item A closed branch contains an unsatisfiable formula.
+  \item The expansion rules maintain satisfiability.
+\end{enumerate}
+To see that these are sufficient, consider a closed tableau for a formula \lstinline!$\phi$! but assume that \lstinline!$\phi$! is satisfiable.
+The second property ensures that all formulae are satisfiable in each branch of the tableau.
+But this is a contradiction since the closedness of each branch implies that one of these formulae must be unsatisfiable.
+Thus, \lstinline!$\phi$! must be unsatisfiable.
+
+We formalise the first property in Isabelle as follows.
 \begin{lstlisting}[belowskip=0pt,label={lst:bclosed_sound}]
@@
   for each \lstinline!i $\in$ {0,$\,\ldots\,$,k}!.
-  Therefore, the transitive closure of the membership relation \lstinline!{(x, y). x $\in$ elts y}$^\text{+}$! contains a cycle. 
+  Therefore, the transitive closure of the membership relation
+\begin{lstlisting}
+  {(x, y). x $\in$ elts y}$^\text{+}$ 
+\end{lstlisting}
+  contains a cycle.
   This, however, is a contradiction because the membership relation of \lstinline!V! is well-founded (see \autoref{sec:semantics}) and well-foundedness is closed under transitive closure.
@@
 Moreover, the domain predicate holds for \lstinline!b' @ b! since the assumptions of the modified induction hypothesis are fulfilled.
-Ultimately, we use the corresponding introduction rule for the domain predicate to obtain \lstinline!mlss_proc_branch b!.
+Ultimately, we use the corresponding introduction rule
+\begin{lstlisting}
+  $\llbracket$ $\neg$ lin_sat b; mlss_proc_branch_dom (b' @ b) $\rrbracket$
+  $\Longrightarrow$ mlss_proc_branch_dom b
+\end{lstlisting}
+for the domain predicate to obtain \lstinline!mlss_proc_branch b!.
 
@@
 
-\section{Future Work}
+\section{Conclusion and Future Work}
+By closely following a paper by \citeauthor{new_fast_tableau}~\cite{new_fast_tableau}, we developed a formalisation of a tableau calculus for a ground fragment of set theory called \MLSS{}.
+The formalisation includes an abstract description of a decision procedure that builds on the calculus.
+
+The refinement of the abstract description to an executable specification from which code can be exported is left for future work.
+This executable specification should generate certificates that can be replayed through Isabelle's inference kernel in order to facilitate the integration of the procedure into Isabelle.
+
 Challenges in obtaining an executable specification: membership cycle test unnecessary, fast branch membership test, proof reconstruction.
@@
 
-\section{Conclusion}
-
 \begin{acks}

### 2022-09-26 paper/paper.tex
Finish paper

@@
 \begin{abstract}
-  Using Isabelle/HOL, we verify the state-of-the-art decision procedure for multi-level syllogistics with singleton (\MLSS{} for short), which is a ground fragment of set theory.
+  Using Isabelle/HOL, we verify the state-of-the-art decision procedure for multi-level syllogistic with singleton (\MLSS{} for short), which is a ground fragment of set theory.
   We formalise the syntax and semantics of \MLSS{} as well as a sound and complete tableau calculus for it.
@@
 
-\listoftodos{}
+%%%%%%%%%\listoftodos{}
 
@@
 In Isabelle/HOL, there are specialised procedures for dealing with e.g.\ natural numbers, linear arithmetic, and metric spaces.
-Some of these procedures have been verified in Isabelle/HOL such as a procedure for presburger arithmetic~\cite{presburger} that was later extended to mixed real-integer arithmetic~\cite{arithmetic}.
+Some of these procedures have been verified in Isabelle/HOL such as a procedure for Presburger arithmetic~\cite{presburger} that was later extended to mixed real-integer arithmetic~\cite{arithmetic}.
 This procedure, though, uses reflection to work on goals in Isabelle/HOL, which, during execution, either sacrifices speed by going through the simplifier or soundness by trusting generated code.
 More recently, \citeauthor{orders}~\cite{orders} presented a verified decision procedure for orders that produces certificates. 
-This approach offers efficient execution by using generated code as well as soundness because the certificates are replayed through Isabelle's inference kernel.
+This approach offers efficient execution by using generated code as well as soundness as the certificates are replayed through Isabelle's inference kernel.
 
 The focus of this paper is another ubiquitous structure in mathematics, namely sets.
-In particular, we consider a ground fragment of set theory which \citeauthor{new_fast_tableau}~\cite{new_fast_tableau} call multi-level syllogistics with singleton (\MLSS{}).
+In particular, we consider a ground fragment of set theory which \citeauthor{new_fast_tableau}~\cite{new_fast_tableau} call multi-level syllogistic with singleton (\MLSS{}).
 The fragment includes the usual set operations of union, intersection, difference, membership, equality and, in addition, it allows the construction of singleton sets.
@@
 \begin{itemize}
-  \item We deliver the omitted proof of Lemma~2 in the paper~\cite{new_fast_tableau}, which is key for the completeness proof of the calculus.
+  \item We deliver the omitted proof of Lemma~2 in the paper~\cite{new_fast_tableau}, which is a key building block for the completeness proof of the calculus.
   \item The formal proof of completeness lead to the discovery that the calculus was missing a rule for eliminating double negation. 
@@
 
-All in all, the abstract specification of the decision procedure provides a stepping stone towards a refinement to an executable specification from which code can be generated.
-
 \subsection{Related Work}
-AFP entry~\cite{hybrid_logic_afp} and article~\cite{hybrid_logic} and workshop paper in Isabelle 2022 workshop.
-A Verified Proof of Craig Interpolation for Basic Modal Logic via Tableaux in Lean.
-A Tableau for Ecumenical Propositional Logic.
-Automatization of the semantic tableau method.
-Formalisation de la logique temporelle dans le coq.
-Procedure was later generalised to cartesian product etc.
-Procedure was also generate to add monotone functions.
-\todo[inline]{Look at all articles of From}
-
-First time MLSS~\cite{mlss_first}.
-MLSS is in NP~\cite{mlss_np}.
-First MLSS Tableau~\cite{mlss_first_tableau}.
-Beckert~\cite{tableau_quantifier_free}, Diploma~\cite{tableau_quantifier_free_diploma}.
-Tableau-based MLSS with functions and quantification~\cite{mlss_functions_quantification}.
-MLSS with monotone functions~\cite{mlss_monotone_functions}, reduction to MLSS.
-Another generalisation with cartesion product and some map constructs~\cite{mlss_cartesian_map}.
+Since the literature on decidable fragments of set theory is vast, we only focus on \MLSS{} here.
+The fragment was first shown to be decidable by \citeauthor{mlss_first}~\cite{mlss_first}.
+Subsequent work~\cite{mlss_np} found the decision problem to be \textbf{NP}-complete.
+To obtain a practical decision procedure, \citeauthor{mlss_first_tableau}~\cite{mlss_first_tableau} proposed a tableau calculus, which was later improved by \citeauthor{tableau_quantifier_free}~\cite{tableau_quantifier_free}.
+Both of these procedures construct a model during execution that is used to guide the proof search.
+\citeauthor{tableau_quantifier_free} also cover an extension of the calculus with uninterpreted functions, which was revisited by \citeauthor{mlss_quantification}~\cite{mlss_quantification} who avoided the construction of a model during execution.
+In this paper, we consider a version of the latter procedure due to \citeauthor{new_fast_tableau}~\cite{new_fast_tableau} that is specialised to \MLSS{} and where the branching rules of the calculus are set up in a way to guarantee mutual exclusivity of the branches.
+Later extensions of the calculus added certain interpreted functions such as monotone functions~\cite{mlss_monotone_functions} and the inverse of a function~\cite{mlss_cartesian_map}.
+The latter extension notably includes the Cartesian product.
+Those extensions, though, did not improve upon the tableau calculus for \MLSS{}.
+
+While a lot of work has gone into the verification of decision procedures, seemingly little has been put towards the verification of tableau calculi.
+In Lean, there is a verified tableau calculus for basic modal logic~\cite{modal_logic_lean}.
+In Coq, there is a formalisation of a tableau calculus for ecumenical logic~\cite{ecumenical_logic}; it is, however, missing a completeness proof.
+Concerning Isabelle/HOL, there is formalisation of a sound and complete tableau calculus for hybrid logic~\cite{hybrid_logic} available in the \textit{Archive of Formal Proofs}~\cite{hybrid_logic_afp}.
+The termination proof is part of ongoing work~\cite{hybrid_logic_workshop}.
+% Formalisation de la logique temporelle dans le coq.
 
@@
 Isabelle/HOL~\cite{isabelle} conforms to everyday mathematical notation for the most part.
-For the benefit of the reader that is unfamiliar with Isabelle/HOL, we establish notation and in particular some essential datatypes together with their primitive operations that are specific to Isabelle/HOL.
+We establish notation and in particular some essential data types together with their primitive operations that are specific to Isabelle/HOL.
 
@@
 and \lstinline!'a $\Rightarrow$ 'b!
-for the space of total functions from \lstinline!'a!
-to \lstinline!'b!.
+for the space of total functions from type \lstinline!'a!
+to type \lstinline!'b!.
 
 Sets with elements of type \lstinline!'a! have the type \lstinline!'a set!.
-The cardinality of a set is determined with \lstinline!card! and \lstinline!f ` A! denotes the image of \lstinline!A! under \lstinline!f!.
-
-Analogously to sets, we use \lstinline!'a list!
-to describe the type of lists, which are constructed as the empty list \lstinline![]!
-or with the infix constructor \lstinline!#!, and are appended with the infix operator \lstinline!@!.
+The cardinality of a set \lstinline!A! is denoted by \lstinline!card A! and the image of \lstinline!A! under \lstinline!f! by \lstinline!f ` A!.
+
+We use \lstinline!'a list!
+to describe the type of lists, which are constructed using the empty list \lstinline![]!
+constructor or the infix cons constructor \lstinline!#!, and are appended with the infix operator \lstinline!@!.
 The function \lstinline!set! converts a list into a set.
 
-We remark that \textbf{iff} is equivalent to \lstinline!$=$!
-on type of booleans \lstinline!bool!
+We remark that $\longleftrightarrow$ is equivalent to \lstinline!$=$!
+on the type of Booleans \lstinline!bool!
 and \lstinline!$\equiv$!
 is definitional equality of the meta-logic of Isabelle/HOL, which is called Isabelle/Pure.
-The other important constant of the meta-logic, apart from \lstinline!$\equiv$!, is the meta-implication \lstinline!$\Longrightarrow$!.
-A chain of implications
+Meta-implication is denoted by \lstinline!$\Longrightarrow$! and a chain of implications
 \begin{lstlisting}
@@
 \end{lstlisting}
-can be denoted more concisely by
+can be abbreviated by 
 \begin{lstlisting}
@@
 The only restriction on the type of variables is that it needs to be infinite.
-Isabelle's data type package automatically defines a function that gives us the set of variables in a set term, which we name \lstinline!vars!.
+Isabelle/HOL's data type package automatically defines a function that gives us the set of variables in a set term, which we name \lstinline!vars!.
 In what follows, we will overload the function \lstinline!vars! to also work on set atoms, formulae, and branches.
@@
 
-Similarly to the function \lstinline!vars!, the automatically defined function \lstinline!atoms :: 'a fm $\Rightarrow$ 'a set! retrieves all set atoms in a formula.
-We combine these functions to extract all the variables occuring in a set formula.
+Like the function \lstinline!vars!, the automatically defined function \lstinline!atoms :: 'a fm $\Rightarrow$ 'a set! retrieves all set atoms in a formula.
+We combine these functions to extract all the variables occurring in a set formula.
 \begin{lstlisting}
@@
 
-In addition to the variable that appear in a formula, we will later also need to obtain those set terms that occur as subterms in a formula. 
-Like \lstinline!vars!, we overload the constant \lstinline!subterms!.
+Similarly to \lstinline!vars!, we introduce the overloaded constant \lstinline!subterms!.
 \begin{lstlisting}
@@
 \end{lstlisting}
-The subterms of an atom are then the subterms of those terms that make up the atom.
+The subterms of an atom are then the subterms of its arguments.
 \begin{lstlisting}
@@
 \end{lstlisting}
-Ultimately, we lift this to formulae by using \lstinline!atoms! again.
+Finally, we lift this to formulae by using \lstinline!atoms! again.
 \begin{lstlisting}
@@
 
-Finally, we consider the subformulae of a formula and define a function that computes them.
+Lastly, we consider the subformulae of a formula and define a function that computes them.
 \begin{lstlisting}
@@
 \end{lstlisting}
-The \textit{literals} are a type of subformulae of particular importance so we define a predicate for them.
+\textit{Literals} are a type of subformulae of particular importance, so we define a predicate for them.
 \begin{lstlisting}
@@
 \subsection{Semantics}
-Formulae in \MLSS{} contain arbitrarily nested set terms.
-We account for this nesting by basing the semantics on the von Neumann hierarchy $\mathcal{V}$ of sets that is defined inductively as
+We base the semantics of \MLSS{} on the von Neumann hierarchy $\mathcal{V}$ of sets that is defined inductively as
   \[
@@
 where $\mathcal{P}(S)$ is the powerset of $S$.
-The sets in $\mathcal{V}$ are well-founded meaning that there can be no membership cycle in $\mathcal{V}$.
+The sets in $\mathcal{V}$ are well-founded, i.e.\ there can be no membership cycle in $\mathcal{V}$.
 
 In Isabelle/HOL, which is simply typed, this definition is not accepted, though.
-A way to work around this limitiation is provided by Lawrence in an AFP entry~\cite{zfc_in_hol_afp}, which adds $\mathcal{V}$ to Isabelle/HOL by way of axiomatisation.
+A way to work around this limitation is provided by Lawrence in an entry~\cite{zfc_in_hol_afp} of the \textit{Archive of Formal Proofs}, which adds $\mathcal{V}$ to Isabelle/HOL by way of axiomatisation.
 Two articles~\cite{zfc_in_hol1,zfc_in_hol2} have been written that provide further context to this entry.
@@
   \item The function \lstinline!elts :: V $\Rightarrow$ V set! maps a set to its elements.
-      Note that even though we use the typed sets provided by Isabelle/HOL here, \lstinline!elts! actually returns a class.
-      This class only constitutes a set if the class is small enough.
+      Note that even though we use the typed sets provided by Isabelle/HOL here, \lstinline!elts! actually returns a class (relative to \lstinline!V!).
+      This class only constitutes a set (relative to \lstinline!V!) if the class is small.
   \item The predicate \lstinline!small :: 'a set $\Rightarrow$ bool!, which is polymorphic in \lstinline!'a!, indicates whether a class is small.
@@
 \item The usual set operations such as equality ($=$), union ($\sqcup$), intersection ($\sqcap$), and difference ($-$) are defined.
-\item Finally, the empty set coincides with the ordinal $0$ so it is denoted by \lstinline!0 :: V!.
+\item Finally, the empty set coincides with the ordinal $0$, so it is denoted by \lstinline!0 :: V!.
 \end{itemize}
@@
 
-For formulae, we write \lstinline!M $\models$ $\phi$! for the judgement that $\phi$ holds under the valuation function \lstinline!M!.
-The implementation of $\models$ coincides with the interpretation function of \citeauthor{lqe}~\cite{lqe}, so omit it for brevity.
+We write \lstinline!M $\models$ $\phi$! for the judgement that the formula $\phi$ holds under the valuation function \lstinline!M!.
+The implementation of $\models$ coincides with the interpretation function of \citeauthor{lqe}~\cite{lqe}.
 As usual, a formula \lstinline!$\phi$! is called \textit{satisfiable} if there exists a model \lstinline!$M$! with \lstinline!$M$ $\models$ $\phi$!.
@@
 Inspired by the formalisation of a tableau calculus for hybrid logic by \citeauthor{hybrid_logic_afp}~\cite{hybrid_logic_afp}, we simply use a list to represent a branch of the tableau tree.
-Note that formulae are added to the front of the list during branch expansion so \lstinline!last b! for a branch \lstinline!b! is always the formula that we are trying to disprove with the tableau. 
+Note that formulae are added to the front of the list during branch expansion, so \lstinline!last b! for a branch \lstinline!b! is always the formula that we are trying to disprove with the tableau. 
 \begin{lstlisting}
@@
 \begin{lstlisting}
+definition vars :: 'a branch $\Rightarrow$ 'a set 
+  vars b $\equiv$ $\bigcup$(vars ` set b)
+
 definition subterms ::
@@
   subterms b $\equiv$ $\bigcup$(subterms ` set b)
-
-definition vars :: 'a branch $\Rightarrow$ 'a set 
-  vars b $\equiv$ $\bigcup$(vars ` set b)
 \end{lstlisting}
@@
 with new formulae \lstinline[breaklines=true]!b' = [$\psi_1$,$\ldots$,$\psi_n$]!, which we denote by \lstinline!&\lexpands{b'}{b}&!.
-\autoref{tab:lexpands} shows all linear expansion rules.
-Note that, in the first two rules for \lstinline!$\eqS$!, \lstinline!l! is a literal occuring in the branch.
-Furthermore, the term-for-term substitution \lstinline!l{s/t}! is restricted to the top-level set terms of \lstinline!l!;
+\autoref{tab:lexpands} shows the linear expansion rules.
+Note that in the first two rules for \lstinline!$\eqS$!, \lstinline!l! is a literal occurring in the branch.
+Furthermore, the term-for-term substitution \lstinline!l{s/t}!
+is restricted to the top-level set terms of \lstinline!l!, i.e.\ the set terms that occur directly under one of the atom constructors \lstinline!$\inS$!
+or \lstinline!$\eqS$!;
 for example, given the literal
@@
 \end{lstlisting}
-is saturated --- neither linear nor branching rules apply --- and open but there clearly is no model for the initial formula \lstinline!&\fmAnd{\fmNeg{\fmNeg{\fmNeg{p}}}}{p}&!.
+is saturated --- neither linear nor branching rules apply --- and open, but there clearly is no model for the initial formula \lstinline!&\fmAnd{\fmNeg{\fmNeg{\fmNeg{p}}}}{p}&!.
 
@@
 \end{table*}
-After running out of linear rules apply, only the branching rules shown in \autoref{tab:bexpands} remain.
+After running out of linear rules to apply, only the branching rules shown in \autoref{tab:bexpands} remain.
 A rule is applicable if its \textit{precondition} is met and, to prevent unnecessary branching, it is not subsumed as indicated by the \textit{subsumption condition}.
-These rules create multiple branches in the tableau so we represent the different possibilities \lstinline!bs'! to expand a branch \lstinline!b! as a set and write \lstinline!&\fexpands{bs'}{b}&!. 
+These rules create multiple branches in the tableau, so we represent the different possibilities \lstinline!bs'! to expand a branch \lstinline!b! as a set and write \lstinline!&\fexpands{bs'}{b}&!. 
 Accordingly, we get a new branch \lstinline!b' @ b! in the tableau for each \lstinline!b' $\in$ set bs'!.
@@
 Note that even branching rules are defined such that they never create new subterms with exception of the last rule, which adds a new variable to the branch.
-These variables serve to manifest an inequality so we call them \textit{witnesses}.
+These variables serve to manifest an inequality; hence, we call them \textit{witnesses}.
 \begin{lstlisting}
@@
   $\neg$ lin_sat b
-  $\Longrightarrow$ mlss_proc_branch = mlss_proc_branch ((SOME b'. &\lexpands{b'}{b}& $\land$ set b $\subset$ set (b' @ b)) @ b)
+  $\Longrightarrow$ mlss_proc_branch b = mlss_proc_branch ((SOME b'. &\lexpands{b'}{b}& $\land$ set b $\subset$ set (b' @ b)) @ b)
 | $\llbracket$ lin_sat b; bclosed b $\rrbracket$ $\Longrightarrow$ mlss_proc_branch b = True
@@
 
-\noindent The mechanics of the decision procedure are characteristical for a procedure based on a tableau calculus:
-it determines the satisfiability of a given formula \lstinline!$\phi$! by determining whether the formula has a closed tableau.
+\noindent The mechanics of the decision procedure are typical for a procedure based on a tableau calculus:
+it decides the satisfiability of a given formula \lstinline!$\phi$! by determining whether the formula has a closed tableau.
 More specifically, it initialises the tableau with the singleton branch \lstinline![$\phi$]! and checks whether this branch can be expanded to a closed tableau.
@@
 We abstractly specify the behaviour of the decision procedure, leaving an executable specification from which code can be generated to future work.
-The implementation as displayed in \autoref{fig:alg} uses a couple features of Isabelle's function package:
+The implementation as displayed in \autoref{fig:alg} uses a couple features of Isabelle/HOL's function package:
 instead of defining the function via pattern matching, we specify the equations of the function as conditional rewrite rules.
 This requires us to prove that the assumptions of the equations are non-overlapping, which is done by automation.
-The other concern is that Isabelle requires functions to be total so a recursive function needs to terminate in order for it to be well-defined;
+The other concern is that Isabelle/HOL requires functions to be total, so a recursive function needs to terminate in order for it to be well-defined;
 nevertheless, the termination proof is separated from the definition of the function for modularity.
@@
 which characterises for which arguments the function terminates.
-Each equation of the function receives the assumption that the predicate holds for the argument.
+Each equation of the function is guarded by an assumption that the predicate holds for the argument.
 To be able to reason about this predicate, we pass the \lstinline!domintros! flag that instructs the function package to derive introduction rules for it.
-Before we go into more detail on how the termination is handled, we discuss the definition of \lstinline!mlss_proc_branch!.
+In \autoref{sec:correct}, we will use these rules to discharge the guarding assumptions for the context that \lstinline!mlss_proc_branch! is called in.
+Before we go into more detail on how the termination is proved, we discuss its definition.
 
@@
 Otherwise, we choose an applicable branching rule and recursively check whether all newly created branches can be closed.
-The final equation applies once no further branch expansion is possible, in which case we just tests for closedness of the branch.
+The final equation applies once no further branch expansion is possible, in which case we just test for closedness of the branch.
 
@@
 We call this kind of branches \textit{well-formed}.
-In the definition, the expression \lstinline!&\expandss{b'}{b}&! denotes that \lstinline!b'! is one of the branches that results from applying (potentially zero) expansion rules to \lstinline!b!.
+In the definition below, the expression \lstinline!&\expandss{b'}{b}&!
+denotes that \lstinline!b'!
+is one of the branches that results from applying (potentially zero) expansion rules to \lstinline!b!.
 \begin{lstlisting}
@@
 
-In particular, we will use this notion in \autoref{sec:bound} where we derive an upper bound for the cardinality of a well-formed branch.
+In particular, we will use this notion in \autoref{sec:bound} where we derive an upper bound for the cardinality of well-formed branches.
 This is then used in \autoref{sec:correct} to justify the termination of the decision procedure. 
 Before we come to that, though, we prove soundness and completeness in Section~\ref{sec:sound} and~\ref{sec:complete}, respectively.
-In \autoref{sec:correct}, we also show that both properties easily transfer to the decision procedure which, together with termination, gives us total correctness.
+In \autoref{sec:correct}, we also show that both properties easily transfer to \lstinline!mlss_proc! which, together with termination, establishes that it is a decision procedure.
 
@@
 is in \lstinline!b!.
-Note that we use \citeauthor{graph_theory_afp}'s graph library~\cite{graph_theory_afp} where a graph is represented as a record of vertices, arcs (directed edges), and two functions \lstinline!tail!
+Note that we use \citeauthor{graph_theory_afp}'s graph library~\cite{graph_theory_afp} that represents a graph as a record of vertices, arcs (directed edges), and two functions \lstinline!tail!
 and \lstinline!head!
@@
 We can always choose such a function since we assume an infinite universe of variables.
-Then, we return the singleton set \lstinline!vset {I x}!, which, together with the cardinality constraint, ensures that realisations are distinct between pure witnesses themselves as well as between pure witnesses and set terms.
+Then, we return the singleton set \lstinline!vset {I x}!, which, together with the cardinality constraint, guarantees that realisations are distinct between pure witnesses themselves as well as between pure witnesses and set terms.
 The notation \lstinline!u $\rightarrow_\text{G}$ s! in the definition below indicates that there is an edge from \lstinline!u! to \lstinline!s! in the graph \lstinline!G!.
@@
 
-Again, we need to ensure that the assumptions of the equations are non-overlapping and that function terminates.
+Again, we need to ensure that the assumptions of the equations are non-overlapping and that the function terminates.
 The former is taken care of by automation, leaving us to prove termination.
@@
 They are also a proper subset of \lstinline!av! since \lstinline!p! is an ancestor of \lstinline!v! but not of \lstinline!p! itself.
-Together with the fact that \lstinline!bgraph b! is finite, it follows that the cardinality of the ancestors decreases in each recursive call thus proving the termination of \lstinline!realise!.
+Together with the fact that \lstinline!bgraph b! is finite, it follows that the cardinality of the ancestors decreases in each recursive call, thus proving the termination of \lstinline!realise!.
 
@@
 \subsection{Characterisation of the Pure Witnesses\label{sec:pwits}}
-Recall that the pure witnesses for a branch \lstinline!b! are those witnesses that are not related to other subterms in \lstinline!last b! by equality.
+Recall that the pure witnesses of a branch \lstinline!b! are those witnesses that are not related to other subterms in \lstinline!last b! by equality.
 In the context of a well-formed branch, this characterisation can be strengthened to any set term and, in addition, we also get that there is no membership literal where a pure witness is on the right-hand side.
@@
 To illustrate the usefulness of \hyperref[lst:lemma_2]{\lstinline!lemma_2!}, we prove Proposition~\ref{it:at_eq}.
-The proofs of all propositions translate well into Isabelle so we refer to the original paper~\cite{new_fast_tableau} for the remaining proofs. 
+The proofs of all propositions translate well into Isabelle, so we refer to the original paper~\cite{new_fast_tableau} for the remaining proofs. 
 \begin{proof}[Proof of Proposition~\ref{it:at_eq}]
@@
 \begin{lstlisting}
-  &\phantom{$\Longrightarrow$}& realise s $\neq$ realise t
-  $\Longrightarrow$ elts (realise s) $\neq$ elts (realise t).
-\end{lstlisting}
-  Without loss of generality --- the other case is symmetric ---, we obtain an \lstinline!e! such that
+  realise s $\neq$ realise t
+\end{lstlisting}
+  which implies that
+\begin{lstlisting}
+  elts (realise s) $\neq$ elts (realise t).
+\end{lstlisting}
+  Without loss of generality --- the other case is symmetric --- we obtain an \lstinline!e! such that
 \begin{lstlisting}
@@
 and the definition of \lstinline!realise!, we obtain an \lstinline!u!
-such that \lstinline!e = realise u!
+where \lstinline!e = realise u!
 and \lstinline!u $\rightarrow_\text{bgraph b}$ s!.
@@
   To see that this is true, recall that \lstinline!b! is saturated and thus the subsumption condition of the fifth branching rule must hold.
-  This implies that either \lstinline!u $\inS$ t!
-or \lstinline!u $\notinS$ t!
-is in \lstinline!b!
-but the latter being in \lstinline!b!
-would be a contradiction to \lstinline!u $\nrightarrow_\text{bgraph b}$ t!
-meaning that the former must hold.
+  This implies that either \lstinline!u $\inS$ t! or \lstinline!u $\notinS$ t! is in \lstinline!b!.
+  But the latter being in \lstinline!b! would be a contradiction to \lstinline!u $\nrightarrow_\text{bgraph b}$ t! meaning that the former must hold.
   Ultimately, we infer that \lstinline!u $\in$ s $\diffS$ t! is in \lstinline!b! by invoking saturation again.
@@
 
-\noindent The final step towards obtaining a proper model is to connect the realisation function to the semantics as defined in \autoref{sec:semantics}.
+\noindent The final step to obtain a proper model is to connect the realisation function to the semantics as defined in \autoref{sec:semantics}.
 For set terms, we can use the Propositions~\ref{it:empty}--\ref{it:single} to prove the lemma below by induction on \lstinline!t!.
@@
 \section{Soundness of the Calculus\label{sec:sound}}
-A tableau calculus is sound if, for any closed tableau, the corresponding formula is unsatisfiable. 
+A tableau calculus is sound if for any closed tableau, the corresponding formula is unsatisfiable. 
 We prove the following two properties to establish soundness.
@@
 \end{enumerate}
-To see that these are sufficient, consider a closed tableau for a formula \lstinline!$\phi$! but assume that \lstinline!$\phi$! is satisfiable.
-The second property ensures that all formulae are satisfiable in each branch of the tableau.
-But this is a contradiction since the closedness of each branch implies that one of these formulae must be unsatisfiable.
-Thus, \lstinline!$\phi$! must be unsatisfiable.
 
@@
   We perform a case analysis on the derivation \lstinline!&\fexpands{bs'}{b}&!.
-  As all cases but that of the last rule are straightforward, we only consider this last case where we have 
+  As all cases but that of the last rule are straightforward, we only consider this case.
+  We have 
 \begin{lstlisting}
@@
   for some fresh variable \lstinline!x!, i.e.\ \lstinline!x $\notin$ vars b! and in particular \lstinline!x $\notin$ vars s $\cup$ vars t!.
-  Another consequence of the last rule being applied is that \lstinline!s $\neqS$ t!
-is in \lstinline!b!, which, because \lstinline!M!
-is a model, implies that \lstinline!&\Ist& M s $\neq$ &\Ist& M t!.
+  Another consequence of the last rule being applied is that \lstinline!s $\neqS$ t! is in \lstinline!b!, which implies that \lstinline!&\Ist& M s $\neq$ &\Ist& M t! because \lstinline!M! is a model.
   This inequality manifests itself through some \lstinline!y! with
@@
 
-\section{Upper Bounding the Cardinality of a Branch\label{sec:bound}}
-Soundness and completeness of the calculus are the building blocks with which we establish partial correctness of the algorithm in \autoref{fig:alg}.
-Consequently, we still need to give a termination proof to ascertain total correctness.
+\section{Bounding the Cardinality of a Branch\label{sec:bound}}
+We still need to give a termination proof to ascertain total correctness.
 The core of our argument is that the number of distinct formulae that can be derived with the expansion rules is finite.
@@
 Concerning the second inequality, first observe that each literal contains an atom, implying that the set of literals is at most twice as large as the set of atoms in a branch. 
-In addition, an atom is formed by combining two subterms occuring in the branch with either \lstinline!$\inS$! or \lstinline!$\eqS$!, ultimately yielding
+In addition, an atom is formed by combining two subterms occurring in the branch with either \lstinline!$\inS$! or \lstinline!$\eqS$!;
+hence, we have
 \begin{lstlisting}
@@
 \begin{lstlisting}
-  $\llbracket$ expandss b' b; set b $\subset$ set b' $\rrbracket$
+  $\llbracket$ &\expandss{b'}{b}&; set b $\subset$ set b' $\rrbracket$
   $\Longrightarrow$ mlss_proc_branch_dom b'
@@
 for all \lstinline!b'!.
-The former strengthening is admissable since, by definition, every expansion of a well-formed branch is again well-formed while the latter ensures that the cardinality of \lstinline!b'!
+The former strengthening is admissible since, by definition, every expansion of a well-formed branch is again well-formed while the latter ensures that the cardinality of \lstinline!b'!
 is larger than the cardinality of \lstinline!b!, implying that the difference to the upper bound decreases.
@@
 If \lstinline!b! is not saturated, we can expand \lstinline!b! using a linear or a branching rule, which adds new formulae to the resulting branch(es).
-In the first case, we can thus, using \lstinline!SOME!, choose a branch \lstinline!b'! that linearly expands \lstinline!b!.
+In the first case, we can thus choose a branch \lstinline!b'! with \lstinline!SOME! that linearly expands \lstinline!b!.
 Moreover, the domain predicate holds for \lstinline!b' @ b! since the assumptions of the modified induction hypothesis are fulfilled.
@@
 \begin{proof}
-  We prove the completeness by computation induction on \lstinline!b!.
-
-  The case of the first and third equation, where the branch can be expanded, are immediately resolved with the induction hypothesis.
+  We proceed to prove the completeness by computation induction on \lstinline!b!.
+
+  The cases of the first and third equation, where the branch can be expanded, are immediately resolved with the induction hypothesis.
 
@@
 \begin{lstlisting}
-  expandss b' b $\land$
+  &\expandss{b'}{b}& $\land$
   ($\exists$M. $\forall$$\psi$ $\in$ set b'. M $\models$ $\psi$) $\land$ bclosed b'
@@
 \begin{lstlisting}
-  expandss b'' (b' @ b) $\land$
+  &\expandss{b''}{(b' @ b)}& $\land$
   ($\exists$M. $\forall$$\psi$ $\in$ set b''. M $\models$ $\psi$) $\land$ bclosed b''.
 \end{lstlisting}
-  Due to the transitivity of \lstinline!expandss!, the above also holds if we replace \lstinline!b' @ b! with \lstinline!b!, which is our goal.
+  Due to the transitivity of \lstinline!$\expandssSym$!, the above also holds if we replace \lstinline!b' @ b! with \lstinline!b!, which is our goal.
 
@@
 
-\noindent To finish of the proof of total correctness, note that every singleton branch is trivially well-formed so termination, completeness, and soundness easily transfer to \lstinline!mlss_proc!.
+\noindent To finish of the proof of total correctness, note that every singleton branch is trivially well-formed; thus, termination, completeness, and soundness easily transfer to \lstinline!mlss_proc!.
 \begin{lstlisting}
@@
 
-Challenges in obtaining an executable specification: membership cycle test unnecessary, fast branch membership test, proof reconstruction.
-Additional constructs such as finiteness and pairs.
-Optimisations for the algorithm.
+%Challenges in obtaining an executable specification: membership cycle test unnecessary, fast branch membership test, proof reconstruction.
+%Additional constructs such as finiteness and pairs.
+%Optimisations for the algorithm.
 

### 2022-09-28 paper/paper.tex
Sources, symbols, pdflatex

@@
-\documentclass[sigplan,10pt,anonymous,review]{acmart}
+\documentclass[
+  sigplan,
+  10pt,
+  %anonymous,
+  %review,
+  ]{acmart}
 \settopmatter{printfolios=true,printccs=false,printacmref=false}
@@
 \usepackage{pgfplots}
-\pgfplotsset{compat=1.16}
 
@@
 
-% % LuaLatex on Arxiv: https://tex.stackexchange.com/questions/372154/lualatex-how-to-produce-pdf-acceptable-by-arxiv 
-% \hypersetup{%
+% For arXiv: LuaLatex on Arxiv: https://tex.stackexchange.com/questions/372154/lualatex-how-to-produce-pdf-acceptable-by-arxiv 
+%\hypersetup{%
 %  pdfcreator = {},
 %  pdfproducer = {}
-% }
-% \pdfvariable suppressoptionalinfo \numexpr 1+2+4+8+16+32+64+128+256+512 \relax
+%}
+%\pdfvariable suppressoptionalinfo \numexpr 1+2+4+8+16+32+64+128+256+512 \relax
+
+% For arXiv: Remove copyright information
+% \setcopyright{none}
+% \renewcommand\footnotetextcopyrightpermission[1]{}
 
@@
 \settowidth{\trianglewidth}{\(\triangleleft\)}
-\newcommand{\lefttrianglebar}{\mathrel{\tikz[baseline]{\draw (1ex, 0.75ex) -- (0, 1.25ex) -- (0, 0.25ex) -- cycle; \draw (0, 0.75ex) -- (1ex, 0.75ex);}}}
-\newcommand{\lefttriangle}{\mathrel{\tikz[baseline]{\draw (1ex, 0.75ex) -- (0, 1.25ex) -- (0, 0.25ex) -- cycle;}}}
+\newcommand{\lefttrianglebar}{\mathrel{\includegraphics{./symbols/lefttrianglebar.pdf}}}
+\newcommand{\lefttriangle}{\mathrel{\includegraphics{./symbols/lefttriangle.pdf}}}
 
@@
 \begin{document}
+
+% For arXiv: Remove conference/author in the header
+% \pagestyle{plain}
 
@@
 The termination proof is part of ongoing work~\cite{hybrid_logic_workshop}.
-% Formalisation de la logique temporelle dans le coq.
+% TODO: Formalisation de la logique temporelle dans le coq.
 
@@
 \begin{acks}
-  \todo[inline]{Acknowledge}
+  The author thanks Kevin Kappelmann for his comments on a draft version of the paper.
 \end{acks}

### 2022-11-23 paper/paper.tex
Work in comments from reviewers

@@
 
-%%%%%%%%%\listoftodos{}
+\listoftodos{}
 
@@
 In contrast to pen-and-paper proofs where we expect the reader to fill in trivial details, interactive theorem provers (ITPs) require us to justify each proof step rigorously.
+\todo[inline]{Reviewer: Reformulate below because we can use cuts}
 Manually reducing each proof step down to the axioms would be impractical, though;
@@
 In Isabelle/HOL, which is simply typed, this definition is not accepted, though.
-A way to work around this limitation is provided by Lawrence in an entry~\cite{zfc_in_hol_afp} of the \textit{Archive of Formal Proofs}, which adds $\mathcal{V}$ to Isabelle/HOL by way of axiomatisation.
+A way to work around this limitation is provided by \citeauthor{zfc_in_hol_afp} in an entry~\cite{zfc_in_hol_afp} of the \textit{Archive of Formal Proofs}, which adds $\mathcal{V}$ to Isabelle/HOL by way of axiomatisation.
 Two articles~\cite{zfc_in_hol1,zfc_in_hol2} have been written that provide further context to this entry.
@@
 
-    \lstinline!@\fmNeg{(\fmOr{p}{q})}@, p! & $\Longrightarrow$ & \lstinline!@\fmNeg{q}@! &&
+    \lstinline!@\fmNeg{(\fmAnd{p}{q})}@, p! & $\Longrightarrow$ & \lstinline!@\fmNeg{q}@! &&
     \lstinline!s $\inS$ $\ton$ $\unionS$ $\ttw$, s $\notinS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$! \\
 
-    \lstinline!@\fmNeg{(\fmOr{p}{q})}@, q! & $\Longrightarrow$ & \lstinline!@\fmNeg{p}@! &&
+    \lstinline!@\fmNeg{(\fmAnd{p}{q})}@, q! & $\Longrightarrow$ & \lstinline!@\fmNeg{p}@! &&
     \lstinline!s $\notinS$ $\ton$, s $\notinS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$ $\unionS$ $\ttw$! \\
@@
 We abstractly specify the behaviour of the decision procedure, leaving an executable specification from which code can be generated to future work.
-The implementation as displayed in \autoref{fig:alg} uses a couple features of Isabelle/HOL's function package:
+The implementation as displayed in \autoref{fig:alg} uses a couple of features of Isabelle/HOL's function package:
 instead of defining the function via pattern matching, we specify the equations of the function as conditional rewrite rules.
@@
 A subset of the witnesses, which we call \textit{pure} witnesses, receives special treatment from the realisation function for reasons that will become apparent in \autoref{sec:pwits}.
-The set terms of a branch can thus be partitioned into two collections as defined below.
-\begin{lstlisting}
-definition wits' :: 'a branch $\Rightarrow$ 'a set where
-  wits' b $\equiv$
+The collection of set terms of a branch can thus be partitioned into two collections as defined below.
+\begin{lstlisting}
+definition pwits :: 'a branch $\Rightarrow$ 'a set where
+  pwits b $\equiv$
     {c $\in$ wits b. $\forall$t $\in$ subterms (last b).
@@
     subterms (last b) $\cup$ 
-    Var ` (wits b - wits' b)
+    Var ` (wits b - pwits b)
 \end{lstlisting}
@@
 and \lstinline!head!
-that map an arc to its source respectively target vertex.
+that map an arc to its source, respectively target, vertex.
 \begin{lstlisting}
 definition bgraph b $\equiv$
-  let vs = Var ` wits' b $\cup$ subterms' b
+  let vs = Var ` pwits b $\cup$ subterms' b
   in $\llparenthesis$ verts = vs,
@@
 The realisation function is defined relative to this graph.
-As mentioned before, the pure witnesses \lstinline!Var ` wits' b! and the rest of the set terms \lstinline!subterms' b! are treated differently by the realisation function.
+As mentioned before, the pure witnesses \lstinline!Var ` pwits b! and the rest of the set terms \lstinline!subterms' b! are treated differently by the realisation function.
 Terms in the latter set are evaluated in accordance to the structure of the graph, i.e.\ the realisation of a vertex is defined as the union of the realisations of the parent vertices. 
-For the former set, we choose a function \lstinline!I! that assigns the witnesses in \lstinline!Var ` wits' b! pairwise distinct sets with cardinality greater than that of the vertices.
+For the former set, we choose a function \lstinline!I! that assigns the witnesses in \lstinline!Var ` pwits b! pairwise distinct sets with cardinality greater than that of the vertices.
 We can always choose such a function since we assume an infinite universe of variables.
@@
 function realise :: 'a pset_term $\Rightarrow$ V where
-  x $\in$ Var ` wits' b $\Longrightarrow$ realise x = vset {I x}
+  x $\in$ Var ` pwits b $\Longrightarrow$ realise x = vset {I x}
 | x $\in$ subterms' b
@@
   assumes wf_branch b
-  assumes c $\in$ wits' b
+  assumes c $\in$ pwits b
   shows (Var c $\eqS$ t) $\notin$ set b
@@
   Assume that \lstinline!s $\eqS$ t! is in \lstinline!b!.
-  If there exists a \lstinline!c $\in$ wits' b! where \lstinline!s = Var c! or \lstinline!t = Var c!, we arrive at a contradiction due to \hyperref[lst:lemma_2]{\lstinline!lemma_2!}.
-  Therefore, both \lstinline!s $\in$ subterms' b! and \lstinline!t $\inS$ subterms' b! must hold.
+  If there exists a \lstinline!c $\in$ pwits b! where \lstinline!s = Var c! or \lstinline!t = Var c!, we arrive at a contradiction due to \hyperref[lst:lemma_2]{\lstinline!lemma_2!}.
+  Therefore, both \lstinline!s $\in$ subterms' b! and \lstinline!t $\in$ subterms' b! must hold.
   Now, assume for contradiction that
@@
 
-\noindent To finish of the proof of total correctness, note that every singleton branch is trivially well-formed; thus, termination, completeness, and soundness easily transfer to \lstinline!mlss_proc!.
+\noindent To finish off the proof of total correctness, note that every singleton branch is trivially well-formed; thus, termination, completeness, and soundness easily transfer to \lstinline!mlss_proc!.
 \begin{lstlisting}

### 2023-02-08 paper/CADE/paper.tex
Begin with CADE paper

--- 2022-11-23 paper/paper.tex
@@
-\documentclass[
-  sigplan,
-  10pt,
-  %anonymous,
-  %review,
-  ]{acmart}
-\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[runningheads]{llncs}
+
+\usepackage[T1]{fontenc}
+
+\usepackage{graphicx}
 
 \usepackage[british]{babel}
+
+% citeauthor
+\usepackage[numbers]{natbib}
+\bibliographystyle{splncs04nat}
+
+\usepackage{amsmath}
+\usepackage{amssymb}
+
+\usepackage{booktabs}
+\usepackage{multirow}
 
@@
 
-% Autoref names
+% Autoref
+\usepackage[hidelinks]{hyperref}
 \renewcommand{\sectionautorefname}{Section}
@@
 \newcommand{\fmNegSymbol}{\boldsymbol{\neg}}
-\newcommand{\fmNeg}[1]{$\fmNegSymbol$ #1}
+\newcommand{\fmNeg}[1]{$\fmNegSymbol\:$#1}
 \newcommand{\fmAtom}{\textbf{A}}
@@
 \title{Towards a Verified Tableau Prover for a Ground Fragment of Set Theory}
-\author{Lukas Stevens}
-\orcid{0000-0003-0222-6858}
-\affiliation{%
-  \institution{Technical University of Munich}
-  \department{Department of Informatics}
-  \streetaddress{Boltzmannstr. 3}
-  \city{Garching}
-  \postcode{85748}
-  \country{Germany}
-}
-\email{lukas.stevens@in.tum.de}
+\author{Lukas Stevens\orcidID{0000-0003-0222-6858}}
+\institute{Technical University of Munich, Boltzmannstr.\ 3, 85748 Garching, Germany
+  \email{lukas.stevens@in.tum.de}}
+
+\maketitle
 
@@
   We also provide an abstract specification of a decision procedure that applies the rules of the calculus exhaustively and prove its termination.
+  \keywords{First keyword \and Second keyword}
 \end{abstract}
-
-\maketitle
 
@@
   $\emptyset$ | Var 'a | Single ('a pset_term)
-| 'a pset_term $\unionS$ 'a pset_term
-| 'a pset_term $\interS$ 'a pset_term
+| 'a pset_term $\unionS$ 'a pset_term | 'a pset_term $\interS$ 'a pset_term
 | 'a pset_term $\diffS$ 'a pset_term
@@
 datatype (vars: 'a) pset_atom =
-  'a pset_term $\inS$ 'a pset_term
-| 'a pset_term $\eqS$ 'a pset_term
+  'a pset_term $\inS$ 'a pset_term | 'a pset_term $\eqS$ 'a pset_term
 \end{lstlisting}
@@
 datatype (atoms: 'a) fm =
-  &\fmAtom& 'a | &\fmNeg{('a fm)}&
-| &\fmAnd{'a fm}{'a fm}& | &\fmOr{'a fm}{'a fm}&
+  &\fmAtom& 'a | &\fmNeg{('a fm)}& | &\fmAnd{'a fm}{'a fm}& | &\fmOr{'a fm}{'a fm}&
 
@@
 \begin{lstlisting}
-fun subterms :: 'a pset_term
-             $\Rightarrow$ 'a pset_term set where
+fun subterms :: 'a pset_term $\Rightarrow$ 'a pset_term set where
   subterms $\emptyset$ = {$\emptyset$}
 | subterms (Var x) = {Var x}
-| subterms (s $\unionS$ t) =
-    {s $\unionS$ t} $\cup$ subterms s $\cup$ subterms t
-| subterms (s $\interS$ t) =
-    {s $\interS$ t} $\cup$ subterms s $\cup$ subterms t
-| subterms (s $\diffS$ t) =
-    {s $\diffS$ t} $\cup$ subterms s $\cup$ subterms t
-| subterms (Single t) =
-    {Single t} $\cup$ subterms t
+| subterms (s $\unionS$ t) = {s $\unionS$ t} $\cup$ subterms s $\cup$ subterms t
+| subterms (s $\interS$ t) = {s $\interS$ t} $\cup$ subterms s $\cup$ subterms t
+| subterms (s $\diffS$ t) = {s $\diffS$ t} $\cup$ subterms s $\cup$ subterms t
+| subterms (Single t) = {Single t} $\cup$ subterms t
 \end{lstlisting}
@@
 \begin{lstlisting}
-fun subterms :: 'a pset_atom
-             $\Rightarrow$ 'a pset_term set where
+fun subterms :: 'a pset_atom $\Rightarrow$ 'a pset_term set where
   subterms (s $\inS$ t) = subterms s $\cup$ subterms t
@@
 \begin{lstlisting}
-definition subterms ::
-    'a pset_fm $\Rightarrow$ 'a pset_term set where
+definition subterms :: 'a pset_fm $\Rightarrow$ 'a pset_term set where
   subterms $\phi$ $\equiv$ $\bigcup$(subterms ` atoms $\phi$)
@@
   subfms (&\fmAtom& a) = {&\fmAtom& a}
-| subfms (&\fmAnd{p}{q}&) =
-    {&\fmAnd{p}{q}&} $\cup$ subfms p $\cup$ subfms q
-| subfms (&\fmOr{p}{q}&) =
-    {&\fmOr{p}{q}&} $\cup$ subfms p $\cup$ subfms q
+| subfms (&\fmAnd{p}{q}&) = {&\fmAnd{p}{q}&} $\cup$ subfms p $\cup$ subfms q
+| subfms (&\fmOr{p}{q}&) = {&\fmOr{p}{q}&} $\cup$ subfms p $\cup$ subfms q
 | subfms (&\fmNeg{q}&) = {&\fmNeg{q}&} $\cup$ subfms q
@@
 fun is_literal :: 'a fm $\Rightarrow$ bool where
-  is_literal (&\fmAtom& _) = True
-| is_literal (&\fmNeg{(\fmAtom{} \_)}&) = True
+  is_literal (&\fmAtom& _) = True | is_literal (&\fmNeg{(\fmAtom{} \_)}&) = True
 | is_literal _ = False
@@
 \begin{lstlisting}
-fun &\Isa& :: ('a $\Rightarrow$ V) $\Rightarrow$ 'a pset_atom $\Rightarrow$ bool
-where
+fun &\Isa& :: ('a $\Rightarrow$ V) $\Rightarrow$ 'a pset_atom $\Rightarrow$ bool where
   &\Isa& M (s $\inS$ t) $\longleftrightarrow$ &\Ist& M s $\in$ elts (&\Ist& M t)
@@
 
-definition subterms ::
-    'a branch $\Rightarrow$ 'a pset_term set where
+definition subterms :: 'a branch $\Rightarrow$ 'a pset_term set where
   subterms b $\equiv$ $\bigcup$(subterms ` set b)
@@
 | (t $\neqS$ t) $\in$ set b $\Longrightarrow$ bclosed b
-| $\llbracket$ member_cycle cs; set cs $\subseteq$ set b$\rrbracket$
-    $\Longrightarrow$ bclosed b
+| $\llbracket$ member_cycle cs; set cs $\subseteq$ set b$\rrbracket$ $\Longrightarrow$ bclosed b
 
@@
 \subsection{Linear Expansion Rules}
-\begin{table*}
+\begin{table*}[t]
+  \centering
   \lstset{
@@
   \newcommand{\ttw}{\text{t}_\text{2}}
-  \begin{tabular}{p{0.21\textwidth}cp{0.18\textwidth} p{0.01\textwidth} p{0.21\textwidth}cp{0.18\textwidth}}
+  \begin{tabular}{>{\centering\arraybackslash}p{0.48\textwidth}m{0.02\textwidth}>{\centering\arraybackslash}p{0.48\textwidth}}
     \toprule
-    \multicolumn{3}{c}{Propositional Rules} && \multicolumn{3}{c}{Rules for \lstinline!$\unionS$!} \\
-    \cmidrule{1-3}\cmidrule{5-7}
-    \lstinline!@\fmAnd{p}{q}@! & $\Longrightarrow$ & \lstinline!p, q! &&
-    \lstinline!s $\notinS$ $\ton$ $\unionS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$, s $\notinS$ $\ttw$! \\
-
-    \lstinline!@\fmNeg{(\fmOr{p}{q})}@! & $\Longrightarrow$ & \lstinline!@\fmNeg{p}@, @\fmNeg{q}@! &&
+    Propositional Rules & & Rules for \lstinline!$\unionS$! \\
+    \cmidrule{1-1}\cmidrule{3-3}
+    \begin{tabular}{p{0.22\textwidth}m{0.06\textwidth}p{0.2\textwidth}}
+      \lstinline!@\fmAnd{p}{q}@! & $\Longrightarrow$ & \lstinline!p, q! \\
+      \lstinline!@\fmNeg{(\fmOr{p}{q})}@! & $\Longrightarrow$ & \lstinline!@\fmNeg{p}@, @\fmNeg{q}@! \\
+      \lstinline!@\fmOr{p}{q}, \fmNeg{p}@! & $\Longrightarrow$ & \lstinline!q! \\
+      \lstinline!@\fmOr{p}{q}, \fmNeg{q}@! & $\Longrightarrow$ & \lstinline!p! \\
+      \lstinline!@\fmNeg{(\fmAnd{p}{q})}@, p! & $\Longrightarrow$ & \lstinline!@\fmNeg{q}@! \\
+      \lstinline!@\fmNeg{(\fmAnd{p}{q})}@, q! & $\Longrightarrow$ & \lstinline!@\fmNeg{p}@! \\
+      \lstinline!@\fmNeg{(\fmNeg{p})}@! & $\Longrightarrow$ & \lstinline!p!\\[6ex]
+    \end{tabular}
+                        & &
+    \begin{tabular}{p{0.22\textwidth}m{0.06\textwidth}p{0.2\textwidth}}
+    \lstinline!s $\notinS$ $\ton$ $\unionS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$!, \newline \lstinline!s $\notinS$ $\ttw$! \\
     \lstinline!s $\inS$ $\ton$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$ $\unionS$ $\ttw$! \\
-
-    \lstinline!@\fmOr{p}{q}, \fmNeg{p}@! & $\Longrightarrow$ & \lstinline!q! &&
     \lstinline!s $\in$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$ $\unionS$ $\ttw$! \\
-
-    \lstinline!@\fmOr{p}{q}, \fmNeg{q}@! & $\Longrightarrow$ & \lstinline!p! &&
-    \lstinline!s $\inS$ $\ton$ $\unionS$ $\ttw$, s $\notinS$ $\ton$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ttw$! \\
-
-    \lstinline!@\fmNeg{(\fmAnd{p}{q})}@, p! & $\Longrightarrow$ & \lstinline!@\fmNeg{q}@! &&
-    \lstinline!s $\inS$ $\ton$ $\unionS$ $\ttw$, s $\notinS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$! \\
-
-    \lstinline!@\fmNeg{(\fmAnd{p}{q})}@, q! & $\Longrightarrow$ & \lstinline!@\fmNeg{p}@! &&
+    \lstinline!s $\inS$ $\ton$ $\unionS$ $\ttw$!, \newline \lstinline!s $\notinS$ $\ton$! & ~\newline $\Longrightarrow$ & ~\newline \lstinline!s $\inS$ $\ttw$! \\
+    \lstinline!s $\inS$ $\ton$ $\unionS$ $\ttw$!, \newline \lstinline!s $\notinS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$! \\
     \lstinline!s $\notinS$ $\ton$, s $\notinS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$ $\unionS$ $\ttw$! \\
-    \lstinline!@\fmNeg{(\fmNeg{p})}@! & $\Longrightarrow$ & \lstinline!p! &&&& \\[3ex]
-
-    \multicolumn{3}{c}{Rules for \lstinline!$\interS$!} && \multicolumn{3}{c}{Rules for \lstinline!$\diffS$!} \\
-    \cmidrule{1-3}\cmidrule{5-7}
-    \lstinline!s $\inS$ $\ton$ $\interS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$, s $\inS$ $\ttw$! &&
-    \lstinline!s $\inS$ $\ton$ $\diffS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$, s $\notinS$ $\ttw$! \\
-
-    \lstinline!s $\notinS$ $\ton$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$ $\interS$ $\ttw$! && 
+    \end{tabular}
+    \\
+    \\
+
+    Rules for \lstinline!$\interS$! && Rules for \lstinline!$\diffS$! \\
+    \cmidrule{1-1}\cmidrule{3-3}
+    \begin{tabular}{p{0.22\textwidth}m{0.06\textwidth}p{0.2\textwidth}}
+      \lstinline!s $\inS$ $\ton$ $\interS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$!, \newline \lstinline!s $\inS$ $\ttw$! \\
+      \lstinline!s $\notinS$ $\ton$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$ $\interS$ $\ttw$! \\ 
+      \lstinline!s $\notinS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$ $\interS$ $\ttw$! \\
+      \lstinline!s $\notinS$ $\ton$ $\interS$ $\ttw$!, \newline \lstinline!s $\inS$ $\ton$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ttw$! \\ 
+      \lstinline!s $\notinS$ $\ton$ $\interS$ $\ttw$! \newline \lstinline!s $\inS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$! \\ 
+      \lstinline!s $\inS$ $\ton$, s $\inS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$ $\interS$ $\ttw$! \\
+    \end{tabular}
+    & &
+    \begin{tabular}{p{0.22\textwidth}m{0.06\textwidth}p{0.2\textwidth}}
+      \lstinline!s $\inS$ $\ton$ $\diffS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$!, \newline \lstinline!s $\notinS$ $\ttw$! \\
       \lstinline!s $\notinS$ $\ton$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$ $\diffS$ $\ttw$! \\
-
-    \lstinline!s $\notinS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$ $\interS$ $\ttw$! &&
       \lstinline!s $\inS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$ $\diffS$ $\ttw$! \\
-
-    \lstinline!s $\notinS$ $\ton$ $\interS$ $\ttw$, s $\inS$ $\ton$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ttw$! && 
-    \lstinline!s $\notinS$ $\ton$ $\diffS$ $\ttw$, s $\inS$ $\ton$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ttw$! \\
-
-    \lstinline!s $\notinS$ $\ton$ $\interS$ $\ttw$, s $\inS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$! && 
-    \lstinline!s $\notinS$ $\ton$ $\diffS$ $\ttw$, s $\notinS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$! \\
-
-    \lstinline!s $\inS$ $\ton$, s $\inS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$ $\interS$ $\ttw$! &&
-    \lstinline!s $\inS$ $\ton$, s $\notinS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$ $\diffS$ $\ttw$! \\[3ex]
-
-    \multicolumn{3}{c}{Rules for \lstinline!Single!} && \multicolumn{3}{c}{Rules for \lstinline!$\eqS$!} \\
-    \cmidrule{1-3}\cmidrule{5-7}
-    & $\Longrightarrow$ & \lstinline!s $\inS$ Single s! &&
+      \lstinline!s $\notinS$ $\ton$ $\diffS$ $\ttw$!, \newline \lstinline!s $\inS$ $\ton$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ttw$! \\
+      \lstinline!s $\notinS$ $\ton$ $\diffS$ $\ttw$!, \newline \lstinline!s $\notinS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$! \\
+      \lstinline!s $\inS$ $\ton$, s $\notinS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$ $\diffS$ $\ttw$! \\
+    \end{tabular}
+    \\
+    \\
+
+    Rules for \lstinline!Single! && Rules for \lstinline!$\eqS$! \\
+    \cmidrule{1-1} \cmidrule{3-3}
+    \begin{tabular}{p{0.22\textwidth}m{0.06\textwidth}p{0.2\textwidth}}
+      & $\Longrightarrow$ & \lstinline!s $\inS$ Single s! \\
+      \lstinline!s $\inS$ Single t! & $\Longrightarrow$ & \lstinline!s $\eqS$ t! \\
+      \lstinline!s $\notinS$ Single t! & $\Longrightarrow$ & \lstinline!s $\neqS$ t! \\
+    \end{tabular}
+    & &
+    \begin{tabular}{p{0.22\textwidth}m{0.06\textwidth}p{0.2\textwidth}}
       \lstinline!$\ton$ $\eqS$ $\ttw$, l! & $\Longrightarrow$ & \lstinline!l{$\ttw$/$\ton$}! \\
-
-    \lstinline!s $\inS$ Single t! & $\Longrightarrow$ & \lstinline!s $\eqS$ t! &&
       \lstinline!$\ton$ $\eqS$ $\ttw$, l! & $\Longrightarrow$ & \lstinline!l{$\ton$/$\ttw$}! \\
-
-    \lstinline!s $\notinS$ Single t! & $\Longrightarrow$ & \lstinline!s $\neqS$ t! &&
       \lstinline!s$_\text{1}$ $\inS$ t, s$_\text{2}$ $\notinS$ t! & $\Longrightarrow$ & \lstinline!s$_\text{1}$ $\neqS$ s$_\text{2}$! \\
+    \end{tabular}
+    \\
     \bottomrule
@@
 \begin{lstlisting}
-definition lin_sat b $\equiv$
-  $\forall$b'. &\lexpands{b'}{b}& $\longrightarrow$ set b' $\subseteq$ set b
+definition lin_sat b $\equiv$ $\forall$b'. &\lexpands{b'}{b}& $\longrightarrow$ set b' $\subseteq$ set b
 \end{lstlisting}
@@
   }
+  \centering
   \newcommand{\ton}{\text{t}_\text{1}}
@@
   \newcommand{\lpipe}{\rule[-0.4ex]{0.6pt}{2.3ex}}
-  \begin{tabular}{>{\centering\arraybackslash}p{0.28\textwidth} >{\centering\arraybackslash}p{0.34\textwidth} >{\centering\arraybackslash}p{0.3\textwidth}}
+  \begin{tabular}{>{\centering\arraybackslash}p{0.34\textwidth} >{\centering\arraybackslash}p{0.32\textwidth} >{\centering\arraybackslash}p{0.3\textwidth}}
     \toprule
@@
     \end{tabular}
-         & \lstinline!@\fmOr{p}{q}@ $\in$ set b! & \lstinline!p $\in$ set b $\lor$ @\fmNeg{p}@ $\in$ set b! \\
+         & \lstinline!@\fmOr{p}{q}@ $\in$ set b! &
+         \begin{tabular}{rl}
+         & \lstinline!p $\in$ set b $\lor$! \\
+         $\fmNegSymbol$ & \lstinline!p $\in$ set b! \\
+         \end{tabular} \\
 
@@
     \end{tabular}
-         & \lstinline!@\fmNeg{(\fmAnd{p}{q})}@ $\in$ set b!
-         & \lstinline!@\fmNeg{p}@ $\in$ set b $\lor$ p $\in$ set b! \\
+         & \lstinline!@\fmNeg{(\fmAnd{p}{q})}@ $\in$ set b! &
+          \begin{tabular}{rl}
+            & \lstinline!p $\in$ set b $\lor$! \\
+            $\fmNegSymbol$ & \lstinline!p $\in$ set b! \\
+          \end{tabular} \\
 
@@
            \lstinline!(s $\inS$ $\ton$ $\unionS$ $\ttw$) $\in$ set  b! \\
-           \lstinline!$\ton$ $\unionS$ $\ttw$ $\in$ subterms (last b)! \\
+           \lstinline!$\ton$ $\unionS$ $\ttw$ $\in$ subterms $\phi$! \\
          \end{tabular}
@@
            \lstinline!(s $\inS$ $\ton$) $\in$ set  b! \\
-           \lstinline!($\ton$ $\interS$ $\ttw$) $\in$ subterms (last b)! \\
+           \lstinline!($\ton$ $\interS$ $\ttw$) $\in$ subterms $\phi$! \\
          \end{tabular}
@@
            \lstinline!(s $\inS$ $\ton$) $\in$ set  b! \\
-           \lstinline!($\ton$ $\diffS$ $\ttw$) $\in$ subterms (last b)! \\
+           \lstinline!($\ton$ $\diffS$ $\ttw$) $\in$ subterms $\phi$! \\
          \end{tabular}
@@
 
-    \begin{tabular}{@{}>{\raggedleft\arraybackslash}p{0.11\textwidth}c>{\raggedright\arraybackslash}p{0.11\textwidth}@{}}
+         \begin{tabular}{@{}>{\raggedleft\arraybackslash}p{0.165\textwidth}c>{\raggedright\arraybackslash}p{0.165\textwidth}@{}}
       \\
@@
            \lstinline!($\ton$ $\neqS$ $\ttw$) $\in$ set b! \\
-           \lstinline!$\ton$ $\in$ subterms (last b)! \\
-           \lstinline!$\ttw$ $\in$ subterms (last b)! \\
+           \lstinline!$\ton$ $\in$ subterms $\phi$! \\
+           \lstinline!$\ttw$ $\in$ subterms $\phi$! \\
            \lstinline!x $\notin$ vars b! \\
@@
 
-           \lstinline!$\exists$s.! & \lstinline!(s $\inS$ $\ton$) $\in$ set b $\land$! \\
-                                   & \lstinline!(s $\notinS$ $\ttw$) $\in$ set b! \\
+           \lstinline!$\exists$s.! & \lstinline!(s $\inS$ $\ton$) $\in$ set b! \\
+               \lstinline!$\land$! & \lstinline!(s $\notinS$ $\ttw$) $\in$ set b! \\
            \multicolumn{2}{c}{$\lor$} \\
-           \lstinline!$\exists$s.! & \lstinline!(s $\notinS$ $\ton$) $\in$ set b $\land$! \\
-                                   & \lstinline!(s $\inS$ $\ttw$) $\in$ set b! \\
+           \lstinline!$\exists$s.! & \lstinline!(s $\notinS$ $\ton$) $\in$ set b! \\
+               \lstinline!$\land$! & \lstinline!(s $\inS$ $\ttw$) $\in$ set b! \\
          \end{tabular}
@@
   \bigskip
-  \caption{Branching expansion rules.\label{tab:bexpands}}
+  \caption{Branching expansion rules.\label{tab:bexpands} We write \lstinline!$\phi$! for \lstinline!last b! here.}
 \end{table*}
@@
 \section{A Decision Procedure for MLSS}
-\begin{figure*}
-  \centering
 \begin{lstlisting}
 function (domintros) mlss_proc_branch :: 'a branch $\Rightarrow$ bool where
-  $\neg$ lin_sat b
-  $\Longrightarrow$ mlss_proc_branch b = mlss_proc_branch ((SOME b'. &\lexpands{b'}{b}& $\land$ set b $\subset$ set (b' @ b)) @ b)
+  $\neg$ lin_sat b $\Longrightarrow$ mlss_proc_branch b =
+  mlss_proc_branch ((SOME b'. &\lexpands{b'}{b}& $\land$
+                              set b $\subset$ set (b' @ b)) @ b)
 | $\llbracket$ lin_sat b; bclosed b $\rrbracket$ $\Longrightarrow$ mlss_proc_branch b = True
-| $\llbracket$ $\neg$ sat b; bopen b; lin_sat b $\rrbracket$
-  $\Longrightarrow$ mlss_proc_branch b = ($\forall$b' $\in$ (SOME bs. &\fexpands{bs}{b}&). mlss_proc_branch (b' @ b))
+| $\llbracket$ $\neg$ sat b; bopen b; lin_sat b $\rrbracket$ $\Longrightarrow$ mlss_proc_branch b =
+  ($\forall$b' $\in$ (SOME bs. &\fexpands{bs}{b}&). mlss_proc_branch (b' @ b))
 | $\llbracket$ lin_sat b; sat b $\rrbracket$ $\Longrightarrow$ mlss_proc_branch b = bclosed b
@@
 \end{lstlisting}
-\caption{Definition of the decision procedure through \lstinline|mlss_proc_branch| and \lstinline|mlss_proc|.\label{fig:alg}}
-\end{figure*}
 
@@
 We abstractly specify the behaviour of the decision procedure, leaving an executable specification from which code can be generated to future work.
-The implementation as displayed in \autoref{fig:alg} uses a couple of features of Isabelle/HOL's function package:
+The implementation as displayed in TODO uses a couple of features of Isabelle/HOL's function package:
 instead of defining the function via pattern matching, we specify the equations of the function as conditional rewrite rules.
@@
 definition pwits :: 'a branch $\Rightarrow$ 'a set where
-  pwits b $\equiv$
-    {c $\in$ wits b. $\forall$t $\in$ subterms (last b).
+  pwits b $\equiv$ {c $\in$ wits b. $\forall$t $\in$ subterms (last b).
                           AT (Var c $\eqS$ t) $\notin$ set b $\land$
@@
 
-definition subterms' ::
-    'a branch $\Rightarrow$ 'a pset_term set where
-  subterms' b $\equiv$
-    subterms (last b) $\cup$ 
-    Var ` (wits b - pwits b)
+definition subterms' :: 'a branch $\Rightarrow$ 'a pset_term set where
+  subterms' b $\equiv$ subterms (last b) $\cup$ Var ` (wits b - pwits b)
 \end{lstlisting}
@@
   let vs = Var ` pwits b $\cup$ subterms' b
-  in $\llparenthesis$ verts = vs,
-     &$\phantom{\llparenthesis}$& arcs = {(s, t). (s $\inS$ t) $\in$ set b},
+  in $\llparenthesis$ verts = vs, arcs = {(s, t). (s $\inS$ t) $\in$ set b},
      &$\phantom{\llparenthesis}$& tail = fst, head = snd $\rrparenthesis$
@@
 | x $\in$ subterms' b
-  $\Longrightarrow$ realise t =
-  &\phantom{$\Longrightarrow$ }&  vset {realise ` parents (bgraph b) s}
+  $\Longrightarrow$ realise t = vset {realise ` parents (bgraph b) s}
 | x $\notin$ verts G $\Longrightarrow$ realise x = 0
@@
 lemma lemma_2:
-  assumes wf_branch b
-  assumes c $\in$ pwits b
-  shows (Var c $\eqS$ t) $\notin$ set b
-    and (t $\eqS$ Var c) $\notin$ set b
+  assumes wf_branch b and c $\in$ pwits b
+  shows (Var c $\eqS$ t) $\notin$ set b and (t $\eqS$ Var c) $\notin$ set b
     and (t $\inS$ Var c) $\notin$ set b
@@
 \begin{lstlisting}[columns=fixed]
-  realise u $\in$ elts (realise t)
-        = e $\in$ elts (realise t)
+  realise u $\in$ elts (realise t) = e $\in$ elts (realise t)
 \end{lstlisting}
@@
     realise u $\in$ elts (realise (s $\diffS$ t))
-    &\textover[r]{=}{realise}& e $\in$ elts (realise (s $\diffS$ t)).
+          = e $\in$ elts (realise (s $\diffS$ t)).
   \end{lstlisting}
@@
 lemma coherence:
-  assumes $\phi$ $\in$ set b
-  shows ($\lambda$x. realise (Var x)) $\models$ $\phi$
+  assumes $\phi$ $\in$ set b shows ($\lambda$x. realise (Var x)) $\models$ $\phi$
 \end{lstlisting}
@@
 lemma lexpands_sound:
-  assumes &\lexpands{b'}{b}& and $\phi$ $\in$ set b'
-  assumes $\bigwedge$$\psi$. $\psi$ $\in$ set b $\Longrightarrow$ M $\models$ $\psi$
+  assumes &\lexpands{b'}{b}& and $\phi$ $\in$ set b' and $\bigwedge$$\psi$. $\psi$ $\in$ set b $\Longrightarrow$ M $\models$ $\psi$
   shows M $\models$ $\phi$
@@
 lemma bexpands_sound:
-  assumes &\fexpands{bs'}{b}&
-  assumes $\bigwedge$$\psi$. $\psi$ $\in$ set b $\Longrightarrow$ M $\models$ $\psi$
-  shows $\exists$M'. $\exists$b' $\in$ bs'.
-          $\forall$$\psi$ $\in$ set (b' @ b). M' $\models$ $\psi$
+  assumes &\fexpands{bs'}{b}& and $\bigwedge$$\psi$. $\psi$ $\in$ set b $\Longrightarrow$ M $\models$ $\psi$
+  shows $\exists$M'. $\exists$b' $\in$ bs'. $\forall$$\psi$ $\in$ set (b' @ b). M' $\models$ $\psi$
 \end{lstlisting}
@@
 \begin{lstlisting}
-  bs' = {[Var x $\inS$ s, Var x $\notinS$ t],
-         [Var x $\inS$ t, Var x $\notinS$ s]}
+  bs' = {[Var x $\inS$ s, Var x $\notinS$ t], [Var x $\inS$ t, Var x $\notinS$ s]}
 \end{lstlisting}
@@
   &\phantom{$\leq$}& card {$\psi$ $\in$ set b. is_literal $\psi$}
-  $\leq$ 2 * card (subfms (last b)) +
-  &\phantom{$\leq$}& card {$\psi$ $\in$ set b. is_literal $\psi$}
-  $\leq$ 2 * card (subfms (last b)) +
-  &\phantom{$\leq$}& 16 * (card (subterms (last b)))$^4$
+  $\leq$ 2 * card (subfms (last b)) + card {$\psi$ $\in$ set b. is_literal $\psi$}
+  $\leq$ 2 * card (subfms (last b)) + 16 * (card (subterms (last b)))$^4$
 \end{lstlisting}
@@
   &\phantom{$\leq$}& card {$\psi$ $\in$ set b. $\neg$ is_literal $\psi$}
-  $\leq$ card (subfms (last b)) +
-  &\phantom{$\leq$}& card ($\fmNegSymbol$ ` subfms (last b))
+  $\leq$ card (subfms (last b)) + card ($\fmNegSymbol$ ` subfms (last b))
   $\leq$ 2 * card (subfms (last b))
@@
   &\phantom{$\leq$}& card {$\psi$ $\in$ set b. is_literal $\psi$}
-  $\leq$ 2 * card ($\bigcup$(atoms ` set b))
-  $\leq$ 2 * 2 * (card (subterms b))$^\text{2}$.
+  $\leq$ 2 * card ($\bigcup$(atoms ` set b)) $\leq$ 2 * 2 * (card (subterms b))$^\text{2}$.
 \end{lstlisting}
@@
   &\phantom{$\leq$}& 4 * (card (subterms b))$^\text{2}$
-  $\leq$ 4 * (card (subterms (last b))
-  &\phantom{$\leq$}&      + card (wits b))$^\text{2}$
-  $\leq$ 4 * (card (subterms (last b))
-  &\phantom{$\leq$}&      + (card (subterms (last b))$^\text{2}$)$^\text{2}$
+  $\leq$ 4 * (card (subterms (last b)) + card (wits b))$^\text{2}$
+  $\leq$ 4 * (card (subterms (last b)) + (card (subterms (last b))$^\text{2}$)$^\text{2}$
   $\leq$ 4 * 4 * (card (subterms (last b)))$^\text{4}$.
@@
 lemma mlss_proc_branch_dom_if_wf_branch:
-  assumes wf_branch b
-  shows mlss_proc_branch_dom b
+  assumes wf_branch b shows mlss_proc_branch_dom b
 \end{lstlisting}
@@
 \begin{lstlisting}
-  $\llbracket$ &\expandss{b'}{b}&; set b $\subset$ set b' $\rrbracket$
-  $\Longrightarrow$ mlss_proc_branch_dom b'
+  $\llbracket$ &\expandss{b'}{b}&; set b $\subset$ set b' $\rrbracket$ $\Longrightarrow$ mlss_proc_branch_dom b'
 \end{lstlisting}
@@
   fixes b :: 'a branch
-  assumes wf_branch b
-  assumes $\neg$ mlss_proc_branch b
+  assumes wf_branch b and $\neg$ mlss_proc_branch b
   assumes infinite (UNIV :: 'a set)
@@
 lemma mlss_proc_branch_sound:
-  assumes wf_branch b
-  assumes $\forall$$\psi$ $\in$ set b. M $\models$ $\psi$
+  assumes wf_branch b and $\forall$$\psi$ $\in$ set b. M $\models$ $\psi$
   shows $\neg$ mlss_proc_branch b
@@
 \begin{lstlisting}
-  &\expandss{b'}{b}& $\land$
-  ($\exists$M. $\forall$$\psi$ $\in$ set b'. M $\models$ $\psi$) $\land$ bclosed b'
+  &\expandss{b'}{b}& $\land$ ($\exists$M. $\forall$$\psi$ $\in$ set b'. M $\models$ $\psi$) $\land$ bclosed b'
 \end{lstlisting}
@@
   fixes $\phi$ :: 'a pset_fm
-  assumes $\neg$ mlss_proc $\phi$
-  assumes infinite (UNIV :: 'a set)
+  assumes $\neg$ mlss_proc $\phi$ and infinite (UNIV :: 'a set)
   shows $\exists$M. M $\models$ $\phi$
@@
 theorem mlss_proc_sound:
-  assumes M $\models$ $\phi$
-  shows $\neg$ mlss_proc $\phi$
+  assumes M $\models$ $\phi$ shows $\neg$ mlss_proc $\phi$
 \end{lstlisting}
@@
 
-\begin{acks}
+\subsubsection{Acknowledgements}
 The author thanks Kevin Kappelmann for his comments on a draft version of the paper.
-\end{acks}
-
-\bibliographystyle{ACM-Reference-Format}
+
 \bibliography{sources}

### 2023-02-14 paper/CADE/paper.tex
Continue with paper

@@
 \bibliographystyle{splncs04nat}
+
+\usepackage{appendix}
 
@@
 
+\newcommand{\hfmem}{\boldsymbol{\in}}
+\newcommand{\hfnotmem}{\boldsymbol{\notin}}
+
 \newcommand{\fmAnd}[2]{#1 $\boldsymbol{\land}$ #2}
@@
   We also provide an abstract specification of a decision procedure that applies the rules of the calculus exhaustively and prove its termination.
-  \keywords{First keyword \and Second keyword}
+  \keywords{First TODO \and Second TODO}
 \end{abstract}
@@
   \item We derive an explicit upper bound for the number of formulae in a branch of the tableau.
+  \item We restrict one branching rule TODO.
 \end{itemize}
@@
 The termination proof is part of ongoing work~\cite{hybrid_logic_workshop}.
+TODO: Mention SteP, Check reviewer comment.
 % TODO: Formalisation de la logique temporelle dans le coq.
@@
 Sets with elements of type \lstinline!'a! have the type \lstinline!'a set!.
-The cardinality of a set \lstinline!A! is denoted by \lstinline!card A! and the image of \lstinline!A! under \lstinline!f! by \lstinline!f ` A!.
+The cardinality of a set \lstinline!A! is denoted by \lstinline!|A|! and the image of \lstinline!A! under \lstinline!f! by \lstinline!f ` A!.
 
@@
 
-Similarly to \lstinline!vars!, we introduce the overloaded constant \lstinline!subterms!.
-\begin{lstlisting}
-fun subterms :: 'a pset_term $\Rightarrow$ 'a pset_term set where
-  subterms $\emptyset$ = {$\emptyset$}
-| subterms (Var x) = {Var x}
-| subterms (s $\unionS$ t) = {s $\unionS$ t} $\cup$ subterms s $\cup$ subterms t
-| subterms (s $\interS$ t) = {s $\interS$ t} $\cup$ subterms s $\cup$ subterms t
-| subterms (s $\diffS$ t) = {s $\diffS$ t} $\cup$ subterms s $\cup$ subterms t
-| subterms (Single t) = {Single t} $\cup$ subterms t
-\end{lstlisting}
-The subterms of an atom are then the subterms of its arguments.
-\begin{lstlisting}
-fun subterms :: 'a pset_atom $\Rightarrow$ 'a pset_term set where
-  subterms (s $\inS$ t) = subterms s $\cup$ subterms t
-| subterms (s $\eqS$ t) = subterms s $\cup$ subterms t
-\end{lstlisting}
-Finally, we lift this to formulae by using \lstinline!atoms! again.
-\begin{lstlisting}
-definition subterms :: 'a pset_fm $\Rightarrow$ 'a pset_term set where
-  subterms $\phi$ $\equiv$ $\bigcup$(subterms ` atoms $\phi$)
-\end{lstlisting}
-
-Lastly, we consider the subformulae of a formula and define a function that computes them.
-\begin{lstlisting}
-fun subfms :: 'a fm $\Rightarrow$ 'a fm set where
-  subfms (&\fmAtom& a) = {&\fmAtom& a}
-| subfms (&\fmAnd{p}{q}&) = {&\fmAnd{p}{q}&} $\cup$ subfms p $\cup$ subfms q
-| subfms (&\fmOr{p}{q}&) = {&\fmOr{p}{q}&} $\cup$ subfms p $\cup$ subfms q
-| subfms (&\fmNeg{q}&) = {&\fmNeg{q}&} $\cup$ subfms q
-\end{lstlisting}
+Similarly to \lstinline!vars!, we introduce the overloaded constant \lstinline!subterms :: 'b $\Rightarrow$ 'a pset_term set!, which returns all the set terms that are subterms of a set term, set atom, or formula, respectively.
+Lastly, we consider the subformulae of a formula and define a function \lstinline!subfms :: 'a fm $\Rightarrow$ 'a fm set!
+that computes them.
+The definition of \lstinline!subterms! and \lstinline!subfms! is standard so we omit them here and refer to the appendix.
+
 \textit{Literals} are a type of subformulae of particular importance, so we define a predicate for them.
@@
 \end{lstlisting}
+
 \subsection{Semantics}
-We base the semantics of \MLSS{} on the von Neumann hierarchy $\mathcal{V}$ of sets that is defined inductively as
-  \[
-    \begin{array}{rclr}
-      \mathcal{V}_0 & = & \emptyset, \\
-      \mathcal{V}_{\alpha + 1} & = & \mathcal{P}(\mathcal{V}_\alpha) & \text{for each ordinal $\alpha$}, \\
-      \mathcal{V}_{\lambda} & = & \bigcup_{\mu < \lambda} \mathcal{V}_\mu & \text{for each limit ordinal $\lambda$}, \\
-      \mathcal{V} & = & \bigcup_\alpha \mathcal{V}_\alpha, \\
-    \end{array}
-  \]
-where $\mathcal{P}(S)$ is the powerset of $S$.
-The sets in $\mathcal{V}$ are well-founded, i.e.\ there can be no membership cycle in $\mathcal{V}$.
-
-In Isabelle/HOL, which is simply typed, this definition is not accepted, though.
-A way to work around this limitation is provided by \citeauthor{zfc_in_hol_afp} in an entry~\cite{zfc_in_hol_afp} of the \textit{Archive of Formal Proofs}, which adds $\mathcal{V}$ to Isabelle/HOL by way of axiomatisation.
-Two articles~\cite{zfc_in_hol1,zfc_in_hol2} have been written that provide further context to this entry.
-The entry declares a type \lstinline!V! that comes with the following functionality:
+The original paper~\cite{new_fast_tableau} bases the semantics of \MLSS{} on the von Neumann hierarchy of sets $\mathcal{V}$. 
+While there is an entry~\cite{zfc_in_hol_afp} in the \textit{Archive of Formal Proofs} (AFP) that axiomatises this hierarchy,
+we instead use the hierarchy of \textit{hereditarily finite sets} (HF sets) which fulfil all the same axioms as $\mathcal{V}$ except for the axiom of infinity.
+The HF sets are sufficient for our purposes as \MLSS{} is not powerful enough to force infinite models~\cite[Chapter~14.2]{set_theory}.
+In contrast to $\mathcal{V}$, the HF sets are directly representable in Isabelle/HOL, and indeed, there is an AFP entry~\cite{hf_AFP} that formalises them.
+The entry defines a type \lstinline!hf! that comes with the following functionality:
 \begin{itemize}
-  \item The function \lstinline!elts :: V $\Rightarrow$ V set! maps a set to its elements.
-      Note that even though we use the typed sets provided by Isabelle/HOL here, \lstinline!elts! actually returns a class (relative to \lstinline!V!).
-      This class only constitutes a set (relative to \lstinline!V!) if the class is small.
-  \item The predicate \lstinline!small :: 'a set $\Rightarrow$ bool!, which is polymorphic in \lstinline!'a!, indicates whether a class is small.
-  \item The function \lstinline!vset :: 'a set $\Rightarrow$ V! turns a class into a set of type \lstinline!V! if it is small. 
-\item The usual set operations such as equality ($=$), union ($\sqcup$), intersection ($\sqcap$), and difference ($-$) are defined.
-\item Finally, the empty set coincides with the ordinal $0$, so it is denoted by \lstinline!0 :: V!.
+  \item The function \lstinline!HF :: hf set $\Rightarrow$ set! that converts a finite set of HF sets into an HF set.
+\item The usual set operations such as equality ($=$), membership ($\hfmem$), union ($\sqcup$), intersection ($\sqcap$), and difference ($-$) are defined.
+\item Finally, the empty set coincides with the ordinal $0$, so it is denoted by \lstinline!0 :: hf!.
 \end{itemize}
 
-Equipped with the above, we define the interpretation function \lstinline!&\Ist&!
-for set terms that interprets a set term with respect to a valuation function \lstinline!M :: 'a $\Rightarrow$ V!
-for variables.
-\begin{lstlisting}
-fun &\Ist& :: ('a $\Rightarrow$ V) $\Rightarrow$ 'a pset_term $\Rightarrow$ V where
-  &\Ist& M $\emptyset$ = 0
-| &\Ist& M (Var x) = M x
-| &\Ist& M (Single s) = vset {&\Ist& M s}
-| &\Ist& M (s $\unionS$ t) = &\Ist& M s $\sqcup$ &\Ist& M t 
-| &\Ist& M (s $\interS$ t) = &\Ist& M s $\sqcap$ &\Ist& M t 
-| &\Ist& M (s $\diffS$ t) = &\Ist& M s $-$ &\Ist& M t 
-\end{lstlisting}
-The interpretation function \lstinline!I$_\text{sa}$! for set atoms is straightforward as well.
-\begin{lstlisting}
-fun &\Isa& :: ('a $\Rightarrow$ V) $\Rightarrow$ 'a pset_atom $\Rightarrow$ bool where
-  &\Isa& M (s $\inS$ t) $\longleftrightarrow$ &\Ist& M s $\in$ elts (&\Ist& M t)
-| &\Isa& M (s $\eqS$ t) $\longleftrightarrow$ &\Ist& M s = &\Ist& M t
-\end{lstlisting}
+Equipped with the above, we define the interpretation functions
+\begin{lstlisting}
+  &\Ist& :: ('a $\Rightarrow$ hf) $\Rightarrow$ 'a pset_term $\Rightarrow$ hf &\textnormal{and}&
+  &\Isa& :: ('a $\Rightarrow$ hf) $\Rightarrow$ 'a pset_atom $\Rightarrow$ hf
+\end{lstlisting}
+in the standard way, i.e.\ by mapping each syntactic construct to the corresponding operation on HF sets and interpreting variables with respect to a given valuation function \lstinline!M :: 'a $\Rightarrow$ hf!.
+For the concrete definition we refer to the appendix.
 
@@
 To facilitate inductive reasoning, we show a stronger statement by constructing a model \lstinline!M! such that \lstinline!M $\models$ $\phi$! for all \lstinline!$\phi$ $\in$ set b!.
-At the core of the model, there is a \textit{realisation} function that maps set terms to sets of type \lstinline!V!.
+At the core of the model, there is a \textit{realisation} function that maps set terms to sets of type \lstinline!hf!.
 A subset of the witnesses, which we call \textit{pure} witnesses, receives special treatment from the realisation function for reasons that will become apparent in \autoref{sec:pwits}.
@@
 and \lstinline!head!
-that map an arc to its source, respectively target, vertex.
+that map an arc to its source and target vertex, respectively.
 \begin{lstlisting}
@@
 We can always choose such a function since we assume an infinite universe of variables.
-Then, we return the singleton set \lstinline!vset {I x}!, which, together with the cardinality constraint, guarantees that realisations are distinct between pure witnesses themselves as well as between pure witnesses and set terms.
+Then, we return the singleton set \lstinline!HF {I x}!, which, together with the cardinality constraint, guarantees that realisations are distinct between pure witnesses themselves as well as between pure witnesses and set terms.
 The notation \lstinline!u $\rightarrow_\text{G}$ s! in the definition below indicates that there is an edge from \lstinline!u! to \lstinline!s! in the graph \lstinline!G!.
@@
 function realise :: 'a pset_term $\Rightarrow$ V where
-  x $\in$ Var ` pwits b $\Longrightarrow$ realise x = vset {I x}
+  x $\in$ Var ` pwits b $\Longrightarrow$ realise x = HF {I x}
 | x $\in$ subterms' b
-  $\Longrightarrow$ realise t = vset {realise ` parents (bgraph b) s}
+  $\Longrightarrow$ realise t = HF {realise ` parents (bgraph b) s}
 | x $\notin$ verts G $\Longrightarrow$ realise x = 0
@@
 The assumption that \lstinline!b! is open implies that there are no membership cycles and thus \lstinline!bgraph b! is acyclic.
+TODO: Shorten this proof.
 Now, consider the ancestors \lstinline!av! of a vertex \lstinline!v!, which are those vertices from which we can reach \lstinline!v!.
@@
 So why are pure witnesses treated differently?
-According to the definition of \lstinline!realise!, the pure witnesses would be evaluated to the empty set \lstinline!0 :: V!, were they not treated separately. 
+According to the definition of \lstinline!realise!, the pure witnesses would be evaluated to the empty set \lstinline!0 :: hf!, were they not treated separately. 
 To see that this is a problem, consider the branch
@@
 
-The proof of \hyperref[lst:lemma_2]{\lstinline!lemma_2!} is more technical than interesting so we refer the reader to the formalisation.
+The proof of \hyperref[lst:lemma_2]{\lstinline!lemma_2!} is more technical than interesting so we refer the reader to the formalisation or paper TODO.
 
@@
 \begin{enumerate}[label={(\roman*)}]
-  \item\label{it:at_mem} We have \lstinline!realise s $\in$ elts (realise t)! if it holds that \lstinline!s $\inS$ t! is in \lstinline!b!. 
+  \item\label{it:at_mem} We have \lstinline!realise s $\hfmem$ realise t! if it holds that \lstinline!s $\inS$ t! is in \lstinline!b!. 
   \item\label{it:at_eq} We have \lstinline!realise s $=$ realise t! if \lstinline!s $\eqS$ t! is in \lstinline!b!.
   \item\label{it:af_eq} We have \lstinline!realise s $\neq$ realise t! if \lstinline!s $\neqS$ t! is in \lstinline!b!.
-  \item\label{it:af_mem} We have \lstinline!realise s $\notin$ elts (realise t)! if it holds that \lstinline!s $\notinS$ t! is in \lstinline!b!. 
+  \item\label{it:af_mem} We have \lstinline!realise s $\hfnotmem$ realise t! if it holds that \lstinline!s $\notinS$ t! is in \lstinline!b!. 
 \end{enumerate}
@@
 \begin{lstlisting}
-  realise s $\neq$ realise t
-\end{lstlisting}
-  which implies that
-\begin{lstlisting}
-  elts (realise s) $\neq$ elts (realise t).
+  realise s $\neq$ realise t.
 \end{lstlisting}
@@
 \begin{lstlisting}
-  e $\in$ elts (realise s) $\land$ e $\notin$ elts (realise t). 
+  e $\hfmem$ realise s $\land$ e $\hfnotmem$ realise t. 
 \end{lstlisting}
@@
 \begin{lstlisting}[columns=fixed]
-  realise u $\in$ elts (realise t) = e $\in$ elts (realise t)
-\end{lstlisting}
-  using Proposition~\ref{it:at_mem}, which stands in contradiction to the assumption \lstinline!e $\notin$ elts (realise t)!.
+  realise u $\hfmem$ realise t = e $\hfmem$ realise t
+\end{lstlisting}
+  using Proposition~\ref{it:at_mem}, which stands in contradiction to the assumption \lstinline!e $\hfnotmem$ realise t!.
 \end{proof}
@@
 \noindent The results on literals can now be lowered to set terms.
+All of the proofs are straightforward so we refer the reader to the formalisation.
 \begin{enumerate}[label=(\alph*)]
@@
     \begin{center}
-    \lstinline!realise (Single t) = vset {realise t}!.
+    \lstinline!realise (Single t) = HF {realise t}!.
     \end{center}
 \end{enumerate}
-\begin{proof}[Proof of Proposition~\ref{it:empty}]
-  By definition, \lstinline!realise $\emptyset$! is equal to \lstinline!0! if \lstinline!$\emptyset$! is not a vertex of \lstinline!bgraph b!.
-  Otherwise, \lstinline!$\emptyset$! must be in \lstinline!subterms' b! as it is not a witness.
-  Moreover, since \lstinline!b!
-is open, we know \lstinline!s $\inS$ $\emptyset$!
-is not in \lstinline!b!
-for any \lstinline!s!, implying that the term \lstinline!$\emptyset$!
-has no parents in \lstinline!bgraph b!.
-  Ultimately, we obtain \lstinline!realise $\emptyset$ = 0!.
-\end{proof}
-\begin{proof}[Proof of Proposition~\ref{it:op} for $\star_\texttt{s} = -_\texttt{s}$]
-  For equality, it suffices to prove subset inclusion in both directions.
-
-  \textbf{Direction $\subseteq$. } Fix some \lstinline!e! in \lstinline!elts (realise (s $\diffS$ t))!.
-  Again, we obtain an \lstinline!u! with 
-\begin{lstlisting}
-  e = realise u $\land$ u $\rightarrow_\text{bgraph b}$ (s $\diffS$ t).
-\end{lstlisting}
-  We therefore have \lstinline!(u $\inS$ s $\diffS$ t) $\in$ set b! which, due to \lstinline!b! being saturated, implies that \lstinline!u $\inS$ s! is in \lstinline!b!.
-  Furthermore, the fifth branching rule (see \autoref{tab:bexpands}) must have its subsumption condition met due to saturation of \lstinline!b!, meaning that either one of \lstinline!u $\inS$ t!
-or \lstinline!u $\notinS$ t!
-must be in the branch.
-  Only the latter, however, can be the case since the branch is open.
-  By applying Proposition~\ref{it:at_mem} and~\ref{it:af_mem} we obtain
-\begin{lstlisting}
-  realise u $\in$ realise s $\land$ realise u $\notin$ realise t
-\end{lstlisting}
-  and together with \lstinline!e = realise u! we arrive at the goal 
-\begin{lstlisting}
-  e $\in$ elts (realise s - realise t).
-\end{lstlisting}
-
-\textbf{Direction $\supseteq$. } Let \lstinline[columns=fullflexible]!e $\in$ elts (realise s - realise t)! for some fixed \lstinline!e!.
-  As before, we obtain an \lstinline!u! such that \lstinline!e! is the realisation of \lstinline!u!, i.e.\ \lstinline!e = realisation u!, and
-\begin{lstlisting}
-  u $\rightarrow_\text{bgraph b}$ s $\land$ u $\nrightarrow_\text{bgraph b}$ t.
-\end{lstlisting}
-  An immediate consequence is that \lstinline!u $\inS$ s! is in \lstinline!b!.
-  Moreover, we claim that \lstinline!u $\notinS$ t! is also in \lstinline!b!.
-  To see that this is true, recall that \lstinline!b! is saturated and thus the subsumption condition of the fifth branching rule must hold.
-  This implies that either \lstinline!u $\inS$ t! or \lstinline!u $\notinS$ t! is in \lstinline!b!.
-  But the latter being in \lstinline!b! would be a contradiction to \lstinline!u $\nrightarrow_\text{bgraph b}$ t! meaning that the former must hold.
-  Ultimately, we infer that \lstinline!u $\in$ s $\diffS$ t! is in \lstinline!b! by invoking saturation again.
-  The last step is to apply Proposition~\ref{it:at_mem} in order to obtain the goal
-  \begin{lstlisting}
-    realise u $\in$ elts (realise (s $\diffS$ t))
-          = e $\in$ elts (realise (s $\diffS$ t)).
-  \end{lstlisting}
-\end{proof}
-
-\noindent The final step to obtain a proper model is to connect the realisation function to the semantics as defined in \autoref{sec:semantics}.
+
+The final step to obtain a proper model is to connect the realisation function to the semantics as defined in \autoref{sec:semantics}.
 For set terms, we can use the Propositions~\ref{it:empty}--\ref{it:single} to prove the lemma below by induction on \lstinline!t!.
@@
 \begin{lstlisting}
-  &\Ist& M t$_\text{i}$ $\in$ elts (&\Ist& M t$_\text{i + 1 mod k + 1}$)
+  &\Ist& M t$_\text{i}$ $\hfmem$ &\Ist& M t$_\text{i + 1 mod k + 1}$
 \end{lstlisting}
@@
 \begin{lstlisting}
-  {(x, y). x $\in$ elts y}$^\text{+}$ 
+  {(x, y). x $\hfmem$ y}$^\text{+}$ 
 \end{lstlisting}
   contains a cycle.
-  This, however, is a contradiction because the membership relation of \lstinline!V! is well-founded (see \autoref{sec:semantics}) and well-foundedness is closed under transitive closure.
+  This, however, is a contradiction because the membership relation of \lstinline!hf! is well-founded (see \autoref{sec:semantics} TODO) and well-foundedness is closed under transitive closure.
 \end{proof}
@@
 \begin{lstlisting}
-  &\phantom{$\lor$}& y $\in$ elts (&\Ist& M s) $\land$ y $\notin$ elts (&\Ist& M t)
-  $\lor$ y $\in$ elts (&\Ist& M t) $\land$ y $\notin$ elts (&\Ist& M s).
+  &\phantom{$\lor$}& y $\hfmem$ &\Ist& M s $\land$ y $\hfnotmem$ &\Ist& M t
+  $\lor$ y $\hfmem$ &\Ist& M t $\land$ y $\hfnotmem$ &\Ist& M s.
 \end{lstlisting}
@@
 \section{Bounding the Cardinality of a Branch\label{sec:bound}}
-We still need to give a termination proof to ascertain total correctness.
-The core of our argument is that the number of distinct formulae that can be derived with the expansion rules is finite.
-In particular, we claim that the following upper bound holds.
-\begin{lstlisting}
-  &\phantom{$\leq$}& card (set b)
-  &\textover{$=$}{$\leq$}& card {$\psi$ $\in$ set b. $\neg$ is_literal $\psi$} +
-  &\phantom{$\leq$}& card {$\psi$ $\in$ set b. is_literal $\psi$}
-  $\leq$ 2 * card (subfms (last b)) + card {$\psi$ $\in$ set b. is_literal $\psi$}
-  $\leq$ 2 * card (subfms (last b)) + 16 * (card (subterms (last b)))$^4$
-\end{lstlisting}
-We treat both inequalities separately, starting with the first one whose justification hinges on an invariant: 
-every new composite subformula that is introduced by an expansion rule is the negation of a subformula that already occurs in \lstinline!last b!.
-Written as a formula, we have
-\begin{lstlisting}
-  &\phantom{$\subseteq$}& {$\psi$ $\in$ set b. $\neg$ is_literal $\psi$}
-  $\subseteq$ subfms (last b) $\cup$ $\fmNegSymbol$ ` subfms (last b).
-\end{lstlisting}
-Confirming that this is an invariant is an easy exercise of going through each of the propositional rules in \autoref{tab:lexpands} and the first two branching rules in \autoref{tab:bexpands}.
-Thus, we have
-\begin{lstlisting}
-  &\phantom{$\leq$}& card {$\psi$ $\in$ set b. $\neg$ is_literal $\psi$}
-  $\leq$ card (subfms (last b)) + card ($\fmNegSymbol$ ` subfms (last b))
-  $\leq$ 2 * card (subfms (last b))
-\end{lstlisting}
-which is the first inequality we asserted.
-
-Concerning the second inequality, first observe that each literal contains an atom, implying that the set of literals is at most twice as large as the set of atoms in a branch. 
-In addition, an atom is formed by combining two subterms occurring in the branch with either \lstinline!$\inS$! or \lstinline!$\eqS$!;
-hence, we have
-\begin{lstlisting}
-  &\phantom{$\leq$}& card {$\psi$ $\in$ set b. is_literal $\psi$}
-  $\leq$ 2 * card ($\bigcup$(atoms ` set b)) $\leq$ 2 * 2 * (card (subterms b))$^\text{2}$.
-\end{lstlisting}
-Since no expansion rule introduces new subterms, we know the cardinality of the subterms in a branch \lstinline!b!
-is limited by the subterms of \lstinline!last b!
-plus the number of witnesses added to the branch.
-Note that the last branching rule requires \lstinline!s $\neqS$ t! to be in \lstinline!b! for some \lstinline!s,t $\in$ subterms (last b)!.
-The subsumption condition then prohibits that the rule can be applied again for the combination of \lstinline!s!
-and \lstinline!t!, thus yielding an upper bound of \lstinline!(card (subterms (last b)))$^\text{2}$!
-for the witnesses.
-This allows us to finish the calculation that justifies the second inequality:
-\begin{lstlisting}
-  &\phantom{$\leq$}& 4 * (card (subterms b))$^\text{2}$
-  $\leq$ 4 * (card (subterms (last b)) + card (wits b))$^\text{2}$
-  $\leq$ 4 * (card (subterms (last b)) + (card (subterms (last b))$^\text{2}$)$^\text{2}$
-  $\leq$ 4 * 4 * (card (subterms (last b)))$^\text{4}$.
-\end{lstlisting}
 All in all, we get the following lemma.
@@
   assumes wf_branch b
-  shows card (set b)
-        $\leq$ 2 * card (subfms (last b)) +
-        &\phantom{$\leq$}& 16 * (card (subterms (last b)))$^\text{4}$
+  shows |set b|
+      $\leq$ 2 * |subfms (last b)| + 16 * |subterms (last b)|$^\text{4}$
 \end{lstlisting}
@@
 \begin{proof}
-  We proceed by strong induction on the difference between the upper bound due to \hyperref[lst:card_branch]{\lstinline!card_wf_branch_ub!} and \lstinline!card b!.
+  We proceed by strong induction on the difference between the upper bound due to \hyperref[lst:card_branch]{\lstinline!card_wf_branch_ub!} and \lstinline!|b|!.
 
@@
 
+\appendix
+\clearpage\chapter*{Appendix}
+
+\section{Subterms and Subformulas}
+
+Similarly to \lstinline!vars!, we introduce the overloaded constant \lstinline!subterms!.
+\begin{lstlisting}
+fun subterms :: 'a pset_term $\Rightarrow$ 'a pset_term set where
+  subterms $\emptyset$ = {$\emptyset$}
+| subterms (Var x) = {Var x}
+| subterms (s $\unionS$ t) = {s $\unionS$ t} $\cup$ subterms s $\cup$ subterms t
+| subterms (s $\interS$ t) = {s $\interS$ t} $\cup$ subterms s $\cup$ subterms t
+| subterms (s $\diffS$ t) = {s $\diffS$ t} $\cup$ subterms s $\cup$ subterms t
+| subterms (Single t) = {Single t} $\cup$ subterms t
+\end{lstlisting}
+The subterms of an atom are then the subterms of its arguments.
+\begin{lstlisting}
+fun subterms :: 'a pset_atom $\Rightarrow$ 'a pset_term set where
+  subterms (s $\inS$ t) = subterms s $\cup$ subterms t
+| subterms (s $\eqS$ t) = subterms s $\cup$ subterms t
+\end{lstlisting}
+Finally, we lift this to formulae by using \lstinline!atoms! again.
+\begin{lstlisting}
+definition subterms :: 'a pset_fm $\Rightarrow$ 'a pset_term set where
+  subterms $\phi$ $\equiv$ $\bigcup$(subterms ` atoms $\phi$)
+\end{lstlisting}
+
+Lastly, we consider the subformulae of a formula and define a function that computes them.
+\begin{lstlisting}
+fun subfms :: 'a fm $\Rightarrow$ 'a fm set where
+  subfms (&\fmAtom& a) = {&\fmAtom& a}
+| subfms (&\fmAnd{p}{q}&) = {&\fmAnd{p}{q}&} $\cup$ subfms p $\cup$ subfms q
+| subfms (&\fmOr{p}{q}&) = {&\fmOr{p}{q}&} $\cup$ subfms p $\cup$ subfms q
+| subfms (&\fmNeg{q}&) = {&\fmNeg{q}&} $\cup$ subfms q
+\end{lstlisting}
+
+\section{Definition of the Interpretation Functions}
+
+\begin{lstlisting}
+fun &\Ist& :: ('a $\Rightarrow$ hf) $\Rightarrow$ 'a pset_term $\Rightarrow$ hf where
+  &\Ist& M $\emptyset$ = 0
+| &\Ist& M (Var x) = M x
+| &\Ist& M (Single s) = HF {&\Ist& M s}
+| &\Ist& M (s $\unionS$ t) = &\Ist& M s $\sqcup$ &\Ist& M t 
+| &\Ist& M (s $\interS$ t) = &\Ist& M s $\sqcap$ &\Ist& M t 
+| &\Ist& M (s $\diffS$ t) = &\Ist& M s $-$ &\Ist& M t 
+\end{lstlisting}
+The interpretation function \lstinline!I$_\text{sa}$! for set atoms is straightforward as well.
+\begin{lstlisting}
+fun &\Isa& :: ('a $\Rightarrow$ hf) $\Rightarrow$ 'a pset_atom $\Rightarrow$ bool where
+  &\Isa& M (s $\inS$ t) $\longleftrightarrow$ &\Ist& M s $\hfmem$ &\Ist& M t
+| &\Isa& M (s $\eqS$ t) $\longleftrightarrow$ &\Ist& M s = &\Ist& M t
+\end{lstlisting}
+
+\section{Upper-Bounding the Cardinality of a Branch}
+
+We still need to give a termination proof to ascertain total correctness.
+The core of our argument is that the number of distinct formulae that can be derived with the expansion rules is finite.
+In particular, we claim that the following upper bound holds.
+\begin{lstlisting}
+  &\phantom{$\leq$}& |set b|
+  &\textover{$=$}{$\leq$}& |{$\psi$ $\in$ set b. $\neg$ is_literal $\psi$}| +
+  &\phantom{$\leq$}& |{$\psi$ $\in$ set b. is_literal $\psi$}|
+  $\leq$ 2 * |subfms (last b)| + |{$\psi$ $\in$ set b. is_literal $\psi$}|
+  $\leq$ 2 * |subfms (last b)| + 16 * |subterms (last b)|$^4$
+\end{lstlisting}
+We treat both inequalities separately, starting with the first one whose justification hinges on an invariant: 
+every new composite subformula that is introduced by an expansion rule is the negation of a subformula that already occurs in \lstinline!last b!.
+Written as a formula, we have
+\begin{lstlisting}
+  &\phantom{$\subseteq$}& {$\psi$ $\in$ set b. $\neg$ is_literal $\psi$}
+  $\subseteq$ subfms (last b) $\cup$ $\fmNegSymbol$ ` subfms (last b).
+\end{lstlisting}
+Confirming that this is an invariant is an easy exercise of going through each of the propositional rules in \autoref{tab:lexpands} and the first two branching rules in \autoref{tab:bexpands}.
+Thus, we have
+\begin{lstlisting}
+  &\phantom{$\leq$}& |{$\psi$ $\in$ set b. $\neg$ is_literal $\psi$}|
+  $\leq$ |subfms (last b)| + |$\fmNegSymbol$ ` subfms (last b)|
+  $\leq$ 2 * |subfms (last b)|
+\end{lstlisting}
+which is the first inequality we asserted.
+
+Concerning the second inequality, first observe that each literal contains an atom, implying that the set of literals is at most twice as large as the set of atoms in a branch. 
+In addition, an atom is formed by combining two subterms occurring in the branch with either \lstinline!$\inS$! or \lstinline!$\eqS$!;
+hence, we have
+\begin{lstlisting}
+  &\phantom{$\leq$}& |{$\psi$ $\in$ set b. is_literal $\psi$}|
+  $\leq$ 2 * |$\bigcup$(atoms ` set b)| $\leq$ 2 * 2 * |subterms b|$^\text{2}$.
+\end{lstlisting}
+Since no expansion rule introduces new subterms, we know the cardinality of the subterms in a branch \lstinline!b!
+is limited by the subterms of \lstinline!last b!
+plus the number of witnesses added to the branch.
+Note that the last branching rule requires \lstinline!s $\neqS$ t! to be in \lstinline!b! for some \lstinline!s,t $\in$ subterms (last b)!.
+The subsumption condition then prohibits that the rule can be applied again for the combination of \lstinline!s!
+and \lstinline!t!, thus yielding an upper bound of \lstinline!|subterms (last b)|$^\text{2}$!
+for the witnesses.
+This allows us to finish the calculation that justifies the second inequality:
+\begin{lstlisting}
+  &\phantom{$\leq$}& 4 * |subterms b|$^\text{2}$
+  $\leq$ 4 * |subterms (last b)| + |wits b|$^\text{2}$
+  $\leq$ 4 * |subterms (last b)| + (|subterms (last b)|$^\text{2}$)$^\text{2}$
+  $\leq$ 4 * 4 * |subterms (last b)|$^\text{4}$.
+\end{lstlisting}
+All in all, we get the following lemma.
+\begin{lstlisting}[label={lst:card_branch}]
+lemma card_wf_branch_ub:
+  assumes wf_branch b
+  shows |set b|
+      $\leq$ 2 * |subfms (last b)| + 16 * |subterms (last b)|$^\text{4}$
+\end{lstlisting}
+
 \end{document}

### 2023-02-15 paper/CADE/paper.tex
Continue

@@
 In contrast to pen-and-paper proofs where we expect the reader to fill in trivial details, interactive theorem provers (ITPs) require us to justify each proof step rigorously.
-\todo[inline]{Reviewer: Reformulate below because we can use cuts}
 Manually reducing each proof step down to the axioms would be impractical, though;
@@
 Meta-implication is denoted by \lstinline!$\Longrightarrow$! and a chain of implications
-\begin{lstlisting}
-  A$_\text{1}$ $\Longrightarrow$ $\cdots$ $\Longrightarrow$ A$_\text{k}$ $\Longrightarrow$ C
-\end{lstlisting}
+\lstinline!A$_\text{1}$ $\Longrightarrow$ $\cdots$ $\Longrightarrow$ A$_\text{k}$ $\Longrightarrow$ C!
 can be abbreviated by 
-\begin{lstlisting}
-  $\llbracket$ A$_\text{1}$;$\,\ldots\,$;A$_\text{k}$ $\rrbracket$ $\Longrightarrow$ C.
-\end{lstlisting}
+\lstinline!$\llbracket$ A$_\text{1}$;$\,\ldots\,$;A$_\text{k}$ $\rrbracket$ $\Longrightarrow$ C!.
 
@@
 \begin{lstlisting}
-definition vars :: 'a pset_fm $\Rightarrow$ 'a set where
-  vars $\phi$ $\equiv$ $\bigcup$(vars ` atoms $\phi$)
+definition vars $\phi$ $\equiv$ $\bigcup$(vars ` atoms $\phi$)
 \end{lstlisting}
@@
 that computes them.
-The definition of \lstinline!subterms! and \lstinline!subfms! is standard so we omit them here and refer to the appendix.
+The functions \lstinline!subterms! and \lstinline!subfms! are defined in the usual way (cf. Appendix TODO).
 
@@
 While there is an entry~\cite{zfc_in_hol_afp} in the \textit{Archive of Formal Proofs} (AFP) that axiomatises this hierarchy,
-we instead use the hierarchy of \textit{hereditarily finite sets} (HF sets) which fulfil all the same axioms as $\mathcal{V}$ except for the axiom of infinity.
+we instead use the hierarchy of \textit{hereditarily finite sets} (HF sets) which fulfil all the same axioms as $\mathcal{V}$ --- that is, the axioms of ZF --- except for the axiom of infinity.\todo{Membership is well-founded.}
 The HF sets are sufficient for our purposes as \MLSS{} is not powerful enough to force infinite models~\cite[Chapter~14.2]{set_theory}.
@@
 Equipped with the above, we define the interpretation functions
-\begin{lstlisting}
-  &\Ist& :: ('a $\Rightarrow$ hf) $\Rightarrow$ 'a pset_term $\Rightarrow$ hf &\textnormal{and}&
-  &\Isa& :: ('a $\Rightarrow$ hf) $\Rightarrow$ 'a pset_atom $\Rightarrow$ hf
-\end{lstlisting}
+\begin{itemize}
+  \item \lstinline!&\Ist& :: ('a $\Rightarrow$ hf) $\Rightarrow$ 'a pset_term $\Rightarrow$ hf! and 
+  \item \lstinline!&\Isa& :: ('a $\Rightarrow$ hf) $\Rightarrow$ 'a pset_atom $\Rightarrow$ hf!
+\end{itemize}
 in the standard way, i.e.\ by mapping each syntactic construct to the corresponding operation on HF sets and interpreting variables with respect to a given valuation function \lstinline!M :: 'a $\Rightarrow$ hf!.
@@
 \end{lstlisting}
-We now lift the functions that extract the variables respectively subterms of formulae to branches.
-\begin{lstlisting}
-definition vars :: 'a branch $\Rightarrow$ 'a set 
-  vars b $\equiv$ $\bigcup$(vars ` set b)
-
-definition subterms :: 'a branch $\Rightarrow$ 'a pset_term set where
-  subterms b $\equiv$ $\bigcup$(subterms ` set b)
-\end{lstlisting}
+
+\noindent The functions \lstinline!vars! and \lstinline!subterms! are lifted to branches in the expected way.
+
 In the standard tableau calculus for propositional logic as \citeauthor{tableau}~\cite{tableau} describes it, a branch is called \textit{closed} if it contains both the negation of a formula and the formula itself;
@@
 For \MLSS{}, we extend the notion of closedness with three additional rules; the first two are straightforward while the last one states that a branch is closed when the branch contains a membership cycle
-\begin{lstlisting}
-  t$_\text{0}$ $\inS$ t$_\text{1}$, t$_\text{1}$ $\inS$ t$_\text{2}$, $\ldots$, t$_\text{k}$ $\inS$ t$_\text{0}$.
-\end{lstlisting}
+\lstinline!t$_\text{0}$ $\inS$ t$_\text{1}$, t$_\text{1}$ $\inS$ t$_\text{2}$, $\ldots$, t$_\text{k}$ $\inS$ t$_\text{0}$!.
 
@@
 \begin{table*}[t]
+  \centering
+  \lstset{
+    escapeinside={@}{@}
+  }
+  \newcommand{\ton}{\text{t}_\text{1}}
+  \newcommand{\ttw}{\text{t}_\text{2}}
+  \begin{tabular}{ccc}
+    \toprule
+    Propositional Rules & {\hskip 6pt} & Rules for \lstinline!$\unionS$! \\
+    \cmidrule(l{1pt}){1-1}\cmidrule(r{1pt}){3-3}
+    \begin{tabular}{p{0.18\textwidth}m{0.05\textwidth}p{0.175\textwidth}}
+      \lstinline!@\fmAnd{p}{q}@! & $\Longrightarrow$ & \lstinline!p, q! \\
+      \lstinline!@\fmNeg{(\fmOr{p}{q})}@! & $\Longrightarrow$ & \lstinline!@\fmNeg{p}@, @\fmNeg{q}@! \\
+      \lstinline!@\fmOr{p}{q}, \fmNeg{p}@! & $\Longrightarrow$ & \lstinline!q! \\
+      \lstinline!@\fmOr{p}{q}, \fmNeg{q}@! & $\Longrightarrow$ & \lstinline!p! \\
+      \lstinline!@\fmNeg{(\fmAnd{p}{q})}@, p! & $\Longrightarrow$ & \lstinline!@\fmNeg{q}@! \\
+      \lstinline!@\fmNeg{(\fmAnd{p}{q})}@, q! & $\Longrightarrow$ & \lstinline!@\fmNeg{p}@! \\
+      \lstinline!@\fmNeg{(\fmNeg{p})}@! & $\Longrightarrow$ & \lstinline!p!\\[6ex]
+    \end{tabular}
+                        & &
+    \begin{tabular}{p{0.22\textwidth}m{0.05\textwidth}p{0.22\textwidth}}
+    \lstinline!s $\notinS$ $\ton$ $\unionS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$, s $\notinS$ $\ttw$! \\
+    \lstinline!s $\inS$ $\ton$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$ $\unionS$ $\ttw$! \\
+    \lstinline!s $\in$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$ $\unionS$ $\ttw$! \\
+    \lstinline!s $\inS$ $\ton$ $\unionS$ $\ttw$!, \newline \lstinline!s $\notinS$ $\ton$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ttw$! \\
+    \lstinline!s $\inS$ $\ton$ $\unionS$ $\ttw$!, \newline \lstinline!s $\notinS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$! \\
+    \lstinline!s $\notinS$ $\ton$, s $\notinS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$ $\unionS$ $\ttw$! \\
+    \end{tabular}
+    \\
+    \\
+
+    Rules for \lstinline!$\interS$! && Rules for \lstinline!$\diffS$! \\
+    \cmidrule(l{1pt}){1-1}\cmidrule(r{1pt}){3-3}
+    \begin{tabular}{c}
+      $\vdots$ \\
+    \end{tabular}
+    & &
+    \begin{tabular}{c}
+      $\vdots$ \\
+    \end{tabular}
+    \\
+    \\
+
+    Rules for \lstinline!Single! && Rules for \lstinline!$\eqS$! \\
+    \cmidrule(l{1pt}){1-1}\cmidrule(r{1pt}){3-3}
+    \begin{tabular}{p{0.18\textwidth}m{0.05\textwidth}p{0.175\textwidth}}
+      & $\Longrightarrow$ & \lstinline!s $\inS$ Single s! \\
+      \lstinline!s $\inS$ Single t! & $\Longrightarrow$ & \lstinline!s $\eqS$ t! \\
+      \lstinline!s $\notinS$ Single t! & $\Longrightarrow$ & \lstinline!s $\neqS$ t! \\
+    \end{tabular}
+    & &
+    \begin{tabular}{p{0.22\textwidth}m{0.05\textwidth}p{0.22\textwidth}}
+      \lstinline!$\ton$ $\eqS$ $\ttw$, l! & $\Longrightarrow$ & \lstinline!l{$\ttw$/$\ton$}! \\
+      \lstinline!$\ton$ $\eqS$ $\ttw$, l! & $\Longrightarrow$ & \lstinline!l{$\ton$/$\ttw$}! \\
+      \lstinline!s$_\text{1}$ $\inS$ t, s$_\text{2}$ $\notinS$ t! & $\Longrightarrow$ & \lstinline!s$_\text{1}$ $\neqS$ s$_\text{2}$! \\
+    \end{tabular}
+    \\
+    \bottomrule
+  \end{tabular}
+  \bigskip
+  \caption{Linear expansion rules.\label{tab:lexpands}}
+\end{table*}
+The calculus considers two kinds of branch expansion rules: \textit{linear} and \textit{branching} rules.
+As the name suggests, branching rules lead to the creation of new branches in the tableau while linear rules only extend a branch \lstinline!b!
+with new formulae \lstinline[breaklines=true]!b' = [$\psi_1$,$\ldots$,$\psi_n$]!, which we denote by \lstinline!&\lexpands{b'}{b}&!.
+\autoref{tab:lexpands} shows the linear expansion rules.
+Note that in the first two rules for \lstinline!$\eqS$!, \lstinline!l! is a literal occurring in the branch.
+Furthermore, the term-for-term substitution \lstinline!l{s/t}!
+is restricted to replace \lstinline!t!
+only if it occurs as a top-level set terms of \lstinline!l!, i.e.\ it occurs directly under one of the atom constructors \lstinline!$\inS$!
+or \lstinline!$\eqS$!.
+
+The more crucial restriction of the linear rules is that that no new subterm may be created by their application;
+for instance, the second rule for \lstinline!$\unionS$! is
+\lstinline!s $\inS$ t$_\text{1}$ $\Longrightarrow$ s $\inS$ t$_\text{1}$ $\unionS$ t$_\text{2}$!,
+which formally represents 
+\begin{lstlisting}
+  (s $\inS$ t$_\text{1}$) $\in$ set b $\Longrightarrow$ &\lexpands{[s $\inS$ t$_\text{1}$ $\unionS$ t$_\text{2}$]}{b}&,
+\end{lstlisting}
+may only be used under the condition
+\lstinline[breaklines]!t$_\text{1}$ $\unionS$ t$_\text{2}$ $\in$ subterms (last b)!.
+The purpose of this restriction is to prevent unbounded expansion of the branch.
+In fact, we give an explicit upper bound for the number of formulae in a branch in \autoref{sec:bound}.
+
+Due to boundedness, repeated expansion with linear rules eventually results in a \textit{linearly saturated} branch, i.e.\ a branch where no application of linear rules would produce new formulae.
+\begin{lstlisting}
+definition lin_sat b $\equiv$ $\forall$b'. &\lexpands{b'}{b}& $\longrightarrow$ set b' $\subseteq$ set b
+\end{lstlisting}
+
+Finally, we remark that the original paper~\cite{new_fast_tableau} is missing the last propositional rule dealing with double negation.
+This rule is required for completeness, though, considering that the branch
+\lstinline![&\fmNeg{\fmNeg{\fmNeg{p}}}&, p, &\fmAnd{\fmNeg{\fmNeg{\fmNeg{p}}}}{p}&]!
+is saturated --- neither linear nor branching rules apply --- and open, but there clearly is no model for the initial formula \lstinline!&\fmAnd{\fmNeg{\fmNeg{\fmNeg{p}}}}{p}&!.
+
+\subsection{Branching Rules\label{sec:branching}}
+\begin{table*}
+  \lstset{
+    escapeinside={@}{@}
+  }
+  \centering
+  \newcommand{\ton}{\text{t}_\text{1}}
+  \newcommand{\ttw}{\text{t}_\text{2}}
+  \newcommand{\lpipe}{\rule[-0.4ex]{0.6pt}{2.3ex}}
+  \begin{tabular}{>{\centering\arraybackslash}p{0.34\textwidth} >{\centering\arraybackslash}p{0.32\textwidth} >{\centering\arraybackslash}p{0.3\textwidth}}
+    \toprule
+    Rule & Precondition & Subsumption condition \\
+    \midrule
+    \begin{tabular}{@{}>{\raggedleft\arraybackslash}p{0.11\textwidth}c>{\raggedright\arraybackslash}p{0.11\textwidth}@{}}
+      \\
+      \midrule
+      \lstinline!p! & \lpipe & \lstinline!@\fmNeg{p}@! \\
+    \end{tabular}
+         & \lstinline!@\fmOr{p}{q}@ $\in$ set b! &
+         \begin{tabular}{rl}
+         & \lstinline!p $\in$ set b $\lor$! \\
+         $\fmNegSymbol$ & \lstinline!p $\in$ set b! \\
+         \end{tabular} \\
+
+    \begin{tabular}{@{}>{\raggedleft\arraybackslash}p{0.11\textwidth}c>{\raggedright\arraybackslash}p{0.11\textwidth}@{}}
+      \\
+      \midrule
+      \lstinline!s $\inS$ $\ton$! & \lpipe & \lstinline!s $\notinS$ $\ton$! \\
+    \end{tabular}
+         &
+         \begin{tabular}{c}
+           \lstinline!(s $\inS$ $\ton$ $\unionS$ $\ttw$) $\in$ set  b! \\
+           \lstinline!$\ton$ $\unionS$ $\ttw$ $\in$ subterms $\phi$! \\
+         \end{tabular}
+         &
+         \begin{tabular}{@{}rl@{}}
+                  & \lstinline!(s $\inS$ $\ton$) $\in$ set b! \\
+           $\lor$ & \lstinline!(s $\notinS$ $\ton$) $\in$ set b!
+         \end{tabular}
+         \\
+
+      $\vdots$ & $\vdots$ & $\vdots$ \\
+
+      \addlinespace[0.6em]
+
+      \begin{tabular}{@{}>{\raggedleft\arraybackslash}p{0.165\textwidth}c>{\raggedright\arraybackslash}p{0.165\textwidth}@{}}
+      \\
+      \midrule
+      \lstinline!Var x $\inS$ $\ton$! \lstinline!Var x $\notinS$ $\ttw$! & \multirow{2}{*}{\rule[-0.4ex]{0.6pt}{4.8ex}} & \lstinline!Var x $\notinS$ $\ton$! \lstinline!Var x $\inS$ $\ttw$! \\
+    \end{tabular}
+         & 
+         \begin{tabular}{c}
+           \lstinline!($\ton$ $\neqS$ $\ttw$) $\in$ set b! \\
+           \lstinline!$\ton$ $\in$ subterms $\phi$! \\
+           \lstinline!$\ttw$ $\in$ subterms $\phi$! \\
+           \lstinline!x $\notin$ vars b! \\
+         \end{tabular}
+         &
+         \begin{tabular}{r@{\hspace{0.5em}}l}
+
+           \lstinline!$\exists$s.! & \lstinline!(s $\inS$ $\ton$) $\in$ set b! \\
+               \lstinline!$\land$! & \lstinline!(s $\notinS$ $\ttw$) $\in$ set b! \\
+           \multicolumn{2}{c}{$\lor$} \\
+           \lstinline!$\exists$s.! & \lstinline!(s $\notinS$ $\ton$) $\in$ set b! \\
+               \lstinline!$\land$! & \lstinline!(s $\inS$ $\ttw$) $\in$ set b! \\
+         \end{tabular}
+         \\
+         \bottomrule
+  \end{tabular}
+  \bigskip
+  \caption{Branching expansion rules.\label{tab:bexpands} We write \lstinline!$\phi$! for \lstinline!last b! here.}
+\end{table*}
+After running out of linear rules to apply, only the branching rules shown in \autoref{tab:bexpands} remain.
+A rule is applicable if its \textit{precondition} is met and, to prevent unnecessary branching, it is not subsumed as indicated by the \textit{subsumption condition}.
+These rules create multiple branches in the tableau, so we represent the different possibilities \lstinline!bs'! to expand a branch \lstinline!b! as a set and write \lstinline!&\fexpands{bs'}{b}&!. 
+Accordingly, we get a new branch \lstinline!b' @ b! in the tableau for each \lstinline!b' $\in$ set bs'!.
+
+A linearly saturated branch where no further branching is possible is called a \textit{saturated} branch. 
+\begin{lstlisting}
+definition sat b $\equiv$ lin_sat b $\land$ ($\nexists$bs'. &\fexpands{bs'}{b}&)
+\end{lstlisting}
+
+Note that even branching rules are defined such that they never create new subterms with exception of the last rule, which adds a new variable to the branch.
+These variables serve to manifest an inequality; hence, we call them \textit{witnesses}.
+\begin{lstlisting}
+definition wits b $\equiv$ vars b - vars (last b)
+\end{lstlisting}
+
+\section{A Decision Procedure for MLSS}
+\begin{lstlisting}
+function mlss_proc_branch :: 'a branch $\Rightarrow$ bool where
+  $\neg$ lin_sat b $\Longrightarrow$ mlss_proc_branch b =
+  mlss_proc_branch ((SOME b'. &\lexpands{b'}{b}& $\land$
+                              set b $\subset$ set (b' @ b)) @ b)
+| $\llbracket$ lin_sat b; bclosed b $\rrbracket$ $\Longrightarrow$ mlss_proc_branch b = True
+| $\llbracket$ $\neg$ sat b; bopen b; lin_sat b $\rrbracket$ $\Longrightarrow$ mlss_proc_branch b =
+  ($\forall$b' $\in$ (SOME bs. &\fexpands{bs}{b}&). mlss_proc_branch (b' @ b))
+| $\llbracket$ lin_sat b; sat b $\rrbracket$ $\Longrightarrow$ mlss_proc_branch b = bclosed b
+
+definition mlss_proc :: 'a pset_fm $\Rightarrow$ bool where
+  mlss_proc $\phi$ $\equiv$ mlss_proc_branch [$\phi$]
+\end{lstlisting}
+
+\noindent The mechanics of the decision procedure are typical for a procedure based on a tableau calculus:
+it decides the satisfiability of a given formula \lstinline!$\phi$! by determining whether the formula has a closed tableau.
+More specifically, it initialises the tableau with the singleton branch \lstinline![$\phi$]! and checks whether this branch can be expanded to a closed tableau.
+
+We abstractly specify the behaviour of the decision procedure, leaving an executable specification from which code can be generated to future work.
+The implementation as displayed in TODO uses a couple of features of Isabelle/HOL's function package:
+instead of defining the function via pattern matching, we specify the equations of the function as conditional rewrite rules.
+This requires us to prove that the assumptions of the equations are non-overlapping, which is done by automation.
+The other concern is that Isabelle/HOL requires functions to be total, so a recursive function needs to terminate in order for it to be well-defined;
+nevertheless, the termination proof is separated from the definition of the function for modularity.
+The function package maintains the soundness of the definition by introducing a so-called domain predicate \lstinline!mlss_proc_branch_dom!
+which characterises for which arguments the function terminates.
+Each equation of the function is guarded by an assumption that the predicate holds for the argument.
+In \autoref{sec:correct}, we will show that the domain predicate holds for the context that \lstinline!mlss_proc_branch! is called in.
+Before we go into more detail on how the termination is proved, we discuss its definition.
+
+The purpose of the function is to determine whether a given branch can be expanded to a closed tableau.
+As stated before, we first use linear expansion rules in order to prevent premature branching;
+to this end, we recursively expand the branch with linear rules until the branch is linearly saturated.
+Note that we use Hilbert's $\varepsilon$-operator in the form of \lstinline!SOME! to choose some rule that actually adds new formulae to the branch. 
+As soon as the branch is linearly saturated, we terminate if the branch is closed as shown in the second equation.
+Otherwise, we choose an applicable branching rule and recursively check whether all newly created branches can be closed.
+The final equation applies once no further branch expansion is possible, in which case we just test for closedness of the branch.
+
+The procedure \lstinline!mlss_proc! then calls \lstinline!mlss_proc_branch! with a singleton branch \lstinline![$\phi$]! to determine the satisfiability of a given formula \lstinline!$\phi$!.
+
+This means that \lstinline!mlss_proc_branch! is only applied to branches that result from applying the expansion rules.
+We call this kind of branches \textit{well-formed}.
+In the definition below, the expression \lstinline!&\expandss{b'}{b}&!
+denotes that \lstinline!b'!
+is one of the branches that results from applying (potentially zero) expansion rules to \lstinline!b!.
+\begin{lstlisting}
+definition wf_branch b $\equiv$ $\exists$$\phi$. &\expandss{b}{[$\phi$]}&
+\end{lstlisting}
+
+In particular, we will use this notion in \autoref{sec:bound} where we derive an upper bound for the cardinality of well-formed branches.
+This is then used in \autoref{sec:correct} to justify the termination of the decision procedure. 
+Before we come to that, though, we prove soundness and completeness in Section~\ref{sec:sound} and~\ref{sec:complete}, respectively.
+In \autoref{sec:correct}, we also show that both properties easily transfer to \lstinline!mlss_proc! which, together with termination, establishes that it is a decision procedure.
+
+\section{Completeness of the Calculus\label{sec:complete}}
+For completeness of the calculus, we need to show that every unsatisfiable formula has a closed tableau or, conversely, that the formula is satisfiable if there is a saturated and open branch in the tableau.
+To facilitate inductive reasoning, we show a stronger statement by constructing a model \lstinline!M! such that \lstinline!M $\models$ $\phi$! for all \lstinline!$\phi$ $\in$ set b!.
+At the core of the model, there is a \textit{realisation} function that maps set terms to sets of type \lstinline!hf!.
+A subset of the witnesses, which we call \textit{pure} witnesses, receives special treatment from the realisation function for reasons that will become apparent in \autoref{sec:pwits}.
+The collection of set terms of a branch can thus be partitioned into two collections as defined below.
+\begin{lstlisting}
+definition pwits :: 'a branch $\Rightarrow$ 'a set where
+  pwits b $\equiv$ {c $\in$ wits b. $\forall$t $\in$ subterms (last b).
+                          AT (Var c $\eqS$ t) $\notin$ set b $\land$
+                          AT (t $\eqS$ Var c) $\notin$ set b}
+
+definition subterms' :: 'a branch $\Rightarrow$ 'a pset_term set where
+  subterms' b $\equiv$ subterms (last b) $\cup$ Var ` (wits b - pwits b)
+\end{lstlisting}
+
+We aim to construct a syntactic model that is derived from the membership literals \lstinline!s $\inS$ t! in the branch.
+To this end, we construct a graph whose vertices are the disjoint union of the sets above and there is an edge from \lstinline!s!
+to \lstinline!t!
+in the graph if, and only if, \lstinline!s $\inS$ t!
+is in \lstinline!b!.
+Note that we use \citeauthor{graph_theory_afp}'s graph library~\cite{graph_theory_afp} that represents a graph as a record of vertices, arcs (directed edges), and two functions \lstinline!tail!
+and \lstinline!head!
+that map an arc to its source and target vertex, respectively.
+\begin{lstlisting}
+definition bgraph b $\equiv$ let vs = Var ` pwits b $\cup$ subterms' b
+  in $\llparenthesis$ verts = vs, arcs = {(s, t). (s $\inS$ t) $\in$ set b},
+     &$\phantom{\llparenthesis}$& tail = fst, head = snd $\rrparenthesis$
+\end{lstlisting}
+
+The realisation function is defined relative to this graph.
+As mentioned before, the pure witnesses \lstinline!Var ` pwits b! and the rest of the set terms \lstinline!subterms' b! are treated differently by the realisation function.
+Terms in the latter set are evaluated in accordance to the structure of the graph, i.e.\ the realisation of a vertex is defined as the union of the realisations of the parent vertices. 
+For the former set, we choose a function \lstinline!I! that assigns the witnesses in \lstinline!Var ` pwits b! pairwise distinct sets with cardinality greater than that of the vertices.
+We can always choose such a function since we assume an infinite universe of variables.
+Then, we return the singleton set \lstinline!HF {I x}!, which, together with the cardinality constraint, guarantees that realisations are distinct between pure witnesses themselves as well as between pure witnesses and set terms.
+The notation \lstinline!u $\rightarrow_\text{G}$ s! in the definition below indicates that there is an edge from \lstinline!u! to \lstinline!s! in the graph \lstinline!G!.
+\begin{lstlisting}
+abbreviation parents G s $\equiv$ {u. u $\rightarrow_\text{G}$ s}
+
+function realise :: 'a pset_term $\Rightarrow$ V where
+  x $\in$ Var ` pwits b $\Longrightarrow$ realise x = HF {I x}
+| x $\in$ subterms' b
+  $\Longrightarrow$ realise t = HF {realise ` parents (bgraph b) s}
+| x $\notin$ verts G $\Longrightarrow$ realise x = 0
+\end{lstlisting}
+
+Again, we need to ensure that the assumptions of the equations are non-overlapping and that the function terminates.
+The former is taken care of by automation, leaving us to prove termination.
+The assumption that \lstinline!b! is open implies that there are no membership cycles and thus \lstinline!bgraph b! is acyclic.
+Furthermore, the graph is finite by definition.
+Thus, we can use the cardinality of the set of ancestors as a measure that decreases in each recursive call.
+
+Before we prove that the realisation function constitutes a model in \autoref{sec:real}, we will first explain the significance of the pure witnesses.
+
+\subsection{Characterisation of the Pure Witnesses\label{sec:pwits}}
+Recall that the pure witnesses of a branch \lstinline!b! are those witnesses that are not related to other subterms in \lstinline!last b! by equality.
+In the context of a well-formed branch, this characterisation can be strengthened to any set term and, in addition, we also get that there is no membership literal where a pure witness is on the right-hand side.
+Intuitively speaking, the realisation of a pure witness does not depend on the realisation of any other set term.
+\begin{lstlisting}[label={lst:lemma_2}]
+lemma lemma_2:
+  assumes wf_branch b and c $\in$ pwits b
+  shows (Var c $\eqS$ t) $\notin$ set b and (t $\eqS$ Var c) $\notin$ set b
+    and (t $\inS$ Var c) $\notin$ set b
+\end{lstlisting}
+So why are pure witnesses treated differently?
+According to the definition of \lstinline!realise!, the pure witnesses would be evaluated to the empty set \lstinline!0 :: hf!, were they not treated separately. 
+To see that this is a problem, consider the branch
+\lstinline!b = [Var s $\neqS$ Var t, Var t $\neqS$ Var u]!
+which expands to several open and saturated branches, one of which is
+\begin{lstlisting}
+  [Var x $\neqS$ Var y, Var x $\inS$ Var s, Var x $\notinS$ Var t,
+  &\phantom{[Var x $\neqS$ Var y,}& Var y $\inS$ Var t, Var y $\notinS$ Var u] @ b
+\end{lstlisting}
+for some fresh \lstinline!x! and \lstinline!y!.
+Assigning both \lstinline!Var x! and \lstinline!Var y! a value of \lstinline!0! would contradict the literal \lstinline!Var x $\neqS$ Var y!.
+To prevent this, we assign the pure witnesses pairwise different values.
+
+The proof of \hyperref[lst:lemma_2]{\lstinline!lemma_2!} is more technical than interesting so we refer the reader to the formalisation or paper TODO.
+
+\subsection{Realisation of an Open Branch\label{sec:real}}
+Remember that for completeness, we need to show that the realisation function for an open and saturated branch \lstinline!b! actually constitutes a model for all formulae in the branch.
+We start by verifying that the realisation function models all literals in the branch; more formally, the following propositions hold:
+\begin{enumerate}[label={(\roman*)}]
+  \item\label{it:at_mem} We have \lstinline!realise s $\hfmem$ realise t! if it holds that \lstinline!s $\inS$ t! is in \lstinline!b!. 
+  \item\label{it:at_eq} We have \lstinline!realise s $=$ realise t! if \lstinline!s $\eqS$ t! is in \lstinline!b!.
+  \item\label{it:af_eq} We have \lstinline!realise s $\neq$ realise t! if \lstinline!s $\neqS$ t! is in \lstinline!b!.
+  \item\label{it:af_mem} We have \lstinline!realise s $\hfnotmem$ realise t! if it holds that \lstinline!s $\notinS$ t! is in \lstinline!b!. 
+\end{enumerate}
+To illustrate the usefulness of \hyperref[lst:lemma_2]{\lstinline!lemma_2!}, we prove Proposition~\ref{it:at_eq}.
+The proofs of all propositions translate well into Isabelle, so we refer to the original paper~\cite{new_fast_tableau} for the remaining proofs. 
+\begin{proof}[Proof of Proposition~\ref{it:at_eq}]
+  Assume that \lstinline!s $\eqS$ t! is in \lstinline!b!.
+  If there exists a \lstinline!c $\in$ pwits b! where \lstinline!s = Var c! or \lstinline!t = Var c!, we arrive at a contradiction due to \hyperref[lst:lemma_2]{\lstinline!lemma_2!}.
+  Therefore, both \lstinline!s $\in$ subterms' b! and \lstinline!t $\in$ subterms' b! must hold.
+  Now, assume for contradiction that \lstinline!realise s $\neq$ realise t!.
+  Without loss of generality --- the other case is symmetric --- we obtain an \lstinline!e! such that
+  \lstinline!e $\hfmem$ realise s! and \lstinline!e $\hfnotmem$ realise t!. 
+  Considering that \lstinline!s $\in$ subterms' b!
+and the definition of \lstinline!realise!, we obtain an \lstinline!u!
+where \lstinline!e = realise u!
+and \lstinline!u $\rightarrow_\text{bgraph b}$ s!.
+  This, in turn, yields that \lstinline!u $\inS$ s! must be in \lstinline!b!.
+  Together with the assumption \lstinline!(s $\eqS$ t) $\in$ set b! and the saturation of \lstinline!b!, it follows that \lstinline!u $\inS$ t! must also be in \lstinline!b!.
+  But then we have
+  \lstinline!realise u $\hfmem$ realise t = e $\hfmem$ realise t!
+  using Proposition~\ref{it:at_mem}, which stands in contradiction to the assumption \lstinline!e $\hfnotmem$ realise t!.
+\end{proof}
+
+\noindent The results on literals can now be lowered to set terms.
+All of the proofs are straightforward so we refer the reader to the formalisation.
+\begin{enumerate}[label=(\alph*)]
+  \item\label{it:empty} It holds that \lstinline!realise $\emptyset$ = 0!.
+  \item\label{it:op} Let \lstinline!$\star_\text{s}$ $\in$ {$\unionS$, $\diffS$, $\interS$}!. If the term \lstinline!s $\star_\text{s}$ t! occurs in \lstinline!subterms b!, then
+    \begin{center}
+    \lstinline!realise (s $\star_\text{s}$ t) = realise s $\star$ realise t!.
+    \end{center}
+  \item\label{it:single} If \lstinline!Single t $\in$ subterms b!, then
+    \begin{center}
+    \lstinline!realise (Single t) = HF {realise t}!.
+    \end{center}
+\end{enumerate}
+
+The final step to obtain a proper model is to connect the realisation function to the semantics as defined in \autoref{sec:semantics}.
+For set terms, we can use the Propositions~\ref{it:empty}--\ref{it:single} to prove the lemma below by induction on \lstinline!t!.
+\begin{lstlisting}
+lemma assumes t $\in$ subterms b
+      shows &\Ist& ($\lambda$x. realise (Var x)) t = realise t
+\end{lstlisting}
+Lifting the above result to formulae yields the coherence of \lstinline!b! as the original paper~\cite{new_fast_tableau} calls it.
+The proof is a tedious but straightforward induction on the the size of the formulae.
+\begin{lstlisting}[label={lst:coherence}]
+lemma coherence:
+  assumes $\phi$ $\in$ set b shows ($\lambda$x. realise (Var x)) $\models$ $\phi$
+\end{lstlisting}
+The coherence property finishes the proof of completeness of the calculus as it gives us a model for every formula in an open and saturated branch.
+
+\section{Soundness of the Calculus\label{sec:sound}}
+A tableau calculus is sound if for any closed tableau, the corresponding formula is unsatisfiable. 
+We prove the following two properties to establish soundness.
+\begin{enumerate}[label={(\roman*)}]
+  \item A closed branch contains an unsatisfiable formula.
+  \item The expansion rules maintain satisfiability.
+\end{enumerate}
+
+We formalise the first property in Isabelle below.
+
+\begin{lstlisting}[belowskip=0pt,label={lst:bclosed_sound}]
+lemma bclosed_sound:
+  assumes bclosed b shows $\exists$$\phi$ $\in$ set b. M $\not\models$ $\phi$
+\end{lstlisting}
+\begin{proof}
+  It is clear that, for any \lstinline!s!, neither does \lstinline!M! model \lstinline!s $\in$ $\emptyset$! nor \lstinline!s $\neqS$ s!.
+  Furthermore, no model can satisfy both \lstinline!$\phi$! and \lstinline!&\fmNeg{$\phi$}&! at the same time.
+  Lastly, a membership cycle is impossible since the membership relation of \lstinline!hf! is well-founded (see \autoref{sec:semantics} TODO).
+\end{proof}
+
+\noindent We are left with showing that both linear and branching expansion rules preserve satisfiability.
+As for the linear rules, a straightforward proof by case analysis on \lstinline!&\lexpands{b'}{b}&! suffices to obtain the lemma below.
+\begin{lstlisting}[label={lst:lexpands_sound}]
+lemma lexpands_sound:
+  assumes &\lexpands{b'}{b}& and $\phi$ $\in$ set b' and $\bigwedge$$\psi$. $\psi$ $\in$ set b $\Longrightarrow$ M $\models$ $\psi$
+  shows M $\models$ $\phi$
+\end{lstlisting}
+A similar argument would work for the branching rules if it were not for the last rule adding new variables.
+Those variables need to be assigned specific values; hence, we modify the model as shown in the proof below.
+\begin{lstlisting}[belowskip=0pt, label={lst:bexpands_sound}]
+lemma bexpands_sound:
+  assumes &\fexpands{bs'}{b}& and $\bigwedge$$\psi$. $\psi$ $\in$ set b $\Longrightarrow$ M $\models$ $\psi$
+  shows $\exists$M'. $\exists$b' $\in$ bs'. $\forall$$\psi$ $\in$ set (b' @ b). M' $\models$ $\psi$
+\end{lstlisting}
+\begin{proof}
+  We only consider the case where \lstinline!&\fexpands{bs'}{b}&! was proved by applying the last rule to \lstinline!s $\neqS$ t! for some \lstinline!s! and \lstinline!t!.
+  We have 
+\begin{lstlisting}
+  bs' = {[Var x $\inS$ s, Var x $\notinS$ t], [Var x $\inS$ t, Var x $\notinS$ s]}
+\end{lstlisting}
+  for some fresh variable \lstinline!x!.
+  Since \lstinline!s $\neqS$ t! is in \lstinline!b!, we have that \lstinline!&\Ist& M s $\neq$ &\Ist& M t! because \lstinline!M! is a model.
+  Without loss of generality this inequality manifests itself through some \lstinline!y! with
+  \lstinline!y $\hfmem$ &\Ist& M s! and \lstinline!y $\hfnotmem$ &\Ist& M t!.
+  We update \lstinline!M! such that it maps \lstinline!x! to \lstinline!y! to obtain the assignment \lstinline!M'!.
+  Note that \lstinline!M'! is still a model for formulae in \lstinline!b! because \lstinline!x! is fresh with respect to \lstinline!b!.
+  Furthermore, it is also a model for the first branch in \lstinline!bs'!, which finishes the proof.
+\end{proof}
+
+\section{Bounding the Cardinality of a Branch\label{sec:bound}}
+All in all, we get the following lemma.
+\begin{lstlisting}[label={lst:card_branch}]
+lemma card_wf_branch_ub:
+  assumes wf_branch b
+  shows |set b|
+      $\leq$ 2 * |subfms (last b)| + 16 * |subterms (last b)|$^\text{4}$
+\end{lstlisting}
+
+\section{Total Correctness of the Decision Procedure\label{sec:correct}}
+We first prove the termination of the procedure for well-formed branches, i.e.\ every well-formed branch is in the domain of \lstinline!mlss_proc_branch!.
+\begin{lstlisting}[belowskip=0pt]
+lemma mlss_proc_branch_dom_if_wf_branch:
+  assumes wf_branch b shows mlss_proc_branch_dom b
+\end{lstlisting}
+\begin{proof}
+  We proceed by strong induction on the difference between the upper bound due to \hyperref[lst:card_branch]{\lstinline!card_wf_branch_ub!} and \lstinline!|b|!.
+
+  We only have one inductive case where the induction hypothesis holds for all well-formed branches \lstinline!b'!
+where the difference between the upper bound and the cardinality is smaller than that of \lstinline!b!.
+  We strengthen the assumptions of the induction hypothesis to obtain
+\begin{lstlisting}
+  $\llbracket$ &\expandss{b'}{b}&; set b $\subset$ set b' $\rrbracket$ $\Longrightarrow$ mlss_proc_branch_dom b'
+\end{lstlisting}
+for all \lstinline!b'!.
+The former strengthening is admissible since, by definition, every expansion of a well-formed branch is again well-formed while the latter ensures that the cardinality of \lstinline!b'!
+is larger than the cardinality of \lstinline!b!, implying that the difference to the upper bound decreases.
+
+Now, consider the case where \lstinline!b! is saturated.
+Then, the procedure terminates without recursive calls and we immediately get \lstinline!mlss_proc_branch_dom b! using the corresponding introduction rule for the domain predicate.
+
+If \lstinline!b! is not saturated, we can expand \lstinline!b! using a linear or a branching rule, which adds new formulae to the resulting branch(es).
+In the first case, we can thus choose a branch \lstinline!b'! with \lstinline!SOME! that linearly expands \lstinline!b!.
+Moreover, the domain predicate holds for \lstinline!b' @ b! since the assumptions of the modified induction hypothesis are fulfilled.
+Ultimately, we use the corresponding introduction rule
+\begin{lstlisting}
+  $\llbracket$ $\neg$ lin_sat b; mlss_proc_branch_dom (b' @ b) $\rrbracket$
+  $\Longrightarrow$ mlss_proc_branch_dom b
+\end{lstlisting}
+for the domain predicate to obtain \lstinline!mlss_proc_branch b!.
+
+The case of a branching rule is analogous.
+\end{proof}
+
+\noindent The above lemma allows us to use the computation induction rule of \lstinline!mlss_proc_branch! on well-formed branches, which we use to prove soundness and completeness.
+As both proofs are essentially an application of soundness, respectively completeness, of the calculus, we refer the reader to the formalisation.
+\begin{lstlisting}
+lemma mlss_proc_branch_complete:
+  fixes b :: 'a branch
+  assumes wf_branch b and $\neg$ mlss_proc_branch b
+  assumes infinite (UNIV :: 'a set)
+  shows $\exists$M. M $\models$ last b
+
+lemma mlss_proc_branch_sound:
+  assumes wf_branch b and $\forall$$\psi$ $\in$ set b. M $\models$ $\psi$
+  shows $\neg$ mlss_proc_branch b
+\end{lstlisting}
+
+To finish off the proof of total correctness, note that every singleton branch is trivially well-formed; thus, termination, completeness, and soundness easily transfer to \lstinline!mlss_proc!.
+\begin{lstlisting}
+theorem mlss_proc_complete:
+  fixes $\phi$ :: 'a pset_fm
+  assumes $\neg$ mlss_proc $\phi$ and infinite (UNIV :: 'a set)
+  shows $\exists$M. M $\models$ $\phi$
+
+theorem mlss_proc_sound:
+  assumes M $\models$ $\phi$ shows $\neg$ mlss_proc $\phi$
+\end{lstlisting}
+
+\section{Conclusion and Future Work}
+By closely following a paper by \citeauthor{new_fast_tableau}~\cite{new_fast_tableau}, we developed a formalisation of a tableau calculus for a ground fragment of set theory called \MLSS{}.
+The formalisation includes an abstract description of a decision procedure that builds on the calculus.
+
+The refinement of the abstract description to an executable specification from which code can be exported is left for future work.
+This executable specification should generate certificates that can be replayed through Isabelle's inference kernel in order to facilitate the integration of the procedure into Isabelle.
+
+%Challenges in obtaining an executable specification: membership cycle test unnecessary, fast branch membership test, proof reconstruction.
+%Additional constructs such as finiteness and pairs.
+%Optimisations for the algorithm.
+
+\subsubsection{Acknowledgements}
+The author thanks Kevin Kappelmann for his comments on a draft version of the paper.
+
+\bibliography{sources}
+
+\appendix
+\clearpage\chapter*{Appendix}
+
+\section{Subterms and Subformulas}
+
+Similarly to \lstinline!vars!, we introduce the overloaded constant \lstinline!subterms!.
+\begin{lstlisting}
+fun subterms :: 'a pset_term $\Rightarrow$ 'a pset_term set where
+  subterms $\emptyset$ = {$\emptyset$}
+| subterms (Var x) = {Var x}
+| subterms (s $\unionS$ t) = {s $\unionS$ t} $\cup$ subterms s $\cup$ subterms t
+| subterms (s $\interS$ t) = {s $\interS$ t} $\cup$ subterms s $\cup$ subterms t
+| subterms (s $\diffS$ t) = {s $\diffS$ t} $\cup$ subterms s $\cup$ subterms t
+| subterms (Single t) = {Single t} $\cup$ subterms t
+\end{lstlisting}
+The subterms of an atom are then the subterms of its arguments.
+\begin{lstlisting}
+fun subterms :: 'a pset_atom $\Rightarrow$ 'a pset_term set where
+  subterms (s $\inS$ t) = subterms s $\cup$ subterms t
+| subterms (s $\eqS$ t) = subterms s $\cup$ subterms t
+\end{lstlisting}
+Finally, we lift this to formulae by using \lstinline!atoms! again.
+\begin{lstlisting}
+definition subterms :: 'a pset_fm $\Rightarrow$ 'a pset_term set where
+  subterms $\phi$ $\equiv$ $\bigcup$(subterms ` atoms $\phi$)
+\end{lstlisting}
+
+Lastly, we consider the subformulae of a formula and define a function that computes them.
+\begin{lstlisting}
+fun subfms :: 'a fm $\Rightarrow$ 'a fm set where
+  subfms (&\fmAtom& a) = {&\fmAtom& a}
+| subfms (&\fmAnd{p}{q}&) = {&\fmAnd{p}{q}&} $\cup$ subfms p $\cup$ subfms q
+| subfms (&\fmOr{p}{q}&) = {&\fmOr{p}{q}&} $\cup$ subfms p $\cup$ subfms q
+| subfms (&\fmNeg{q}&) = {&\fmNeg{q}&} $\cup$ subfms q
+\end{lstlisting}
+
+\begin{lstlisting}
+definition vars :: 'a branch $\Rightarrow$ 'a set 
+  vars b $\equiv$ $\bigcup$(vars ` set b)
+
+definition subterms :: 'a branch $\Rightarrow$ 'a pset_term set where
+  subterms b $\equiv$ $\bigcup$(subterms ` set b)
+\end{lstlisting}
+
+\section{Definition of the Interpretation Functions}
+
+\begin{lstlisting}
+fun &\Ist& :: ('a $\Rightarrow$ hf) $\Rightarrow$ 'a pset_term $\Rightarrow$ hf where
+  &\Ist& M $\emptyset$ = 0
+| &\Ist& M (Var x) = M x
+| &\Ist& M (Single s) = HF {&\Ist& M s}
+| &\Ist& M (s $\unionS$ t) = &\Ist& M s $\sqcup$ &\Ist& M t 
+| &\Ist& M (s $\interS$ t) = &\Ist& M s $\sqcap$ &\Ist& M t 
+| &\Ist& M (s $\diffS$ t) = &\Ist& M s $-$ &\Ist& M t 
+\end{lstlisting}
+The interpretation function \lstinline!I$_\text{sa}$! for set atoms is straightforward as well.
+\begin{lstlisting}
+fun &\Isa& :: ('a $\Rightarrow$ hf) $\Rightarrow$ 'a pset_atom $\Rightarrow$ bool where
+  &\Isa& M (s $\inS$ t) $\longleftrightarrow$ &\Ist& M s $\hfmem$ &\Ist& M t
+| &\Isa& M (s $\eqS$ t) $\longleftrightarrow$ &\Ist& M s = &\Ist& M t
+\end{lstlisting}
+
+\section{Expansion Rules}
+
+\begin{table*}[h!]
   \centering
@@
 \end{table*}
-The calculus considers two kinds of branch expansion rules: \textit{linear} and \textit{branching} rules.
-As the name suggests, branching rules lead to the creation of new branches in the tableau while linear rules only extend a branch \lstinline!b!
-with new formulae \lstinline[breaklines=true]!b' = [$\psi_1$,$\ldots$,$\psi_n$]!, which we denote by \lstinline!&\lexpands{b'}{b}&!.
-\autoref{tab:lexpands} shows the linear expansion rules.
-Note that in the first two rules for \lstinline!$\eqS$!, \lstinline!l! is a literal occurring in the branch.
-Furthermore, the term-for-term substitution \lstinline!l{s/t}!
-is restricted to the top-level set terms of \lstinline!l!, i.e.\ the set terms that occur directly under one of the atom constructors \lstinline!$\inS$!
-or \lstinline!$\eqS$!;
-for example, given the literal
-\begin{lstlisting}
-  l = &\fmNeg{((s $\unionS$ u) $\diffS$ s $\eqS$ s $\unionS$ u)}&
-\end{lstlisting}
-we have
-\begin{lstlisting}
-  &\phantom{=}& (&\fmNeg{((s $\unionS$ u) $\diffS$ s $\eqS$ s $\unionS$ u)}&){t/s $\unionS$ u}
-  = &\fmNeg{((s $\unionS$ u) $\diffS$ s $\eqS$ t)}&.
-\end{lstlisting}
-The perhaps more crucial restriction of the linear rules is that that no new subterm may be created by their application;
-for instance, the second rule for \lstinline!$\unionS$!
-\begin{lstlisting}
-  s $\inS$ t$_\text{1}$ $\Longrightarrow$ s $\inS$ t$_\text{1}$ $\unionS$ t$_\text{2}$,
-\end{lstlisting}
-which formally represents 
-\begin{lstlisting}
-  (s $\inS$ t$_\text{1}$) $\in$ set b $\Longrightarrow$ &\lexpands{[s $\inS$ t$_\text{1}$ $\unionS$ t$_\text{2}$]}{b}&,
-\end{lstlisting}
-may only be used under the condition
-\begin{lstlisting}
-  t$_\text{1}$ $\unionS$ t$_\text{2}$ $\in$ subterms (last b).
-\end{lstlisting}
-The purpose of this restriction is to prevent unbounded expansion of the branch.
-In fact, we give an explicit upper bound for the number of formulae in a branch in \autoref{sec:bound}.
-
-Due to boundedness, repeated expansion with linear rules eventually results in a \textit{linearly saturated} branch, i.e.\ a branch where no application of linear rules would produce new formulae.
-\begin{lstlisting}
-definition lin_sat b $\equiv$ $\forall$b'. &\lexpands{b'}{b}& $\longrightarrow$ set b' $\subseteq$ set b
-\end{lstlisting}
-
-Finally, we remark that the original paper~\cite{new_fast_tableau} is missing the last propositional rule dealing with double negation.
-This rule is required for completeness, though, considering that the branch
-\begin{lstlisting}
-  [&\fmNeg{\fmNeg{\fmNeg{p}}}&, p, &\fmAnd{\fmNeg{\fmNeg{\fmNeg{p}}}}{p}&]
-\end{lstlisting}
-is saturated --- neither linear nor branching rules apply --- and open, but there clearly is no model for the initial formula \lstinline!&\fmAnd{\fmNeg{\fmNeg{\fmNeg{p}}}}{p}&!.
-
-\subsection{Branching Rules\label{sec:branching}}
+
 \begin{table*}
@@
 \end{table*}
-After running out of linear rules to apply, only the branching rules shown in \autoref{tab:bexpands} remain.
-A rule is applicable if its \textit{precondition} is met and, to prevent unnecessary branching, it is not subsumed as indicated by the \textit{subsumption condition}.
-These rules create multiple branches in the tableau, so we represent the different possibilities \lstinline!bs'! to expand a branch \lstinline!b! as a set and write \lstinline!&\fexpands{bs'}{b}&!. 
-Accordingly, we get a new branch \lstinline!b' @ b! in the tableau for each \lstinline!b' $\in$ set bs'!.
-
-A linearly saturated branch where no further branching is possible is called a \textit{saturated} branch. 
-\begin{lstlisting}
-definition sat b $\equiv$ lin_sat b $\land$ ($\nexists$bs'. &\fexpands{bs'}{b}&)
-\end{lstlisting}
-
-Note that even branching rules are defined such that they never create new subterms with exception of the last rule, which adds a new variable to the branch.
-These variables serve to manifest an inequality; hence, we call them \textit{witnesses}.
-\begin{lstlisting}
-definition wits :: 'a branch $\Rightarrow$ 'a set where
-  wits b $\equiv$ vars b - vars (last b)
-\end{lstlisting}
-
-\section{A Decision Procedure for MLSS}
-\begin{lstlisting}
-function (domintros) mlss_proc_branch :: 'a branch $\Rightarrow$ bool where
-  $\neg$ lin_sat b $\Longrightarrow$ mlss_proc_branch b =
-  mlss_proc_branch ((SOME b'. &\lexpands{b'}{b}& $\land$
-                              set b $\subset$ set (b' @ b)) @ b)
-| $\llbracket$ lin_sat b; bclosed b $\rrbracket$ $\Longrightarrow$ mlss_proc_branch b = True
-| $\llbracket$ $\neg$ sat b; bopen b; lin_sat b $\rrbracket$ $\Longrightarrow$ mlss_proc_branch b =
-  ($\forall$b' $\in$ (SOME bs. &\fexpands{bs}{b}&). mlss_proc_branch (b' @ b))
-| $\llbracket$ lin_sat b; sat b $\rrbracket$ $\Longrightarrow$ mlss_proc_branch b = bclosed b
-
-definition mlss_proc :: 'a pset_fm $\Rightarrow$ bool where
-  mlss_proc $\phi$ $\equiv$ mlss_proc_branch [$\phi$]
-\end{lstlisting}
-
-\noindent The mechanics of the decision procedure are typical for a procedure based on a tableau calculus:
-it decides the satisfiability of a given formula \lstinline!$\phi$! by determining whether the formula has a closed tableau.
-More specifically, it initialises the tableau with the singleton branch \lstinline![$\phi$]! and checks whether this branch can be expanded to a closed tableau.
-
-We abstractly specify the behaviour of the decision procedure, leaving an executable specification from which code can be generated to future work.
-The implementation as displayed in TODO uses a couple of features of Isabelle/HOL's function package:
-instead of defining the function via pattern matching, we specify the equations of the function as conditional rewrite rules.
-This requires us to prove that the assumptions of the equations are non-overlapping, which is done by automation.
-The other concern is that Isabelle/HOL requires functions to be total, so a recursive function needs to terminate in order for it to be well-defined;
-nevertheless, the termination proof is separated from the definition of the function for modularity.
-The function package maintains the soundness of the definition by introducing a so-called domain predicate \lstinline!mlss_proc_branch_dom!
-which characterises for which arguments the function terminates.
-Each equation of the function is guarded by an assumption that the predicate holds for the argument.
-To be able to reason about this predicate, we pass the \lstinline!domintros! flag that instructs the function package to derive introduction rules for it.
-In \autoref{sec:correct}, we will use these rules to discharge the guarding assumptions for the context that \lstinline!mlss_proc_branch! is called in.
-Before we go into more detail on how the termination is proved, we discuss its definition.
-
-The purpose of the function is to determine whether a given branch can be expanded to a closed tableau.
-As stated before, we first use linear expansion rules in order to prevent premature branching;
-to this end, we recursively expand the branch with linear rules until the branch is linearly saturated.
-Note that we use Hilbert's $\varepsilon$-operator in the form of \lstinline!SOME! to choose some rule that actually adds new formulae to the branch. 
-As soon as the branch is linearly saturated, we terminate if the branch is closed as shown in the second equation.
-Otherwise, we choose an applicable branching rule and recursively check whether all newly created branches can be closed.
-The final equation applies once no further branch expansion is possible, in which case we just test for closedness of the branch.
-
-The procedure \lstinline!mlss_proc! then calls \lstinline!mlss_proc_branch! with a singleton branch \lstinline![$\phi$]! to determine the satisfiability of a given formula \lstinline!$\phi$!.
-
-This means that \lstinline!mlss_proc_branch! is only applied to branches that result from applying the expansion rules.
-We call this kind of branches \textit{well-formed}.
-In the definition below, the expression \lstinline!&\expandss{b'}{b}&!
-denotes that \lstinline!b'!
-is one of the branches that results from applying (potentially zero) expansion rules to \lstinline!b!.
-\begin{lstlisting}
-definition wf_branch b $\equiv$ $\exists$$\phi$. &\expandss{b}{[$\phi$]}&
-\end{lstlisting}
-
-In particular, we will use this notion in \autoref{sec:bound} where we derive an upper bound for the cardinality of well-formed branches.
-This is then used in \autoref{sec:correct} to justify the termination of the decision procedure. 
-Before we come to that, though, we prove soundness and completeness in Section~\ref{sec:sound} and~\ref{sec:complete}, respectively.
-In \autoref{sec:correct}, we also show that both properties easily transfer to \lstinline!mlss_proc! which, together with termination, establishes that it is a decision procedure.
-
-\section{Completeness of the Calculus\label{sec:complete}}
-For completeness of the calculus, we need to show that every unsatisfiable formula has a closed tableau or, conversely, that the formula is satisfiable if there is a saturated and open branch in the tableau.
-To facilitate inductive reasoning, we show a stronger statement by constructing a model \lstinline!M! such that \lstinline!M $\models$ $\phi$! for all \lstinline!$\phi$ $\in$ set b!.
-At the core of the model, there is a \textit{realisation} function that maps set terms to sets of type \lstinline!hf!.
-A subset of the witnesses, which we call \textit{pure} witnesses, receives special treatment from the realisation function for reasons that will become apparent in \autoref{sec:pwits}.
-The collection of set terms of a branch can thus be partitioned into two collections as defined below.
-\begin{lstlisting}
-definition pwits :: 'a branch $\Rightarrow$ 'a set where
-  pwits b $\equiv$ {c $\in$ wits b. $\forall$t $\in$ subterms (last b).
-                          AT (Var c $\eqS$ t) $\notin$ set b $\land$
-                          AT (t $\eqS$ Var c) $\notin$ set b}
-
-definition subterms' :: 'a branch $\Rightarrow$ 'a pset_term set where
-  subterms' b $\equiv$ subterms (last b) $\cup$ Var ` (wits b - pwits b)
-\end{lstlisting}
-
-We aim to construct a syntactic model that is derived from the membership literals \lstinline!s $\inS$ t! in the branch.
-To this end, we construct a graph whose vertices are the disjoint union of the sets above and there is an edge from \lstinline!s!
-to \lstinline!t!
-in the graph if, and only if, \lstinline!s $\inS$ t!
-is in \lstinline!b!.
-Note that we use \citeauthor{graph_theory_afp}'s graph library~\cite{graph_theory_afp} that represents a graph as a record of vertices, arcs (directed edges), and two functions \lstinline!tail!
-and \lstinline!head!
-that map an arc to its source and target vertex, respectively.
-\begin{lstlisting}
-definition bgraph b $\equiv$
-  let vs = Var ` pwits b $\cup$ subterms' b
-  in $\llparenthesis$ verts = vs, arcs = {(s, t). (s $\inS$ t) $\in$ set b},
-     &$\phantom{\llparenthesis}$& tail = fst, head = snd $\rrparenthesis$
-\end{lstlisting}
-
-The realisation function is defined relative to this graph.
-As mentioned before, the pure witnesses \lstinline!Var ` pwits b! and the rest of the set terms \lstinline!subterms' b! are treated differently by the realisation function.
-Terms in the latter set are evaluated in accordance to the structure of the graph, i.e.\ the realisation of a vertex is defined as the union of the realisations of the parent vertices. 
-For the former set, we choose a function \lstinline!I! that assigns the witnesses in \lstinline!Var ` pwits b! pairwise distinct sets with cardinality greater than that of the vertices.
-We can always choose such a function since we assume an infinite universe of variables.
-Then, we return the singleton set \lstinline!HF {I x}!, which, together with the cardinality constraint, guarantees that realisations are distinct between pure witnesses themselves as well as between pure witnesses and set terms.
-The notation \lstinline!u $\rightarrow_\text{G}$ s! in the definition below indicates that there is an edge from \lstinline!u! to \lstinline!s! in the graph \lstinline!G!.
-\begin{lstlisting}
-abbreviation parents G s $\equiv$ {u. u $\rightarrow_\text{G}$ s}
-
-function realise :: 'a pset_term $\Rightarrow$ V where
-  x $\in$ Var ` pwits b $\Longrightarrow$ realise x = HF {I x}
-| x $\in$ subterms' b
-  $\Longrightarrow$ realise t = HF {realise ` parents (bgraph b) s}
-| x $\notin$ verts G $\Longrightarrow$ realise x = 0
-\end{lstlisting}
-
-Again, we need to ensure that the assumptions of the equations are non-overlapping and that the function terminates.
-The former is taken care of by automation, leaving us to prove termination.
-The assumption that \lstinline!b! is open implies that there are no membership cycles and thus \lstinline!bgraph b! is acyclic.
-TODO: Shorten this proof.
-Now, consider the ancestors \lstinline!av! of a vertex \lstinline!v!, which are those vertices from which we can reach \lstinline!v!.
-Due to acyclicity, it holds that the ancestors \lstinline!ap! of a parent \lstinline!p! of \lstinline!v! are a subset of \lstinline!av!.
-They are also a proper subset of \lstinline!av! since \lstinline!p! is an ancestor of \lstinline!v! but not of \lstinline!p! itself.
-Together with the fact that \lstinline!bgraph b! is finite, it follows that the cardinality of the ancestors decreases in each recursive call, thus proving the termination of \lstinline!realise!.
-
-Before we prove that the realisation function constitutes a model in \autoref{sec:real}, we will first explain the significance of the pure witnesses.
-
-\subsection{Characterisation of the Pure Witnesses\label{sec:pwits}}
-Recall that the pure witnesses of a branch \lstinline!b! are those witnesses that are not related to other subterms in \lstinline!last b! by equality.
-In the context of a well-formed branch, this characterisation can be strengthened to any set term and, in addition, we also get that there is no membership literal where a pure witness is on the right-hand side.
-Intuitively speaking, the realisation of a pure witness does not depend on the realisation of any other set term.
-\begin{lstlisting}[label={lst:lemma_2}]
-lemma lemma_2:
-  assumes wf_branch b and c $\in$ pwits b
-  shows (Var c $\eqS$ t) $\notin$ set b and (t $\eqS$ Var c) $\notin$ set b
-    and (t $\inS$ Var c) $\notin$ set b
-\end{lstlisting}
-So why are pure witnesses treated differently?
-According to the definition of \lstinline!realise!, the pure witnesses would be evaluated to the empty set \lstinline!0 :: hf!, were they not treated separately. 
-To see that this is a problem, consider the branch
-\begin{lstlisting}
-  b = [Var s $\neqS$ Var t, Var t $\neqS$ Var u]
-\end{lstlisting}
-which expands to several open and saturated branches, one of which is
-\begin{lstlisting}
-  [Var x $\neqS$ Var y,
-   Var x $\inS$ Var s, Var x $\notinS$ Var t,
-   Var y $\inS$ Var t, Var y $\notinS$ Var u] @ b
-\end{lstlisting}
-for some fresh \lstinline!x! and \lstinline!y!.
-Assigning both \lstinline!Var x! and \lstinline!Var y! a value of \lstinline!0! would contradict the literal \lstinline!Var x $\neqS$ Var y!.
-To prevent this, we assign the pure witnesses pairwise different values.
-
-The proof of \hyperref[lst:lemma_2]{\lstinline!lemma_2!} is more technical than interesting so we refer the reader to the formalisation or paper TODO.
-
-\subsection{Realisation of an Open Branch\label{sec:real}}
-Remember that for completeness, we need to show that the realisation function for an open and saturated branch \lstinline!b! actually constitutes a model for all formulae in the branch.
-We start by verifying that the realisation function models all literals in the branch; more formally, the following propositions hold:
-\begin{enumerate}[label={(\roman*)}]
-  \item\label{it:at_mem} We have \lstinline!realise s $\hfmem$ realise t! if it holds that \lstinline!s $\inS$ t! is in \lstinline!b!. 
-  \item\label{it:at_eq} We have \lstinline!realise s $=$ realise t! if \lstinline!s $\eqS$ t! is in \lstinline!b!.
-  \item\label{it:af_eq} We have \lstinline!realise s $\neq$ realise t! if \lstinline!s $\neqS$ t! is in \lstinline!b!.
-  \item\label{it:af_mem} We have \lstinline!realise s $\hfnotmem$ realise t! if it holds that \lstinline!s $\notinS$ t! is in \lstinline!b!. 
-\end{enumerate}
-To illustrate the usefulness of \hyperref[lst:lemma_2]{\lstinline!lemma_2!}, we prove Proposition~\ref{it:at_eq}.
-The proofs of all propositions translate well into Isabelle, so we refer to the original paper~\cite{new_fast_tableau} for the remaining proofs. 
-\begin{proof}[Proof of Proposition~\ref{it:at_eq}]
-  Assume that \lstinline!s $\eqS$ t! is in \lstinline!b!.
-  If there exists a \lstinline!c $\in$ pwits b! where \lstinline!s = Var c! or \lstinline!t = Var c!, we arrive at a contradiction due to \hyperref[lst:lemma_2]{\lstinline!lemma_2!}.
-  Therefore, both \lstinline!s $\in$ subterms' b! and \lstinline!t $\in$ subterms' b! must hold.
-  Now, assume for contradiction that
-\begin{lstlisting}
-  realise s $\neq$ realise t.
-\end{lstlisting}
-  Without loss of generality --- the other case is symmetric --- we obtain an \lstinline!e! such that
-\begin{lstlisting}
-  e $\hfmem$ realise s $\land$ e $\hfnotmem$ realise t. 
-\end{lstlisting}
-  Considering the fact that \lstinline!s $\in$ subterms' b!
-and the definition of \lstinline!realise!, we obtain an \lstinline!u!
-where \lstinline!e = realise u!
-and \lstinline!u $\rightarrow_\text{bgraph b}$ s!.
-  This, in turn, yields that \lstinline!u $\inS$ s! must be in \lstinline!b!.
-  Together with the assumption \lstinline!(s $\eqS$ t) $\in$ set b! and the saturation of \lstinline!b!, it follows that \lstinline!u $\inS$ t! must also be in \lstinline!b!.
-  But then we have
-\begin{lstlisting}[columns=fixed]
-  realise u $\hfmem$ realise t = e $\hfmem$ realise t
-\end{lstlisting}
-  using Proposition~\ref{it:at_mem}, which stands in contradiction to the assumption \lstinline!e $\hfnotmem$ realise t!.
-\end{proof}
-
-\noindent The results on literals can now be lowered to set terms.
-All of the proofs are straightforward so we refer the reader to the formalisation.
-\begin{enumerate}[label=(\alph*)]
-  \item\label{it:empty} It holds that \lstinline!realise $\emptyset$ = 0!.
-  \item\label{it:op} Let \lstinline!$\star_\text{s}$ $\in$ {$\unionS$, $\diffS$, $\interS$}!. If the term \lstinline!s $\star_\text{s}$ t! occurs in \lstinline!subterms b!, then
-    \begin{center}
-    \lstinline!realise (s $\star_\text{s}$ t) = realise s $\star$ realise t!.
-    \end{center}
-  \item\label{it:single} If \lstinline!Single t $\in$ subterms b!, then
-    \begin{center}
-    \lstinline!realise (Single t) = HF {realise t}!.
-    \end{center}
-\end{enumerate}
-
-The final step to obtain a proper model is to connect the realisation function to the semantics as defined in \autoref{sec:semantics}.
-For set terms, we can use the Propositions~\ref{it:empty}--\ref{it:single} to prove the lemma below by induction on \lstinline!t!.
-\begin{lstlisting}
-lemma
-  assumes t $\in$ subterms b
-  shows &\Ist& ($\lambda$x. realise (Var x)) t = realise t
-\end{lstlisting}
-Lifting the above result to formulae yields the coherence of \lstinline!b! as the original paper~\cite{new_fast_tableau} calls it.
-The proof is a tedious but straightforward induction on the the size of the formulae.
-\begin{lstlisting}[label={lst:coherence}]
-lemma coherence:
-  assumes $\phi$ $\in$ set b shows ($\lambda$x. realise (Var x)) $\models$ $\phi$
-\end{lstlisting}
-The coherence property finishes the proof of completeness of the calculus as it gives us a model for every formula in an open and saturated branch.
-
-\section{Soundness of the Calculus\label{sec:sound}}
-A tableau calculus is sound if for any closed tableau, the corresponding formula is unsatisfiable. 
-We prove the following two properties to establish soundness.
-\begin{enumerate}[label={(\roman*)}]
-  \item A closed branch contains an unsatisfiable formula.
-  \item The expansion rules maintain satisfiability.
-\end{enumerate}
-
-We formalise the first property in Isabelle as follows.
-\begin{lstlisting}[belowskip=0pt,label={lst:bclosed_sound}]
-lemma bclosed_sound:
-  assumes bclosed b shows $\exists$$\phi$ $\in$ set b. M $\not\models$ $\phi$
-\end{lstlisting}
-\begin{proof}
-  Assume for contradiction that \lstinline!b! is closed but we have a model, i.e.\ it holds that \lstinline!$\forall$$\phi$ $\in$ set b. M $\models$ $\phi$!.
-  We perform a case analysis on \lstinline!bclosed b!.
-  It is clear that, for any \lstinline!s!, neither does \lstinline!M! model \lstinline!s $\in$ $\emptyset$! nor \lstinline!s $\neqS$ s!.
-  Furthermore, no model can satisfy both \lstinline!$\phi$! and \lstinline!&\fmNeg{$\phi$}&! at the same time.
-  This leaves us with the case where \lstinline!b! contains a membership cycle
-\begin{lstlisting}
-  t$_\text{0}$ $\inS$ t$_\text{1}$, t$_\text{1}$ $\inS$ t$_\text{2}$, $\ldots$, t$_\text{k}$ $\inS$ t$_\text{0}$.
-\end{lstlisting}
-  By assumption, each atom of the membership cycle holds under the model \lstinline!M!; in other words, we have
-\begin{lstlisting}
-  &\Ist& M t$_\text{i}$ $\hfmem$ &\Ist& M t$_\text{i + 1 mod k + 1}$
-\end{lstlisting}
-  for each \lstinline!i $\in$ {0,$\,\ldots\,$,k}!.
-  Therefore, the transitive closure of the membership relation
-\begin{lstlisting}
-  {(x, y). x $\hfmem$ y}$^\text{+}$ 
-\end{lstlisting}
-  contains a cycle.
-  This, however, is a contradiction because the membership relation of \lstinline!hf! is well-founded (see \autoref{sec:semantics} TODO) and well-foundedness is closed under transitive closure.
-\end{proof}
-
-\noindent We are left with showing that both linear and branching expansion rules preserve satisfiability.
-As for the linear rules, a straightforward proof by case analysis on \lstinline!&\lexpands{b'}{b}&! suffices to obtain the lemma below.
-\begin{lstlisting}[label={lst:lexpands_sound}]
-lemma lexpands_sound:
-  assumes &\lexpands{b'}{b}& and $\phi$ $\in$ set b' and $\bigwedge$$\psi$. $\psi$ $\in$ set b $\Longrightarrow$ M $\models$ $\psi$
-  shows M $\models$ $\phi$
-\end{lstlisting}
-A similar argument would work for the branching rules if it were not for the last rule adding new variables.
-Those variables need to be assigned specific values; hence, we modify the model as shown in the proof below.
-\begin{lstlisting}[belowskip=0pt, label={lst:bexpands_sound}]
-lemma bexpands_sound:
-  assumes &\fexpands{bs'}{b}& and $\bigwedge$$\psi$. $\psi$ $\in$ set b $\Longrightarrow$ M $\models$ $\psi$
-  shows $\exists$M'. $\exists$b' $\in$ bs'. $\forall$$\psi$ $\in$ set (b' @ b). M' $\models$ $\psi$
-\end{lstlisting}
-\begin{proof}
-  We perform a case analysis on the derivation \lstinline!&\fexpands{bs'}{b}&!.
-  As all cases but that of the last rule are straightforward, we only consider this case.
-  We have 
-\begin{lstlisting}
-  bs' = {[Var x $\inS$ s, Var x $\notinS$ t], [Var x $\inS$ t, Var x $\notinS$ s]}
-\end{lstlisting}
-  for some fresh variable \lstinline!x!, i.e.\ \lstinline!x $\notin$ vars b! and in particular \lstinline!x $\notin$ vars s $\cup$ vars t!.
-  Another consequence of the last rule being applied is that \lstinline!s $\neqS$ t! is in \lstinline!b!, which implies that \lstinline!&\Ist& M s $\neq$ &\Ist& M t! because \lstinline!M! is a model.
-  This inequality manifests itself through some \lstinline!y! with
-\begin{lstlisting}
-  &\phantom{$\lor$}& y $\hfmem$ &\Ist& M s $\land$ y $\hfnotmem$ &\Ist& M t
-  $\lor$ y $\hfmem$ &\Ist& M t $\land$ y $\hfnotmem$ &\Ist& M s.
-\end{lstlisting}
-  We now update the assignment \lstinline!M! such that it maps \lstinline!x! to \lstinline!y! to obtain the assignment \lstinline!M'!.
-  Note that values of both \lstinline!s! and \lstinline!t! are unchanged under this update as \lstinline!x $\notin$ vars s! and \lstinline!x $\notin$ vars t!.
-  Considering our assumption about \lstinline!y! and the fact that \lstinline!&\Ist& M' (Var x) = y!, we infer that
-\begin{lstlisting}
-  $\exists$b' $\in$ bs'. $\forall$$\psi$ $\in$ set b'. M' $\models$ $\psi$.
-\end{lstlisting}
-  In other words, due to the disjunction, it holds that \lstinline!M'! is a model for one \lstinline!b' $\in$ bs!.
-  Thus, we are left with proving that \lstinline!M'! is a model for \lstinline!b!.
-  This is indeed the case since the assumption \lstinline!M $\models$ $\psi$!
-is equivalent to \lstinline!M' $\models$ $\psi$!
-for all \lstinline!$\psi$ $\in$ set b!
-given that \lstinline!x!
-is fresh with respect to b.
-\end{proof}
-
-\section{Bounding the Cardinality of a Branch\label{sec:bound}}
-All in all, we get the following lemma.
-\begin{lstlisting}[label={lst:card_branch}]
-lemma card_wf_branch_ub:
-  assumes wf_branch b
-  shows |set b|
-      $\leq$ 2 * |subfms (last b)| + 16 * |subterms (last b)|$^\text{4}$
-\end{lstlisting}
-
-\section{Total Correctness of the Decision Procedure\label{sec:correct}}
-We first prove the termination of the procedure for well-formed branches, i.e.\ every well-formed branch is in the domain of \lstinline!mlss_proc_branch!.
-\begin{lstlisting}[belowskip=0pt]
-lemma mlss_proc_branch_dom_if_wf_branch:
-  assumes wf_branch b shows mlss_proc_branch_dom b
-\end{lstlisting}
-\begin{proof}
-  We proceed by strong induction on the difference between the upper bound due to \hyperref[lst:card_branch]{\lstinline!card_wf_branch_ub!} and \lstinline!|b|!.
-
-  We only have one inductive case where the induction hypothesis holds for all well-formed branches \lstinline!b'!
-where the difference between the upper bound and the cardinality is smaller than that of \lstinline!b!.
-  We strengthen the assumptions of the induction hypothesis to obtain
-\begin{lstlisting}
-  $\llbracket$ &\expandss{b'}{b}&; set b $\subset$ set b' $\rrbracket$ $\Longrightarrow$ mlss_proc_branch_dom b'
-\end{lstlisting}
-for all \lstinline!b'!.
-The former strengthening is admissible since, by definition, every expansion of a well-formed branch is again well-formed while the latter ensures that the cardinality of \lstinline!b'!
-is larger than the cardinality of \lstinline!b!, implying that the difference to the upper bound decreases.
-
-Now, consider the case where \lstinline!b! is saturated.
-Then, the procedure terminates without recursive calls and we immediately get \lstinline!mlss_proc_branch_dom b! using the corresponding introduction rule for the domain predicate.
-
-If \lstinline!b! is not saturated, we can expand \lstinline!b! using a linear or a branching rule, which adds new formulae to the resulting branch(es).
-In the first case, we can thus choose a branch \lstinline!b'! with \lstinline!SOME! that linearly expands \lstinline!b!.
-Moreover, the domain predicate holds for \lstinline!b' @ b! since the assumptions of the modified induction hypothesis are fulfilled.
-Ultimately, we use the corresponding introduction rule
-\begin{lstlisting}
-  $\llbracket$ $\neg$ lin_sat b; mlss_proc_branch_dom (b' @ b) $\rrbracket$
-  $\Longrightarrow$ mlss_proc_branch_dom b
-\end{lstlisting}
-for the domain predicate to obtain \lstinline!mlss_proc_branch b!.
-
-The case of a branching rule is analogous.
-\end{proof}
-
-\noindent The above lemma allows us to use the computation induction rule of \lstinline!mlss_proc_branch!
-on well-formed branches, which we do to prove soundness and completeness, starting with completeness.
-\begin{lstlisting}[belowskip=0pt]
-lemma mlss_proc_branch_complete:
-  fixes b :: 'a branch
-  assumes wf_branch b and $\neg$ mlss_proc_branch b
-  assumes infinite (UNIV :: 'a set)
-  shows $\exists$M. M $\models$ last b
-\end{lstlisting}
-\begin{proof}
-  We proceed to prove the completeness by computation induction on \lstinline!b!.
-
-  The cases of the first and third equation, where the branch can be expanded, are immediately resolved with the induction hypothesis.
-
-  We evaluate \lstinline!mlss_proc_branch b! to \lstinline!True! in the case of the third equation, yielding a contradiction with the assumption \lstinline!$\neg$ mlss_proc_branch b!.
-
-  This leaves us with the final case where we assume that the branch is saturated and \lstinline!mlss_proc_branch b! evaluates to \lstinline!bclosed b!.
-  The branch, however, must be open due to the assumption \lstinline!$\neg$ mlss_proc_branch b!.
-  This allows us to use the Lemma~\hyperref[lst:coherence]{\lstinline!coherence!} to obtain the model
-\begin{lstlisting}
-  ($\lambda$x. realise (Var x)) $\models$ last b.
-\end{lstlisting}
-\end{proof}
-
-\noindent With completeness out of the way, only soundness remains to be proven. 
-\begin{lstlisting}[belowskip=0pt]
-lemma mlss_proc_branch_sound:
-  assumes wf_branch b and $\forall$$\psi$ $\in$ set b. M $\models$ $\psi$
-  shows $\neg$ mlss_proc_branch b
-\end{lstlisting}
-\begin{proof}
-  We proceed by assuming \lstinline!mlss_proc_branch b! and deriving \lstinline!False!.
-  Using the assumptions, we can prove that there exists \lstinline!b'! such that
-\begin{lstlisting}
-  &\expandss{b'}{b}& $\land$ ($\exists$M. $\forall$$\psi$ $\in$ set b'. M $\models$ $\psi$) $\land$ bclosed b'
-\end{lstlisting}
-  by computation induction on \lstinline!b!.
-
-  In the case where we choose a branching rule to expand \lstinline!b!
-with the branches in \lstinline!bs'!, we use Lemma~\hyperref[lst:lexpands_sound]{\lstinline!bexpands_sound!} to obtain \lstinline!b' $\in$ bs'!
-and \lstinline!M'!
-with
-\begin{lstlisting}
-  $\forall$$\psi$ $\in$ set (b' @ b). M' $\models$ $\psi$.
-\end{lstlisting}
-  Together with the induction hypothesis, we obtain a \lstinline!b''! where 
-\begin{lstlisting}
-  &\expandss{b''}{(b' @ b)}& $\land$
-  ($\exists$M. $\forall$$\psi$ $\in$ set b''. M $\models$ $\psi$) $\land$ bclosed b''.
-\end{lstlisting}
-  Due to the transitivity of \lstinline!$\expandssSym$!, the above also holds if we replace \lstinline!b' @ b! with \lstinline!b!, which is our goal.
-
-  The case of a linear expansion is dealt with in an analogous manner using Lemma~\hyperref[lst:lexpands_sound]{\lstinline!lexpands_sound!}.
-
-  The two remaining cases are trivial as the branch is already closed, finishing the proof of the proposition.
-
-  This proposition, however, is a contradiction to the statement of Lemma~\hyperref[lst:bclosed_sound]{\lstinline!bclosed_sound!}: no closed branch can have a model. 
-\end{proof}
-
-\noindent To finish off the proof of total correctness, note that every singleton branch is trivially well-formed; thus, termination, completeness, and soundness easily transfer to \lstinline!mlss_proc!.
-\begin{lstlisting}
-theorem mlss_proc_complete:
-  fixes $\phi$ :: 'a pset_fm
-  assumes $\neg$ mlss_proc $\phi$ and infinite (UNIV :: 'a set)
-  shows $\exists$M. M $\models$ $\phi$
-
-theorem mlss_proc_sound:
-  assumes M $\models$ $\phi$ shows $\neg$ mlss_proc $\phi$
-\end{lstlisting}
-
-\section{Conclusion and Future Work}
-By closely following a paper by \citeauthor{new_fast_tableau}~\cite{new_fast_tableau}, we developed a formalisation of a tableau calculus for a ground fragment of set theory called \MLSS{}.
-The formalisation includes an abstract description of a decision procedure that builds on the calculus.
-
-The refinement of the abstract description to an executable specification from which code can be exported is left for future work.
-This executable specification should generate certificates that can be replayed through Isabelle's inference kernel in order to facilitate the integration of the procedure into Isabelle.
-
-%Challenges in obtaining an executable specification: membership cycle test unnecessary, fast branch membership test, proof reconstruction.
-%Additional constructs such as finiteness and pairs.
-%Optimisations for the algorithm.
-
-\subsubsection{Acknowledgements}
-The author thanks Kevin Kappelmann for his comments on a draft version of the paper.
-
-\bibliography{sources}
-
-\appendix
-\clearpage\chapter*{Appendix}
-
-\section{Subterms and Subformulas}
-
-Similarly to \lstinline!vars!, we introduce the overloaded constant \lstinline!subterms!.
-\begin{lstlisting}
-fun subterms :: 'a pset_term $\Rightarrow$ 'a pset_term set where
-  subterms $\emptyset$ = {$\emptyset$}
-| subterms (Var x) = {Var x}
-| subterms (s $\unionS$ t) = {s $\unionS$ t} $\cup$ subterms s $\cup$ subterms t
-| subterms (s $\interS$ t) = {s $\interS$ t} $\cup$ subterms s $\cup$ subterms t
-| subterms (s $\diffS$ t) = {s $\diffS$ t} $\cup$ subterms s $\cup$ subterms t
-| subterms (Single t) = {Single t} $\cup$ subterms t
-\end{lstlisting}
-The subterms of an atom are then the subterms of its arguments.
-\begin{lstlisting}
-fun subterms :: 'a pset_atom $\Rightarrow$ 'a pset_term set where
-  subterms (s $\inS$ t) = subterms s $\cup$ subterms t
-| subterms (s $\eqS$ t) = subterms s $\cup$ subterms t
-\end{lstlisting}
-Finally, we lift this to formulae by using \lstinline!atoms! again.
-\begin{lstlisting}
-definition subterms :: 'a pset_fm $\Rightarrow$ 'a pset_term set where
-  subterms $\phi$ $\equiv$ $\bigcup$(subterms ` atoms $\phi$)
-\end{lstlisting}
-
-Lastly, we consider the subformulae of a formula and define a function that computes them.
-\begin{lstlisting}
-fun subfms :: 'a fm $\Rightarrow$ 'a fm set where
-  subfms (&\fmAtom& a) = {&\fmAtom& a}
-| subfms (&\fmAnd{p}{q}&) = {&\fmAnd{p}{q}&} $\cup$ subfms p $\cup$ subfms q
-| subfms (&\fmOr{p}{q}&) = {&\fmOr{p}{q}&} $\cup$ subfms p $\cup$ subfms q
-| subfms (&\fmNeg{q}&) = {&\fmNeg{q}&} $\cup$ subfms q
-\end{lstlisting}
-
-\section{Definition of the Interpretation Functions}
-
-\begin{lstlisting}
-fun &\Ist& :: ('a $\Rightarrow$ hf) $\Rightarrow$ 'a pset_term $\Rightarrow$ hf where
-  &\Ist& M $\emptyset$ = 0
-| &\Ist& M (Var x) = M x
-| &\Ist& M (Single s) = HF {&\Ist& M s}
-| &\Ist& M (s $\unionS$ t) = &\Ist& M s $\sqcup$ &\Ist& M t 
-| &\Ist& M (s $\interS$ t) = &\Ist& M s $\sqcap$ &\Ist& M t 
-| &\Ist& M (s $\diffS$ t) = &\Ist& M s $-$ &\Ist& M t 
-\end{lstlisting}
-The interpretation function \lstinline!I$_\text{sa}$! for set atoms is straightforward as well.
-\begin{lstlisting}
-fun &\Isa& :: ('a $\Rightarrow$ hf) $\Rightarrow$ 'a pset_atom $\Rightarrow$ bool where
-  &\Isa& M (s $\inS$ t) $\longleftrightarrow$ &\Ist& M s $\hfmem$ &\Ist& M t
-| &\Isa& M (s $\eqS$ t) $\longleftrightarrow$ &\Ist& M s = &\Ist& M t
-\end{lstlisting}
 

### 2023-02-17 paper/CADE/paper.tex
Continue with paper

@@
 
-\newcommand{\fmAnd}[2]{#1 $\boldsymbol{\land}$ #2}
+\newcommand{\fmAndSymbol}{\boldsymbol{\land}}
+\newcommand{\fmAnd}[2]{#1 $\fmAndSymbol$ #2}
 \newcommand{\fmOr}[2]{#1 $\boldsymbol{\lor}$ #2}
@@
   We also provide an abstract specification of a decision procedure that applies the rules of the calculus exhaustively and prove its termination.
-  \keywords{First TODO \and Second TODO}
+  \keywords{First TODO \and Second TODO\todo{keywords}}
 \end{abstract}
@@
 \listoftodos{}
+
+\clearpage
 
@@
   \item We derive an explicit upper bound for the number of formulae in a branch of the tableau.
-  \item We restrict one branching rule TODO.
+  \item We restrict one branching rule\todo{Elaborate}.
 \end{itemize}
@@
 The termination proof is part of ongoing work~\cite{hybrid_logic_workshop}.
-TODO: Mention SteP, Check reviewer comment.
+\todo[inline]{Mention SteP, Check reviewr comment}
 % TODO: Formalisation de la logique temporelle dans le coq.
@@
 that computes them.
-The functions \lstinline!subterms! and \lstinline!subfms! are defined in the usual way (cf. Appendix TODO).
+The functions \lstinline!subterms! and \lstinline!subfms! are defined in the usual way (cf. Appendix \todo{reference Appendix}).
 
@@
 We abstractly specify the behaviour of the decision procedure, leaving an executable specification from which code can be generated to future work.
-The implementation as displayed in TODO uses a couple of features of Isabelle/HOL's function package:
+The implementation as displayed in \todo{How to reference algorithm?} uses a couple of features of Isabelle/HOL's function package:
 instead of defining the function via pattern matching, we specify the equations of the function as conditional rewrite rules.
@@
 
-In particular, we will use this notion in \autoref{sec:bound} where we derive an upper bound for the cardinality of well-formed branches.
-This is then used in \autoref{sec:correct} to justify the termination of the decision procedure. 
+In particular, we will use this notion in \autoref{sec:correct} to state an upper bound for the cardinality of well-formed branches, with which we justify the termination of the decision procedure. 
 Before we come to that, though, we prove soundness and completeness in Section~\ref{sec:sound} and~\ref{sec:complete}, respectively.
@@
 
-The proof of \hyperref[lst:lemma_2]{\lstinline!lemma_2!} is more technical than interesting so we refer the reader to the formalisation or paper TODO.
+The proof of \hyperref[lst:lemma_2]{\lstinline!lemma_2!} is more technical than interesting so we refer the reader to the formalisation.
 
@@
   Furthermore, no model can satisfy both \lstinline!$\phi$! and \lstinline!&\fmNeg{$\phi$}&! at the same time.
-  Lastly, a membership cycle is impossible since the membership relation of \lstinline!hf! is well-founded (see \autoref{sec:semantics} TODO).
+  Lastly, a membership cycle is impossible since the membership relation of \lstinline!hf! is well-founded.
 \end{proof}
@@
 
-\section{Bounding the Cardinality of a Branch\label{sec:bound}}
-All in all, we get the following lemma.
+\section{Total Correctness of the Decision Procedure\label{sec:correct}}
+We first demonstrate the termination of the procedure for well-formed branches, i.e.\ every well-formed branch is in the domain of \lstinline!mlss_proc_branch!.
+To this end, we derive an upper bound for the number of distinct formulae in a branch whose proof we omit here for brevity (cf. Appendix\todo{Reference Appendix}).
+We should point out that this bound is not to be construed as the complexity of the procedure as it may create exponentially many branches in general.
 \begin{lstlisting}[label={lst:card_branch}]
@@
   assumes wf_branch b
-  shows |set b|
-      $\leq$ 2 * |subfms (last b)| + 16 * |subterms (last b)|$^\text{4}$
-\end{lstlisting}
-
-\section{Total Correctness of the Decision Procedure\label{sec:correct}}
-We first prove the termination of the procedure for well-formed branches, i.e.\ every well-formed branch is in the domain of \lstinline!mlss_proc_branch!.
-\begin{lstlisting}[belowskip=0pt]
+  shows |set b| $\leq$ 2$\:$*$\,$|subfms (last b)| + 16$\:$*$\,$|subterms (last b)|$^\text{4}$
+\end{lstlisting}
+Remember that \lstinline!mlss_proc_branch! only applies a linear expansion rule to a branch if the application results in new formulae.   
+Moreover, the subsumption conditions of the branching expansion rules ensure that each of the newly created branches contain new formulae.
+Ultimately, we conclude that the procedure must terminate because the number of formualae increases in each step but it is also bounded. 
+\begin{lstlisting}
 lemma mlss_proc_branch_dom_if_wf_branch:
@@
 \end{lstlisting}
-\begin{proof}
-  We proceed by strong induction on the difference between the upper bound due to \hyperref[lst:card_branch]{\lstinline!card_wf_branch_ub!} and \lstinline!|b|!.
-
-  We only have one inductive case where the induction hypothesis holds for all well-formed branches \lstinline!b'!
-where the difference between the upper bound and the cardinality is smaller than that of \lstinline!b!.
-  We strengthen the assumptions of the induction hypothesis to obtain
-\begin{lstlisting}
-  $\llbracket$ &\expandss{b'}{b}&; set b $\subset$ set b' $\rrbracket$ $\Longrightarrow$ mlss_proc_branch_dom b'
-\end{lstlisting}
-for all \lstinline!b'!.
-The former strengthening is admissible since, by definition, every expansion of a well-formed branch is again well-formed while the latter ensures that the cardinality of \lstinline!b'!
-is larger than the cardinality of \lstinline!b!, implying that the difference to the upper bound decreases.
-
-Now, consider the case where \lstinline!b! is saturated.
-Then, the procedure terminates without recursive calls and we immediately get \lstinline!mlss_proc_branch_dom b! using the corresponding introduction rule for the domain predicate.
-
-If \lstinline!b! is not saturated, we can expand \lstinline!b! using a linear or a branching rule, which adds new formulae to the resulting branch(es).
-In the first case, we can thus choose a branch \lstinline!b'! with \lstinline!SOME! that linearly expands \lstinline!b!.
-Moreover, the domain predicate holds for \lstinline!b' @ b! since the assumptions of the modified induction hypothesis are fulfilled.
-Ultimately, we use the corresponding introduction rule
-\begin{lstlisting}
-  $\llbracket$ $\neg$ lin_sat b; mlss_proc_branch_dom (b' @ b) $\rrbracket$
-  $\Longrightarrow$ mlss_proc_branch_dom b
-\end{lstlisting}
-for the domain predicate to obtain \lstinline!mlss_proc_branch b!.
-
-The case of a branching rule is analogous.
-\end{proof}
-
-\noindent The above lemma allows us to use the computation induction rule of \lstinline!mlss_proc_branch! on well-formed branches, which we use to prove soundness and completeness.
+
+The above lemma allows us to use the computation induction rule of \lstinline!mlss_proc_branch! on well-formed branches, which we use to prove soundness and completeness.
 As both proofs are essentially an application of soundness, respectively completeness, of the calculus, we refer the reader to the formalisation.
@@
 \end{lstlisting}
+
+\section{Dealing with Urelements}
+In the introduction, we motivated the verification of \lstinline!mlss_proc! with the prospect of integrating it as a tactic into Isabelle.
+For this to work, we need to replay a proof in the tableau calculus through Isabelle/HOL, i.e.\ we need to map every branch expansion rule to a corresponding theorem in Isabelle/HOL.
+This is straightforward for all expansion rules except for the last branching expansion rule.
+To illustrate, suppose that we are to disprove a statement of the form \lstinline!s $\neq$ (t :: 'a) $\land$ s $\in$ (A :: 'a set) $\cup$ B $\land$ $\ldots$! in Isabelle/HOL
+that corresponds to \lstinline!s' $\neqS$ t' $\fmAndSymbol$ s' $\inS$ A' $\unionS$ B' $\fmAndSymbol$ $\ldots$! in our set syntax.
+Applying the decision procedure to this goal might result in a tableau proof containing an application of the last branching rule rule to \lstinline!(s' $\neqS$ t') $\in$ set b!, which creates the branches \lstinline![Var x $\inS$ s', Var x $\notinS$ t'] @ b!
+and \lstinline![Var x $\notinS$ s', Var x $\inS$ t'] @ b!.
+But this can not be converted to a proof in Isabelle/HOL since \lstinline!s! and \lstinline!t! are not sets.
+
+To deal with this problem, we introduce a light-weight type system that, given a mapping from variables to natural numbers, assigns a natural number to each set term.
+
+\begin{itemize}
+  \item Give example from HOL
+  \item Show how branch is expanded
+  \item Mention implemented constraint solver
+  \item Show how definition of realise needs to be changed
+  \item Syntax needs to be changed for empty set
+\end{itemize}
 
@@
   \bigskip
-  \caption{Linear expansion rules.\label{tab:lexpands}}
+  \caption{Linear expansion rules}
 \end{table*}
@@
   \bigskip
-  \caption{Branching expansion rules.\label{tab:bexpands} We write \lstinline!$\phi$! for \lstinline!last b! here.}
+  \caption{Branching expansion rules. We write \lstinline!$\phi$! for \lstinline!last b! here.}
 \end{table*}

### 2023-02-20 paper/CADE/paper.tex
Continue with paper

@@
 
-\usepackage{enumitem}
+\usepackage[inline]{enumitem}
 
@@
 
-\textit{Literals} are a type of subformulae of particular importance, so we define a predicate for them.
-\begin{lstlisting}
-fun is_literal :: 'a fm $\Rightarrow$ bool where
-  is_literal (&\fmAtom& _) = True | is_literal (&\fmNeg{(\fmAtom{} \_)}&) = True
-| is_literal _ = False
-\end{lstlisting}
-
 \subsection{Semantics}
@@
 The implementation of $\models$ coincides with the interpretation function of \citeauthor{lqe}~\cite{lqe}.
-As usual, a formula \lstinline!$\phi$! is called \textit{satisfiable} if there exists a model \lstinline!$M$! with \lstinline!$M$ $\models$ $\phi$!.
+As usual, a formula \lstinline!$\phi$! is called \textit{satisfiable} if there exists a model \lstinline!M! with \lstinline!M $\models$ $\phi$!.
 Otherwise, \lstinline!$\phi$! is called \textit{unsatisfiable}.
@@
 The realisation function is defined relative to this graph.
-As mentioned before, the pure witnesses \lstinline!Var ` pwits b! and the rest of the set terms \lstinline!subterms' b! are treated differently by the realisation function.
+As mentioned before, the pure witnesses are treated differently than the rest of the set terms by the realisation function.
 Terms in the latter set are evaluated in accordance to the structure of the graph, i.e.\ the realisation of a vertex is defined as the union of the realisations of the parent vertices. 
-For the former set, we choose a function \lstinline!I! that assigns the witnesses in \lstinline!Var ` pwits b! pairwise distinct sets with cardinality greater than that of the vertices.
+For the former set, we choose a function \lstinline!I! that assigns the pure witnesses pairwise distinct sets with cardinality greater than that of the vertices.
 We can always choose such a function since we assume an infinite universe of variables.
@@
 \end{lstlisting}
+\newcommand{\reflemmatwo}{\hyperref[lst:lemma_2]{\lstinline!lemma_2!}}
 So why are pure witnesses treated differently?
@@
 
-The proof of \hyperref[lst:lemma_2]{\lstinline!lemma_2!} is more technical than interesting so we refer the reader to the formalisation.
+The proof of \reflemmatwo is more technical than interesting so we refer the reader to the formalisation.
 
@@
 We start by verifying that the realisation function models all literals in the branch; more formally, the following propositions hold:
-\begin{enumerate}[label={(\roman*)}]
+\begin{enumerate}[label={(\arabic*)}]
   \item\label{it:at_mem} We have \lstinline!realise s $\hfmem$ realise t! if it holds that \lstinline!s $\inS$ t! is in \lstinline!b!. 
@@
 \end{enumerate}
-To illustrate the usefulness of \hyperref[lst:lemma_2]{\lstinline!lemma_2!}, we prove Proposition~\ref{it:at_eq}.
+To illustrate the usefulness of \reflemmatwo, we prove Proposition~\ref{it:at_eq}.
 The proofs of all propositions translate well into Isabelle, so we refer to the original paper~\cite{new_fast_tableau} for the remaining proofs. 
@@
   Assume that \lstinline!s $\eqS$ t! is in \lstinline!b!.
-  If there exists a \lstinline!c $\in$ pwits b! where \lstinline!s = Var c! or \lstinline!t = Var c!, we arrive at a contradiction due to \hyperref[lst:lemma_2]{\lstinline!lemma_2!}.
+  If there exists a \lstinline!c $\in$ pwits b! where \lstinline!s = Var c! or \lstinline!t = Var c!, we arrive at a contradiction due to \reflemmatwo.
   Therefore, both \lstinline!s $\in$ subterms' b! and \lstinline!t $\in$ subterms' b! must hold.
@@
 A tableau calculus is sound if for any closed tableau, the corresponding formula is unsatisfiable. 
-We prove the following two properties to establish soundness.
-\begin{enumerate}[label={(\roman*)}]
+We prove the following two properties to establish soundness:
+\begin{enumerate*}[label={(\arabic*)}]
   \item A closed branch contains an unsatisfiable formula.
   \item The expansion rules maintain satisfiability.
-\end{enumerate}
+\end{enumerate*}
 
@@
 Moreover, the subsumption conditions of the branching expansion rules ensure that each of the newly created branches contain new formulae.
-Ultimately, we conclude that the procedure must terminate because the number of formualae increases in each step but it is also bounded. 
+Ultimately, we conclude that the procedure must terminate because the number of formulae increases in each step but is also bounded. 
 \begin{lstlisting}
@@
 To illustrate, suppose that we are to disprove a statement of the form \lstinline!s $\neq$ (t :: 'a) $\land$ s $\in$ (A :: 'a set) $\cup$ B $\land$ $\ldots$! in Isabelle/HOL
-that corresponds to \lstinline!s' $\neqS$ t' $\fmAndSymbol$ s' $\inS$ A' $\unionS$ B' $\fmAndSymbol$ $\ldots$! in our set syntax.
-Applying the decision procedure to this goal might result in a tableau proof containing an application of the last branching rule rule to \lstinline!(s' $\neqS$ t') $\in$ set b!, which creates the branches \lstinline![Var x $\inS$ s', Var x $\notinS$ t'] @ b!
+that corresponds to \lstinline!s' $\neqS$ t' $\fmAndSymbol$ s' $\inS$ A' $\unionS$ B' $\fmAndSymbol$ $\ldots$!
+in our set syntax for some \lstinline!s'!, \lstinline!t'!, \lstinline!A'!, and \lstinline!B'!.
+Applying the decision procedure to this formula might result in a tableau proof containing an application of the last branching rule rule to \lstinline!(s' $\neqS$ t') $\in$ set b!, which creates the branches \lstinline![Var x $\inS$ s', Var x $\notinS$ t'] @ b!
 and \lstinline![Var x $\notinS$ s', Var x $\inS$ t'] @ b!.
@@
 
-To deal with this problem, we introduce a light-weight type system that, given a mapping from variables to natural numbers, assigns a natural number to each set term.
-
-\begin{itemize}
-  \item Give example from HOL
-  \item Show how branch is expanded
-  \item Mention implemented constraint solver
-  \item Show how definition of realise needs to be changed
-  \item Syntax needs to be changed for empty set
-\end{itemize}
+To deal with this problem, we introduce a light-weight type system that, given a mapping $\Gamma$ from variables to natural numbers, assigns a natural number \lstinline!l!, which we call level, to each set term \lstinline!t'!.
+We write \lstinline!$\Gamma$ $\vdash$ t' : l! for this judgement.
+Intuitively speaking, this indicates that the corresponding term \lstinline!t! in Isabelle/HOL has type
+\[
+  \texttt{'a}\ \underbrace{\texttt{set}\ \ldots\ \texttt{set}}_{\texttt{l}\ \text{times}}
+\]
+for some \lstinline!'a!.
+We can extend this to a judgement \lstinline!$\Gamma$ $\vdash$ a!
+for atoms \lstinline!a!, where we require the levels of \lstinline!s'!
+and \lstinline!t'!
+to be equal if \lstinline!a = s' $\eqS$ t'!
+and the level of \lstinline!t'!
+to be the successor of that of \lstinline!s'!
+if \lstinline!a = s' $\inS$ t'!.
+By only judging the atoms, we can extend this to formulas for which we use the same notation.
+
+We can now define the urelements with respect to a formula, i.e.\ those set terms whose corresponding type in Isabelle/HOL might not be a set.
+\begin{lstlisting}
+definition urelem :: 'a pset_fm $\Rightarrow$ 'a pset_term $\Rightarrow$ bool where
+  urelem $\phi$ t $\equiv$ $\exists\Gamma$. $\Gamma$ $\vdash$ $\phi$ $\land$ $\Gamma$ $\vdash$ t : 0
+\end{lstlisting}
+Using this definition, we make two changes to the specification of the calculus: 
+\begin{enumerate*}[label=(\arabic*)]
+  \item First and foremost, we require that neither \lstinline!s! nor \lstinline!t! is an urelement in the precondition of the last branching expansion rule.
+  \item We add an argument to the $\emptyset$ constructor that specifies the level of each instance of $\emptyset$.
+    This argument is only used for the typing judgement; it has no impact on the semantics.
+\end{enumerate*}
+
+Soundness, of course, is not affected by these changes but we have to make a few amendments to maintain completeness:
+\begin{enumerate*}[label=(\arabic*)]
+  \item The first equation of the realisation function now also needs to account for the urelements.
+    In particular, it needs to ensure that urelements receive pairwise different values unless they are related through equality atoms. 
+    This has no effect on pure witnesses since they can not be related through equality atoms due to \reflemmatwo.
+  \item We need to adjust the existing completeness proof in those places where it directly refers to the definition of \lstinline!realise! to account for the case where a given term is an urelement.
+  \item The completeness theorem receives the additional assumption that \lstinline!$\Gamma$ $\vdash$ $\phi$! holds for the initial formula $\phi$. 
+  \item For the completeness proof, we need to show that the typing judgement is invariant under branch expansion.
+\end{enumerate*}
+
+All in all, having the formalisation allows us to confidently modify the calculus as Isabelle immediately shows us where the proofs break.
+
+TODO: Reference type inference and wrap up.
 
@@
 \clearpage\chapter*{Appendix}
+\todo[inline]{Clean up Appendix}
 
@@
 \section{Upper-Bounding the Cardinality of a Branch}
+\textit{Literals} are a type of subformulae of particular importance, so we define a predicate for them.
+\begin{lstlisting}
+fun is_literal :: 'a fm $\Rightarrow$ bool where
+  is_literal (&\fmAtom& _) = True | is_literal (&\fmNeg{(\fmAtom{} \_)}&) = True
+| is_literal _ = False
+\end{lstlisting}
 

### 2023-02-21 paper/CADE/paper.tex
Continue

@@
 \usepackage{pgfplots}
+
+% Set float spacing
+%\setlength{\abovecaptionskip}{0pt}
+%\setlength{\textfloatsep}{14pt plus 2pt minus 2pt}
 
@@
 
-\listoftodos{}
-
-\clearpage
-
 \section{Introduction}
@@
 Otherwise, \lstinline!$\phi$! is called \textit{unsatisfiable}.
+\todo[inline]{Introduce the initial formula}
 
@@
 
-The proof of \reflemmatwo is more technical than interesting so we refer the reader to the formalisation.
+The proof of \reflemmatwo{} is more technical than interesting so we refer the reader to the formalisation.
 
@@
 \end{enumerate}
-To illustrate the usefulness of \reflemmatwo, we prove Proposition~\ref{it:at_eq}.
+To illustrate the usefulness of \reflemmatwo{}, we prove Proposition~\ref{it:at_eq}.
 The proofs of all propositions translate well into Isabelle, so we refer to the original paper~\cite{new_fast_tableau} for the remaining proofs. 
@@
   Assume that \lstinline!s $\eqS$ t! is in \lstinline!b!.
-  If there exists a \lstinline!c $\in$ pwits b! where \lstinline!s = Var c! or \lstinline!t = Var c!, we arrive at a contradiction due to \reflemmatwo.
+  If there exists a \lstinline!c $\in$ pwits b! where \lstinline!s = Var c! or \lstinline!t = Var c!, we arrive at a contradiction due to \reflemmatwo{}.
   Therefore, both \lstinline!s $\in$ subterms' b! and \lstinline!t $\in$ subterms' b! must hold.
@@
   But then we have
-  \lstinline!realise u $\hfmem$ realise t = e $\hfmem$ realise t!
-  using Proposition~\ref{it:at_mem}, which stands in contradiction to the assumption \lstinline!e $\hfnotmem$ realise t!.
+  \lstinline!realise u $\hfmem$ realise t $\longleftrightarrow$ e $\hfmem$ realise t!
+  using Proposition~\ref{it:at_mem}, which is a contradiction to the assumption \lstinline!e $\hfnotmem$ realise t!.
 \end{proof}
@@
 
-To deal with this problem, we introduce a light-weight type system that, given a mapping $\Gamma$ from variables to natural numbers, assigns a natural number \lstinline!l!, which we call level, to each set term \lstinline!t'!.
-We write \lstinline!$\Gamma$ $\vdash$ t' : l! for this judgement.
-Intuitively speaking, this indicates that the corresponding term \lstinline!t! in Isabelle/HOL has type
+\begin{figure}[t]
+  \centering
+  \lstset{
+    escapeinside={@}{@}
+  }
+  \begin{tabular}{c}
+    \begin{tabular}{ccccc}
+      &\qquad\quad&&\qquad\quad& \lstinline!$\Gamma$ $\vdash$ t : l! \\
+      \cmidrule{1-1}\cmidrule{3-3}\cmidrule{5-5}
+      \lstinline!$\Gamma$ $\vdash$ $\emptyset$ n : Suc n! & &
+      \lstinline!$\Gamma$ $\vdash$ Var x : $\Gamma$ x! & &
+      \lstinline!$\Gamma$ $\vdash$ Single t : Suc l! \\
+    \end{tabular}
+    \\[2em]
+    \begin{tabular}{c}
+      \lstinline!$\star_\text{s}$ $\in$ {$\unionS$,$\,\interS$,$\,\diffS$}! \quad
+      \lstinline!$\Gamma$ $\vdash$ s : l! \quad
+      \lstinline!$\Gamma$ $\vdash$ t : l! \quad 
+      \lstinline!l $\neq$ 0! \\
+      \midrule
+      \lstinline!$\Gamma$ $\vdash$ s $\star_\text{s}$ t : l! \\
+    \end{tabular}
+    \\[2.5em]
+    \begin{tabular}{ccc}
+      \lstinline!$\Gamma$ $\vdash$ s : l! \quad \lstinline!$\Gamma$ $\vdash$ t : l! &\qquad\quad&
+      \lstinline!$\Gamma$ $\vdash$ s : l! \quad \lstinline!$\Gamma$ $\vdash$ t : Suc l!
+      \\
+      \cmidrule{1-1}\cmidrule{3-3}
+      \lstinline!$\Gamma$ $\vdash$ s $\eqS$ t! && \lstinline!$\Gamma$ $\vdash$ s $\inS$ t! \\
+    \end{tabular}
+  \end{tabular}
+  \caption{The type system for set terms and atoms.\label{fig:types}}
+\end{figure}
+
+To deal with this problem, we introduce a light-weight type system as displayed in \autoref{fig:types}. 
+The type of a set term in this system is just a natural number which we call level.
+Intuitively speaking, the level indicates that the corresponding term \lstinline!t! in Isabelle/HOL has type
 \[
@@
 for some \lstinline!'a!.
-We can extend this to a judgement \lstinline!$\Gamma$ $\vdash$ a!
-for atoms \lstinline!a!, where we require the levels of \lstinline!s'!
-and \lstinline!t'!
-to be equal if \lstinline!a = s' $\eqS$ t'!
-and the level of \lstinline!t'!
-to be the successor of that of \lstinline!s'!
-if \lstinline!a = s' $\inS$ t'!.
-By only judging the atoms, we can extend this to formulas for which we use the same notation.
+Note that the constructor $\emptyset$ now receives an additional argument that indicates the level of each instance of $\emptyset$.
+
+Moreover, the typing judgement extends to set atoms by matching up the levels of its component set terms.
+
+Ultimately, we define \lstinline!$\Gamma$ $\vdash$ $\psi$ $\equiv$ $\forall$a $\in$ atoms $\phi$. $\Gamma$ $\vdash$ a! in order to type formulae.
 
@@
   \item First and foremost, we require that neither \lstinline!s! nor \lstinline!t! is an urelement in the precondition of the last branching expansion rule.
-  \item We add an argument to the $\emptyset$ constructor that specifies the level of each instance of $\emptyset$.
+  \item As mentioned above, we add an argument to the $\emptyset$ constructor.
     This argument is only used for the typing judgement; it has no impact on the semantics.
@@
 \begin{enumerate*}[label=(\arabic*)]
-  \item The first equation of the realisation function now also needs to account for the urelements.
+  \item The first equation of \lstinline!realise! now also needs to account for the urelements.
     In particular, it needs to ensure that urelements receive pairwise different values unless they are related through equality atoms. 
-    This has no effect on pure witnesses since they can not be related through equality atoms due to \reflemmatwo.
+    This has no effect on pure witnesses since they can not be related through equality atoms due to \reflemmatwo{}.
   \item We need to adjust the existing completeness proof in those places where it directly refers to the definition of \lstinline!realise! to account for the case where a given term is an urelement.
@@
 
-TODO: Reference type inference and wrap up.
+The modifications above ensure that the proof can be replayed through Isabelle/HOL.
+To actually use the calculus, we need to determine the urelements of the initial formula $\phi$, though.
+In other words, we need to implement an inference algorithm for our light-weight type system.
+The algorithm is, in essence, a simplified version of Hindley-Milney type inference so it has the same two phases:
+it generates constraints using syntax directed rules and then passes them to a constraint solver.
+
+Since we are only interested in the level of a term, we can encode all constraints into the theory consisting of $0$, the successor function \lstinline!S!, and equality (but no disequality).
+Note that constraints of the form \lstinline!l $\neq$ 0! can be replaced by \lstinline!l $=$ S i! with \lstinline!i! being a fresh variable.
+A solver for this theory is straightforward to implement;
+nevertheless, we have to be careful that it computes the minimum assignment $\Gamma$ from variables to levels that fulfils the constraints. 
+This is to guarantee that a set term \lstinline!t! is not an urelement \textbf{iff} \lstinline!$\Gamma$ t > 0!.
+Conversely, all terms \lstinline!s! with \lstinline!$\Gamma$ s $=$ 0! are urelements.
 
@@
 
+%\clearpage\listoftodos
 \appendix
+
 \clearpage\chapter*{Appendix}

### 2023-02-22 paper/CADE/paper.tex
Continue

@@
 
-\title{Towards a Verified Tableau Prover for a Ground Fragment of Set Theory}
+\title{Towards a Verified Tableau Prover for an Unquantified Fragment of Set Theory}
+\titlerunning{Towards a Verified Tableau Prover for MLSS}
 \author{Lukas Stevens\orcidID{0000-0003-0222-6858}}
@@
 \begin{abstract}
-  Using Isabelle/HOL, we verify the state-of-the-art decision procedure for multi-level syllogistic with singleton (\MLSS{} for short), which is a ground fragment of set theory.
+  Using Isabelle/HOL, we verify the state-of-the-art decision procedure for multi-level syllogistic with singleton (\MLSS{} for short), which is an unquantified fragment of set theory.
   We formalise the syntax and semantics of \MLSS{} as well as a sound and complete tableau calculus for it.
@@
 \section{Introduction}
-In contrast to pen-and-paper proofs where we expect the reader to fill in trivial details, interactive theorem provers (ITPs) require us to justify each proof step rigorously.
-Manually reducing each proof step down to the axioms would be impractical, though;
+In contrast to pen-and-paper proofs where we expect the reader to fill in trivial details, interactive theorem provers (ITPs) require us to justify each proof step rigorously;
 hence, ITPs usually offer a wide array of automated methods to deal with the proof goals that arise during interactive use.
@@
 More recently, \citeauthor{orders}~\cite{orders} presented a verified decision procedure for orders that produces certificates. 
-This approach offers efficient execution by using generated code as well as soundness as the certificates are replayed through Isabelle's inference kernel.
+This approach offers efficient execution by using generated code as well as soundness because the certificates are replayed through Isabelle's inference kernel.
 
 The focus of this paper is another ubiquitous structure in mathematics, namely sets.
-In particular, we consider a ground fragment of set theory which \citeauthor{new_fast_tableau}~\cite{new_fast_tableau} call multi-level syllogistic with singleton (\MLSS{}).
+To the best of our knowledge, we present the first formally verified decision procedure for (a fragment of) set theory.
+In particular, we consider an unquantified fragment which \citeauthor{new_fast_tableau}~\cite{new_fast_tableau} call multi-level syllogistic with singleton (\MLSS{}).
 The fragment includes the usual set operations of union, intersection, difference, membership, equality and, in addition, it allows the construction of singleton sets.
-Due to the omnipresence of sets, we think that automation for this fragment will benefit the users of Isabelle, especially when integrated into the simplifier like the aforementioned order solver.
-We also expect that the generation of certificates will be straightforward since \MLSS{} admits a tableau calculus.
+
+Since \MLSS{} admits a tableau calculus, we expect that the generation of certificates will be straightforward.
+Like with the aforementioned order solver, this would pave the way towards an integration of the decision procedure into Isabelle, adding to its growing body of verified decision procedures.
 
 \subsection{Contributions}
-In this paper, we present a formalisation in Isabelle/HOL of a tableau calculus for \MLSS{} due to \citeauthor{new_fast_tableau}~\cite{new_fast_tableau}\cite[Chapter 14]{set_theory}.
+This paper is about the formalisation in Isabelle/HOL of a tableau calculus for \MLSS{} due to \citeauthor{new_fast_tableau}~\cite{new_fast_tableau}\cite[Chapter 14]{set_theory}.
 We prove soundness and completeness of the calculus and give an abstract specification of a decision procedure that applies the rules of the calculus exhaustively.
 To obtain total correctness of the procedure, we prove its termination.
-The formalisation follows the paper closely but gives a more thorough account of some important details:
+The formalisation initially follows the paper but gives a more thorough account of some important details:
 \begin{itemize}
@@
   \item We derive an explicit upper bound for the number of formulae in a branch of the tableau.
-  \item We restrict one branching rule\todo{Elaborate}.
 \end{itemize}
+
+In the context of Isabelle/HOL, there is one important aspect that requires us to modify the calculus in the paper:
+the calculus works under the assumption that every variable is a set;
+however, this is not the case in Isabelle/HOL, e.g.\ consider the expression \lstinline!n $\in$ A! where \lstinline!n! is a natural number.
+In order to deal with these urelements, we extend the calculus with a light-weight type system and a verified inference algorithm that identifies the urelements.
+
+The modification of the calculus required non-trivial changes to the completeness proof.
+Here, the formalisation was instrumental because Isabelle immediately revealed broken proofs that needed fixing.
+This illustrates the usefulness of ITPs for developing logic calculi: they allow us to confidently make modifications without compromising correctness.
 
@@
 
+Abstract framework to prove completeness for first-order tableau systems~\cite{completeness_coinductive}.
+Verified sequent calculus prover for first-order logic~\cite{sequent_first_order}.
+Tableau calculus for temporal logic CTL in Coq~\cite{ctl_tableau}
+Abstract framework for synthetic completeness in AFP~\cite{synthetic_completeness_afp}.
+
 While a lot of work has gone into the verification of decision procedures, seemingly little has been put towards the verification of tableau calculi.
@@
 The termination proof is part of ongoing work~\cite{hybrid_logic_workshop}.
-\todo[inline]{Mention SteP, Check reviewr comment}
 % TODO: Formalisation de la logique temporelle dans le coq.
@@
 or \lstinline!$\eqS$!.
+As an example, \lstinline!u! and \lstinline!s $\interS$ t! are top-level terms of \lstinline!u $\inS$ s $\interS$ t! but \lstinline!s! and \lstinline!t! are not.
 
@@
 
-\section{Dealing with Urelements}
+\section{Dealing with Urelements\label{sec:urelements}}
 In the introduction, we motivated the verification of \lstinline!mlss_proc! with the prospect of integrating it as a tactic into Isabelle.
@@
 This is straightforward for all expansion rules except for the last branching expansion rule.
-To illustrate, suppose that we are to disprove a statement of the form \lstinline!s $\neq$ (t :: 'a) $\land$ s $\in$ (A :: 'a set) $\cup$ B $\land$ $\ldots$! in Isabelle/HOL
-that corresponds to \lstinline!s' $\neqS$ t' $\fmAndSymbol$ s' $\inS$ A' $\unionS$ B' $\fmAndSymbol$ $\ldots$!
+To illustrate, suppose that we are to disprove a statement of the form \lstinline!s $\neq$ (t :: 'a) $\land$ s $\in$ (A :: 'a set) $\cup$ B $\land$ $\ldots$! in Isabelle/HOL.
+By way of reification, we convert this statement to \lstinline!s' $\neqS$ t' $\fmAndSymbol$ s' $\inS$ A' $\unionS$ B' $\fmAndSymbol$ $\ldots$!
 in our set syntax for some \lstinline!s'!, \lstinline!t'!, \lstinline!A'!, and \lstinline!B'!.
@@
 
-We can now define the urelements with respect to a formula, i.e.\ those set terms whose corresponding type in Isabelle/HOL might not be a set.
+We can now define the urelements with respect to a formula.
+An urelement is a set term whose corresponding type in Isabelle/HOL might not be a set.
 \begin{lstlisting}
@@
     This has no effect on pure witnesses since they can not be related through equality atoms due to \reflemmatwo{}.
-  \item We need to adjust the existing completeness proof in those places where it directly refers to the definition of \lstinline!realise! to account for the case where a given term is an urelement.
+  \item We need to adjust the completeness proof in those places where it directly refers to the definition of \lstinline!realise! to account for the case where a given term is an urelement.
   \item The completeness theorem receives the additional assumption that \lstinline!$\Gamma$ $\vdash$ $\phi$! holds for the initial formula $\phi$. 
@@
 
-All in all, having the formalisation allows us to confidently modify the calculus as Isabelle immediately shows us where the proofs break.
+All in all, having the formalisation allows us to confidently modify the calculus as Isabelle immediately shows us where the proofs break.\todo{Longer description}
 
@@
 nevertheless, we have to be careful that it computes the minimum assignment $\Gamma$ from variables to levels that fulfils the constraints. 
-This is to guarantee that a set term \lstinline!t! is not an urelement \textbf{iff} \lstinline!$\Gamma$ t > 0!.
+This is to guarantee that a set term \lstinline!t! is not an urelement \textbf{iff} \lstinline!$\Gamma$ t $>$ 0!.
 Conversely, all terms \lstinline!s! with \lstinline!$\Gamma$ s $=$ 0! are urelements.
@@
 \section{Conclusion and Future Work}
-By closely following a paper by \citeauthor{new_fast_tableau}~\cite{new_fast_tableau}, we developed a formalisation of a tableau calculus for a ground fragment of set theory called \MLSS{}.
+By closely following a paper by \citeauthor{new_fast_tableau}~\cite{new_fast_tableau}, we developed a formalisation of a tableau calculus for an unquantified fragment of set theory called \MLSS{}.
 The formalisation includes an abstract description of a decision procedure that builds on the calculus.

### 2023-02-23 paper/CADE/paper.tex
Continue

@@
   Using Isabelle/HOL, we verify the state-of-the-art decision procedure for multi-level syllogistic with singleton (\MLSS{} for short), which is an unquantified fragment of set theory.
-  We formalise the syntax and semantics of \MLSS{} as well as a sound and complete tableau calculus for it.
+  We formalise its syntax and semantics as well as a sound and complete tableau calculus for it.
   We also provide an abstract specification of a decision procedure that applies the rules of the calculus exhaustively and prove its termination.
+  Furthermore, we extend the calculus with a light-weight type system that paves the way for an integration of the procedure into Isabelle/HOL.
   \keywords{First TODO \and Second TODO\todo{keywords}}
@@
 \subsection{Contributions}
-This paper is about the formalisation in Isabelle/HOL of a tableau calculus for \MLSS{} due to \citeauthor{new_fast_tableau}~\cite{new_fast_tableau}\cite[Chapter 14]{set_theory}.
+We present a formalisation in Isabelle/HOL of a tableau calculus for \MLSS{} due to \citeauthor{new_fast_tableau}~\cite{new_fast_tableau}\cite[Chapter 14]{set_theory}.
 We prove soundness and completeness of the calculus and give an abstract specification of a decision procedure that applies the rules of the calculus exhaustively.
@@
 
-Abstract framework to prove completeness for first-order tableau systems~\cite{completeness_coinductive}.
-Verified sequent calculus prover for first-order logic~\cite{sequent_first_order}.
-Tableau calculus for temporal logic CTL in Coq~\cite{ctl_tableau}
-Abstract framework for synthetic completeness in AFP~\cite{synthetic_completeness_afp}.
-
-While a lot of work has gone into the verification of decision procedures, seemingly little has been put towards the verification of tableau calculi.
-In Lean, there is a verified tableau calculus for basic modal logic~\cite{modal_logic_lean}.
-In Coq, there is a formalisation of a tableau calculus for ecumenical logic~\cite{ecumenical_logic}; it is, however, missing a completeness proof.
-Concerning Isabelle/HOL, there is formalisation of a sound and complete tableau calculus for hybrid logic~\cite{hybrid_logic} available in the \textit{Archive of Formal Proofs}~\cite{hybrid_logic_afp}.
-The termination proof is part of ongoing work~\cite{hybrid_logic_workshop}.
+There is a large body of work at the intersection of ITPs and tableau methods but to keep with the theme of the paper we only consider formalisations of correctness here.
+For first-order logic, there are abstract completeness proofs using the \emph{Beth-Hintikka style} of possibly infinite derivation trees~\cite{completeness_coinductive} as well as the \emph{Henkin style} of maximally consistent sets~\cite{synthetic_completeness_afp}.
+Both are abstract enough to be instantiated with a wide range of concrete calculi.
+A more concrete formalisation~\cite{sequent_first_order} verifies a sequent calculus for first-order logic whose completeness proof is via a translation to semantic tableau.
+
+Beyond completeness, we target decidability which is more attainable for propositional logic.  
+There is a verified tableau calculus for the modal logic S5~\cite{lean_modal_logic} in Lean and one for hybrid logic~\cite{hybrid_logic} in Isabelle/HOL.
+Both of these formalisation do not prove termination but there is a formalisation of a tableau calculus for the temporal logic CTL in Coq~\cite{ctl_tableau} that does.
 % TODO: Formalisation de la logique temporelle dans le coq.
@@
 We will often drop the atom constructor \lstinline!&\fmAtom{}&! to reduce clutter.
-Additionally, we use the abbreviations \lstinline!s $\notinS$ t!
-and \lstinline!s $\neqS$ t!
-to denote \lstinline!&\fmNeg{\fmAtom{} (s $\inS$ t)}&!
-and \lstinline!&\fmNeg{\fmAtom{} (s $\eqS$ t)}&!, respectively.
-
-Like the function \lstinline!vars!, the automatically defined function \lstinline!atoms :: 'a fm $\Rightarrow$ 'a set! retrieves all set atoms in a formula.
+Additionally, we use \lstinline!s $\notinS$ t! and \lstinline!s $\neqS$ t! to denote \lstinline!&\fmNeg{\fmAtom{} (s $\inS$ t)}&! and \lstinline!&\fmNeg{\fmAtom{} (s $\eqS$ t)}&!, respectively.
+
+Similarly to \lstinline!vars!, we get the function \lstinline!atoms :: 'a fm $\Rightarrow$ 'a set! for free which retrieves all set atoms in a formula.
 We combine these functions to extract all the variables occurring in a set formula.
@@
 
-Similarly to \lstinline!vars!, we introduce the overloaded constant \lstinline!subterms :: 'b $\Rightarrow$ 'a pset_term set!, which returns all the set terms that are subterms of a set term, set atom, or formula, respectively.
+Likewise, we fix the constant \lstinline!subterms :: 'b $\Rightarrow$ 'a pset_term set!.
+We overload this constant to return the set terms that are subterms of a set term, set atom, or formula, respectively.
 Lastly, we consider the subformulae of a formula and define a function \lstinline!subfms :: 'a fm $\Rightarrow$ 'a fm set!
 that computes them.
-The functions \lstinline!subterms! and \lstinline!subfms! are defined in the usual way (cf. Appendix \todo{reference Appendix}).
+The functions \lstinline!subterms! and \lstinline!subfms! are defined in the usual way (cf.\ \appendixname~\ref{appendix:subterms}).
 
@@
 The original paper~\cite{new_fast_tableau} bases the semantics of \MLSS{} on the von Neumann hierarchy of sets $\mathcal{V}$. 
-While there is an entry~\cite{zfc_in_hol_afp} in the \textit{Archive of Formal Proofs} (AFP) that axiomatises this hierarchy,
-we instead use the hierarchy of \textit{hereditarily finite sets} (HF sets) which fulfil all the same axioms as $\mathcal{V}$ --- that is, the axioms of ZF --- except for the axiom of infinity.\todo{Membership is well-founded.}
+While there is an entry~\cite{zfc_in_hol_afp} in the \emph{Archive of Formal Proofs} (AFP) that axiomatises this hierarchy,
+we instead use the hierarchy of \emph{hereditarily finite sets} (HF sets) which fulfil all the same axioms as $\mathcal{V}$ --- that is, the axioms of ZF --- except for the axiom of infinity.
+In particular, the membership relation is well-founded.
 The HF sets are sufficient for our purposes as \MLSS{} is not powerful enough to force infinite models~\cite[Chapter~14.2]{set_theory}.
@@
 in the standard way, i.e.\ by mapping each syntactic construct to the corresponding operation on HF sets and interpreting variables with respect to a given valuation function \lstinline!M :: 'a $\Rightarrow$ hf!.
-For the concrete definition we refer to the appendix.
+For the concrete definition we refer to formalisation (cf.\ \appendixname~\ref{appendix:semantics}).
 
@@
 The implementation of $\models$ coincides with the interpretation function of \citeauthor{lqe}~\cite{lqe}.
-As usual, a formula \lstinline!$\phi$! is called \textit{satisfiable} if there exists a model \lstinline!M! with \lstinline!M $\models$ $\phi$!.
-Otherwise, \lstinline!$\phi$! is called \textit{unsatisfiable}.
-\todo[inline]{Introduce the initial formula}
+As usual, a formula \lstinline!$\phi$! is called \emph{satisfiable} if there exists a model \lstinline!M! with \lstinline!M $\models$ $\phi$!.
+Otherwise, \lstinline!$\phi$! is called \emph{unsatisfiable}.
 
@@
 Note that formulae are added to the front of the list during branch expansion, so \lstinline!last b! for a branch \lstinline!b! is always the formula that we are trying to disprove with the tableau.
+We sometimes call this formula the \emph{initial formula}.
 \begin{lstlisting}
@@
 
-In the standard tableau calculus for propositional logic as \citeauthor{tableau}~\cite{tableau} describes it, a branch is called \textit{closed} if it contains both the negation of a formula and the formula itself;
-conversely, it is called \textit{open} if it is not closed.
+In the standard tableau calculus for propositional logic as \citeauthor{tableau}~\cite{tableau} describes it, a branch is called \emph{closed} if it contains both the negation of a formula and the formula itself;
+conversely, it is called \emph{open} if it is not closed.
 For \MLSS{}, we extend the notion of closedness with three additional rules; the first two are straightforward while the last one states that a branch is closed when the branch contains a membership cycle
@@
 \end{lstlisting}
-A tableau is called \textit{closed} if all of its branches are closed.
+A tableau is called \emph{closed} if all of its branches are closed.
 
@@
     \lstinline!s $\inS$ $\ton$ $\unionS$ $\ttw$!, \newline \lstinline!s $\notinS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$! \\
-    \lstinline!s $\notinS$ $\ton$, s $\notinS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$ $\unionS$ $\ttw$! \\
-    \end{tabular}
-    \\
+    \lstinline!s $\notinS$ $\ton$, s $\notinS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$ $\unionS$ $\ttw$! \\[0.9em]
+    \end{tabular}
     \\
@@
     \begin{tabular}{c}
-      $\vdots$ \\
+      $\vdots$
     \end{tabular}
@@
     \end{tabular}
-    \\
-    \\
+    \\[0.9em]
 
@@
   \bigskip
-  \caption{Linear expansion rules.\label{tab:lexpands}}
+  \caption{Linear expansion rules.\label{tab:lexpands} We omit the rules for $\interS$ and $\diffS$ as they coincide with the original paper (cf.\ \appendixname~\ref{appendix:rules}).}
 \end{table*}
-The calculus considers two kinds of branch expansion rules: \textit{linear} and \textit{branching} rules.
+
+The calculus considers two kinds of branch expansion rules: \emph{linear} and \emph{branching} rules.
 As the name suggests, branching rules lead to the creation of new branches in the tableau while linear rules only extend a branch \lstinline!b!
@@
 Note that in the first two rules for \lstinline!$\eqS$!, \lstinline!l! is a literal occurring in the branch.
-Furthermore, the term-for-term substitution \lstinline!l{s/t}!
-is restricted to replace \lstinline!t!
-only if it occurs as a top-level set terms of \lstinline!l!, i.e.\ it occurs directly under one of the atom constructors \lstinline!$\inS$!
-or \lstinline!$\eqS$!.
+Furthermore, the term-for-term substitution \lstinline!l{s/t}! is restricted to replace \lstinline!t! only if it occurs as a top-level set terms of \lstinline!l!.
 As an example, \lstinline!u! and \lstinline!s $\interS$ t! are top-level terms of \lstinline!u $\inS$ s $\interS$ t! but \lstinline!s! and \lstinline!t! are not.
@@
 The purpose of this restriction is to prevent unbounded expansion of the branch.
-In fact, we give an explicit upper bound for the number of formulae in a branch in \autoref{sec:bound}.
-
-Due to boundedness, repeated expansion with linear rules eventually results in a \textit{linearly saturated} branch, i.e.\ a branch where no application of linear rules would produce new formulae.
+In fact, we give an explicit upper bound for the number of formulae in a branch in \autoref{sec:correct}.
+
+Due to boundedness, repeated expansion with linear rules eventually results in a \emph{linearly saturated} branch, i.e.\ a branch where no application of linear rules would produce new formulae.
 \begin{lstlisting}
@@
 \subsection{Branching Rules\label{sec:branching}}
-\begin{table*}
+\begin{table*}[t]
   \lstset{
@@
   \bigskip
-  \caption{Branching expansion rules.\label{tab:bexpands} We write \lstinline!$\phi$! for \lstinline!last b! here.}
+  \caption{Branching expansion rules.\label{tab:bexpands} We write \lstinline!$\phi$!
+for \lstinline!last b!
+here.
+The omitted rules coincide with the original paper (cf.\ \appendixname~\ref{appendix:rules}).}
 \end{table*}
 After running out of linear rules to apply, only the branching rules shown in \autoref{tab:bexpands} remain.
-A rule is applicable if its \textit{precondition} is met and, to prevent unnecessary branching, it is not subsumed as indicated by the \textit{subsumption condition}.
+A rule is applicable if its \emph{precondition} is met and, to prevent unnecessary branching, it is not subsumed as indicated by the \emph{subsumption condition}.
 These rules create multiple branches in the tableau, so we represent the different possibilities \lstinline!bs'! to expand a branch \lstinline!b! as a set and write \lstinline!&\fexpands{bs'}{b}&!. 
@@
 
-A linearly saturated branch where no further branching is possible is called a \textit{saturated} branch. 
+A linearly saturated branch where no further branching is possible is called a \emph{saturated} branch. 
 \begin{lstlisting}
@@
 Note that even branching rules are defined such that they never create new subterms with exception of the last rule, which adds a new variable to the branch.
-These variables serve to manifest an inequality; hence, we call them \textit{witnesses}.
+These variables serve to manifest an inequality; hence, we call them \emph{witnesses}.
 \begin{lstlisting}
@@
 \section{A Decision Procedure for MLSS}
+The mechanics of the decision procedure are typical for a procedure based on a tableau calculus:
+it decides the satisfiability of a given formula \lstinline!$\phi$! by determining whether the formula has a closed tableau.
+More specifically, it initialises the tableau with the singleton branch \lstinline![$\phi$]! and checks whether this branch can be expanded to a closed tableau.
+
+We abstractly specify the behaviour of the decision procedure, leaving an executable specification from which code can be generated to future work.
+The implementation uses a couple of features of Isabelle/HOL's function package:
+instead of defining the function via pattern matching, we specify the equations of the function as conditional rewrite rules.
+This requires us to prove that the assumptions of the equations are non-overlapping, which is done by automation.
+The other concern is that Isabelle/HOL requires functions to be total, so a recursive function needs to terminate in order for it to be well-defined;
+nevertheless, the termination proof is separated from the definition of the function for modularity.
+The function package maintains the soundness of the definition by introducing a so-called domain predicate \lstinline!mlss_proc_branch_dom!
+which characterises for which arguments the function terminates.
+Each equation of the function is guarded by an assumption that the predicate holds for the argument.
+In \autoref{sec:correct}, we will show that the domain predicate holds for the context that \lstinline!mlss_proc_branch! is called in.
+Before we go into more detail on how the termination is proved, we discuss its definition as shown below.
+
 \begin{lstlisting}
@@
 
-\noindent The mechanics of the decision procedure are typical for a procedure based on a tableau calculus:
-it decides the satisfiability of a given formula \lstinline!$\phi$! by determining whether the formula has a closed tableau.
-More specifically, it initialises the tableau with the singleton branch \lstinline![$\phi$]! and checks whether this branch can be expanded to a closed tableau.
-
-We abstractly specify the behaviour of the decision procedure, leaving an executable specification from which code can be generated to future work.
-The implementation as displayed in \todo{How to reference algorithm?} uses a couple of features of Isabelle/HOL's function package:
-instead of defining the function via pattern matching, we specify the equations of the function as conditional rewrite rules.
-This requires us to prove that the assumptions of the equations are non-overlapping, which is done by automation.
-The other concern is that Isabelle/HOL requires functions to be total, so a recursive function needs to terminate in order for it to be well-defined;
-nevertheless, the termination proof is separated from the definition of the function for modularity.
-The function package maintains the soundness of the definition by introducing a so-called domain predicate \lstinline!mlss_proc_branch_dom!
-which characterises for which arguments the function terminates.
-Each equation of the function is guarded by an assumption that the predicate holds for the argument.
-In \autoref{sec:correct}, we will show that the domain predicate holds for the context that \lstinline!mlss_proc_branch! is called in.
-Before we go into more detail on how the termination is proved, we discuss its definition.
-
 The purpose of the function is to determine whether a given branch can be expanded to a closed tableau.
@@
 This means that \lstinline!mlss_proc_branch! is only applied to branches that result from applying the expansion rules.
-We call this kind of branches \textit{well-formed}.
+We call this kind of branches \emph{well-formed}.
 In the definition below, the expression \lstinline!&\expandss{b'}{b}&!
@@
 To facilitate inductive reasoning, we show a stronger statement by constructing a model \lstinline!M! such that \lstinline!M $\models$ $\phi$! for all \lstinline!$\phi$ $\in$ set b!.
-At the core of the model, there is a \textit{realisation} function that maps set terms to sets of type \lstinline!hf!.
-A subset of the witnesses, which we call \textit{pure} witnesses, receives special treatment from the realisation function for reasons that will become apparent in \autoref{sec:pwits}.
+At the core of the model, there is a \emph{realisation} function that maps set terms to sets of type \lstinline!hf!.
+A subset of the witnesses, which we call \emph{pure} witnesses, receives special treatment from the realisation function for reasons that will become apparent in \autoref{sec:pwits}.
 The collection of set terms of a branch can thus be partitioned into two collections as defined below.
@@
   pwits b $\equiv$ {c $\in$ wits b. $\forall$t $\in$ subterms (last b).
-                          AT (Var c $\eqS$ t) $\notin$ set b $\land$
-                          AT (t $\eqS$ Var c) $\notin$ set b}
+         $\phantom{\equiv}$     AT (Var c $\eqS$ t) $\notin$ set b $\land$ AT (t $\eqS$ Var c) $\notin$ set b}
 
@@
   \lstinline!e $\hfmem$ realise s! and \lstinline!e $\hfnotmem$ realise t!. 
-  Considering that \lstinline!s $\in$ subterms' b!
-and the definition of \lstinline!realise!, we obtain an \lstinline!u!
-where \lstinline!e = realise u!
-and \lstinline!u $\rightarrow_\text{bgraph b}$ s!.
-  This, in turn, yields that \lstinline!u $\inS$ s! must be in \lstinline!b!.
-  Together with the assumption \lstinline!(s $\eqS$ t) $\in$ set b! and the saturation of \lstinline!b!, it follows that \lstinline!u $\inS$ t! must also be in \lstinline!b!.
+  Considering that \lstinline!s $\in$ subterms' b! and the definition of \lstinline!realise!, we obtain a \lstinline!d! \lstinline!e = realise d! and \lstinline!d $\rightarrow_\text{bgraph b}$ s!.
+  This, in turn, yields that \lstinline!d $\inS$ s! must be in \lstinline!b!.
+  Together with the assumption \lstinline!(s $\eqS$ t) $\in$ set b! and the saturation of \lstinline!b!, it follows that \lstinline!d $\inS$ t! must also be in \lstinline!b!.
   But then we have
-  \lstinline!realise u $\hfmem$ realise t $\longleftrightarrow$ e $\hfmem$ realise t!
+  \lstinline!realise d $\hfmem$ realise t $\longleftrightarrow$ e $\hfmem$ realise t!
   using Proposition~\ref{it:at_mem}, which is a contradiction to the assumption \lstinline!e $\hfnotmem$ realise t!.
@@
 We first demonstrate the termination of the procedure for well-formed branches, i.e.\ every well-formed branch is in the domain of \lstinline!mlss_proc_branch!.
-To this end, we derive an upper bound for the number of distinct formulae in a branch whose proof we omit here for brevity (cf. Appendix\todo{Reference Appendix}).
+To this end, we derive an upper bound for the number of distinct formulae in a branch whose proof we omit here for brevity (cf.\ \appendixname~\ref{appendix:bound}).
 We should point out that this bound is not to be construed as the complexity of the procedure as it may create exponentially many branches in general.
@@
 Moreover, the subsumption conditions of the branching expansion rules ensure that each of the newly created branches contain new formulae.
-Ultimately, we conclude that the procedure must terminate because the number of formulae increases in each step but is also bounded. 
-\begin{lstlisting}
-lemma mlss_proc_branch_dom_if_wf_branch:
-  assumes wf_branch b shows mlss_proc_branch_dom b
-\end{lstlisting}
-
-The above lemma allows us to use the computation induction rule of \lstinline!mlss_proc_branch! on well-formed branches, which we use to prove soundness and completeness.
+Ultimately, we conclude that the procedure must terminate for well-formed branches because the number of formulae increases in each step but is also bounded. 
+\begin{lstlisting}
+lemma assumes wf_branch b shows mlss_proc_branch_dom b
+\end{lstlisting}
+
+The above lemma allows us to utilise the computation induction rule of \lstinline!mlss_proc_branch! on well-formed branches, which we use to prove soundness and completeness.
 As both proofs are essentially an application of soundness, respectively completeness, of the calculus, we refer the reader to the formalisation.
@@
 \section{Dealing with Urelements\label{sec:urelements}}
-In the introduction, we motivated the verification of \lstinline!mlss_proc! with the prospect of integrating it as a tactic into Isabelle.
-For this to work, we need to replay a proof in the tableau calculus through Isabelle/HOL, i.e.\ we need to map every branch expansion rule to a corresponding theorem in Isabelle/HOL.
+In the introduction, we stated the goal of integrating \lstinline!mlss_proc! as a tactic into Isabelle.
+For this to work, we need to map every branch expansion rule to a corresponding theorem in Isabelle/HOL.
 This is straightforward for all expansion rules except for the last branching expansion rule.
-To illustrate, suppose that we are to disprove a statement of the form \lstinline!s $\neq$ (t :: 'a) $\land$ s $\in$ (A :: 'a set) $\cup$ B $\land$ $\ldots$! in Isabelle/HOL.
-By way of reification, we convert this statement to \lstinline!s' $\neqS$ t' $\fmAndSymbol$ s' $\inS$ A' $\unionS$ B' $\fmAndSymbol$ $\ldots$!
+To illustrate, suppose that we are to disprove a statement of the form
+\begin{lstlisting}
+  s $\neq$ (t :: 'a) $\land$ s $\in$ (A :: 'a set) $\cup$ B $\land$ $\ldots$
+\end{lstlisting}
+in Isabelle/HOL.
+By way of reification, we convert this to a formula of the shape \lstinline!s' $\neqS$ t' $\fmAndSymbol$ s' $\inS$ A' $\unionS$ B' $\fmAndSymbol$ $\ldots$!
 in our set syntax for some \lstinline!s'!, \lstinline!t'!, \lstinline!A'!, and \lstinline!B'!.
-Applying the decision procedure to this formula might result in a tableau proof containing an application of the last branching rule rule to \lstinline!(s' $\neqS$ t') $\in$ set b!, which creates the branches \lstinline![Var x $\inS$ s', Var x $\notinS$ t'] @ b!
-and \lstinline![Var x $\notinS$ s', Var x $\inS$ t'] @ b!.
-But this can not be converted to a proof in Isabelle/HOL since \lstinline!s! and \lstinline!t! are not sets.
-
-\begin{figure}[t]
+When we apply the decision procedure to this formula, we might get back a tableau proof that contains an application of the last branching rule to \lstinline!(s' $\neqS$ t') $\in$ set b!.
+This results in two branches, one of which is \lstinline![Var x $\inS$ s', Var x $\notinS$ t'] @ b!;
+however, there is no matching rule in Isabelle/HOL since \lstinline!s! and \lstinline!t! are not sets.
+
+\begin{figure}[b]
   \centering
@@
 
-All in all, having the formalisation allows us to confidently modify the calculus as Isabelle immediately shows us where the proofs break.\todo{Longer description}
-
 The modifications above ensure that the proof can be replayed through Isabelle/HOL.
@@
 
-\section{Subterms and Subformulas}
+\section{Subterms and Subformulas\label{appendix:subterms}}
 
@@
 
-\section{Definition of the Interpretation Functions}
+\section{Definition of the Interpretation Functions\label{appendix:semantics}}
 
@@
 
-\section{Expansion Rules}
+\section{Expansion Rules\label{appendix:rules}}
 
@@
 
-\section{Upper-Bounding the Cardinality of a Branch}
-\textit{Literals} are a type of subformulae of particular importance, so we define a predicate for them.
+\section{Upper-Bounding the Cardinality of a Branch\label{appendix:bound}}
+\emph{Literals} are a type of subformulae of particular importance, so we define a predicate for them.
 \begin{lstlisting}

### 2023-02-24 paper/CADE/paper.tex
Finish paper for now

@@
   Furthermore, we extend the calculus with a light-weight type system that paves the way for an integration of the procedure into Isabelle/HOL.
-  \keywords{First TODO \and Second TODO\todo{keywords}}
+  \keywords{Decision procedures \and Tableau \and Proof assistants \and Set theory}
 \end{abstract}
@@
 The modification of the calculus required non-trivial changes to the completeness proof.
-Here, the formalisation was instrumental because Isabelle immediately revealed broken proofs that needed fixing.
+Here, the formalisation was instrumental because Isabelle immediately revealed which proofs had been broken.
 This illustrates the usefulness of ITPs for developing logic calculi: they allow us to confidently make modifications without compromising correctness.
@@
 Later extensions of the calculus added certain interpreted functions such as monotone functions~\cite{mlss_monotone_functions} and the inverse of a function~\cite{mlss_cartesian_map}.
-The latter extension notably includes the Cartesian product.
+The second extension notably includes the Cartesian product.
 Those extensions, though, did not improve upon the tableau calculus for \MLSS{}.
@@
 We overload this constant to return the set terms that are subterms of a set term, set atom, or formula, respectively.
-Lastly, we consider the subformulae of a formula and define a function \lstinline!subfms :: 'a fm $\Rightarrow$ 'a fm set!
-that computes them.
-The functions \lstinline!subterms! and \lstinline!subfms! are defined in the usual way (cf.\ \appendixname~\ref{appendix:subterms}).
+Lastly, we compute the subformulae of a formula with the function \lstinline!subfms :: 'a fm $\Rightarrow$ 'a fm set!.
+The functions \lstinline!subterms! and \lstinline!subfms! are defined in the  way (cf.\ \appendixname~\ref{appendix:subterms}).
 
@@
   \newcommand{\ttw}{\text{t}_\text{2}}
+  \caption{Linear expansion rules.\label{tab:lexpands} We omit the rules for $\interS$ and $\diffS$ as they coincide with the original paper (cf.\ \appendixname~\ref{appendix:linear_rules}).}
   \begin{tabular}{ccc}
@@
   \end{tabular}
-  \bigskip
-  \caption{Linear expansion rules.\label{tab:lexpands} We omit the rules for $\interS$ and $\diffS$ as they coincide with the original paper (cf.\ \appendixname~\ref{appendix:rules}).}
 \end{table*}
@@
 As the name suggests, branching rules lead to the creation of new branches in the tableau while linear rules only extend a branch \lstinline!b!
-with new formulae \lstinline[breaklines=true]!b' = [$\psi_1$,$\ldots$,$\psi_n$]!, which we denote by \lstinline!&\lexpands{b'}{b}&!.
+with new formulae \lstinline!b' = [$\psi_1$,$\ldots$,$\psi_n$]!, which we denote by \lstinline!&\lexpands{b'}{b}&!.
 \autoref{tab:lexpands} shows the linear expansion rules.
@@
   \newcommand{\lpipe}{\rule[-0.4ex]{0.6pt}{2.3ex}}
+  \caption{Branching expansion rules.\label{tab:bexpands} We write \lstinline!$\phi$!
+for \lstinline!last b!
+here.
+The omitted rules coincide with the original paper (cf.\ \appendixname~\ref{appendix:branching_rules}).}
   \begin{tabular}{>{\centering\arraybackslash}p{0.34\textwidth} >{\centering\arraybackslash}p{0.32\textwidth} >{\centering\arraybackslash}p{0.3\textwidth}}
@@
   \end{tabular}
-  \bigskip
-  \caption{Branching expansion rules.\label{tab:bexpands} We write \lstinline!$\phi$!
-for \lstinline!last b!
-here.
-The omitted rules coincide with the original paper (cf.\ \appendixname~\ref{appendix:rules}).}
 \end{table*}
@@
 The function package maintains the soundness of the definition by introducing a so-called domain predicate \lstinline!mlss_proc_branch_dom!
-which characterises for which arguments the function terminates.
+which characterises the arguments for which the function terminates.
 Each equation of the function is guarded by an assumption that the predicate holds for the argument.
@@
   \lstinline!e $\hfmem$ realise s! and \lstinline!e $\hfnotmem$ realise t!. 
-  Considering that \lstinline!s $\in$ subterms' b! and the definition of \lstinline!realise!, we obtain a \lstinline!d! \lstinline!e = realise d! and \lstinline!d $\rightarrow_\text{bgraph b}$ s!.
+  Considering that \lstinline!s $\in$ subterms' b!
+and the definition of \lstinline!realise!, we obtain a \lstinline!d!
+with \lstinline!e = realise d!
+and \lstinline!d $\rightarrow_\text{bgraph b}$ s!.
   This, in turn, yields that \lstinline!d $\inS$ s! must be in \lstinline!b!.
@@
 \begin{proof}
-  We only consider the case where \lstinline!&\fexpands{bs'}{b}&! was proved by applying the last rule to \lstinline!s $\neqS$ t! for some \lstinline!s! and \lstinline!t!.
+  We only consider the case where \lstinline!&\fexpands{bs'}{b}&! was proved by applying the last branching expansion rule to \lstinline!s $\neqS$ t! for some \lstinline!s! and \lstinline!t!.
   We have 
@@
 The formalisation includes an abstract description of a decision procedure that builds on the calculus.
+To make the decision procedure compatible with Isabelle/HOL, we extended the calculus with a light-weight type system while maintaining completeness.
 
@@
 
-\subsubsection{Acknowledgements}
-The author thanks Kevin Kappelmann for his comments on a draft version of the paper.
+%\subsubsection{Acknowledgements}
+%The author thanks Kevin Kappelmann for his comments on a draft version of the paper.
 
@@
 \clearpage\chapter*{Appendix}
-\todo[inline]{Clean up Appendix}
-
 \section{Subterms and Subformulas\label{appendix:subterms}}
 
-Similarly to \lstinline!vars!, we introduce the overloaded constant \lstinline!subterms!.
+We introduce the function \lstinline!subterms :: 'b $\Rightarrow$ 'a pset_term! that is overloaded in its argument.
+We first overload the function for set terms.
 \begin{lstlisting}
@@
 
-Lastly, we consider the subformulae of a formula and define a function that computes them.
+We also consider the subformulae of a formula and define a function that computes them.
 \begin{lstlisting}
@@
 
+Finally, we lift the functions \lstinline!vars! and \lstinline!subterms! to branches.
 \begin{lstlisting}
@@
 
+\clearpage
+
 \section{Definition of the Interpretation Functions\label{appendix:semantics}}
 
+We define the interpretation function \lstinline!&\Ist&! for set atoms.
 \begin{lstlisting}
@@
 
-\section{Expansion Rules\label{appendix:rules}}
+\section{Upper-Bounding the Cardinality of a Branch\label{appendix:bound}}
+The argument below argues about the \emph{literals} of a branch, which are defined as follows.
+\begin{lstlisting}
+fun is_literal :: 'a fm $\Rightarrow$ bool where
+  is_literal (&\fmAtom& _) = True | is_literal (&\fmNeg{(\fmAtom{} \_)}&) = True
+| is_literal _ = False
+\end{lstlisting}
+
+We still need to give a termination proof to ascertain total correctness.
+The core of our argument is that the number of distinct formulae that can be derived with the expansion rules is finite.
+In particular, we claim that the following upper bound holds.
+\begin{lstlisting}
+  &\phantom{$\leq$}& |set b|
+  &\textover{$=$}{$\leq$}& |{$\psi$ $\in$ set b. $\neg$ is_literal $\psi$}| +
+  &\phantom{$\leq$}& |{$\psi$ $\in$ set b. is_literal $\psi$}|
+  $\leq$ 2 * |subfms (last b)| + |{$\psi$ $\in$ set b. is_literal $\psi$}|
+  $\leq$ 2 * |subfms (last b)| + 16 * |subterms (last b)|$^4$
+\end{lstlisting}
+We treat both inequalities separately, starting with the first one whose justification hinges on an invariant: 
+every new composite subformula that is introduced by an expansion rule is the negation of a subformula that already occurs in \lstinline!last b!.
+Written as a formula, we have
+\begin{lstlisting}
+  &\phantom{$\subseteq$}& {$\psi$ $\in$ set b. $\neg$ is_literal $\psi$}
+  $\subseteq$ subfms (last b) $\cup$ $\fmNegSymbol$ ` subfms (last b).
+\end{lstlisting}
+Confirming that this is an invariant is an easy exercise of going through each of the propositional rules in \autoref{tab:lexpands} and the first two branching rules in \autoref{tab:bexpands}.
+Thus, we have
+\begin{lstlisting}
+  &\phantom{$\leq$}& |{$\psi$ $\in$ set b. $\neg$ is_literal $\psi$}|
+  $\leq$ |subfms (last b)| + |$\fmNegSymbol$ ` subfms (last b)|
+  $\leq$ 2 * |subfms (last b)|
+\end{lstlisting}
+which is the first inequality we asserted.
+
+Concerning the second inequality, first observe that each literal contains an atom, implying that the set of literals is at most twice as large as the set of atoms in a branch. 
+In addition, an atom is formed by combining two subterms occurring in the branch with either \lstinline!$\inS$! or \lstinline!$\eqS$!;
+hence, we have
+\begin{lstlisting}
+  &\phantom{$\leq$}& |{$\psi$ $\in$ set b. is_literal $\psi$}|
+  $\leq$ 2 * |$\bigcup$(atoms ` set b)| $\leq$ 2 * 2 * |subterms b|$^\text{2}$.
+\end{lstlisting}
+Since no expansion rule introduces new subterms, we know the cardinality of the subterms in a branch \lstinline!b!
+is limited by the subterms of \lstinline!last b!
+plus the number of witnesses added to the branch.
+Note that the last branching rule requires \lstinline!s $\neqS$ t! to be in \lstinline!b! for some \lstinline!s,t $\in$ subterms (last b)!.
+The subsumption condition then prohibits that the rule can be applied again for the combination of \lstinline!s!
+and \lstinline!t!, thus yielding an upper bound of \lstinline!|subterms (last b)|$^\text{2}$!
+for the witnesses.
+This allows us to finish the calculation that justifies the second inequality:
+\begin{lstlisting}
+  &\phantom{$\leq$}& 4 * |subterms b|$^\text{2}$
+  $\leq$ 4 * |subterms (last b)| + |wits b|$^\text{2}$
+  $\leq$ 4 * |subterms (last b)| + (|subterms (last b)|$^\text{2}$)$^\text{2}$
+  $\leq$ 4 * 4 * |subterms (last b)|$^\text{4}$.
+\end{lstlisting}
+All in all, we get the following lemma.
+\begin{lstlisting}[label={lst:card_branch}]
+lemma card_wf_branch_ub:
+  assumes wf_branch b
+  shows |set b|
+      $\leq$ 2 * |subfms (last b)| + 16 * |subterms (last b)|$^\text{4}$
+\end{lstlisting}
+
+\clearpage
+\section{Linear Expansion Rules\label{appendix:linear_rules}}
 
@@
   \newcommand{\ttw}{\text{t}_\text{2}}
+  \caption{Linear expansion rules.}
   \begin{tabular}{>{\centering\arraybackslash}p{0.48\textwidth}m{0.02\textwidth}>{\centering\arraybackslash}p{0.48\textwidth}}
@@
   \end{tabular}
-  \bigskip
-  \caption{Linear expansion rules}
 \end{table*}
 
-\begin{table*}
+\clearpage
+\section{Branching Expansion Rules\label{appendix:branching_rules}}
+
+\begin{table*}[h]
   \lstset{
@@
   \newcommand{\lpipe}{\rule[-0.4ex]{0.6pt}{2.3ex}}
+  \caption{Branching expansion rules. We write \lstinline!$\phi$! for \lstinline!last b! here.}
   \begin{tabular}{>{\centering\arraybackslash}p{0.34\textwidth} >{\centering\arraybackslash}p{0.32\textwidth} >{\centering\arraybackslash}p{0.3\textwidth}}
@@
   \end{tabular}
-  \bigskip
-  \caption{Branching expansion rules. We write \lstinline!$\phi$! for \lstinline!last b! here.}
 \end{table*}
 
-\section{Upper-Bounding the Cardinality of a Branch\label{appendix:bound}}
-\emph{Literals} are a type of subformulae of particular importance, so we define a predicate for them.
-\begin{lstlisting}
-fun is_literal :: 'a fm $\Rightarrow$ bool where
-  is_literal (&\fmAtom& _) = True | is_literal (&\fmNeg{(\fmAtom{} \_)}&) = True
-| is_literal _ = False
-\end{lstlisting}
-
-We still need to give a termination proof to ascertain total correctness.
-The core of our argument is that the number of distinct formulae that can be derived with the expansion rules is finite.
-In particular, we claim that the following upper bound holds.
-\begin{lstlisting}
-  &\phantom{$\leq$}& |set b|
-  &\textover{$=$}{$\leq$}& |{$\psi$ $\in$ set b. $\neg$ is_literal $\psi$}| +
-  &\phantom{$\leq$}& |{$\psi$ $\in$ set b. is_literal $\psi$}|
-  $\leq$ 2 * |subfms (last b)| + |{$\psi$ $\in$ set b. is_literal $\psi$}|
-  $\leq$ 2 * |subfms (last b)| + 16 * |subterms (last b)|$^4$
-\end{lstlisting}
-We treat both inequalities separately, starting with the first one whose justification hinges on an invariant: 
-every new composite subformula that is introduced by an expansion rule is the negation of a subformula that already occurs in \lstinline!last b!.
-Written as a formula, we have
-\begin{lstlisting}
-  &\phantom{$\subseteq$}& {$\psi$ $\in$ set b. $\neg$ is_literal $\psi$}
-  $\subseteq$ subfms (last b) $\cup$ $\fmNegSymbol$ ` subfms (last b).
-\end{lstlisting}
-Confirming that this is an invariant is an easy exercise of going through each of the propositional rules in \autoref{tab:lexpands} and the first two branching rules in \autoref{tab:bexpands}.
-Thus, we have
-\begin{lstlisting}
-  &\phantom{$\leq$}& |{$\psi$ $\in$ set b. $\neg$ is_literal $\psi$}|
-  $\leq$ |subfms (last b)| + |$\fmNegSymbol$ ` subfms (last b)|
-  $\leq$ 2 * |subfms (last b)|
-\end{lstlisting}
-which is the first inequality we asserted.
-
-Concerning the second inequality, first observe that each literal contains an atom, implying that the set of literals is at most twice as large as the set of atoms in a branch. 
-In addition, an atom is formed by combining two subterms occurring in the branch with either \lstinline!$\inS$! or \lstinline!$\eqS$!;
-hence, we have
-\begin{lstlisting}
-  &\phantom{$\leq$}& |{$\psi$ $\in$ set b. is_literal $\psi$}|
-  $\leq$ 2 * |$\bigcup$(atoms ` set b)| $\leq$ 2 * 2 * |subterms b|$^\text{2}$.
-\end{lstlisting}
-Since no expansion rule introduces new subterms, we know the cardinality of the subterms in a branch \lstinline!b!
-is limited by the subterms of \lstinline!last b!
-plus the number of witnesses added to the branch.
-Note that the last branching rule requires \lstinline!s $\neqS$ t! to be in \lstinline!b! for some \lstinline!s,t $\in$ subterms (last b)!.
-The subsumption condition then prohibits that the rule can be applied again for the combination of \lstinline!s!
-and \lstinline!t!, thus yielding an upper bound of \lstinline!|subterms (last b)|$^\text{2}$!
-for the witnesses.
-This allows us to finish the calculation that justifies the second inequality:
-\begin{lstlisting}
-  &\phantom{$\leq$}& 4 * |subterms b|$^\text{2}$
-  $\leq$ 4 * |subterms (last b)| + |wits b|$^\text{2}$
-  $\leq$ 4 * |subterms (last b)| + (|subterms (last b)|$^\text{2}$)$^\text{2}$
-  $\leq$ 4 * 4 * |subterms (last b)|$^\text{4}$.
-\end{lstlisting}
-All in all, we get the following lemma.
-\begin{lstlisting}[label={lst:card_branch}]
-lemma card_wf_branch_ub:
-  assumes wf_branch b
-  shows |set b|
-      $\leq$ 2 * |subfms (last b)| + 16 * |subterms (last b)|$^\text{4}$
-\end{lstlisting}
-
 \end{document}

### 2023-03-01 paper/CADE/paper.tex
Continue

@@
 
-\title{Towards a Verified Tableau Prover for an Unquantified Fragment of Set Theory}
+\title{Towards a Verified Tableau Prover for a Quantifier-Free Fragment of Set Theory}
 \titlerunning{Towards a Verified Tableau Prover for MLSS}
@@
 \begin{abstract}
-  Using Isabelle/HOL, we verify the state-of-the-art decision procedure for multi-level syllogistic with singleton (\MLSS{} for short), which is an unquantified fragment of set theory.
+  Using Isabelle/HOL, we verify the state-of-the-art decision procedure for multi-level syllogistic with singleton (\MLSS{} for short), which is a quantifier-free fragment of set theory.
   We formalise its syntax and semantics as well as a sound and complete tableau calculus for it.
@@
 \section{Introduction}
-In contrast to pen-and-paper proofs where we expect the reader to fill in trivial details, interactive theorem provers (ITPs) require us to justify each proof step rigorously;
-hence, ITPs usually offer a wide array of automated methods to deal with the proof goals that arise during interactive use.
-
 In Isabelle/HOL, there are specialised procedures for dealing with e.g.\ natural numbers, linear arithmetic, and metric spaces.
-Some of these procedures have been verified in Isabelle/HOL such as a procedure for Presburger arithmetic~\cite{presburger} that was later extended to mixed real-integer arithmetic~\cite{arithmetic}.
-This procedure, though, uses reflection to work on goals in Isabelle/HOL, which, during execution, either sacrifices speed by going through the simplifier or soundness by trusting generated code.
+Some of these procedures\todo{Related work?} have been verified in Isabelle/HOL such as a procedure for Presburger arithmetic~\cite{presburger} that was later extended to mixed real-integer arithmetic~\cite{arithmetic}.
+This procedure, though, uses reflection to work on goals in Isabelle/HOL, which, during execution, either sacrifices\todo{Formulation: no sacrifices but trust} speed by going through the simplifier or soundness by trusting the code generator.
 More recently, \citeauthor{orders}~\cite{orders} presented a verified decision procedure for orders that produces certificates. 
@@
 To the best of our knowledge, we present the first formally verified decision procedure for (a fragment of) set theory.
-In particular, we consider an unquantified fragment which \citeauthor{new_fast_tableau}~\cite{new_fast_tableau} call multi-level syllogistic with singleton (\MLSS{}).
+In particular, we consider a quantifier-free fragment which \citeauthor{new_fast_tableau}~\cite{new_fast_tableau} call multi-level syllogistic with singleton (\MLSS{}).
 The fragment includes the usual set operations of union, intersection, difference, membership, equality and, in addition, it allows the construction of singleton sets.
 
-Since \MLSS{} admits a tableau calculus, we expect that the generation of certificates will be straightforward.
+Since \MLSS{} admits a tableau calculus, we expect that the generation of certificates\todo{Formulation} will be straightforward.
 Like with the aforementioned order solver, this would pave the way towards an integration of the decision procedure into Isabelle, adding to its growing body of verified decision procedures.
@@
 
-We use the propositional fragment of formulae due to \citeauthor{lqe}~\cite{lqe} with set atoms as propositional atoms to form the unquantified fragment \MLSS{} of set theory.
+We use the propositional fragment of formulae due to \citeauthor{lqe}~\cite{lqe} with set atoms as propositional atoms to form the quantifier-free fragment \MLSS{} of set theory.
 \begin{lstlisting}
@@
 \section{Conclusion and Future Work}
-By closely following a paper by \citeauthor{new_fast_tableau}~\cite{new_fast_tableau}, we developed a formalisation of a tableau calculus for an unquantified fragment of set theory called \MLSS{}.
+By closely following a paper by \citeauthor{new_fast_tableau}~\cite{new_fast_tableau}, we developed a formalisation of a tableau calculus for a quantifier-free fragment of set theory called \MLSS{}.
 The formalisation includes an abstract description of a decision procedure that builds on the calculus.

### 2023-03-03 paper/CADE/paper.tex
First submitted version of paper

@@
   We formalise its syntax and semantics as well as a sound and complete tableau calculus for it.
-  We also provide an abstract specification of a decision procedure that applies the rules of the calculus exhaustively and prove its termination.
+  We also provide an executable specification of a decision procedure that applies the rules of the calculus exhaustively and prove its termination.
   Furthermore, we extend the calculus with a light-weight type system that paves the way for an integration of the procedure into Isabelle/HOL.
@@
 In Isabelle/HOL, there are specialised procedures for dealing with e.g.\ natural numbers, linear arithmetic, and metric spaces.
-Some of these procedures\todo{Related work?} have been verified in Isabelle/HOL such as a procedure for Presburger arithmetic~\cite{presburger} that was later extended to mixed real-integer arithmetic~\cite{arithmetic}.
-This procedure, though, uses reflection to work on goals in Isabelle/HOL, which, during execution, either sacrifices\todo{Formulation: no sacrifices but trust} speed by going through the simplifier or soundness by trusting the code generator.
+Some of these procedures have been verified in Isabelle/HOL such as a procedure for Presburger arithmetic~\cite{presburger} that was later extended to mixed real-integer arithmetic~\cite{arithmetic}.
+This procedure, though, uses reflection to work on goals in Isabelle/HOL, which, during execution, either sacrifices speed by going through the simplifier or requires trusting the code generator.
 More recently, \citeauthor{orders}~\cite{orders} presented a verified decision procedure for orders that produces certificates. 
@@
 
-Since \MLSS{} admits a tableau calculus, we expect that the generation of certificates\todo{Formulation} will be straightforward.
-Like with the aforementioned order solver, this would pave the way towards an integration of the decision procedure into Isabelle, adding to its growing body of verified decision procedures.
+Since \MLSS{} admits a tableau calculus, the generation of certificates will be straightforward.
+Like with the aforementioned order solver, this paves the way towards an integration of the decision procedure into Isabelle, adding to its growing body of verified decision procedures.
 
@@
 To obtain total correctness of the procedure, we prove its termination.
+We also formalise a refinement to an executable specification from which Isabelle can generate code.
 The formalisation initially follows the paper but gives a more thorough account of some important details:
@@
 This illustrates the usefulness of ITPs for developing logic calculi: they allow us to confidently make modifications without compromising correctness.
+
+All in all, the formalisation amounts to over 6000 lines of theory.
+It is available online~\cite{formalisation}.
 
@@
 
-To deal with this problem, we introduce a light-weight type system as displayed in \autoref{fig:types}. 
+To deal with this problem, we formalise a light-weight type system as displayed in \autoref{fig:types}. 
 The type of a set term in this system is just a natural number which we call level.
@@
 
-Ultimately, we define \lstinline!$\Gamma$ $\vdash$ $\psi$ $\equiv$ $\forall$a $\in$ atoms $\phi$. $\Gamma$ $\vdash$ a! in order to type formulae.
+Ultimately, we define \lstinline!$\Gamma$ $\vdash$ $\phi$ $\equiv$ $\forall$a $\in$ atoms $\phi$. $\Gamma$ $\vdash$ a! in order to type formulae.
 

### 2023-03-03 paper/CADE/paper.tex
A few amendments

@@
 To obtain total correctness of the procedure, we prove its termination.
-We also formalise a refinement to an executable specification from which Isabelle can generate code.
+In order to obtain an executable procedure, we also give a naive refinement to a more concrete specification from which code can be generated.
 The formalisation initially follows the paper but gives a more thorough account of some important details:
@@
 
-We abstractly specify the behaviour of the decision procedure, leaving an executable specification from which code can be generated to future work.
+We only discuss the abstract specification here and refer the reader to the formalisation for the executable specification.
 The implementation uses a couple of features of Isabelle/HOL's function package:
@@
 To make the decision procedure compatible with Isabelle/HOL, we extended the calculus with a light-weight type system while maintaining completeness.
-
-The refinement of the abstract description to an executable specification from which code can be exported is left for future work.
-This executable specification should generate certificates that can be replayed through Isabelle's inference kernel in order to facilitate the integration of the procedure into Isabelle.
+We also refined the abstract specification to an executable specification from which code can be generated.
+
+In future work, we plan to implement an efficient executable specification in the style of a worklist algorithm. 
+This specification should also generate certificates that can be replayed through Isabelle's inference kernel in order to facilitate the integration of the procedure into Isabelle.
 
@@
 All in all, we get the following lemma.
-\begin{lstlisting}[label={lst:card_branch}]
+\begin{lstlisting}
 lemma card_wf_branch_ub:

### 2023-03-06 paper/CADE/paper.tex
paper: AND verify

@@
 Note that constraints of the form \lstinline!l $\neq$ 0! can be replaced by \lstinline!l $=$ S i! with \lstinline!i! being a fresh variable.
-A solver for this theory is straightforward to implement;
+A solver for this theory is straightforward to implement and verify;
 nevertheless, we have to be careful that it computes the minimum assignment $\Gamma$ from variables to levels that fulfils the constraints. 

### 2023-03-06 paper/CADE/paper.tex
paper: comments by Kevin

@@
 \usepackage[hidelinks]{hyperref}
-\renewcommand{\sectionautorefname}{Section}
-\renewcommand{\subsectionautorefname}{Section}
+\usepackage[capitalise]{cleveref}
 
@@
 In Isabelle/HOL, there are specialised procedures for dealing with e.g.\ natural numbers, linear arithmetic, and metric spaces.
-Some of these procedures have been verified in Isabelle/HOL such as a procedure for Presburger arithmetic~\cite{presburger} that was later extended to mixed real-integer arithmetic~\cite{arithmetic}.
+Some of these procedures have been verified in Isabelle/HOL, such as a procedure for Presburger arithmetic~\cite{presburger} that was later extended to mixed real-integer arithmetic~\cite{arithmetic}.
 This procedure, though, uses reflection to work on goals in Isabelle/HOL, which, during execution, either sacrifices speed by going through the simplifier or requires trusting the code generator.
@@
 Later extensions of the calculus added certain interpreted functions such as monotone functions~\cite{mlss_monotone_functions} and the inverse of a function~\cite{mlss_cartesian_map}.
-The second extension notably includes the Cartesian product.
+The latter extension notably includes the Cartesian product.
 Those extensions, though, did not improve upon the tableau calculus for \MLSS{}.
@@
 
-\section{Syntax and Semantics of MLSS\label{sec:semantics}}
+\section{Syntax and Semantics of MLSS\label[section]{sec:semantics}}
 \subsection{Syntax}
@@
 Lastly, we compute the subformulae of a formula with the function \lstinline!subfms :: 'a fm $\Rightarrow$ 'a fm set!.
-The functions \lstinline!subterms! and \lstinline!subfms! are defined in the  way (cf.\ \appendixname~\ref{appendix:subterms}).
+The functions \lstinline!subterms! and \lstinline!subfms! are defined in the expected way (cf.\ \appendixname~\ref{appendix:subterms}).
 
@@
 While there is an entry~\cite{zfc_in_hol_afp} in the \emph{Archive of Formal Proofs} (AFP) that axiomatises this hierarchy,
-we instead use the hierarchy of \emph{hereditarily finite sets} (HF sets) which fulfil all the same axioms as $\mathcal{V}$ --- that is, the axioms of ZF --- except for the axiom of infinity.
+we instead use the hierarchy of \emph{hereditarily finite sets} (HF sets) which fulfil all the same axioms as $\mathcal{V}$ -- that is, the axioms of ZF -- except for the axiom of infinity.
 In particular, the membership relation is well-founded.
@@
   \newcommand{\ttw}{\text{t}_\text{2}}
-  \caption{Linear expansion rules.\label{tab:lexpands} We omit the rules for $\interS$ and $\diffS$ as they coincide with the original paper (cf.\ \appendixname~\ref{appendix:linear_rules}).}
+  \caption{Linear expansion rules.\label[table]{tab:lexpands} We omit the rules for $\interS$ and $\diffS$ as they coincide with the original paper (cf.\ \appendixname~\ref{appendix:linear_rules}).}
   \begin{tabular}{ccc}
@@
 with new formulae \lstinline!b' = [$\psi_1$,$\ldots$,$\psi_n$]!, which we denote by \lstinline!&\lexpands{b'}{b}&!.
-\autoref{tab:lexpands} shows the linear expansion rules.
+\cref{tab:lexpands} shows the linear expansion rules.
 Note that in the first two rules for \lstinline!$\eqS$!, \lstinline!l! is a literal occurring in the branch.
-Furthermore, the term-for-term substitution \lstinline!l{s/t}! is restricted to replace \lstinline!t! only if it occurs as a top-level set terms of \lstinline!l!.
+Furthermore, the term-for-term substitution \lstinline!l{s/t}! is restricted to replace \lstinline!t! only if it occurs as a top-level set term of \lstinline!l!.
 As an example, \lstinline!u! and \lstinline!s $\interS$ t! are top-level terms of \lstinline!u $\inS$ s $\interS$ t! but \lstinline!s! and \lstinline!t! are not.
@@
 The purpose of this restriction is to prevent unbounded expansion of the branch.
-In fact, we give an explicit upper bound for the number of formulae in a branch in \autoref{sec:correct}.
+In fact, we give an explicit upper bound for the number of formulae in a branch in \cref{sec:correct}.
 
@@
 
-\subsection{Branching Rules\label{sec:branching}}
+\subsection{Branching Rules\label[section]{sec:branching}}
 \begin{table*}[t]
@@
   \newcommand{\lpipe}{\rule[-0.4ex]{0.6pt}{2.3ex}}
-  \caption{Branching expansion rules.\label{tab:bexpands} We write \lstinline!$\phi$!
+  \caption{Branching expansion rules.\label[table]{tab:bexpands} We write \lstinline!$\phi$!
 for \lstinline!last b!
 here.
-The omitted rules coincide with the original paper (cf.\ \appendixname~\ref{appendix:branching_rules}).}
+The elided rules coincide with the original paper (cf.\ \appendixname~\ref{appendix:branching_rules}).}
   \begin{tabular}{>{\centering\arraybackslash}p{0.34\textwidth} >{\centering\arraybackslash}p{0.32\textwidth} >{\centering\arraybackslash}p{0.3\textwidth}}
@@
 \end{table*}
-After running out of linear rules to apply, only the branching rules shown in \autoref{tab:bexpands} remain.
+After running out of linear rules to apply, only the branching rules shown in \cref{tab:bexpands} remain.
 A rule is applicable if its \emph{precondition} is met and, to prevent unnecessary branching, it is not subsumed as indicated by the \emph{subsumption condition}.
@@
 Each equation of the function is guarded by an assumption that the predicate holds for the argument.
-In \autoref{sec:correct}, we will show that the domain predicate holds for the context that \lstinline!mlss_proc_branch! is called in.
+In \cref{sec:correct}, we will show that the domain predicate holds for the context that \lstinline!mlss_proc_branch! is called in.
 Before we go into more detail on how the termination is proved, we discuss its definition as shown below.
@@
 
-In particular, we will use this notion in \autoref{sec:correct} to state an upper bound for the cardinality of well-formed branches, with which we justify the termination of the decision procedure. 
+In particular, we use this notion in \cref{sec:correct} to state an upper bound for the cardinality of well-formed branches.
+The upper bound justifies the termination of the decision procedure. 
 Before we come to that, though, we prove soundness and completeness in Section~\ref{sec:sound} and~\ref{sec:complete}, respectively.
-In \autoref{sec:correct}, we also show that both properties easily transfer to \lstinline!mlss_proc! which, together with termination, establishes that it is a decision procedure.
-
-\section{Completeness of the Calculus\label{sec:complete}}
+In \cref{sec:correct}, we also show that both properties easily transfer to \lstinline!mlss_proc! which, together with termination, establishes that it is a decision procedure.
+
+\section{Completeness of the Calculus\label[section]{sec:complete}}
 For completeness of the calculus, we need to show that every unsatisfiable formula has a closed tableau or, conversely, that the formula is satisfiable if there is a saturated and open branch in the tableau.
@@
 At the core of the model, there is a \emph{realisation} function that maps set terms to sets of type \lstinline!hf!.
-A subset of the witnesses, which we call \emph{pure} witnesses, receives special treatment from the realisation function for reasons that will become apparent in \autoref{sec:pwits}.
+A subset of the witnesses, which we call \emph{pure} witnesses, receives special treatment from the realisation function for reasons that will become apparent in \cref{sec:pwits}.
 The collection of set terms of a branch can thus be partitioned into two collections as defined below.
@@
 
-Before we prove that the realisation function constitutes a model in \autoref{sec:real}, we will first explain the significance of the pure witnesses.
-
-\subsection{Characterisation of the Pure Witnesses\label{sec:pwits}}
+Before we prove that the realisation function constitutes a model in \cref{sec:real}, we will first explain the significance of the pure witnesses.
+
+\subsection{Characterisation of the Pure Witnesses\label[section]{sec:pwits}}
 Recall that the pure witnesses of a branch \lstinline!b! are those witnesses that are not related to other subterms in \lstinline!last b! by equality.
@@
 
-\subsection{Realisation of an Open Branch\label{sec:real}}
+\subsection{Realisation of an Open Branch\label[section]{sec:real}}
 Remember that for completeness, we need to show that the realisation function for an open and saturated branch \lstinline!b! actually constitutes a model for all formulae in the branch.
@@
 
-The final step to obtain a proper model is to connect the realisation function to the semantics as defined in \autoref{sec:semantics}.
+The final step to obtain a proper model is to connect the realisation function to the semantics as defined in \cref{sec:semantics}.
 For set terms, we can use the Propositions~\ref{it:empty}--\ref{it:single} to prove the lemma below by induction on \lstinline!t!.
@@
 
-\section{Soundness of the Calculus\label{sec:sound}}
+\section{Soundness of the Calculus\label[section]{sec:sound}}
 A tableau calculus is sound if for any closed tableau, the corresponding formula is unsatisfiable. 
@@
 
-\section{Total Correctness of the Decision Procedure\label{sec:correct}}
+\section{Total Correctness of the Decision Procedure\label[section]{sec:correct}}
 We first demonstrate the termination of the procedure for well-formed branches, i.e.\ every well-formed branch is in the domain of \lstinline!mlss_proc_branch!.
@@
 
-\section{Dealing with Urelements\label{sec:urelements}}
+\section{Dealing with Urelements\label[section]{sec:urelements}}
 In the introduction, we stated the goal of integrating \lstinline!mlss_proc! as a tactic into Isabelle.
@@
   \end{tabular}
-  \caption{The type system for set terms and atoms.\label{fig:types}}
+  \caption{The type system for set terms and atoms.\label[figure]{fig:types}}
 \end{figure}
 
-To deal with this problem, we formalise a light-weight type system as displayed in \autoref{fig:types}. 
+To deal with this problem, we formalise a light-weight type system as displayed in \cref{fig:types}. 
 The type of a set term in this system is just a natural number which we call level.
@@
 nevertheless, we have to be careful that it computes the minimum assignment $\Gamma$ from variables to levels that fulfils the constraints. 
-This is to guarantee that a set term \lstinline!t! is not an urelement \textbf{iff} \lstinline!$\Gamma$ t $>$ 0!.
+This is to guarantee that a set term \lstinline!t! is not an urelement if, and only if, \lstinline!$\Gamma$ t $>$ 0!.
 Conversely, all terms \lstinline!s! with \lstinline!$\Gamma$ s $=$ 0! are urelements.
@@
 \section{Conclusion and Future Work}
-By closely following a paper by \citeauthor{new_fast_tableau}~\cite{new_fast_tableau}, we developed a formalisation of a tableau calculus for a quantifier-free fragment of set theory called \MLSS{}.
+We developed a formalisation of a tableau calculus for a quantifier-free fragment of set theory called \MLSS{} based on a paper by \citeauthor{new_fast_tableau}~\cite{new_fast_tableau}. 
 The formalisation includes an abstract description of a decision procedure that builds on the calculus.
@@
 \end{lstlisting}
-Confirming that this is an invariant is an easy exercise of going through each of the propositional rules in \autoref{tab:lexpands} and the first two branching rules in \autoref{tab:bexpands}.
+Confirming that this is an invariant is an easy exercise of going through each of the propositional rules in \cref{tab:lexpands} and the first two branching rules in \cref{tab:bexpands}.
 Thus, we have

### 2023-05-23 paper/CADE/paper.tex
Start working in feedback

@@
   We also provide an executable specification of a decision procedure that applies the rules of the calculus exhaustively and prove its termination.
-  Furthermore, we extend the calculus with a light-weight type system that paves the way for an integration of the procedure into Isabelle/HOL.
+  Furthermore, we extend the calculus with a lightweight type system that paves the way for an integration of the procedure into Isabelle/HOL.
   \keywords{Decision procedures \and Tableau \and Proof assistants \and Set theory}
@@
   \item We deliver the omitted proof of Lemma~2 in the paper~\cite{new_fast_tableau}, which is a key building block for the completeness proof of the calculus.
-  \item The formal proof of completeness lead to the discovery that the calculus was missing a rule for eliminating double negation. 
-  \item We derive an explicit upper bound for the number of formulae in a branch of the tableau.
+  \item The formal proof of completeness leads to the discovery that the calculus was missing a rule for eliminating double negation. 
+  \item We derive an explicit upper bound for the number of formulas in a branch of the tableau.
 \end{itemize}
@@
 however, this is not the case in Isabelle/HOL, e.g.\ consider the expression \lstinline!n $\in$ A! where \lstinline!n! is a natural number.
-In order to deal with these urelements, we extend the calculus with a light-weight type system and a verified inference algorithm that identifies the urelements.
+We call these variables urelements.
+In order to deal with them, we extend the calculus with a lightweight type system and a verified inference algorithm that identifies the urelements.
 
@@
 \citeauthor{tableau_quantifier_free} also cover an extension of the calculus with uninterpreted functions, which was revisited by \citeauthor{mlss_quantification}~\cite{mlss_quantification} who avoided the construction of a model during execution.
-In this paper, we consider a version of the latter procedure due to \citeauthor{new_fast_tableau}~\cite{new_fast_tableau} that is specialised to \MLSS{} and where the branching rules of the calculus are set up in a way to guarantee mutual exclusivity of the branches.
+In this paper, we consider a version of the latter procedure due to \citeauthor{new_fast_tableau}~\cite{new_fast_tableau} that is specialised to \MLSS{} and where the branching rules of the calculus are set up in a way that guarantees mutual exclusivity of the branches.
 Later extensions of the calculus added certain interpreted functions such as monotone functions~\cite{mlss_monotone_functions} and the inverse of a function~\cite{mlss_cartesian_map}.
@@
 Isabelle/HOL's data type package automatically defines a function that gives us the set of variables in a set term, which we name \lstinline!vars!.
-In what follows, we will overload the function \lstinline!vars! to also work on set atoms, formulae, and branches.
+In what follows, we will overload the function \lstinline!vars! to also work on set atoms, formulas, and branches.
 \begin{lstlisting}
@@
 
-We use the propositional fragment of formulae due to \citeauthor{lqe}~\cite{lqe} with set atoms as propositional atoms to form the quantifier-free fragment \MLSS{} of set theory.
+We use the propositional fragment of formulas due to \citeauthor{lqe}~\cite{lqe} with set atoms as propositional atoms to form the quantifier-free fragment \MLSS{} of set theory.
 \begin{lstlisting}
@@
 
-Likewise, we fix the constant \lstinline!subterms :: 'b $\Rightarrow$ 'a pset_term set!.
+Likewise, we fix the constant \lstinline!subterms :: 'b $\Rightarrow$ 'a pset_term set!.\todo{Explain b'}
 We overload this constant to return the set terms that are subterms of a set term, set atom, or formula, respectively.
-Lastly, we compute the subformulae of a formula with the function \lstinline!subfms :: 'a fm $\Rightarrow$ 'a fm set!.
+Lastly, we compute the subformulas of a formula with the function \lstinline!subfms :: 'a fm $\Rightarrow$ 'a fm set!.
 The functions \lstinline!subterms! and \lstinline!subfms! are defined in the expected way (cf.\ \appendixname~\ref{appendix:subterms}).
@@
 In particular, the membership relation is well-founded.
-The HF sets are sufficient for our purposes as \MLSS{} is not powerful enough to force infinite models~\cite[Chapter~14.2]{set_theory}.
+The HF sets are sufficient for our purposes as \MLSS{} is not powerful enough to force infinite models~\cite[Chapter~14.2]{set_theory}.\cite{Argue power: self-evident}
 In contrast to $\mathcal{V}$, the HF sets are directly representable in Isabelle/HOL, and indeed, there is an AFP entry~\cite{hf_AFP} that formalises them.
@@
 Inspired by the formalisation of a tableau calculus for hybrid logic by \citeauthor{hybrid_logic_afp}~\cite{hybrid_logic_afp}, we simply use a list to represent a branch of the tableau tree.
-Note that formulae are added to the front of the list during branch expansion, so \lstinline!last b! for a branch \lstinline!b! is always the formula that we are trying to disprove with the tableau.
+Note that formulas are added to the front of the list during branch expansion, so \lstinline!last b! for a branch \lstinline!b! is always the formula that we are trying to disprove with the tableau.
 We sometimes call this formula the \emph{initial formula}.
@@
 \end{lstlisting}
-A tableau is called \emph{closed} if all of its branches are closed.
+A tableau is called \emph{closed} if all of its branches are closed.\todo{What is a tableau?}
 
@@
   \newcommand{\ttw}{\text{t}_\text{2}}
-  \caption{Linear expansion rules.\label[table]{tab:lexpands} We omit the rules for $\interS$ and $\diffS$ as they coincide with the original paper (cf.\ \appendixname~\ref{appendix:linear_rules}).}
+  \caption{Linear expansion rules.\label[table]{tab:lexpands} We omit the rules for $\interS$ and $\diffS$ as they coincide with the original paper~\cite{new_fast_tableau} (cf.\ \appendixname~\ref{appendix:linear_rules}).}
   \begin{tabular}{ccc}
@@
 As the name suggests, branching rules lead to the creation of new branches in the tableau while linear rules only extend a branch \lstinline!b!
-with new formulae \lstinline!b' = [$\psi_1$,$\ldots$,$\psi_n$]!, which we denote by \lstinline!&\lexpands{b'}{b}&!.
+with new formulas \lstinline!b' = [$\psi_1$,$\ldots$,$\psi_n$]!, which we denote by \lstinline!&\lexpands{b'}{b}&!.
 \cref{tab:lexpands} shows the linear expansion rules.
@@
 
-The more crucial restriction of the linear rules is that that no new subterm may be created by their application;
+A more crucial restriction of the linear rules is that no new subterm may be created by their application;
 for instance, the second rule for \lstinline!$\unionS$! is
@@
 The purpose of this restriction is to prevent unbounded expansion of the branch.
-In fact, we give an explicit upper bound for the number of formulae in a branch in \cref{sec:correct}.
-
-Due to boundedness, repeated expansion with linear rules eventually results in a \emph{linearly saturated} branch, i.e.\ a branch where no application of linear rules would produce new formulae.
+In fact, we give an explicit upper bound for the number of formulas in a branch in \cref{sec:correct}.
+
+Due to boundedness, repeated expansion with linear rules eventually results in a \emph{linearly saturated} branch, i.e.\ a branch where no application of linear rules would produce new formulas.
 \begin{lstlisting}
@@
 \lstinline![&\fmNeg{\fmNeg{\fmNeg{p}}}&, p, &\fmAnd{\fmNeg{\fmNeg{\fmNeg{p}}}}{p}&]!
-is saturated --- neither linear nor branching rules apply --- and open, but there clearly is no model for the initial formula \lstinline!&\fmAnd{\fmNeg{\fmNeg{\fmNeg{p}}}}{p}&!.
+is saturated --- neither linear nor branching rules apply --- and open, but there clearly is no model for the initial formula \lstinline!&\fmAnd{\fmNeg{\fmNeg{\fmNeg{p}}}}{p}&!.\todo{Split into multiple sentences}
 
@@
 here.
-The elided rules coincide with the original paper (cf.\ \appendixname~\ref{appendix:branching_rules}).}
+The elided rules coincide with the original paper~\cite{new_fast_tableau} (cf.\ \appendixname~\ref{appendix:branching_rules}).}
   \begin{tabular}{>{\centering\arraybackslash}p{0.34\textwidth} >{\centering\arraybackslash}p{0.32\textwidth} >{\centering\arraybackslash}p{0.3\textwidth}}
@@
 After running out of linear rules to apply, only the branching rules shown in \cref{tab:bexpands} remain.
-A rule is applicable if its \emph{precondition} is met and, to prevent unnecessary branching, it is not subsumed as indicated by the \emph{subsumption condition}.
+A rule is applicable if its \emph{precondition} is met and, to prevent unnecessary branching, if it is not subsumed as indicated by the \emph{subsumption condition}.
 These rules create multiple branches in the tableau, so we represent the different possibilities \lstinline!bs'! to expand a branch \lstinline!b! as a set and write \lstinline!&\fexpands{bs'}{b}&!. 
-Accordingly, we get a new branch \lstinline!b' @ b! in the tableau for each \lstinline!b' $\in$ set bs'!.
+Accordingly, we get a new branch \lstinline!b' @ b! in the tableau for each \lstinline!b' $\in$ bs'!.
 
@@
 Each equation of the function is guarded by an assumption that the predicate holds for the argument.
-In \cref{sec:correct}, we will show that the domain predicate holds for the context that \lstinline!mlss_proc_branch! is called in.
-Before we go into more detail on how the termination is proved, we discuss its definition as shown below.
+In \cref{sec:correct}, we will show that the domain predicate holds for the context that the function \lstinline!mlss_proc_branch! is called in.
+Before we go into more detail on how the termination is proved, we discuss the definition of the function as shown below.
 
@@
 to this end, we recursively expand the branch with linear rules until the branch is linearly saturated.
-Note that we use Hilbert's $\varepsilon$-operator in the form of \lstinline!SOME! to choose some rule that actually adds new formulae to the branch. 
+Note that we use Hilbert's $\varepsilon$-operator in the form of \lstinline!SOME! to choose some rule that actually adds new formulas to the branch. 
 As soon as the branch is linearly saturated, we terminate if the branch is closed as shown in the second equation.
@@
 \subsection{Realisation of an Open Branch\label[section]{sec:real}}
-Remember that for completeness, we need to show that the realisation function for an open and saturated branch \lstinline!b! actually constitutes a model for all formulae in the branch.
+Remember that for completeness, we need to show that the realisation function for an open and saturated branch \lstinline!b! actually constitutes a model for all formulas in the branch.
 We start by verifying that the realisation function models all literals in the branch; more formally, the following propositions hold:
@@
 \end{lstlisting}
-Lifting the above result to formulae yields the coherence of \lstinline!b! as the original paper~\cite{new_fast_tableau} calls it.
-The proof is a tedious but straightforward induction on the the size of the formulae.
+Lifting the above result to formulas yields the coherence of \lstinline!b! as the original paper~\cite{new_fast_tableau} calls it.
+The proof is a tedious but straightforward induction on the the size of the formulas.
 \begin{lstlisting}[label={lst:coherence}]
@@
   We update \lstinline!M! such that it maps \lstinline!x! to \lstinline!y! to obtain the assignment \lstinline!M'!.
-  Note that \lstinline!M'! is still a model for formulae in \lstinline!b! because \lstinline!x! is fresh with respect to \lstinline!b!.
+  Note that \lstinline!M'! is still a model for formulas in \lstinline!b! because \lstinline!x! is fresh with respect to \lstinline!b!.
   Furthermore, it is also a model for the first branch in \lstinline!bs'!, which finishes the proof.
@@
 We first demonstrate the termination of the procedure for well-formed branches, i.e.\ every well-formed branch is in the domain of \lstinline!mlss_proc_branch!.
-To this end, we derive an upper bound for the number of distinct formulae in a branch whose proof we omit here for brevity (cf.\ \appendixname~\ref{appendix:bound}).
+To this end, we derive an upper bound for the number of distinct formulas in a branch whose proof we omit here for brevity (cf.\ \appendixname~\ref{appendix:bound}).
 We should point out that this bound is not to be construed as the complexity of the procedure as it may create exponentially many branches in general.
@@
 \end{lstlisting}
-Remember that \lstinline!mlss_proc_branch! only applies a linear expansion rule to a branch if the application results in new formulae.   
-Moreover, the subsumption conditions of the branching expansion rules ensure that each of the newly created branches contain new formulae.
-Ultimately, we conclude that the procedure must terminate for well-formed branches because the number of formulae increases in each step but is also bounded. 
+Remember that \lstinline!mlss_proc_branch! only applies a linear expansion rule to a branch if the application results in new formulas.   
+Moreover, the subsumption conditions of the branching expansion rules ensure that each of the newly created branches contain new formulas.
+Ultimately, we conclude that the procedure must terminate for well-formed branches because the number of formulas increases in each step but is also bounded. 
 \begin{lstlisting}
@@
 
-To deal with this problem, we formalise a light-weight type system as displayed in \cref{fig:types}. 
+To deal with this problem, we formalise a lightweight type system as displayed in \cref{fig:types}. 
 The type of a set term in this system is just a natural number which we call level.
-Intuitively speaking, the level indicates that the corresponding term \lstinline!t! in Isabelle/HOL has type
+Intuitively speaking, the level \texttt{l} means that the corresponding term \lstinline!t! in Isabelle/HOL has type
 \[
@@
 
-Ultimately, we define \lstinline!$\Gamma$ $\vdash$ $\phi$ $\equiv$ $\forall$a $\in$ atoms $\phi$. $\Gamma$ $\vdash$ a! in order to type formulae.
+Ultimately, we define \lstinline!$\Gamma$ $\vdash$ $\phi$ $\equiv$ $\forall$a $\in$ atoms $\phi$. $\Gamma$ $\vdash$ a! in order to type formulas.
 
@@
 To actually use the calculus, we need to determine the urelements of the initial formula $\phi$, though.
-In other words, we need to implement an inference algorithm for our light-weight type system.
+In other words, we need to implement an inference algorithm for our lightweight type system.
 The algorithm is, in essence, a simplified version of Hindley-Milney type inference so it has the same two phases:
@@
 The formalisation includes an abstract description of a decision procedure that builds on the calculus.
-To make the decision procedure compatible with Isabelle/HOL, we extended the calculus with a light-weight type system while maintaining completeness.
+To make the decision procedure compatible with Isabelle/HOL, we extended the calculus with a lightweight type system while maintaining completeness.
 We also refined the abstract specification to an executable specification from which code can be generated.
@@
 \clearpage\chapter*{Appendix}
+\todo[inline]{Remove appendix}
 \section{Subterms and Subformulas\label{appendix:subterms}}
@@
 \end{lstlisting}
-Finally, we lift this to formulae by using \lstinline!atoms! again.
+Finally, we lift this to formulas by using \lstinline!atoms! again.
 \begin{lstlisting}
@@
 
-We also consider the subformulae of a formula and define a function that computes them.
+We also consider the subformulas of a formula and define a function that computes them.
 \begin{lstlisting}
@@
 We still need to give a termination proof to ascertain total correctness.
-The core of our argument is that the number of distinct formulae that can be derived with the expansion rules is finite.
+The core of our argument is that the number of distinct formulas that can be derived with the expansion rules is finite.
 In particular, we claim that the following upper bound holds.

### 2023-05-23 paper/CADE/paper.tex
Realise reviewer comments

@@
 \bibliographystyle{splncs04nat}
-
-\usepackage{appendix}
 
@@
   We formalise its syntax and semantics as well as a sound and complete tableau calculus for it.
-  We also provide an executable specification of a decision procedure that applies the rules of the calculus exhaustively and prove its termination.
+  We also provide an executable specification of a decision procedure that exhaustively applies the rules of the calculus and prove its termination.
   Furthermore, we extend the calculus with a lightweight type system that paves the way for an integration of the procedure into Isabelle/HOL.
-  \keywords{Decision procedures \and Tableau \and Proof assistants \and Set theory}
+  \keywords{Decision procedures \and Semantic tableaux \and Interactive theorem proving \and Set theory}
 \end{abstract}
@@
 We present a formalisation in Isabelle/HOL of a tableau calculus for \MLSS{} due to \citeauthor{new_fast_tableau}~\cite{new_fast_tableau}\cite[Chapter 14]{set_theory}.
-We prove soundness and completeness of the calculus and give an abstract specification of a decision procedure that applies the rules of the calculus exhaustively.
+We prove soundness and completeness of the calculus and give an abstract specification of a decision procedure that exhaustively applies the rules of the calculus.
 To obtain total correctness of the procedure, we prove its termination.
-In order to obtain an executable procedure, we also give a naive refinement to a more concrete specification from which code can be generated.
-The formalisation initially follows the paper but gives a more thorough account of some important details:
+Additionally, we naively refine the abstract to an executable specification from which we can generate code.
+The formalisation initially follows the paper but offers a more thorough account of some important details:
 \begin{itemize}
@@
 All in all, the formalisation amounts to over 6000 lines of theory.
-It is available online~\cite{formalisation}.
+It is part of the \emph{Archive of Formal Proofs} (AFP)~\cite{formalisation}.
+The entry provides an overview theory \href{https://www.isa-afp.org/theories/mlss_decision_proc/#MLSS_Proc_All}{\texttt{MLSS\_Proc\_All.thy}} that highlights the (mostly syntactic) differences between paper and formalisation and references the constants and theorems that are introduced in this paper.
 
@@
     $\emptyset$ | Var 'a | Single ('a pset_term)
-| 'a pset_term $\unionS$ 'a pset_term | 'a pset_term $\interS$ 'a pset_term
+  | 'a pset_term $\unionS$ 'a pset_term
+  | 'a pset_term $\interS$ 'a pset_term
   | 'a pset_term $\diffS$ 'a pset_term
@@
 datatype (vars: 'a) pset_atom =
-  'a pset_term $\inS$ 'a pset_term | 'a pset_term $\eqS$ 'a pset_term
+    'a pset_term $\inS$ 'a pset_term
+  | 'a pset_term $\eqS$ 'a pset_term
 \end{lstlisting}
@@
 datatype (atoms: 'a) fm =
-  &\fmAtom& 'a | &\fmNeg{('a fm)}& | &\fmAnd{'a fm}{'a fm}& | &\fmOr{'a fm}{'a fm}&
+    &\fmAtom& 'a
+  | &\fmNeg{('a fm)}&
+  | &\fmAnd{'a fm}{'a fm}&
+  | &\fmOr{'a fm}{'a fm}&
 
@@
 
-Likewise, we fix the constant \lstinline!subterms :: 'b $\Rightarrow$ 'a pset_term set!.\todo{Explain b'}
+Likewise, we fix the constant \lstinline!subterms :: 'b $\Rightarrow$ 'a pset_term set! that is polymorphic in its argument type \lstinline!'b!.
 We overload this constant to return the set terms that are subterms of a set term, set atom, or formula, respectively.
-Lastly, we compute the subformulas of a formula with the function \lstinline!subfms :: 'a fm $\Rightarrow$ 'a fm set!.
-The functions \lstinline!subterms! and \lstinline!subfms! are defined in the expected way (cf.\ \appendixname~\ref{appendix:subterms}).
+Lastly, we introduce the function \lstinline!subfms :: 'a fm $\Rightarrow$ 'a fm set! that computes the subformulas of a formula.
+The functions \lstinline!subterms! and \lstinline!subfms! are implemented in the expected way.
 
@@
 The original paper~\cite{new_fast_tableau} bases the semantics of \MLSS{} on the von Neumann hierarchy of sets $\mathcal{V}$.
-While there is an entry~\cite{zfc_in_hol_afp} in the \emph{Archive of Formal Proofs} (AFP) that axiomatises this hierarchy,
-we instead use the hierarchy of \emph{hereditarily finite sets} (HF sets) which fulfil all the same axioms as $\mathcal{V}$ -- that is, the axioms of ZF -- except for the axiom of infinity.
+We instead use the hierarchy of \emph{hereditarily finite sets} (HF sets) which fulfil all the same axioms as $\mathcal{V}$ -- that is, the axioms of ZF -- except for the axiom of infinity.
 In particular, the membership relation is well-founded.
-The HF sets are sufficient for our purposes as \MLSS{} is not powerful enough to force infinite models~\cite[Chapter~14.2]{set_theory}.\cite{Argue power: self-evident}
+The HF sets, as we will see, are sufficient to construct a model for any satisfiable \MLSS{} formula.
 In contrast to $\mathcal{V}$, the HF sets are directly representable in Isabelle/HOL, and indeed, there is an AFP entry~\cite{hf_AFP} that formalises them.
@@
 in the standard way, i.e.\ by mapping each syntactic construct to the corresponding operation on HF sets and interpreting variables with respect to a given valuation function \lstinline!M :: 'a $\Rightarrow$ hf!.
-For the concrete definition we refer to formalisation (cf.\ \appendixname~\ref{appendix:semantics}).
+For the concrete definition we refer to formalisation.
 
@@
 We formalise the tableau calculus for \MLSS{} as described by \citeauthor{new_fast_tableau}~\cite{new_fast_tableau}.
-Inspired by the formalisation of a tableau calculus for hybrid logic by \citeauthor{hybrid_logic_afp}~\cite{hybrid_logic_afp}, we simply use a list to represent a branch of the tableau tree.
+Inspired by the formalisation of a tableau calculus for hybrid logic by \citeauthor{hybrid_logic_afp}~\cite{hybrid_logic_afp}, we simply use lists to represent the branches of the tableau tree.
 Note that formulas are added to the front of the list during branch expansion, so \lstinline!last b! for a branch \lstinline!b! is always the formula that we are trying to disprove with the tableau.
@@
 \end{lstlisting}
-A tableau is called \emph{closed} if all of its branches are closed.\todo{What is a tableau?}
+A tableau is called \emph{closed} if all of its branches are closed.
 
@@
   \newcommand{\ttw}{\text{t}_\text{2}}
-  \caption{Linear expansion rules.\label[table]{tab:lexpands} We omit the rules for $\interS$ and $\diffS$ as they coincide with the original paper~\cite{new_fast_tableau} (cf.\ \appendixname~\ref{appendix:linear_rules}).}
+  \caption{Linear expansion rules.\label[table]{tab:lexpands} All rules except the double negation rule coincide with the original paper~\cite{new_fast_tableau}.
+For brevity, we omit the rules for $\interS$ and $\diffS$.}
   \begin{tabular}{ccc}
@@
 Note that in the first two rules for \lstinline!$\eqS$!, \lstinline!l! is a literal occurring in the branch.
-Furthermore, the term-for-term substitution \lstinline!l{s/t}! is restricted to replace \lstinline!t! only if it occurs as a top-level set term of \lstinline!l!.
-As an example, \lstinline!u! and \lstinline!s $\interS$ t! are top-level terms of \lstinline!u $\inS$ s $\interS$ t! but \lstinline!s! and \lstinline!t! are not.
-
+Furthermore, the term-for-term substitution \lstinline!l{s/t}!
+is restricted to the top-level set terms of \lstinline!l!, i.e.\ the set terms that occur directly under one of the atom constructors \lstinline!$\inS$!
+or \lstinline!$\eqS$!;
+for example, given the literal
+\begin{lstlisting}
+  l = &\fmNeg{((s $\unionS$ u) $\diffS$ s $\eqS$ s $\unionS$ u)}&
+\end{lstlisting}
+we have
+\begin{lstlisting}
+  &\phantom{=}& (&\fmNeg{((s $\unionS$ u) $\diffS$ s $\eqS$ s $\unionS$ u)}&){t/s $\unionS$ u}
+  = &\fmNeg{((s $\unionS$ u) $\diffS$ s $\eqS$ t)}&.
+\end{lstlisting}
 A more crucial restriction of the linear rules is that no new subterm may be created by their application;
 for instance, the second rule for \lstinline!$\unionS$! is
-\lstinline!s $\inS$ t$_\text{1}$ $\Longrightarrow$ s $\inS$ t$_\text{1}$ $\unionS$ t$_\text{2}$!,
+\begin{lstlisting}
+  s $\inS$ t$_\text{1}$ $\Longrightarrow$ s $\inS$ t$_\text{1}$ $\unionS$ t$_\text{2}$,
+\end{lstlisting}
 which formally represents 
@@
 \end{lstlisting}
-may only be used under the condition
+and may only be used under the condition
 \lstinline[breaklines]!t$_\text{1}$ $\unionS$ t$_\text{2}$ $\in$ subterms (last b)!.
@@
 \lstinline![&\fmNeg{\fmNeg{\fmNeg{p}}}&, p, &\fmAnd{\fmNeg{\fmNeg{\fmNeg{p}}}}{p}&]!
-is saturated --- neither linear nor branching rules apply --- and open, but there clearly is no model for the initial formula \lstinline!&\fmAnd{\fmNeg{\fmNeg{\fmNeg{p}}}}{p}&!.\todo{Split into multiple sentences}
+is saturated --- neither linear nor branching rules apply --- and open, but there clearly is no model for the initial formula \lstinline!&\fmAnd{\fmNeg{\fmNeg{\fmNeg{p}}}}{p}&!.
 
@@
 here.
-The elided rules coincide with the original paper~\cite{new_fast_tableau} (cf.\ \appendixname~\ref{appendix:branching_rules}).}
+All rules coincide with the original paper~\cite{new_fast_tableau} so we only show an illustrative subset.}
   \begin{tabular}{>{\centering\arraybackslash}p{0.34\textwidth} >{\centering\arraybackslash}p{0.32\textwidth} >{\centering\arraybackslash}p{0.3\textwidth}}
@@
 to this end, we recursively expand the branch with linear rules until the branch is linearly saturated.
-Note that we use Hilbert's $\varepsilon$-operator in the form of \lstinline!SOME! to choose some rule that actually adds new formulas to the branch. 
+Note that we use Hilbert's $\varepsilon$-operator in the form of \lstinline!SOME!\footnote{In the formalisation, the function \lstinline!mlss_proc_branch!
+is actually parametrised by choice functions to allow for refinement.} to choose some rule that actually adds new formulas to the branch.
 As soon as the branch is linearly saturated, we terminate if the branch is closed as shown in the second equation.
@@
 \begin{enumerate*}[label={(\arabic*)}]
-  \item A closed branch contains an unsatisfiable formula.
+  \item It is not possible to satisfy all formulas in a closed branch at once. 
   \item The expansion rules maintain satisfiability.
@@
 We first demonstrate the termination of the procedure for well-formed branches, i.e.\ every well-formed branch is in the domain of \lstinline!mlss_proc_branch!.
-To this end, we derive an upper bound for the number of distinct formulas in a branch whose proof we omit here for brevity (cf.\ \appendixname~\ref{appendix:bound}).
+To this end, we derive an upper bound for the number of distinct formulas in a branch whose proof we omit here for brevity.
 We should point out that this bound is not to be construed as the complexity of the procedure as it may create exponentially many branches in general.
@@
 
-%\subsubsection{Acknowledgements}
-%The author thanks Kevin Kappelmann for his comments on a draft version of the paper.
+\subsubsection{Acknowledgements}
+The author thanks Kevin Kappelmann, Tobias Nipkow, and the anonymous referees for their comments on a draft version of this paper.
 
@@
 
-%\clearpage\listoftodos
-\appendix
-
-\clearpage\chapter*{Appendix}
-\todo[inline]{Remove appendix}
-\section{Subterms and Subformulas\label{appendix:subterms}}
-
-We introduce the function \lstinline!subterms :: 'b $\Rightarrow$ 'a pset_term! that is overloaded in its argument.
-We first overload the function for set terms.
-\begin{lstlisting}
-fun subterms :: 'a pset_term $\Rightarrow$ 'a pset_term set where
-  subterms $\emptyset$ = {$\emptyset$}
-| subterms (Var x) = {Var x}
-| subterms (s $\unionS$ t) = {s $\unionS$ t} $\cup$ subterms s $\cup$ subterms t
-| subterms (s $\interS$ t) = {s $\interS$ t} $\cup$ subterms s $\cup$ subterms t
-| subterms (s $\diffS$ t) = {s $\diffS$ t} $\cup$ subterms s $\cup$ subterms t
-| subterms (Single t) = {Single t} $\cup$ subterms t
-\end{lstlisting}
-The subterms of an atom are then the subterms of its arguments.
-\begin{lstlisting}
-fun subterms :: 'a pset_atom $\Rightarrow$ 'a pset_term set where
-  subterms (s $\inS$ t) = subterms s $\cup$ subterms t
-| subterms (s $\eqS$ t) = subterms s $\cup$ subterms t
-\end{lstlisting}
-Finally, we lift this to formulas by using \lstinline!atoms! again.
-\begin{lstlisting}
-definition subterms :: 'a pset_fm $\Rightarrow$ 'a pset_term set where
-  subterms $\phi$ $\equiv$ $\bigcup$(subterms ` atoms $\phi$)
-\end{lstlisting}
-
-We also consider the subformulas of a formula and define a function that computes them.
-\begin{lstlisting}
-fun subfms :: 'a fm $\Rightarrow$ 'a fm set where
-  subfms (&\fmAtom& a) = {&\fmAtom& a}
-| subfms (&\fmAnd{p}{q}&) = {&\fmAnd{p}{q}&} $\cup$ subfms p $\cup$ subfms q
-| subfms (&\fmOr{p}{q}&) = {&\fmOr{p}{q}&} $\cup$ subfms p $\cup$ subfms q
-| subfms (&\fmNeg{q}&) = {&\fmNeg{q}&} $\cup$ subfms q
-\end{lstlisting}
-
-Finally, we lift the functions \lstinline!vars! and \lstinline!subterms! to branches.
-\begin{lstlisting}
-definition vars :: 'a branch $\Rightarrow$ 'a set 
-  vars b $\equiv$ $\bigcup$(vars ` set b)
-
-definition subterms :: 'a branch $\Rightarrow$ 'a pset_term set where
-  subterms b $\equiv$ $\bigcup$(subterms ` set b)
-\end{lstlisting}
-
-\clearpage
-
-\section{Definition of the Interpretation Functions\label{appendix:semantics}}
-
-We define the interpretation function \lstinline!&\Ist&! for set atoms.
-\begin{lstlisting}
-fun &\Ist& :: ('a $\Rightarrow$ hf) $\Rightarrow$ 'a pset_term $\Rightarrow$ hf where
-  &\Ist& M $\emptyset$ = 0
-| &\Ist& M (Var x) = M x
-| &\Ist& M (Single s) = HF {&\Ist& M s}
-| &\Ist& M (s $\unionS$ t) = &\Ist& M s $\sqcup$ &\Ist& M t 
-| &\Ist& M (s $\interS$ t) = &\Ist& M s $\sqcap$ &\Ist& M t 
-| &\Ist& M (s $\diffS$ t) = &\Ist& M s $-$ &\Ist& M t 
-\end{lstlisting}
-The interpretation function \lstinline!I$_\text{sa}$! for set atoms is straightforward as well.
-\begin{lstlisting}
-fun &\Isa& :: ('a $\Rightarrow$ hf) $\Rightarrow$ 'a pset_atom $\Rightarrow$ bool where
-  &\Isa& M (s $\inS$ t) $\longleftrightarrow$ &\Ist& M s $\hfmem$ &\Ist& M t
-| &\Isa& M (s $\eqS$ t) $\longleftrightarrow$ &\Ist& M s = &\Ist& M t
-\end{lstlisting}
-
-\section{Upper-Bounding the Cardinality of a Branch\label{appendix:bound}}
-The argument below argues about the \emph{literals} of a branch, which are defined as follows.
-\begin{lstlisting}
-fun is_literal :: 'a fm $\Rightarrow$ bool where
-  is_literal (&\fmAtom& _) = True | is_literal (&\fmNeg{(\fmAtom{} \_)}&) = True
-| is_literal _ = False
-\end{lstlisting}
-
-We still need to give a termination proof to ascertain total correctness.
-The core of our argument is that the number of distinct formulas that can be derived with the expansion rules is finite.
-In particular, we claim that the following upper bound holds.
-\begin{lstlisting}
-  &\phantom{$\leq$}& |set b|
-  &\textover{$=$}{$\leq$}& |{$\psi$ $\in$ set b. $\neg$ is_literal $\psi$}| +
-  &\phantom{$\leq$}& |{$\psi$ $\in$ set b. is_literal $\psi$}|
-  $\leq$ 2 * |subfms (last b)| + |{$\psi$ $\in$ set b. is_literal $\psi$}|
-  $\leq$ 2 * |subfms (last b)| + 16 * |subterms (last b)|$^4$
-\end{lstlisting}
-We treat both inequalities separately, starting with the first one whose justification hinges on an invariant: 
-every new composite subformula that is introduced by an expansion rule is the negation of a subformula that already occurs in \lstinline!last b!.
-Written as a formula, we have
-\begin{lstlisting}
-  &\phantom{$\subseteq$}& {$\psi$ $\in$ set b. $\neg$ is_literal $\psi$}
-  $\subseteq$ subfms (last b) $\cup$ $\fmNegSymbol$ ` subfms (last b).
-\end{lstlisting}
-Confirming that this is an invariant is an easy exercise of going through each of the propositional rules in \cref{tab:lexpands} and the first two branching rules in \cref{tab:bexpands}.
-Thus, we have
-\begin{lstlisting}
-  &\phantom{$\leq$}& |{$\psi$ $\in$ set b. $\neg$ is_literal $\psi$}|
-  $\leq$ |subfms (last b)| + |$\fmNegSymbol$ ` subfms (last b)|
-  $\leq$ 2 * |subfms (last b)|
-\end{lstlisting}
-which is the first inequality we asserted.
-
-Concerning the second inequality, first observe that each literal contains an atom, implying that the set of literals is at most twice as large as the set of atoms in a branch. 
-In addition, an atom is formed by combining two subterms occurring in the branch with either \lstinline!$\inS$! or \lstinline!$\eqS$!;
-hence, we have
-\begin{lstlisting}
-  &\phantom{$\leq$}& |{$\psi$ $\in$ set b. is_literal $\psi$}|
-  $\leq$ 2 * |$\bigcup$(atoms ` set b)| $\leq$ 2 * 2 * |subterms b|$^\text{2}$.
-\end{lstlisting}
-Since no expansion rule introduces new subterms, we know the cardinality of the subterms in a branch \lstinline!b!
-is limited by the subterms of \lstinline!last b!
-plus the number of witnesses added to the branch.
-Note that the last branching rule requires \lstinline!s $\neqS$ t! to be in \lstinline!b! for some \lstinline!s,t $\in$ subterms (last b)!.
-The subsumption condition then prohibits that the rule can be applied again for the combination of \lstinline!s!
-and \lstinline!t!, thus yielding an upper bound of \lstinline!|subterms (last b)|$^\text{2}$!
-for the witnesses.
-This allows us to finish the calculation that justifies the second inequality:
-\begin{lstlisting}
-  &\phantom{$\leq$}& 4 * |subterms b|$^\text{2}$
-  $\leq$ 4 * |subterms (last b)| + |wits b|$^\text{2}$
-  $\leq$ 4 * |subterms (last b)| + (|subterms (last b)|$^\text{2}$)$^\text{2}$
-  $\leq$ 4 * 4 * |subterms (last b)|$^\text{4}$.
-\end{lstlisting}
-All in all, we get the following lemma.
-\begin{lstlisting}
-lemma card_wf_branch_ub:
-  assumes wf_branch b
-  shows |set b|
-      $\leq$ 2 * |subfms (last b)| + 16 * |subterms (last b)|$^\text{4}$
-\end{lstlisting}
-
-\clearpage
-\section{Linear Expansion Rules\label{appendix:linear_rules}}
-
-\begin{table*}[h!]
-  \centering
-  \lstset{
-    escapeinside={@}{@}
-  }
-  \newcommand{\ton}{\text{t}_\text{1}}
-  \newcommand{\ttw}{\text{t}_\text{2}}
-  \caption{Linear expansion rules.}
-  \begin{tabular}{>{\centering\arraybackslash}p{0.48\textwidth}m{0.02\textwidth}>{\centering\arraybackslash}p{0.48\textwidth}}
-    \toprule
-    Propositional Rules & & Rules for \lstinline!$\unionS$! \\
-    \cmidrule{1-1}\cmidrule{3-3}
-    \begin{tabular}{p{0.22\textwidth}m{0.06\textwidth}p{0.2\textwidth}}
-      \lstinline!@\fmAnd{p}{q}@! & $\Longrightarrow$ & \lstinline!p, q! \\
-      \lstinline!@\fmNeg{(\fmOr{p}{q})}@! & $\Longrightarrow$ & \lstinline!@\fmNeg{p}@, @\fmNeg{q}@! \\
-      \lstinline!@\fmOr{p}{q}, \fmNeg{p}@! & $\Longrightarrow$ & \lstinline!q! \\
-      \lstinline!@\fmOr{p}{q}, \fmNeg{q}@! & $\Longrightarrow$ & \lstinline!p! \\
-      \lstinline!@\fmNeg{(\fmAnd{p}{q})}@, p! & $\Longrightarrow$ & \lstinline!@\fmNeg{q}@! \\
-      \lstinline!@\fmNeg{(\fmAnd{p}{q})}@, q! & $\Longrightarrow$ & \lstinline!@\fmNeg{p}@! \\
-      \lstinline!@\fmNeg{(\fmNeg{p})}@! & $\Longrightarrow$ & \lstinline!p!\\[6ex]
-    \end{tabular}
-                        & &
-    \begin{tabular}{p{0.22\textwidth}m{0.06\textwidth}p{0.2\textwidth}}
-    \lstinline!s $\notinS$ $\ton$ $\unionS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$!, \newline \lstinline!s $\notinS$ $\ttw$! \\
-    \lstinline!s $\inS$ $\ton$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$ $\unionS$ $\ttw$! \\
-    \lstinline!s $\in$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$ $\unionS$ $\ttw$! \\
-    \lstinline!s $\inS$ $\ton$ $\unionS$ $\ttw$!, \newline \lstinline!s $\notinS$ $\ton$! & ~\newline $\Longrightarrow$ & ~\newline \lstinline!s $\inS$ $\ttw$! \\
-    \lstinline!s $\inS$ $\ton$ $\unionS$ $\ttw$!, \newline \lstinline!s $\notinS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$! \\
-    \lstinline!s $\notinS$ $\ton$, s $\notinS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$ $\unionS$ $\ttw$! \\
-    \end{tabular}
-    \\
-    \\
-
-    Rules for \lstinline!$\interS$! && Rules for \lstinline!$\diffS$! \\
-    \cmidrule{1-1}\cmidrule{3-3}
-    \begin{tabular}{p{0.22\textwidth}m{0.06\textwidth}p{0.2\textwidth}}
-      \lstinline!s $\inS$ $\ton$ $\interS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$!, \newline \lstinline!s $\inS$ $\ttw$! \\
-      \lstinline!s $\notinS$ $\ton$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$ $\interS$ $\ttw$! \\ 
-      \lstinline!s $\notinS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$ $\interS$ $\ttw$! \\
-      \lstinline!s $\notinS$ $\ton$ $\interS$ $\ttw$!, \newline \lstinline!s $\inS$ $\ton$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ttw$! \\ 
-      \lstinline!s $\notinS$ $\ton$ $\interS$ $\ttw$! \newline \lstinline!s $\inS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$! \\ 
-      \lstinline!s $\inS$ $\ton$, s $\inS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$ $\interS$ $\ttw$! \\
-    \end{tabular}
-    & &
-    \begin{tabular}{p{0.22\textwidth}m{0.06\textwidth}p{0.2\textwidth}}
-      \lstinline!s $\inS$ $\ton$ $\diffS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$!, \newline \lstinline!s $\notinS$ $\ttw$! \\
-      \lstinline!s $\notinS$ $\ton$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$ $\diffS$ $\ttw$! \\
-      \lstinline!s $\inS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$ $\diffS$ $\ttw$! \\
-      \lstinline!s $\notinS$ $\ton$ $\diffS$ $\ttw$!, \newline \lstinline!s $\inS$ $\ton$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ttw$! \\
-      \lstinline!s $\notinS$ $\ton$ $\diffS$ $\ttw$!, \newline \lstinline!s $\notinS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\notinS$ $\ton$! \\
-      \lstinline!s $\inS$ $\ton$, s $\notinS$ $\ttw$! & $\Longrightarrow$ & \lstinline!s $\inS$ $\ton$ $\diffS$ $\ttw$! \\
-    \end{tabular}
-    \\
-    \\
-
-    Rules for \lstinline!Single! && Rules for \lstinline!$\eqS$! \\
-    \cmidrule{1-1} \cmidrule{3-3}
-    \begin{tabular}{p{0.22\textwidth}m{0.06\textwidth}p{0.2\textwidth}}
-      & $\Longrightarrow$ & \lstinline!s $\inS$ Single s! \\
-      \lstinline!s $\inS$ Single t! & $\Longrightarrow$ & \lstinline!s $\eqS$ t! \\
-      \lstinline!s $\notinS$ Single t! & $\Longrightarrow$ & \lstinline!s $\neqS$ t! \\
-    \end{tabular}
-    & &
-    \begin{tabular}{p{0.22\textwidth}m{0.06\textwidth}p{0.2\textwidth}}
-      \lstinline!$\ton$ $\eqS$ $\ttw$, l! & $\Longrightarrow$ & \lstinline!l{$\ttw$/$\ton$}! \\
-      \lstinline!$\ton$ $\eqS$ $\ttw$, l! & $\Longrightarrow$ & \lstinline!l{$\ton$/$\ttw$}! \\
-      \lstinline!s$_\text{1}$ $\inS$ t, s$_\text{2}$ $\notinS$ t! & $\Longrightarrow$ & \lstinline!s$_\text{1}$ $\neqS$ s$_\text{2}$! \\
-    \end{tabular}
-    \\
-    \bottomrule
-  \end{tabular}
-\end{table*}
-
-\clearpage
-\section{Branching Expansion Rules\label{appendix:branching_rules}}
-
-\begin{table*}[h]
-  \lstset{
-    escapeinside={@}{@}
-  }
-  \centering
-  \newcommand{\ton}{\text{t}_\text{1}}
-  \newcommand{\ttw}{\text{t}_\text{2}}
-  \newcommand{\lpipe}{\rule[-0.4ex]{0.6pt}{2.3ex}}
-  \caption{Branching expansion rules. We write \lstinline!$\phi$! for \lstinline!last b! here.}
-  \begin{tabular}{>{\centering\arraybackslash}p{0.34\textwidth} >{\centering\arraybackslash}p{0.32\textwidth} >{\centering\arraybackslash}p{0.3\textwidth}}
-    \toprule
-    Rule & Precondition & Subsumption condition \\
-    \midrule
-    \begin{tabular}{@{}>{\raggedleft\arraybackslash}p{0.11\textwidth}c>{\raggedright\arraybackslash}p{0.11\textwidth}@{}}
-      \\
-      \midrule
-      \lstinline!p! & \lpipe & \lstinline!@\fmNeg{p}@! \\
-    \end{tabular}
-         & \lstinline!@\fmOr{p}{q}@ $\in$ set b! &
-         \begin{tabular}{rl}
-         & \lstinline!p $\in$ set b $\lor$! \\
-         $\fmNegSymbol$ & \lstinline!p $\in$ set b! \\
-         \end{tabular} \\
-
-    \begin{tabular}{@{}>{\raggedleft\arraybackslash}p{0.11\textwidth}c>{\raggedright\arraybackslash}p{0.11\textwidth}@{}}
-      \\
-      \midrule
-      \lstinline!@\fmNeg{p}@! & \lpipe & \lstinline!p! \\
-    \end{tabular}
-         & \lstinline!@\fmNeg{(\fmAnd{p}{q})}@ $\in$ set b! &
-          \begin{tabular}{rl}
-            & \lstinline!p $\in$ set b $\lor$! \\
-            $\fmNegSymbol$ & \lstinline!p $\in$ set b! \\
-          \end{tabular} \\
-
-    \begin{tabular}{@{}>{\raggedleft\arraybackslash}p{0.11\textwidth}c>{\raggedright\arraybackslash}p{0.11\textwidth}@{}}
-      \\
-      \midrule
-      \lstinline!s $\inS$ $\ton$! & \lpipe & \lstinline!s $\notinS$ $\ton$! \\
-    \end{tabular}
-         &
-         \begin{tabular}{c}
-           \lstinline!(s $\inS$ $\ton$ $\unionS$ $\ttw$) $\in$ set  b! \\
-           \lstinline!$\ton$ $\unionS$ $\ttw$ $\in$ subterms $\phi$! \\
-         \end{tabular}
-         &
-         \begin{tabular}{@{}rl@{}}
-                  & \lstinline!(s $\inS$ $\ton$) $\in$ set b! \\
-           $\lor$ & \lstinline!(s $\notinS$ $\ton$) $\in$ set b!
-         \end{tabular}
-         \\
-
-    \begin{tabular}{@{}>{\raggedleft\arraybackslash}p{0.11\textwidth}c>{\raggedright\arraybackslash}p{0.11\textwidth}@{}}
-      \\
-      \midrule
-      \lstinline!s $\inS$ $\ttw$! & \lpipe & \lstinline!s $\notinS$ $\ttw$! \\
-    \end{tabular}
-         &
-         \begin{tabular}{c}
-           \lstinline!(s $\inS$ $\ton$) $\in$ set  b! \\
-           \lstinline!($\ton$ $\interS$ $\ttw$) $\in$ subterms $\phi$! \\
-         \end{tabular}
-         &
-         \begin{tabular}{@{}rl@{}}
-                  & \lstinline!(s $\inS$ $\ttw$) $\in$ set b! \\
-           $\lor$ & \lstinline!(s $\notinS$ $\ttw$) $\in$ set b! \\
-        \end{tabular}
-        \\
-
-    \begin{tabular}{@{}>{\raggedleft\arraybackslash}p{0.11\textwidth}c>{\raggedright\arraybackslash}p{0.11\textwidth}@{}}
-      \\
-      \midrule
-      \lstinline!s $\inS$ $\ttw$! & \lpipe & \lstinline!s $\notinS$ $\ttw$! \\
-    \end{tabular}
-         &
-         \begin{tabular}{c}
-           \lstinline!(s $\inS$ $\ton$) $\in$ set  b! \\
-           \lstinline!($\ton$ $\diffS$ $\ttw$) $\in$ subterms $\phi$! \\
-         \end{tabular}
-         &
-         \begin{tabular}{@{}rl@{}}
-                  & \lstinline!(s $\inS$ $\ttw$) $\in$ set b! \\
-           $\lor$ & \lstinline!(s $\notinS$ $\ttw$) $\in$ set b! \\
-         \end{tabular} \\
-         \addlinespace[0.6em]
-
-         \begin{tabular}{@{}>{\raggedleft\arraybackslash}p{0.165\textwidth}c>{\raggedright\arraybackslash}p{0.165\textwidth}@{}}
-      \\
-      \midrule
-      \lstinline!Var x $\inS$ $\ton$! \lstinline!Var x $\notinS$ $\ttw$! & \multirow{2}{*}{\rule[-0.4ex]{0.6pt}{4.8ex}} & \lstinline!Var x $\notinS$ $\ton$! \lstinline!Var x $\inS$ $\ttw$! \\
-    \end{tabular}
-         & 
-         \begin{tabular}{c}
-           \lstinline!($\ton$ $\neqS$ $\ttw$) $\in$ set b! \\
-           \lstinline!$\ton$ $\in$ subterms $\phi$! \\
-           \lstinline!$\ttw$ $\in$ subterms $\phi$! \\
-           \lstinline!x $\notin$ vars b! \\
-         \end{tabular}
-         &
-         \begin{tabular}{r@{\hspace{0.5em}}l}
-
-           \lstinline!$\exists$s.! & \lstinline!(s $\inS$ $\ton$) $\in$ set b! \\
-               \lstinline!$\land$! & \lstinline!(s $\notinS$ $\ttw$) $\in$ set b! \\
-           \multicolumn{2}{c}{$\lor$} \\
-           \lstinline!$\exists$s.! & \lstinline!(s $\notinS$ $\ton$) $\in$ set b! \\
-               \lstinline!$\land$! & \lstinline!(s $\inS$ $\ttw$) $\in$ set b! \\
-         \end{tabular}
-         \\
-         \bottomrule
-  \end{tabular}
-\end{table*}
+\clearpage\listoftodos
 

### 2023-05-24 paper/CADE/paper.tex
Grammarly

@@
 
-The focus of this paper is another ubiquitous structure in mathematics, namely sets.
+This paper focuses on another ubiquitous structure in mathematics, namely sets.
 To the best of our knowledge, we present the first formally verified decision procedure for (a fragment of) set theory.
@@
 
-Since \MLSS{} admits a tableau calculus, the generation of certificates will be straightforward.
-Like with the aforementioned order solver, this paves the way towards an integration of the decision procedure into Isabelle, adding to its growing body of verified decision procedures.
+Since \MLSS{} admits a tableau calculus, generating certificates will be straightforward.
+Like with the aforementioned order solver, this paves the way for an integration of the decision procedure into Isabelle, adding to its growing body of verified decision procedures.
 

### 2023-05-24 paper/CADE/paper.tex
Finish paper

@@
 \begin{itemize}
-  \item We deliver the omitted proof of Lemma~2 in the paper~\cite{new_fast_tableau}, which is a key building block for the completeness proof of the calculus.
-  \item The formal proof of completeness leads to the discovery that the calculus was missing a rule for eliminating double negation. 
-  \item We derive an explicit upper bound for the number of formulas in a branch of the tableau.
+  \item We deliver the omitted proof of Lemma~2 in the paper~\cite{new_fast_tableau}, a key building block for the completeness proof of the calculus.
+  \item The formal proof of completeness reveals that the calculus lacks a rule for eliminating double negation. 
+  \item We derive an explicit upper bound for the number of formulas in a tableau branch.
 \end{itemize}
 
-In the context of Isabelle/HOL, there is one important aspect that requires us to modify the calculus in the paper:
+In the context of Isabelle/HOL, there is one crucial aspect that requires us to modify the calculus in the paper:
 the calculus works under the assumption that every variable is a set;
@@
 We call these variables urelements.
-In order to deal with them, we extend the calculus with a lightweight type system and a verified inference algorithm that identifies the urelements.
+To deal with them, we extend the calculus with a lightweight type system and a verified inference algorithm that identifies the urelements.
 
@@
 Since the literature on decidable fragments of set theory is vast, we only focus on \MLSS{} here.
-The fragment was first shown to be decidable by \citeauthor{mlss_first}~\cite{mlss_first}.
+\citeauthor{mlss_first}~\cite{mlss_first} were the first to show the decidability of the fragment.
 Subsequent work~\cite{mlss_np} found the decision problem to be \textbf{NP}-complete.
 To obtain a practical decision procedure, \citeauthor{mlss_first_tableau}~\cite{mlss_first_tableau} proposed a tableau calculus, which was later improved by \citeauthor{tableau_quantifier_free}~\cite{tableau_quantifier_free}.
-Both of these procedures construct a model during execution that is used to guide the proof search.
-\citeauthor{tableau_quantifier_free} also cover an extension of the calculus with uninterpreted functions, which was revisited by \citeauthor{mlss_quantification}~\cite{mlss_quantification} who avoided the construction of a model during execution.
-In this paper, we consider a version of the latter procedure due to \citeauthor{new_fast_tableau}~\cite{new_fast_tableau} that is specialised to \MLSS{} and where the branching rules of the calculus are set up in a way that guarantees mutual exclusivity of the branches.
-Later extensions of the calculus added certain interpreted functions such as monotone functions~\cite{mlss_monotone_functions} and the inverse of a function~\cite{mlss_cartesian_map}.
+Both of these procedures construct a model during execution that guides the proof search.
+\citeauthor{tableau_quantifier_free} also cover an extension of the calculus with uninterpreted functions, which \citeauthor{mlss_quantification}~\cite{mlss_quantification} later revisited while avoiding the construction of a model during execution.
+In this paper, we consider a version of the latter procedure due to \citeauthor{new_fast_tableau}~\cite{new_fast_tableau} that is specialised to \MLSS{} and where the branching rules of the calculus are set up to guarantee the mutual exclusivity of the branches.
+Later extensions of the calculus added certain interpreted functions, such as monotone functions~\cite{mlss_monotone_functions} and the inverse of a function~\cite{mlss_cartesian_map}.
 The latter extension notably includes the Cartesian product.
@@
 
-There is a large body of work at the intersection of ITPs and tableau methods but to keep with the theme of the paper we only consider formalisations of correctness here.
+There is a large body of work at the intersection of ITPs and tableau methods, but to keep with this paper's theme we only consider formalisations of correctness here.
 For first-order logic, there are abstract completeness proofs using the \emph{Beth-Hintikka style} of possibly infinite derivation trees~\cite{completeness_coinductive} as well as the \emph{Henkin style} of maximally consistent sets~\cite{synthetic_completeness_afp}.
@@
 
-Beyond completeness, we target decidability which is more attainable for propositional logic.  
+Beyond completeness, we target decidability, which is more attainable for propositional logic.  
 There is a verified tableau calculus for the modal logic S5~\cite{lean_modal_logic} in Lean and one for hybrid logic~\cite{hybrid_logic} in Isabelle/HOL.
-Both of these formalisation do not prove termination but there is a formalisation of a tableau calculus for the temporal logic CTL in Coq~\cite{ctl_tableau} that does.
+Both of these do not prove termination but there is a formalisation of a tableau calculus for the temporal logic CTL in Coq~\cite{ctl_tableau} that does.
 % TODO: Formalisation de la logique temporelle dans le coq.
@@
 \subsection{Syntax}
-At the heart of \MLSS{}, we have the type of set terms which is the disjoint union of the empty set and variables as well as the operations union, intersection, set difference, and the singleton set represented by the constructor \lstinline!Single!.
+At the heart of \MLSS{}, we have the type of set terms, which is the disjoint union of the empty set and variables as well as the operations union, intersection, difference, and the singleton set represented by the constructor \lstinline!Single!.
 We keep the type of variables abstract by making it a parameter of the set term data type.
@@
 
-Similarly to \lstinline!vars!, we get the function \lstinline!atoms :: 'a fm $\Rightarrow$ 'a set! for free which retrieves all set atoms in a formula.
+Similarly to \lstinline!vars!, we get the function \lstinline!atoms :: 'a fm $\Rightarrow$ 'a set! for free that retrieves all set atoms in a formula.
 We combine these functions to extract all the variables occurring in a set formula.
@@
 The HF sets, as we will see, are sufficient to construct a model for any satisfiable \MLSS{} formula.
-In contrast to $\mathcal{V}$, the HF sets are directly representable in Isabelle/HOL, and indeed, there is an AFP entry~\cite{hf_AFP} that formalises them.
+In contrast to $\mathcal{V}$, the HF sets are directly representable in Isabelle/HOL, and indeed, an AFP entry~\cite{hf_AFP} formalises them.
 The entry defines a type \lstinline!hf! that comes with the following functionality:
@@
 in the standard way, i.e.\ by mapping each syntactic construct to the corresponding operation on HF sets and interpreting variables with respect to a given valuation function \lstinline!M :: 'a $\Rightarrow$ hf!.
-For the concrete definition we refer to formalisation.
+For the concrete definition we refer to the formalisation.
 
@@
 The implementation of $\models$ coincides with the interpretation function of \citeauthor{lqe}~\cite{lqe}.
-As usual, a formula \lstinline!$\phi$! is called \emph{satisfiable} if there exists a model \lstinline!M! with \lstinline!M $\models$ $\phi$!.
-Otherwise, \lstinline!$\phi$! is called \emph{unsatisfiable}.
+As usual, we call a formula \lstinline!$\phi$! \emph{satisfiable} if there exists a model \lstinline!M! with \lstinline!M $\models$ $\phi$!.
+Otherwise, we say that \lstinline!$\phi$! is \emph{unsatisfiable}.
 
@@
 We formalise the tableau calculus for \MLSS{} as described by \citeauthor{new_fast_tableau}~\cite{new_fast_tableau}.
-Inspired by the formalisation of a tableau calculus for hybrid logic by \citeauthor{hybrid_logic_afp}~\cite{hybrid_logic_afp}, we simply use lists to represent the branches of the tableau tree.
-Note that formulas are added to the front of the list during branch expansion, so \lstinline!last b! for a branch \lstinline!b! is always the formula that we are trying to disprove with the tableau.
+Inspired by the formalisation of a tableau calculus for hybrid logic by \citeauthor{hybrid_logic_afp}~\cite{hybrid_logic_afp}, we use lists to represent the branches of the tableau tree.
+Note that we add formulas to the front of the list during branch expansion, so \lstinline!last b! for a branch \lstinline!b! is always the formula we are trying to disprove with the tableau.
 We sometimes call this formula the \emph{initial formula}.
@@
 
-\noindent The functions \lstinline!vars! and \lstinline!subterms! are lifted to branches in the expected way.
+\noindent We lift the functions \lstinline!vars! and \lstinline!subterms! to branches in the expected way.
 
@@
 
-Note that even branching rules are defined such that they never create new subterms with exception of the last rule, which adds a new variable to the branch.
+Note that even branching rules are defined such that they never create new subterms, except for the last rule that adds a new variable to the branch.
 These variables serve to manifest an inequality; hence, we call them \emph{witnesses}.
@@
 This requires us to prove that the assumptions of the equations are non-overlapping, which is done by automation.
-The other concern is that Isabelle/HOL requires functions to be total, so a recursive function needs to terminate in order for it to be well-defined;
+The other concern is that Isabelle/HOL requires functions to be total, so a recursive function needs to terminate for it to be well-defined;
 nevertheless, the termination proof is separated from the definition of the function for modularity.
@@
 Each equation of the function is guarded by an assumption that the predicate holds for the argument.
-In \cref{sec:correct}, we will show that the domain predicate holds for the context that the function \lstinline!mlss_proc_branch! is called in.
-Before we go into more detail on how the termination is proved, we discuss the definition of the function as shown below.
+In \cref{sec:correct}, we will show that the domain predicate holds for the context in which the function \lstinline!mlss_proc_branch! is called in.
+Before we go into more detail on how the termination is proved, we discuss the definition of the function, as shown below.
 
@@
 
-The purpose of the function is to determine whether a given branch can be expanded to a closed tableau.
+The purpose of the function is to determine whether we can expand a given branch to a closed tableau.
 As stated before, we first use linear expansion rules in order to prevent premature branching;
@@
 is actually parametrised by choice functions to allow for refinement.} to choose some rule that actually adds new formulas to the branch.
-As soon as the branch is linearly saturated, we terminate if the branch is closed as shown in the second equation.
+As soon as the branch is linearly saturated, we terminate if the branch is closed as the second equation shows.
 Otherwise, we choose an applicable branching rule and recursively check whether all newly created branches can be closed.
@@
 
-This means that \lstinline!mlss_proc_branch! is only applied to branches that result from applying the expansion rules.
-We call this kind of branches \emph{well-formed}.
+Thus, we use \lstinline!mlss_proc_branch! is only on branches that result from applying the expansion rules.
+We call this kind of branch \emph{well-formed}.
 In the definition below, the expression \lstinline!&\expandss{b'}{b}&!
@@
 
-In particular, we use this notion in \cref{sec:correct} to state an upper bound for the cardinality of well-formed branches.
+We use this notion in \cref{sec:correct} to state an upper bound for the cardinality of well-formed branches.
 The upper bound justifies the termination of the decision procedure. 
 Before we come to that, though, we prove soundness and completeness in Section~\ref{sec:sound} and~\ref{sec:complete}, respectively.
-In \cref{sec:correct}, we also show that both properties easily transfer to \lstinline!mlss_proc! which, together with termination, establishes that it is a decision procedure.
+In \cref{sec:correct}, we also show that both properties easily transfer to \lstinline!mlss_proc!, which, together with termination, establishes that it is a decision procedure.
 
@@
 A subset of the witnesses, which we call \emph{pure} witnesses, receives special treatment from the realisation function for reasons that will become apparent in \cref{sec:pwits}.
-The collection of set terms of a branch can thus be partitioned into two collections as defined below.
+The collection of set terms of a branch can thus be partitioned into two collections, as defined below.
 \begin{lstlisting}
@@
 
-We aim to construct a syntactic model that is derived from the membership literals \lstinline!s $\inS$ t! in the branch.
+We aim to construct a syntactic model that we derive from the membership literals \lstinline!s $\inS$ t! in the branch.
 To this end, we construct a graph whose vertices are the disjoint union of the sets above and there is an edge from \lstinline!s!
@@
 is in \lstinline!b!.
-Note that we use \citeauthor{graph_theory_afp}'s graph library~\cite{graph_theory_afp} that represents a graph as a record of vertices, arcs (directed edges), and two functions \lstinline!tail!
+Note that we use \citeauthor{graph_theory_afp}'s graph library~\cite{graph_theory_afp} which represents a graph as a record of vertices, arcs (directed edges), and two functions \lstinline!tail!
 and \lstinline!head!
@@
 The realisation function is defined relative to this graph.
-As mentioned before, the pure witnesses are treated differently than the rest of the set terms by the realisation function.
-Terms in the latter set are evaluated in accordance to the structure of the graph, i.e.\ the realisation of a vertex is defined as the union of the realisations of the parent vertices. 
+As mentioned before, the realisation function treats the pure witnesses differently than the rest of the set terms.
+The function evaluates terms in the latter set in accordance to the structure of the graph, i.e.\ the realisation of a vertex is defined as the union of the realisations of the parent vertices. 
 For the former set, we choose a function \lstinline!I! that assigns the pure witnesses pairwise distinct sets with cardinality greater than that of the vertices.
@@
 The former is taken care of by automation, leaving us to prove termination.
-The assumption that \lstinline!b! is open implies that there are no membership cycles and thus \lstinline!bgraph b! is acyclic.
+The assumption that \lstinline!b! is open implies that there are no membership cycles, thus \lstinline!bgraph b! is acyclic.
 Furthermore, the graph is finite by definition.
@@
 Recall that the pure witnesses of a branch \lstinline!b! are those witnesses that are not related to other subterms in \lstinline!last b! by equality.
-In the context of a well-formed branch, this characterisation can be strengthened to any set term and, in addition, we also get that there is no membership literal where a pure witness is on the right-hand side.
+In the context of a well-formed branch, we can strengthen this characterisation to any set term and, in addition, we also get that there is no membership literal where a pure witness is on the right-hand side.
 Intuitively speaking, the realisation of a pure witness does not depend on the realisation of any other set term.
@@
 So why are pure witnesses treated differently?
-According to the definition of \lstinline!realise!, the pure witnesses would be evaluated to the empty set \lstinline!0 :: hf!, were they not treated separately. 
+According to the definition of \lstinline!realise!, it would evaluate the pure witnesses would to the empty set \lstinline!0 :: hf!, were they not treated separately. 
 To see that this is a problem, consider the branch
@@
 
-\noindent The results on literals can now be lowered to set terms.
+\noindent We now lower the results on literals to set terms.
 All of the proofs are straightforward so we refer the reader to the formalisation.
@@
 
-The final step to obtain a proper model is to connect the realisation function to the semantics as defined in \cref{sec:semantics}.
+The final step for obtaining a proper model is to connect the realisation function to the semantics as defined in \cref{sec:semantics}.
 For set terms, we can use the Propositions~\ref{it:empty}--\ref{it:single} to prove the lemma below by induction on \lstinline!t!.
@@
 \end{lstlisting}
-Lifting the above result to formulas yields the coherence of \lstinline!b! as the original paper~\cite{new_fast_tableau} calls it.
+Lifting the above result to formulas yields the coherence of \lstinline!b!, as the original paper~\cite{new_fast_tableau} calls it.
 The proof is a tedious but straightforward induction on the the size of the formulas.
@@
 \section{Soundness of the Calculus\label[section]{sec:sound}}
-A tableau calculus is sound if for any closed tableau, the corresponding formula is unsatisfiable. 
+A tableau calculus is sound if the corresponding formula is unsatisfiable for any closed tableau. 
 We prove the following two properties to establish soundness:
-\begin{enumerate*}[label={(\arabic*)}]
-  \item It is not possible to satisfy all formulas in a closed branch at once. 
+\begin{enumerate}[label={(\arabic*)}]
+  \item It is impossible to satisfy all formulas in a closed branch simultaneously. 
   \item The expansion rules maintain satisfiability.
-\end{enumerate*}
+\end{enumerate}
 
@@
   Since \lstinline!s $\neqS$ t! is in \lstinline!b!, we have that \lstinline!&\Ist& M s $\neq$ &\Ist& M t! because \lstinline!M! is a model.
-  Without loss of generality this inequality manifests itself through some \lstinline!y! with
+  Without loss of generality, this inequality manifests itself through some \lstinline!y! with
   \lstinline!y $\hfmem$ &\Ist& M s! and \lstinline!y $\hfnotmem$ &\Ist& M t!.
-  We update \lstinline!M! such that it maps \lstinline!x! to \lstinline!y! to obtain the assignment \lstinline!M'!.
+  We update \lstinline!M! to map \lstinline!x! to \lstinline!y! to obtain the assignment \lstinline!M'!.
   Note that \lstinline!M'! is still a model for formulas in \lstinline!b! because \lstinline!x! is fresh with respect to \lstinline!b!.
@@
 
-To finish off the proof of total correctness, note that every singleton branch is trivially well-formed; thus, termination, completeness, and soundness easily transfer to \lstinline!mlss_proc!.
+To finish the proof of total correctness, note that every singleton branch is trivially well-formed; thus, termination, completeness, and soundness easily transfer to \lstinline!mlss_proc!.
 \begin{lstlisting}
@@
 In the introduction, we stated the goal of integrating \lstinline!mlss_proc! as a tactic into Isabelle.
-For this to work, we need to map every branch expansion rule to a corresponding theorem in Isabelle/HOL.
+For this to work, we must map every branch expansion rule to a corresponding theorem in Isabelle/HOL.
 This is straightforward for all expansion rules except for the last branching expansion rule.
@@
 in Isabelle/HOL.
-By way of reification, we convert this to a formula of the shape \lstinline!s' $\neqS$ t' $\fmAndSymbol$ s' $\inS$ A' $\unionS$ B' $\fmAndSymbol$ $\ldots$!
+By way of reification, we convert this to a formula of the shape
+\begin{lstlisting}
+ s' $\neqS$ t' $\fmAndSymbol$ s' $\inS$ A' $\unionS$ B' $\fmAndSymbol$ $\ldots$
+\end{lstlisting}
 in our set syntax for some \lstinline!s'!, \lstinline!t'!, \lstinline!A'!, and \lstinline!B'!.
-When we apply the decision procedure to this formula, we might get back a tableau proof that contains an application of the last branching rule to \lstinline!(s' $\neqS$ t') $\in$ set b!.
+When we apply the decision procedure to this formula, it might return a tableau proof that contains an application of the last branching rule to \lstinline!(s' $\neqS$ t') $\in$ set b!.
 This results in two branches, one of which is \lstinline![Var x $\inS$ s', Var x $\notinS$ t'] @ b!;
@@
 for some \lstinline!'a!.
-Note that the constructor $\emptyset$ now receives an additional argument that indicates the level of each instance of $\emptyset$.
+Note that the constructor $\emptyset$ now receives an additional argument indicating the level of each instance of $\emptyset$.
 
@@
   \item As mentioned above, we add an argument to the $\emptyset$ constructor.
-    This argument is only used for the typing judgement; it has no impact on the semantics.
+    This argument is only used for the typing judgement; it does not impact on the semantics.
 \end{enumerate*}
@@
 \begin{enumerate*}[label=(\arabic*)]
-  \item The first equation of \lstinline!realise! now also needs to account for the urelements.
-    In particular, it needs to ensure that urelements receive pairwise different values unless they are related through equality atoms. 
-    This has no effect on pure witnesses since they can not be related through equality atoms due to \reflemmatwo{}.
-  \item We need to adjust the completeness proof in those places where it directly refers to the definition of \lstinline!realise! to account for the case where a given term is an urelement.
+  \item The first equation of \lstinline!realise! now also must account for the urelements.
+    In particular, it has to ensure that urelements receive pairwise different values unless they are related through equality atoms. 
+    This does not affect pure witnesses since they can not be related through equality atoms due to \reflemmatwo{}.
+  \item We must adjust the completeness proof in those places where it directly refers to the definition of \lstinline!realise! to account for the case where a given term is an urelement.
   \item The completeness theorem receives the additional assumption that \lstinline!$\Gamma$ $\vdash$ $\phi$! holds for the initial formula $\phi$. 
-  \item For the completeness proof, we need to show that the typing judgement is invariant under branch expansion.
+  \item For the completeness proof, we must show that the typing judgement is invariant under branch expansion.
 \end{enumerate*}
@@
 The modifications above ensure that the proof can be replayed through Isabelle/HOL.
-To actually use the calculus, we need to determine the urelements of the initial formula $\phi$, though.
-In other words, we need to implement an inference algorithm for our lightweight type system.
+To actually use the calculus, we must determine the urelements of the initial formula $\phi$, though.
+In other words, we have to implement an inference algorithm for our lightweight type system.
 The algorithm is, in essence, a simplified version of Hindley-Milney type inference so it has the same two phases:
@@
 
-Since we are only interested in the level of a term, we can encode all constraints into the theory consisting of $0$, the successor function \lstinline!S!, and equality (but no disequality).
+Since we are only interested in the level of a term, we can encode all constraints into the theory of $0$, the successor function \lstinline!S!, and equality (but no disequality).
 Note that constraints of the form \lstinline!l $\neq$ 0! can be replaced by \lstinline!l $=$ S i! with \lstinline!i! being a fresh variable.
@@
 nevertheless, we have to be careful that it computes the minimum assignment $\Gamma$ from variables to levels that fulfils the constraints. 
-This is to guarantee that a set term \lstinline!t! is not an urelement if, and only if, \lstinline!$\Gamma$ t $>$ 0!.
+This guarantees that a set term \lstinline!t! is not an urelement if, and only if, \lstinline!$\Gamma$ t $>$ 0!.
 Conversely, all terms \lstinline!s! with \lstinline!$\Gamma$ s $=$ 0! are urelements.
@@
 In future work, we plan to implement an efficient executable specification in the style of a worklist algorithm. 
-This specification should also generate certificates that can be replayed through Isabelle's inference kernel in order to facilitate the integration of the procedure into Isabelle.
+This specification should also generate certificates that can be replayed through Isabelle's inference kernel to facilitate the integration of the procedure into Isabelle.
 
@@
 \subsubsection{Acknowledgements}
-The author thanks Kevin Kappelmann, Tobias Nipkow, and the anonymous referees for their comments on a draft version of this paper.
+The author thanks Kevin Kappelmann and Tobias Nipkow for their comments on a draft version of this paper and the anonymous referees for their thorough reviews.
 
@@
 
-\clearpage\listoftodos
+%\clearpage\listoftodos
 
